### 2019-01-22 diagram_NAR.tex
First commit

%\documentclass[tikz]{standalone}

\begin{tikzpicture}[node distance=2cm]

\node (react2) [startend] {Condition 2};
\node (react1) [startend, left = 10pt of react2] {Condition1};
\node (rsup) [startend, right = 10pt of react2] {...};

\node (suu)[above=5pt of react2,IO]{\relsize{+1}\sf Reactivities};

\node(sampling)[Bprocess, below = 20pt of react2]{Stochastic Structural Sampling};
\node (seq)[startend, left =1cm of sampling]{RNA sequence};

\node[IO]  at (seq|-react1) {\sf\bfseries \relsize{+2}Input};

\draw  [arrow] (seq) to (sampling);

\node (sample2) [startend, below= 15pt of sampling] {Sample 2};
\node (sample1) [startend, left = 10pt of sample2] {Sample 1};
\node (samplex) [startend, right = 10pt of sample2] {...};

\node(Merge)[Vprocess, below = 15pt of sample2, ]{{Merge samples}};

\node(clustering)[Bprocess, below = 15pt of Merge]{Clustering with k clusters};
\draw[arrow](Merge) -- (clustering);
\node(dec1) [decision, below=15pt of clustering]{Stopping \\ Criterion};
\draw[arrow] (clustering) to (dec1);
\node(optimaln)[startend, below of = dec1]{Centroid Structures};

\node(n+)[VSprocess, left of = dec1,xshift=-3cm]{\rotatebox{90}{k++}};
\draw[arrow] (n+)  |- (clustering);
\node(dec2) [decision, below=15pt of optimaln]{Pareto\\Optimal?};
\draw[arrow](optimaln) to (dec2);
\node(nnoptim)[VSprocess,left of = dec2, xshift=-3cm ]{\rotatebox{90}{Reject centroid}};

\node(optimalcentroid)[startend, below of = dec2]{Optimal centroid(s)};

\node(output)[right of =optimalcentroid, IO] (outcap) {\sf\bfseries \relsize{+2}Output};
\node(vert)[right of = optimalcentroid, xshift=1cm ]{};

\draw[arrow] (dec1) -- node[anchor=north,arrowcaption] {no} (n+);
\draw[arrow] (dec1) -- node [anchor=west,arrowcaption]{yes} (optimaln);

\draw[arrow] (dec2) -- node[anchor=north,arrowcaption] {no} (nnoptim);
\draw[arrow] (dec2) -- node [anchor=west,arrowcaption]{yes} (optimalcentroid);

%\draw[arrow](Cbz) |- (dec2);

\draw  [arrow] (react1) -- (sampling.north-|react1);
\draw  [arrow] (react2)-- (sampling.north-|react2);
\draw  [arrow] (rsup) -- (sampling.north-|rsup);

\draw[arrow] (sampling.south-|sample1) -- (sample1);
\draw[arrow] (sampling.south-|sample2) -- (sample2);
\draw[arrow] (sampling.south-|samplex) -- (samplex);

\draw[arrow] (sample1) -- (Merge.north-|sample1);
\draw[arrow] (sample2) -- (Merge.north-|sample2);
\draw[arrow] (samplex) -- (Merge.north-|samplex);
%\draw[arrow] (sample1) -- ++(-3cm,0);
%\draw[arrow] (sample2) -- ++(-3cm,0);
%\draw[arrow] (samplei) -- ++(-3cm,0);
%\draw[arrow] (samplex) -- ++(-3cm,0);

\begin{pgfonlayer}{background}
\node[fit=(seq.west|-react1) (react1) (react2) (rsup) (suu), input] (bginput1) {};
\node[fit=(seq)(seq.west|-react1), input] (bginput2) {};
\node[fit=(outcap) (optimalcentroid), output] (bgoutput) {};
\end{pgfonlayer}

\end{tikzpicture}
%\end{document}

### 2019-01-22 graphs/compareSS.py

@@ -1,0 +1,48 @@
+
+
+def parseSS(s):
+    res = []
+    p = []
+    n = len(s)
+    for i,c in enumerate(s):
+        if c =="(":
+            p.append(i)
+        elif c ==")":
+            j = p.pop()
+            res.append((j,i))
+    return res,n
+
+
+def compareSS(s,s_base):
+    (struct,n) = parseSS(s)
+    (base,n) = parseSS(s_base)
+    base = set(base)
+    res = ["." for i in range(n)]
+    for i in range(n):
+        c = s[i]
+        if c == ".":
+            if s_base[i] != ".":
+                res[i] = "\\B{.}"
+            else:
+                res[i] = "\\G{.}"
+    for i,j in struct:
+        if (i,j) not in base:
+            res[i],res[j] = "\\B{(}","\\B{)}"
+        else:
+            res[i],res[j] = "\\G{(}","\\G{)}"
+    return "".join(res)
+
+NATIVE_STRUCT=("NATIVE","((((((((......((((((....)))))).(((....(((.....)))...)))........))))))))........(((((......(((((.....))))).(((....(((....((((....)))).....)))...))).......)))))")
+NO_PROBING_STRUCT=("No probing","(((((((((((((((((.(((((...(......)...))))))))))))))............))))))))........(((((........((((...))).....((...((((....((((....)))).....))))..))).......)))))")
+CMCT_STRUCT=("CMCT","((((((((......((((((....)))))).((((((((......)))))..)))........))))))))(..(..((((...(((((.((((((...))))))...............((((....)))).....)))))))))....)..)....")
+DMS_CMCT_STRUCT=("DMS+CMCT","((((((((......((((((....)))))).((((((((......)))))..)))........)))))))).......((....(((((.((((((...))))))...............((((....)))).....))))).)).............")
+DMS_CMCT_NMIA_STRUCT=("DMS+CMCT+NMIA","((((((((......((((((....)))))).((((((((......)))))..)))........))))))))........(((((......((((((...)))))).(((...((((....((((....)))).....))))..))).......)))))")
+
+OTHERS = [NO_PROBING_STRUCT,CMCT_STRUCT,DMS_CMCT_STRUCT,DMS_CMCT_NMIA_STRUCT]
+
+
+if __name__=="__main__":
+    lbase,base = NATIVE_STRUCT
+    print "%s & {\\tt\\Base{%s}}\\\\"%(lbase,base)
[…]

### 2019-01-22 graphs/diff_probing_boxplots.R

@@ -1,0 +1,55 @@
+library(ggplot2)
+library(ggforce)
+Multiprobing_100WT <-c(0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6134, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6134, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6134, 0.6084, 0.6134, 0.6084, 0.6134, 0.6084, 0.6084, 0.7783, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6134, 0.6084, 0.6084, 0.6084, 0.6084, 0.6134, 0.6084, 0.6084, 0.7665, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6134, 0.6084, 0.6084, 0.7649, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.7597, 0.7531, 0.6134, 0.6084, 0.6084, 0.6084, 0.6084, 0.7597, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6134, 0.6084, 0.6084, 0.6084, 0.6084, 0.6084, 0.6134, 0.6084, 0.6134)
+mean(Multiprobing_100WT)
+min(Multiprobing_1M)
+mean(Multiprobing_1M)
+mean(Multiprobing_2M)
+
+(min(Multiprobing_2MclosetoWT)-min(Multiprobing_2M))*100
[…]

### 2019-01-22 graphs/different_runs.R

@@ -1,0 +1,53 @@
+library(ggplot2)
+
+Noprobing<-c(	0.55	,	0.55	,	0.78	,	0.8	,	0.79	,	0.73	,	0.55	,	0.55	,	0.75	,	0.55		)
+M7ILUMg<-c(	0.84	,	0.84	,	0.79	,	0.85	,	0.84	,	0.85	,	0.85	,	0.8	,	0.85	,	0.85		)
+M7ILU	<-c(	0.84	,	0.84	,	0.84	,	0.84	,	0.84	,	0.84	,	0.84	,	0.84	,	0.84	,	0.84		)
+DMS	<-c(	0.81	,	0.81	,	0.82	,	0.81	,	0.82	,	0.82	,	0.82	,	0.81	,	0.82	,	0.82		)
+NMIAMg	<-c(	0.8	,	0.8	,	0.81	,	0.81	,	0.8	,	0.8	,	0.81	,	0.81	,	0.81	,	0.81		)
+NMIAMgCE	<-c(	0.74	,	0.74	,	0.73	,	0.73	,	0.73	,	0.73	,	0.74	,	0.74	,	0.73	,	0.72		)
+M7	<-c(	0.83	,	0.83	,	0.81	,	0.71	,	0.7	,	0.7	,	0.64	,	0.75	,	0.73	,	0.81		)
+NMIA	<-c(	0.79	,	0.79	,	0.79	,	0.79	,	0.79	,	0.79	,	0.79	,	0.79	,	0.64	,	0.79		)
+M7Mg	<-c(	0.79	,	0.79	,	0.77	,	0.79	,	0.78	,	0.78	,	0.78	,	0.79	,	0.63	,	0.77		)
+M7ILU3Mg	<-c(	0.74	,	0.74	,	0.74	,	0.74	,	0.74	,	0.74	,	0.74	,	0.74	,	0.74	,	0.74		)
+BzCN	<-c(	0.73	,	0.73	,	0.73	,	0.73	,	0.73	,	0.73	,	0.73	,	0.73	,	0.73	,	0.73	)
+NaiMg	<-c(	0.71	,	0.71	,	0.72	,	0.71	,	0.72	,	0.71	,	0.71	,	0.72	,	0.72	,	0.72	)
[…]

### 2019-01-22 graphs/formatStructures.py

@@ -1,0 +1,74 @@
+
+
+# The reference structure, potentially with pseudoknots indicated with brackets and parenthesis
+REF_STRUCT = "((((((((......((((((....)))))).(((....(((.....)))...)))........))))))))........(((((......(((((.....))))).(((....(((....((((....)))).....)))...))).......)))))"
+# The candidate structures (may also contain PKs)
+ALT_STRUCTS = [
+             "(((((((((((((((((.(((((...(......)...))))))))))))))............))))))))........(((((........((((...))).....((...((((....((((....)))).....))))..))).......)))))",
+             "((((((((......((((((....)))))).((((((((......)))))..)))........))))))))(..(..((((...(((((.((((((...))))))...............((((....)))).....)))))))))....)..)....",
+             "((((((((......((((((....)))))).((((((((......)))))..)))........)))))))).......((....(((((.((((((...))))))...............((((....)))).....))))).)).............","((((((((......((((((....)))))).((((((((......)))))..)))........))))))))........(((((......((((((...)))))).(((...((((....((((....)))).....))))..))).......)))))"
+    ]
+
+# Set of DISJOINT intervals -> Colors in the reference structure
+COLORS = {
+    # Format Interval (i,j) -> Color
+      (0,7):"lightseagreen",
+(14,19):"lightseagreen",
+(24,29):"lightseagreen",
+(31,33):"lightseagreen",
+      (38,41):"americanrose",
+(46,48):"americanrose",
+      (52,54):"lightseagreen",
+(63,70):"lightseagreen",
+     (79,83):"lightsalmon",
+(106,108):"lightsalmon",
+(113,115):"lightsalmon",
+ (90,92):"lightsalmon",
+  (93,94):"lightseagreen",
+(100,101):"lightseagreen",
+(102,105):"lightsalmon",
+    (120,123):"lightseagreen",
+(128,131):"lightseagreen", 
+ (137,139):"lightsalmon",
+ (143,145):"lightsalmon",
+(153,157):"lightsalmon",
+    }
+
+def parseSS(s):
+    res = {i:-1 for i in range(len(s))}
+    p = []
+    pk = []
+    for i,c in enumerate(s):
+        if c=="(":
+            p.append(i)
+        elif c==")":
+            j = p.pop()
+            res[i]=j
+            res[j]=i
+        elif c=="[":
[…]

### 2019-01-22 graphs/plot_Arcs.py

@@ -1,0 +1,127 @@
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib as mp
+import matplotlib.cm as cm
+import matplotlib.patches as mpatches
+from matplotlib.pyplot import colorbar
+#lista contains base pairs with their occurence x:b1,yb:2,z:number of time this base pair was detected
+
+def ListBasePairsFromStruct(Struct):  # return dic={structure:[liste de pairs de base ],....}
+    lista = []
+    stack = []
+    for i in range(len(Struct)):  # sequence length
+        if Struct[i] == '(':  # Opening base pair
+            stack.append(i)
+        elif Struct[i] == ')':  # Closing base pair
+            k = stack.pop()
+            lista.append((k, i))
+    return lista
+
+def Compare_structures(s1,s2):
+        lista=[]
+	lista2=[]
+        BP1=ListBasePairsFromStruct(s1)
+	BP2=ListBasePairsFromStruct(s2)
+	commonElements=list(set(BP1)&set(BP2))
+	for elem in BP1:
+		if elem not in commonElements:
+			lista.append((elem[0],elem[1],0.69))
+		else:
+			lista.append((elem[0],elem[1],0.05))
+	for elem in BP2:
+		if elem not in commonElements:
+			lista2.append((elem[0],elem[1],0.45))
+		else:
+			lista2.append((elem[0],elem[1],0.05))
+	#print "lis1",lista,"list2",lista2
+	return lista,lista2
+
+def plotsecodnarystructures(rnaString, lista, lista2, n, reactivities):
+    fig = plt.figure()
+    fg, ax = plt.subplots(1, 1)
+    
+    # for color map:
+    min_val = 0
+    max_val = 1
+    #
+    my_cmap =cm.get_cmap('Set2')#('PiYG')#cm.get_cmap('Greys')
+    norm = matplotlib.colors.Normalize(min_val, max_val)  #
+    x = [elem[0] for elem in lista]
+    y = [elem[1] for elem in lista]
+    z = [elem[2] / float(n) for elem in lista]
+
+    x2 = [elem[0] for elem in lista2]
+    y2 = [elem[1] for elem in lista2]
+    z2 = [elem[2] / float(n) for elem in lista2]
[…]

### 2019-01-22 graphs/plotboxplots.R

@@ -1,0 +1,41 @@
+library(ggplot2)
+Multiprobing_10M<-c(0.8113, 0.6084, 0.7477, 0.6084, 0.5142, 0.6788, 0.7597, 0.6084, 0.7545, 0.6084, 0.6134, 0.8084, 0.838, 0.7597, 0.7597, 0.838, 0.7665, 0.7597, 0.7597, 0.6084, 0.6134, 0.7597, 0.7531, 0.7597, 0.6084, 0.6084, 0.6084, 0.6134, 0.6084, 0.6134, 0.7665, 0.6084, 0.6084, 0.6084, 0.6134, 0.6134, 0.7997, 0.838, 0.6134, 0.6084, 0.8345, 0.6084, 0.7715, 0.6792, 0.6134, 0.6792, 0.838, 0.5364, 0.7477, 0.838, 0.6084, 0.6084, 0.6134, 0.8303, 0.7467, 0.6134, 0.6084, 0.6084, 0.6084, 0.7545, 0.7665, 0.6134, 0.6134, 0.7665, 0.6134, 0.6134, 0.6351, 0.838, 0.6084, 0.6084, 0.6134, 0.6134, 0.6084, 0.8073, 0.7545, 0.6134, 0.6134, 0.838, 0.7649, 0.6084, 0.6134, 0.7545, 0.7665, 0.6084, 0.6863, 0.6134, 0.6084, 0.7411, 0.6134, 0.7347, 0.7997, 0.6084, 0.6134, 0.823, 0.7665, 0.8459, 0.8459, 0.6084, 0.6134, 0.8303)
[…]

### 2019-01-22 graphs/plotfig1.R

@@ -1,0 +1,50 @@
+#R version 3.3.2 
+  
+library(reshape2)
+library(ggplot2)
+
+IPANEMAP<-c(0.945,0.6,0.843,0.836,0.762,0.982,0.851,0.74,0.671,0.761,0.914,0.85, 0.92,0.891,0.668,0.801,0.791,0.894,0.912, 0.827,0.977,0.591,0.667,0.539)
+            
+ Rsample<-c(0.274,0.031,0.576,0.58,0.514,0.749,0.641,0.57,0.548,0.651,0.81,0.755,0.833,0.836,0.624,0.758,0.791,0.894,0.929,0.851,1,0.651,0.727,0.707)
+difference<-IPANEMAP-Rsample
+ 
+mean(IPANEMAP)
+mean(Rsample)
+sd(IPANEMAP)
+sd(Rsample)
+mean(IPANEMAP[1:18])
+mean (Rsample[1:18])
+mean(IPANEMAP[19:24])
+mean (Rsample[19:24])
+mean(IPANEMAP[1:18])/mean (Rsample[1:18])
+mean (Rsample[19:24])/mean (IPANEMAP[19:24])
+Difference<-c(IPANEMAP-Rsample)
+sum(Difference[1:18])
+sum(Difference[19:24])
+
+ k<-c('5SrEcoli [120]','Signal [301]','GroupIIO [412]','HCIRES [336]','tRNAphe [76]','P546 [155]' ,'5d16SrE [530]','GroupI [214]','Telom [47]' ,'RNase [401]' 
+      ,'TPPE [79]','5d16SrHv [473]' ,'Lysine [174]','GroupIIT [425]','SAMI [118]','Fluorider [66]'
+      ,'PreQ1 [34]','MBox [154]','cyclic [97]' ,'5d23SrE [511] ',  'Adenine [71]','tRNAasp [75]','SARS [79]','HIV15 [500]')
+
+ df<- data.frame(Program=rep(c("IPANEMAP", "Rsample"), each=24),
+                  RNAs=rep(c('5SrEcoli [120]','Signal [301]','GroupIIO [412]','HCIRES [336]','tRNAphe [76]','P546 [155]' ,'5d16SrE [530]','GroupI [214]','Telom [47]' ,'RNase [401]' 
+                             ,'TPPE [79]','5d16SrHv [473]' ,'Lysine [174]','GroupIIT [425]','SAMI [118]','Fluorider [66]'
[…]

### 2019-01-22 graphs/plotfig2.R

@@ -1,0 +1,43 @@
+library(reshape2)
+library(ggplot2)
+library(scales)
+
+k<-c('(-)','CMCT','DMS','NMIA','DMS+CMCT','DMS+NMIA','NMIA+CMCT','NMIA+DMS+CMCT')
+df<- data.frame(Program=rep(c("MFE [Rnafold]","MEA [Rnafold]","IPANEMAP"), each=4),
+                Conditions=rep(c('(-)','CMCT','DMS','NMIA')),
+                GM=c(0.6883333333, 0.5791666667,0.6476666667, 0.7431666667,0.6305, 0.6236666667,  0.686,0.6766666667,0.6036666667,0.6988333333,0.7588333333,0.771))
+meanMFE<-mean(0.5791667,0.6476667, 0.7431667)
+meanMEA<-mean(0.6236667,0.6860000,0.6766667)
+meanIPANEMAP<- mean(0.6988333,0.7588333,0.7710000)
+  
+df2<- data.frame(Program=rep(c("IPANEMAP")),
+                Conditions=rep(c('DMS+CMCT','DMS+NMIA','NMIA+CMCT','NMIA+DMS+CMCT')),
+                GM=c(0.7055,0.7681666667,0.7693333333,0.7851666667))
+head(df)
+df<-rbind(df,df2)
+
+df$Conditions<-factor(df$Conditions, levels = k)
+df$Program2<- relevel(df$Program, "IPANEMAP")# to order by legend
+df$x_pos[order(df$Conditions, df$Program2)] <- 1:nrow(df)
+
+ df$x_pos
+ #[1]  3  6  9 12  1  4  7 10  2  5  8 11 13 14 15 16
+ #Adjust position:
+ df$x_pos<- c(3,6.5,10,13.5,2,5.5,9,12.5,1,4.5,8,11.5,15.5,17,18.5,20.5)
+# for bar width modif aes(x=Conditions, y=GM, fill=Program,width=.6)## to keep the same bar width: position_dodge(preserve = "single")
+p<-ggplot(data=df, aes(x=df$x_pos, y=GM, fill=Program2),legend=c("MFE [RNAfold]","IPANEMAP","MEA [RNAfold]"))+
+  geom_hline(yintercept=c(0.6988333333,0.771,0.7588333333),color="grey",size=1,alpha=0.2)+# size for width and alpha for clearness
+  geom_bar(stat="identity",position="dodge", width=1)+
[…]

### 2019-06-12 SuppMat.tex
Updates from Overleaf

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage{adjustbox}   
%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%%%%%%%%%%%%%%%%%%%%%%%%%formatting cells colors
\usepackage[table,x11names,dvipsnames]{xcolor}
 \usepackage{collcell}
 \usepackage{array}
 \usepackage{tikz}
  \usepackage{pgfkeys}
 \usepackage{graphicx}
 \usepackage{amsmath,amssymb}

 \pgfkeys{/heat/.is family, /heat,
      Max colour/.initial = Green4,
      Min colour/.initial = Red1,
      max colour/.initial = SpringGreen3,
      mid colour/.initial = white,
      min colour/.initial = Yellow1,
      text colour/.initial = black,
      Min color/.style = {Min colour=#1},% for our friends who can't spell
      Max color/.style = {Max colour=#1},
      min color/.style = {min colour=#1},
      mid color/.style = {mid colour=#1},
      max color/.style = {max colour=#1},
      text color/.style = {text colour=#1},
      min/.initial = -1,
      mid/.initial = 0,
      max/.initial = 1,
      slider/.code={%
         \tikz{\shade[left color=\HVal{min colour},%
                      right color=\HVal{max colour}]%
            (current page.south west) rectangle ++(#1,12pt);
         }%
      }%
  }
  \newcommand\Heatset[1]{\pgfkeys{/heat, #1}}
  \newcommand\HVal[1]{\pgfkeysvalueof{/heat/#1}}

  \newcolumntype{H}{>{\collectcell\Heat}r<{\endcollectcell}}
  \newcommand\Heat[1]{% \Heat{number in the interval [min, max] }
    \if\relax\detokenize{#1}\relax% empty cell
    \else%
      \pgfmathparse{int(100*(#1-\HVal{min})/(\HVal{max}-\HVal{min}))}% map number to [0,100]
      \ifnum\pgfmathresult>100% too big
        \edef\HeatCell{\noexpand\cellcolor{\HVal{Max colour}}}%
      \else\ifnum\pgfmathresult<0% too small
          \edef\HeatCell{\noexpand\cellcolor{\HVal{Min colour}}}%
      \else\ifnum\pgfmathresult<50% between min and mid
	      \pgfmathparse{int(2*\pgfmathresult)}% map number to [0,100]
		  \edef\HeatCell{\noexpand\cellcolor{\HVal{mid colour}!\pgfmathresult!\HVal{min colour}}}%
		\else% between min and max
	      \pgfmathparse{int(2*(\pgfmathresult-50))}% map number to [0,100]
          \edef\HeatCell{\noexpand\cellcolor{\HVal{max colour}!\pgfmathresult!\HVal{mid colour}}}%
        \fi%
        \fi%
      \fi%
      \HeatCell\textcolor{\HVal{text colour}}{$#1$}%
    \fi%
  }

 \pgfkeys{/heatsec/.is family, /heatsec,
	Max colour/.initial = Green4,
	Min colour/.initial = Red1,
	max colour/.initial = SpringGreen3,
	mid colour/.initial = white,
	min colour/.initial = Yellow1,
	text colour/.initial = black,
	Min color/.style = {Min colour=#1},% for our friends who can't spell
	Max color/.style = {Max colour=#1},
	min color/.style = {min colour=#1},
	mid color/.style = {mid colour=#1},
	max color/.style = {max colour=#1},
	text color/.style = {text colour=#1},
	min/.initial = -1,
	mid/.initial = 0,
	max/.initial = 1,
	slider/.code={%
		\tikz{\shade[left color=\HVal{min colour},%
			right color=\HVal{max colour}]%
			(current page.south west) rectangle ++(#1,12pt);
		}%
	}%
}
\newcommand\HeatSecset[1]{\pgfkeys{/heatsec, #1}}
\newcommand\HSVal[1]{\pgfkeysvalueof{/heatsec/#1}}

\colorlet{BadCol}{Burlywood1!70!red}

\newcolumntype{S}{>{\collectcell\HeatSec}r<{\endcollectcell}}
\newcommand\HeatSec[1]{% \Heat{number in the interval [min, max] }
	\if\relax\detokenize{#1}\relax% empty cell
	\else%
	\pgfmathparse{int(100*(#1-\HSVal{min})/(\HSVal{max}-\HSVal{min}))}% map number to [0,100]
	\ifnum\pgfmathresult>100% too big
	\edef\HeatCell{\noexpand\cellcolor{\HSVal{Max colour}}}%
	\else\ifnum\pgfmathresult<0% too small
	\edef\HeatCell{\noexpand\cellcolor{\HSVal{Min colour}}}%
	\else\ifnum\pgfmathresult<50% between min and mid
	\pgfmathparse{int(2*\pgfmathresult)}% map number to [0,100]
	\edef\HeatCell{\noexpand\cellcolor{\HSVal{mid colour}!\pgfmathresult!\HSVal{min colour}}}%
	\else% between min and max
	\pgfmathparse{int(2*(\pgfmathresult-50))}% map number to [0,100]
	\edef\HeatCell{\noexpand\cellcolor{\HSVal{max colour}!\pgfmathresult!\HSVal{mid colour}}}%
	\fi%
	\fi%
	\fi%
	\HeatCell\textcolor{\HSVal{text colour}}{$#1$}%
	\fi%
}

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{todonotes}
\title{Supplementary materials}
%\author{}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed 
\newcommand{\Software}[1]{\texttt{#1}}
\newcommand{\OurTool}{\Software{IPANEMAP}}
\newcommand{\Start}{s}
\newcommand{\End}{e}

\newcommand{\yp}[1]{{\todo[color=blue!30]{\sf Yann: #1}}}

\begin{document}
\maketitle

\section{ Cordero~\emph{et al} dataset}
\subsection{Structure prediction by RNA}

 \Heatset{min=0.4,  
	max=1,   
	max colour=Aquamarine3, % colour at maximum
	min colour=BadCol,      % colour at minimum
	Min colour=BadCol, % colour for values below min
	Max colour=Aquamarine3   % colour for values above max
}
\begin{table}[h]
\centering
\begin{adjustbox}{max width=1\textwidth}
 \begin{tabular}{lHHHHHHHHHHHHHHHHHHHHH}
 &\multicolumn{8}{l}{ \OurTool{}} & \multicolumn{4}{l}{ RNAfold MFE} & \multicolumn{4}{l}{RNAfold MEA} \\
% & \multicolumn{3}{l}{ GM }  & \multicolumn{3}{l}{ GM } & \multicolumn{3}{l}{ GM }& \multicolumn{3}{l}{ GM }  & \multicolumn{3}{l}{ GM } & \multicolumn{3}{l}{ GM } &\multicolumn{3}{l}{ GM }\\
\toprule
\textbf{RNA }& \multicolumn{1}{l}{(-)} & \multicolumn{1}{l}{CMCT} & \multicolumn{1}{l}{NMIA} & \multicolumn{1}{l}{DMS} & \multicolumn{1}{l}{DMS+NMIA} & \multicolumn{1}{l}{DMS+CMCT}& \multicolumn{1}{l}{NMIA+CMCT}	 & \multicolumn{1}{l}{DMS+NMIA+CMCT} & \multicolumn{1}{l}{(-)}& \multicolumn{1}{l}{DMS }& \multicolumn{1}{l}{CMCT} &\multicolumn{1}{l}{ NMIA}& \multicolumn{1}{l}{(-)} & \multicolumn{1}{l}{DMS} & \multicolumn{1}{l}{CMCT} & \multicolumn{1}{l}{NMIA}\\
\midrule
5SRNA,E.Coli		&	0.25	&	0.238	&	0.247	&	0.244	&	0.247	&	0.247	&	0.254	&	0.241	&	0.241	&	0.686	&	0.254	&	0.686	&	0.241	&	0.686	&	0.269	&	0.686	\\
glycineriboswitch,F.nucleatum		&	0.568	&	0.627	&	0.868	&	0.952	&	0.868	&	0.658	&	0.868	&	0.868	&	0.306	&	0.313	&	0.395	&	0.313	&	0.593	&	0.593	&	0.693	&	0.6	\\
cidGMPriboswitch,V.Cholerae		&	0.654	&	0.77	&	0.654	&	0.77	&	0.654	&	0.77	&	0.654	&	0.77	&	0.77	&	0.77	&	0.667	&	0.77	&	0.77	&	0.72	&	0.667	&	0.77	\\
P4P6domain		&	0.864	&	0.856	&	0.881	&	0.864	&	0.864	&	0.856	&	0.864	&	0.856	&	0.837	&	0.808	&	0.773	&	0.714	&	0.845	&	0.808	&	0.781	&	0.722	\\
adenineriboswitch,add		&	1	&	0.956	&	1	&	0.977	&	1	&	0.956	&	1	&	1	&	1	&	0.333	&	0.41	&	1	&	1	&	0.333	&	0.356	&	0.306	\\
tRNAphenylalanineyeast		&	0.286	&	0.746	&	0.976	&	0.746	&	0.976	&	0.746	&	0.976	&	0.976	&	0.976	&	0.976	&	0.976	&	0.976	&	0.334	&	0.976	&	0.976	&	0.976	\\

\midrule
Average&0.604&0.7&0.77&0.76&0.77&0.705&0.77&0.785&0.69&0.65&0.58&0.74&0.63&0.69&0.62&0.68\\
\bottomrule
\end{tabular}
\end{adjustbox}
\caption{ GM of the predicted structures with \OurTool{} compared to RNAfold MFE and MEA resulting structures}
%( ref. $supplementary\_data/benchmark\_6RNAs$)}
\end{table}
\subsection{Reproducibility}
\begin{table}[h]
\centering
\begin{adjustbox}{max width=1\textwidth}
 \begin{tabular}{lHHHHHHHHHHHHHHHHHHHHH}
 &\multicolumn{3}{l}{ NMIA} & \multicolumn{3}{l}{ DMS} & \multicolumn{3}{l}{CMCT} &\multicolumn{3}{l}{ NMIA+DMS} & \multicolumn{3}{l}{ NMIA+CMCT} & \multicolumn{3}{l}{DMS+CMCT}& \multicolumn{3}{l}{NMIA+DMS+CMCT} \\
% & \multicolumn{3}{l}{ GM }  & \multicolumn{3}{l}{ GM } & \multicolumn{3}{l}{ GM }& \multicolumn{3}{l}{ GM }  & \multicolumn{3}{l}{ GM } & \multicolumn{3}{l}{ GM } &\multicolumn{3}{l}{ GM }\\
\toprule
\textbf{RNA }& \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3} & \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3}& \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3} & \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3}& \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3} & \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3}& \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3}\\
\midrule
5SRNA,E.Coli	&	0.247	&	0.25	&	0.247	&	0.254	&	0.244	&	0.244	&	0.238	&	0.238	&	0.238	&	0.247	&	0.244	&	0.247	&	0.257	&	0.238	&	0.238	&	0.247	&	0.247	&	0.244	&	0.241	&	0.241	&	0.241	\\
glycineriboswitch,F.nucleatum	&	0.868	&	0.868	&	0.868	&	0.952	&	0.952	&	0.952	&	0.868	&	0.634	&	0.627	&	0.868	&	0.868	&	0.868	&	0.868	&	0.868	&	0.868	&	0.667	&	0.658	&	0.658	&	0.868	&	0.868	&	0.868	\\
cidGMPriboswitch,V.Cholerae	&	0.631	&	0.631	&	0.631	&	0.77	&	0.77	&	0.77	&	0.667	&	0.77	&	0.77	&	0.631	&	0.631	&	0.756	&	0.77	&	0.77	&	0.77	&	0.77	&	0.77	&	0.77	&	0.77	&	0.77	&	0.77	\\
P4P6domain	&	0.864	&	0.881	&	0.881	&	0.864	&	0.864	&	0.856	&	0.856	&	0.856	&	0.856	&	0.864	&	0.864	&	0.864	&	0.864	&	0.864	&	0.864	&	0.856	&	0.856	&	0.856	&	0.856	&	0.856	&	0.856	\\
adenineriboswitch,ad	&	1	&	1	&	1	&	0.956	&	0.956	&	0.956	&	1	&	0.956	&	0.956	&	1	&	1	&	1	&	1	&	1	&	1	&	0.956	&	0.956	&	0.956	&	1	&	1	&	1	\\
tRNAphenylalanineyeast	&	0.953	&	0.953	&	0.976	&	0.976	&	0.746	&	0.976	&	0.746	&	0.746	&	0.746	&	0.976	&	0.976	&	0.976	&	0.976	&	0.976	&	0.97	&	0.746	&	0.746	&	0.746	&	0.746	&	0.976	&	0.976	\\

\bottomrule
\end{tabular}
\end{adjustbox}
\caption{ GM of the predicted structures with \OurTool{} over 3 runs, in the presence of 1,2 and 3 sources of probing data}
%( ref. $supplementary\_data/benchmark\_6RNAs$)
\end{table}

%\section{IPANEMAP Vs Rsample}
\section{Hadjin ~\emph{et al} dataset}

  \Heatset{min=0.4,  
	max=1,   
	max colour=Aquamarine3, % colour at maximum
	min colour=BadCol,      % colour at minimum
	Min colour=BadCol, % colour for values below min
	Max colour=Aquamarine3   % colour for values above max
}
\begin{table}
\begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
\begin{tabular}{llllHllH}
\toprule
\textbf{RNA} & Length &  \multicolumn{3}{l}{\Software{Rsample}} &  \multicolumn{3}{l}{\Software{IPANEMAP}} \\
\midrule
	&		&	Sensitivity	&	PPV	&	\multicolumn1c{GM}	&	Sensitivity	&	PPV	&	\multicolumn1c{GM}	\\
A	&	34	&	0.625	&	1	&	0.791	&	0.625	&	1	&	0.791	\\
B	&	530	&	0.6622	&	0.6203	&	0.641	&	0.8986	&	0.8062	&	0.851	\\
C	&	301	&	0.03	&	0.0313	&	0.031	&	0.6	&	0.6	&	0.6	\\
D	&	214	&	0.6349	&	0.6667	&	0.651	&	0.7302	&	0.8519	&	0.789	\\
E	&	500	&	0.6382	&	0.7823	&	0.707	&	0.5263	&	0.5517	&	0.539	\\
F	&	47	&	0.4	&	0.75	&	0.548	&	0.6	&	0.75	&	0.671	\\
G	&	118	&	0.7179	&	0.8	&	0.758	&	0.7692	&	0.8333	&	0.801	\\
H	&	76	&	0.4762	&	0.5556	&	0.514	&	0.7619	&	0.7619	&	0.762	\\
I	&	155	&	0.6786	&	0.8261	&	0.749	&	0.9821	&	0.9821	&	0.982	\\
J	&	79	&	0.7727	&	0.85	&	0.81	&	0.9545	&	0.875	&	0.914	\\
K	&	71	&	1	&	1	&	1	&	1	&	0.9545	&	0.977	\\
L	&	66	&	0.5625	&	0.6923	&	0.624	&	0.625	&	0.7143	&	0.668	\\
M	&	79	&	0.6538	&	0.8095	&	0.727	&	0.6538	&	0.68	&	0.667	\\
N	&	120	&	0.2857	&	0.2632	&	0.274	&	0.9714	&	0.9189	&	0.945	\\
O	&	97	&	0.9286	&	0.9286	&	0.929	&	0.9286	&	0.8966	&	0.912	\\
P	&	511	&	0.9328	&	0.7762	&	0.851	&	0.8908	&	0.7681	&	0.827	\\
Q	&	401	&	0.5826	&	0.5583	&	0.57	&	0.7304	&	0.75	&	0.74	\\
R	&	425	&	0.8397	&	0.8271	&	0.833	&	0.9084	&	0.8815	&	0.895	\\
S	&	336	&	0.5481	&	0.6129	&	0.58	&	0.8077	&	0.866	&	0.836	\\
T	&	473	&	0.7986	&	0.7143	&	0.755	&	0.8819	&	0.8194	&	0.85	\\
U	&	412	&	0.5379	&	0.6174	&	0.576	&	0.75	&	0.8462	&	0.797	\\
V	&	75	&	0.619	&	0.6842	&	0.651	&	0.619	&	0.5652	&	0.591	\\
W	&	174	&	0.8095	&	0.8644	&	0.836	&	0.8095	&	0.9808	&	0.891	\\
X	&	154	&	0.875	&	0.913	&	0.894	&	0.875	&	0.913	&	0.894	\\

\bottomrule
\end{tabular}
\end{adjustbox}
\caption{\OurTool~ prediction performance compared to \Software{Rsample} tool}
\end{table}

 \begin{table}[h]
\centering
\begin{adjustbox}{max width=1\textwidth}
 \begin{tabular}{@{}ll@{}}
 \toprule
Rna ID & Description\\
 \midrule
A	&	PreQ1riboswitchBsubtilis\\
B	&	5domainof16SrRNAEcoli\\
C	&	SignalrecognitonparticleRNAhuman\\
D	&	GroupIintronAzoarcussp\\
E	&	HIV15primepseudoknotdomain\\
F	&	Telomerasepseudoknothuman\\
G	&	SAMIriboswitchTtengcongensis\\
H	&	tRNApheEcoli\\
I	&	P546domainbI3groupIintron\\
J	&	TPPriboswitchEcoli\\
K	&	AdenineriboswitchVvulnificus\\
L	&	FluorideriboswitchPsyringae\\
M	&	SARScoronaviruspseudoknot\\
N	&	5SrRNAEcoli\\
O	&	cyclicdiGMPriboswitchVcholerae\\
P	&	5domainof23SrRNAEcoli\\
Q	&	RNasePBsubtilis\\
R	&	GroupIIntronTthermophila\\
S	&	HepatitisCvirusIRESdomain\\
T	&	5domainof16SrRNAHvolcanii\\
U	&	GroupIIintronOiheyensis\\
V	&	tRNAaspyeast\\
W	&	LysineriboswitchTmaritime\\
X	&	MBoxriboswitchBsubtilis\\
\bottomrule
\end{tabular}
\end{adjustbox}
\caption{24 RNAs from Hadjin \emph{ et al.} dataset.}
\label{clyu}

\end{table}
%  \Heatset{min=0,  
%	max=1,   
%	max colour=Aquamarine3, % colour at maximum
%	min colour=BadCol,      % colour at minimum
%	Min colour=BadCol, % colour for values below min
%	Max colour=Aquamarine3   % colour for values above max
%}

\begin{table}
\begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
\begin{tabular}{llHHHHHHHHHHHlHH}
\toprule
\textbf{RNA} & Length &  \multicolumn1c{Run1} &\multicolumn1c{Run2}&\multicolumn1c{Run3}&\multicolumn1c{Run4}&\multicolumn1c{Run5}&\multicolumn1c{Run6}&\multicolumn1c{Run7}&\multicolumn1c{Run8}&\multicolumn1c{Run9}&\multicolumn1c{Run10}& \multicolumn1c{Average}& \multicolumn1c{Sd}&\multicolumn1c{\textbf{MFE-RNAfold}} & \multicolumn1c{\textbf{MEA-RNAfold}}\\
\midrule

A	&	34	&	0.791	&	0.791	&	0.791	&		0.791	&	0.791	&	0.791	&	0.791	&	0.791	&	0.791	&	0.722	&	0.7841	&	0.0207	&	0.791	&	0.791	\\
B	&	530	&	0.851	&	0.851	&	0.851	&		0.851	&	0.851	&	0.851	&	0.851	&	0.851	&	0.851	&	0.851	&	0.851	&	0	&	0.851	&	0.849	\\
C	&	301	&	0.6	&	0.6	&	0.6	&		0.6	&	0.6	&	0.6	&	0.6	&	0.6	&	0.6	&	0.6	&	0.6	&	0	&	0.6	&	0.597	\\
D	&	214	&	0.789	&	0.789	&	0.789	&		0.781	&	0.781	&	0.781	&	0.781	&	0.781	&	0.781	&	0.781	&	0.7834	&	0.0037	&	0.768	&	0.761	\\
E	&	500	&	0.539	&	0.539	&	0.539	&		0.539	&	0.539	&	0.539	&	0.539	&	0.535	&	0.541	&	0.515	&	0.5364	&	0.0073	&	0.562	&	0.524	\\
F	&	47	&	0.671	&	0.671	&	0.671	&		0.671	&	0.671	&	0.671	&	0.671	&	0.671	&	0.671	&	0.671	&	0.671	&	0	&	0.671	&	0.671	\\
G	&	118	&	0.801	&	0.801	&	0.801	&		0.801	&	0.801	&	0.801	&	0.801	&	0.801	&	0.801	&	0.801	&	0.801	&	0	&	0.892	&	0.879	\\
H	&	76	&	1	&	1	&	1	&		1	&	1	&	1	&	1	&	0.762	&	0.762	&	0.781	&	0.9305	&	0.1063	&	1	&	1	\\
I	&	155	&	0.982	&	0.982	&	0.982	&		0.982	&	0.982	&	0.982	&	0.982	&	0.982	&	0.982	&	0.964	&	0.9802	&	0.0054	&	0.964	&	0.982	\\
J	&	79	&	0.914	&	0.914	&	0.914	&		0.914	&	0.914	&	0.914	&	0.914	&	0.914	&	0.597	&	0.889	&	0.8798	&	0.0946	&	0.914	&	0.914	\\
K	&	71	&	1	&	1	&	1	&		1	&	1	&	1	&	1	&	0.977	&	0.977	&	0.977	&	0.9931	&	0.0105	&	1	&	1	\\
L	&	66	&	0.668	&	0.668	&	0.668	&		0.668	&	0.668	&	0.668	&	0.668	&	0.668	&	0.668	&	0.668	&	0.668	&	0	&	0.668	&	0.668	\\
M	&	79	&	0.706	&	0.706	&	0.706	&		0.706	&	0.706	&	0.706	&	0.721	&	0.721	&	0.667	&	0.745	&	0.709	&	0.0184	&	0.692	&	0.692	\\
N	&	120	&	0.945	&	0.945	&	0.945	&		0.945	&	0.945	&	0.945	&	0.812	&	0.812	&	0.812	&	0.823	&	0.8929	&	0.064	&	0.945	&	0.914	\\
O	&	97	&	0.912	&	0.912	&	0.912	&		0.912	&	0.912	&	0.912	&	0.912	&	0.912	&	0.929	&	0.929	&	0.9154	&	0.0068	&	0.948	&	0.948	\\
P	&	511	&	0.824	&	0.824	&	0.824	&		0.821	&	0.821	&	0.818	&	0.818	&	0.833	&	0.83	&	0.827	&	0.824	&	0.0046	&	0.831	&	0.828	\\
Q	&	401	&	0.74	&	0.74	&	0.74	&		0.74	&	0.74	&	0.74	&	0.74	&	0.74	&	0.74	&	0.74	&	0.74	&	0	&	0.813	&	0.75	\\
R	&	425	&	0.895	&	0.895	&	0.895	&		0.895	&	0.895	&	0.895	&	0.895	&	0.895	&	0.895	&	0.895	&	0.895	&	0	&	0.885	&	0.92	\\
S	&	336	&	0.836	&	0.836	&	0.836	&		0.836	&	0.836	&	0.836	&	0.836	&	0.836	&	0.836	&	0.831	&	0.8355	&	0.0015	&	0.831	&	0.831	\\
T	&	473	&	0.85	&	0.85	&	0.85	&		0.85	&	0.85	&	0.85	&	0.85	&	0.85	&	0.85	&	0.85	&	0.85	&	0	&	0.863	&	0.863	\\
U	&	412	&	0.929	&	0.929	&	0.929	&		0.929	&	0.81	&	0.81	&	0.81	&	0.806	&	0.806	&	0.925	&	0.8683	&	0.06	&	0.929	&	0.843	\\
V	&	75	&	0.591	&	0.591	&	0.591	&		0.591	&	0.591	&	0.591	&	0.591	&	0.591	&	0.591	&	0.901	&	0.622	&	0.093	&	0.591	&	0.591	\\
W	&	174	&	0.891	&	0.891	&	0.891	&		0.891	&	0.891	&	0.823	&	0.823	&	0.823	&	0.823	&	0.823	&	0.857	&	0.034	&	0.842	&	0.823	\\
X	&	154	&	0.894	&	0.894	&	0.894	&		0.894	&	0.894	&	0.894	&	0.894	&	0.894	&	0.894	&	0.894	&	0.894	&	0	&	0.894	&	0.894	\\

\bottomrule
\end{tabular}
\end{adjustbox}
\caption{MCC of predicted structures with \OurTool{} through 10 runs. The Average and the standard deviation (Sd) of the MCC over the 10 runs are reported alongside with the MCC of the MFE and MEA structures obtained with \Software{RNAfold}.}.
\end{table}

  \HeatSecset{min=0,   
	max=.1,   
	max colour=BadCol, % colour at maximum
	min colour=Aquamarine3,      % colour at minimum
	Min colour=Aquamarine3, % colour for values below min
	Max colour=BadCol   % colour for values above max
}

\section{Lariat capping ribozyme in {\itshape Didymium iridis}}

\end{document}

### 2019-06-12 SuppMat.tex
Updates from Overleaf

@@
 \section{ Cordero~\emph{et al} dataset}
-\subsection{Structure prediction by RNA}
+\subsection{Structure prediction ordered by RNA}
 
@@
 
-  \HeatSecset{min=0,   
-	max=.1,   
-	max colour=BadCol, % colour at maximum
-	min colour=Aquamarine3,      % colour at minimum
-	Min colour=Aquamarine3, % colour for values below min
-	Max colour=BadCol   % colour for values above max
-}
+%  \HeatSecset{min=0,   
+%	max=.1,   
+%	max colour=BadCol, % colour at maximum
+%	min colour=Aquamarine3,      % colour at minimum
+%	Min colour=Aquamarine3, % colour for values below min
+%	Max colour=BadCol   % colour for values above max
+%}
 
 \section{Lariat capping ribozyme in {\itshape Didymium iridis}}
-
+\begin{table}
+\begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
+\begin{tabular}{lHHHHHHHHHHHHHHHHH}
+\toprule
+&&\multicolumn{1}{l}{1M7ILUMg}  &\multicolumn{1}{l}{1M7ILU } &\multicolumn{1}{l}{NMIAMg}  &\multicolumn{1}{l}{NMIA } &\multicolumn{1}{l}{1M7ILU3Mg}  &\multicolumn{1}{l}{NMIAMgCE}  &\multicolumn{1}{l}{NaiMg} &\multicolumn{1}{l}{BzCNMg } &\multicolumn{1}{l}{1M7MgCE}&\multicolumn{1}{l}{CMCTMg}  &\multicolumn{1}{l}{1M7ILU3}  &\multicolumn{1}{l}{DMSMg}&\multicolumn{1}{l}{NMIACE }&\multicolumn{1}{l}{1M7CE}&\multicolumn{1}{l}{BzCN}&\multicolumn{1}{l}{Nai}  \\
+\midrule
+&&0.85&0.84&0.81&0.8&0.74&0.73&0.73&0.71&0.7&0.7&0.64&0.63&0.61&0.6&0.6&0.6\\
+1M7ILUMg  &0.85&0.85&0.84&0.8&0.82&0.82&0.82&0.75&0.79&0.78&0.79&0.79&0.85&0.61&0.79&0.79&0.85\\
+1M7ILU  &0.84&&0.84&0.82&0.82&0.7&0.83&0.72&0.82&0.85&0.83&0.84&0.84&0.84&0.6&0.84&0.6\\
+NMIAMg  &0.81&&&0.81&0.8&0.74&0.82&0.82&0.81&0.81&0.79&0.8&0.8&0.61&0.81&0.81&0.6\\
+NMIA  &0.8&&&&0.8&0.74&0.73&0.81&0.78&0.7&0.79&0.63&0.72&0.61&0.7&0.7&0.59\\
+1M7ILU3Mg  &0.74&&&&&0.74&0.74&0.7&0.74&0.74&0.7&0.63&0.74&0.74&0.74&0.74&0.74\\
+NMIAMgCE  &0.73&&&&&&0.73&0.73&0.73&0.73&0.73&0.61&0.73&0.61&0.73&0.73&0.6\\
+NaiMg &0.73&&&&&&&0.73&0.72&0.7&0.73&0.64&0.68&0.61&0.73&0.73&0.59\\
+BzCNMg  &0.71&&&&&&&&0.71&0.65&0.73&0.63&0.72&0.6&0.61&0.6&0.59\\
+1M7MgCE&0.7&&&&&&&&&0.7&0.7&0.7&0.7&0.61&0.61&0.6&0.6\\
+CMCTMg  &0.7&&&&&&&&&&0.7&0.63&0.74&0.61&0.6&0.6&0.6\\
+1M7ILU3  &0.64&&&&&&&&&&&0.64&0.62&0.62&0.65&0.65&0.59\\
+DMSMg&0.63&&&&&&&&&&&&0.63&0.61&0.6&0.6&0.6\\
+NMIACE&0.61&&&&&&&&&&&&&0.61&0.62&0.6&0.6\\
+1M7CE&0.6&&&&&&&&&&&&&&0.6&0.61&0.59\\
+BzCN&0.6&&&&&&&&&&&&&&&0.6&0.6\\
+Nai  &0.6&&&&&&&&&&&&&&&&0.6\\
+\bottomrule
+\end{tabular}
+\end{adjustbox}
+\caption{MCC of predicted structures with \OurTool{} in the case of Bi-probing}.
+\end{table}
+\subsection{Predicted structures with conditions belonging to the same cluster}
+{\centering \includegraphics[width=.4\linewidth]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}\\}
 \end{document}

### 2019-06-12 SuppMat.tex
Updates from Overleaf

@@
 \end{table}
-\subsection{Predicted structures with conditions belonging to the same cluster}
-{\centering \includegraphics[width=.4\linewidth]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}\\}
+
+\begin{table}
+\newcommand{\MyScale}{.2}
+\begin{tabular}{@{}cccc@{}}
+%\toprule
+ \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILU_NMIAMg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7MgCE_NMIAMgCE}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/NaiMg_CMCTMg}\\
+ \midrule
+1M7ILU&1M7ILUMg&1M7MgCE&NaiMg\\
++NMIAMg&+1M7ILU3Mg&+NMIAMgCE&+CMCTMg\\
+\end{tabular}
+\caption{Predicted structures with conditions belonging to the same cluster}
+\end{table}
+
+\subsection{Distribution of MCC values over 100 runs}
+\includegraphics[width=\linewidth]{graphs/histog}
 \end{document}

### 2019-06-14 SuppMat.tex
Updates from Overleaf

@@
 \usepackage{todonotes}
-\title{Supplementary materials}
+\title{\OurTool{}:  Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles\\[1em]
+	Supplementary material
+}
 %\author{}
-%\date{} % Activate to display a given date or no date (if empty),
+\date{} % Activate to display a given date or no date (if empty),
          % otherwise the current date is printed 
@@
 \centering
+\newcommand{\B}{\multicolumn{1}{c}{$\bullet$}}
+\newcommand{\BR}{\multicolumn{1}{c|}{$\bullet$}}
 \begin{adjustbox}{max width=1\textwidth}
- \begin{tabular}{lHHHHHHHHHHHHHHHHHHHHH}
- &\multicolumn{8}{l}{ \OurTool{}} & \multicolumn{4}{l}{ RNAfold MFE} & \multicolumn{4}{l}{RNAfold MEA} \\
-% & \multicolumn{3}{l}{ GM }  & \multicolumn{3}{l}{ GM } & \multicolumn{3}{l}{ GM }& \multicolumn{3}{l}{ GM }  & \multicolumn{3}{l}{ GM } & \multicolumn{3}{l}{ GM } &\multicolumn{3}{l}{ GM }\\
-\toprule
-\textbf{RNA }& \multicolumn{1}{l}{(-)} & \multicolumn{1}{l}{CMCT} & \multicolumn{1}{l}{NMIA} & \multicolumn{1}{l}{DMS} & \multicolumn{1}{l}{DMS+NMIA} & \multicolumn{1}{l}{DMS+CMCT}& \multicolumn{1}{l}{NMIA+CMCT}	 & \multicolumn{1}{l}{DMS+NMIA+CMCT} & \multicolumn{1}{l}{(-)}& \multicolumn{1}{l}{DMS }& \multicolumn{1}{l}{CMCT} &\multicolumn{1}{l}{ NMIA}& \multicolumn{1}{l}{(-)} & \multicolumn{1}{l}{DMS} & \multicolumn{1}{l}{CMCT} & \multicolumn{1}{l}{NMIA}\\
+ \begin{tabular}{r|HHHHHHHH|HHHH|HHHH}\toprule
+ 	                              & \multicolumn{8}{c}{ \OurTool{}}                                                                                                                                                                                                           & \multicolumn{4}{c}{RNAfold MFE}                                                                          & \multicolumn{4}{c}{RNAfold MEA}                                                                         \\
+ 	%                             & \multicolumn{3}{l}{ GM }                                                      & \multicolumn{3}{l}{ GM }                                                              & \multicolumn{3}{l}{ GM }                                                                    & \multicolumn{3}{l}{ GM }                                                        & \multicolumn{3}{l}{ GM }                                                     & \multicolumn{3}{l}{ GM }       & \multicolumn{3}{l}{ GM } \\ \toprule
+\midrule
+\textbf{CMCT} & & \B & & & & \B & \B & \BR & & & \B & & & & \B & \\
+\textbf{NMIA} & & & \B & & \B & & \B & \BR & & & & \BR & & & & \B \\
+\textbf{DMS}  & & & & \B & \B & \B & & \BR & & \B & & & & \B & & \\
  \midrule
@@
  	adenineriboswitch,add         & 1                       & 0.956                    & 1                        & 0.977                   & 1                            & 0.956                        & 1                             & 1                                 & 1                       & 0.333                    & 0.41                     & 1                         & 1                       & 0.333                   & 0.356                    & 0.306                    \\
-tRNAphenylalanineyeast		&	0.286	&	0.746	&	0.976	&	0.746	&	0.976	&	0.746	&	0.976	&	0.976	&	0.976	&	0.976	&	0.976	&	0.976	&	0.334	&	0.976	&	0.976	&	0.976	\\
-
-\midrule
-Average&0.604&0.7&0.77&0.76&0.77&0.705&0.77&0.785&0.69&0.65&0.58&0.74&0.63&0.69&0.62&0.68\\
-\bottomrule
+ 	tRNAphenylalanineyeast        & 0.286                   & 0.746                    & 0.976                    & 0.746                   & 0.976                        & 0.746                        & 0.976                         & 0.976                             & 0.976                   & 0.976                    & 0.976                    & 0.976                     & 0.334                   & 0.976                   & 0.976                    & 0.976                    \\ \midrule
+ 	Average                       & 0.604                   & 0.7                      & 0.77                     & 0.76                    & 0.77                         & 0.705                        & 0.77                          & 0.785                             & 0.69                    & 0.65                     & 0.58                     & 0.74                      & 0.63                    & 0.69                    & 0.62                     & 0.68                     \\ \bottomrule
  \end{tabular}
@@
 \begin{tabular}{@{}cccc@{}}
+
 %\toprule
@@
 \subsection{Distribution of MCC values over 100 runs}
-\includegraphics[width=\linewidth]{graphs/histog}
+%\includegraphics[width=\linewidth]{graphs/histog}
+\includegraphics[width=\linewidth]{graphs/histog1}
 \end{document}

### 2019-06-18 SuppMat.tex
Updates from Overleaf

@@
 
+\usepackage{float}
 \usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
@@
 
-\section{ Cordero~\emph{et al} dataset}
+\section{Detailed results on Cordero~\emph{et al} dataset}
 \subsection{Structure prediction ordered by RNA}
@@
 }
-\begin{table}[h]
+\begin{table}[H]
 \centering
@@
 \subsection{Reproducibility}
-\begin{table}[h]
+\begin{table}[H]
 \centering
@@
 
+\newpage
+
 %\section{IPANEMAP Vs Rsample}
-\section{Hadjin ~\emph{et al} dataset}
+\section{Detailed results on Hadjin~\emph{et al} dataset}
+
+ \begin{table}[H]
+\centering
+\begin{adjustbox}{max width=1\textwidth}
+ \begin{tabular}{@{}ll@{}}
+ \toprule
+RNA ID & Description\\
+ \midrule
+A	&	PreQ1riboswitchBsubtilis\\
+B	&	5domainof16SrRNAEcoli\\
+C	&	SignalrecognitonparticleRNAhuman\\
+D	&	GroupIintronAzoarcussp\\
+E	&	HIV15primepseudoknotdomain\\
+F	&	Telomerasepseudoknothuman\\
+G	&	SAMIriboswitchTtengcongensis\\
+H	&	tRNA phe Ecoli\\
+I	&	P546domainbI3groupIintron\\
+J	&	TPPriboswitchEcoli\\
+K	&	AdenineriboswitchVvulnificus\\
+L	&	FluorideriboswitchPsyringae\\
+M	&	SARScoronaviruspseudoknot\\
+N	&	5SrRNAEcoli\\
+O	&	cyclicdiGMPriboswitchVcholerae\\
+P	&	5domainof23SrRNAEcoli\\
+Q	&	RNasePBsubtilis\\
+R	&	GroupIIntronTthermophila\\
+S	&	HepatitisCvirusIRESdomain\\
+T	&	5domainof16SrRNAHvolcanii\\
+U	&	GroupIIintronOiheyensis\\
+V	&	tRNAaspyeast\\
+W	&	LysineriboswitchTmaritime\\
+X	&	MBoxriboswitchBsubtilis\\
+\bottomrule
+\end{tabular}
+\end{adjustbox}
+\caption{24 RNAs from Hadjin \emph{ et al.} dataset.}
+\label{clyu}
+
+\end{table}
 
@@
 }
-\begin{table}
+\begin{table}[H]
 \begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
@@
 \end{table}
+\newpage
 
- \begin{table}[h]
-\centering
-\begin{adjustbox}{max width=1\textwidth}
- \begin{tabular}{@{}ll@{}}
- \toprule
-Rna ID & Description\\
- \midrule
-A	&	PreQ1riboswitchBsubtilis\\
-B	&	5domainof16SrRNAEcoli\\
-C	&	SignalrecognitonparticleRNAhuman\\
-D	&	GroupIintronAzoarcussp\\
-E	&	HIV15primepseudoknotdomain\\
-F	&	Telomerasepseudoknothuman\\
-G	&	SAMIriboswitchTtengcongensis\\
-H	&	tRNApheEcoli\\
-I	&	P546domainbI3groupIintron\\
-J	&	TPPriboswitchEcoli\\
-K	&	AdenineriboswitchVvulnificus\\
-L	&	FluorideriboswitchPsyringae\\
-M	&	SARScoronaviruspseudoknot\\
-N	&	5SrRNAEcoli\\
-O	&	cyclicdiGMPriboswitchVcholerae\\
-P	&	5domainof23SrRNAEcoli\\
-Q	&	RNasePBsubtilis\\
-R	&	GroupIIntronTthermophila\\
-S	&	HepatitisCvirusIRESdomain\\
-T	&	5domainof16SrRNAHvolcanii\\
-U	&	GroupIIintronOiheyensis\\
-V	&	tRNAaspyeast\\
-W	&	LysineriboswitchTmaritime\\
-X	&	MBoxriboswitchBsubtilis\\
-\bottomrule
-\end{tabular}
-\end{adjustbox}
-\caption{24 RNAs from Hadjin \emph{ et al.} dataset.}
-\label{clyu}
-
-\end{table}
 %  \Heatset{min=0,  
@@
 
-\begin{table}
-\begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
+\begin{table}[H]
+\begin{adjustbox}{max width=.9\textwidth,max totalheight=\textheight,keepaspectratio}
 \begin{tabular}{llHHHHHHHHHHHlHH}
@@
 
-\section{Lariat capping ribozyme in {\itshape Didymium iridis}}
-\begin{table}
+\newpage 
+\section{Detailed results of Lariat capping ribozyme from {\itshape Didymium iridis}}
+
+\begin{table}[H]
+ \Heatset{min=0.2,  
+	max=.5,   
+	max colour=BadCol, % colour at maximum
+	min colour=Aquamarine3,      % colour at minimum
+	Min colour=Aquamarine3, % colour for values below min
+	Max colour=BadCol   % colour for values above max
+}
+\begin{center}
+
+\begin{tabular}{lHH}
+ & \multicolumn2c{Distance of probing profile}\\
+Condition & \multicolumn1c{vs prediction}&\multicolumn1c{vs native}\\
+\toprule
+1M7ILUMg & 0.234 & 0.2872\\
+1M7ILU &	0.2784 & 0.2961\\
+NMIAMg &	0.2606 & 0.2518\\
+NMIA &	0.2961 & 0.3316\\
+1M7ILU3Mg &	0.4291 & 0.3936\\
+NMIAMgCE &	0.4291 & 0.3404\\
+NaiMg 	& 0.3493 & 0.305\\
+BzCNMg & 	0.3759 & 0.2429\\
+1M7MgCE &	0.4379 & 0.3759\\
+CMCTMg &	0.4911 & 0.4113\\
+1M7ILU3 &	0.2961 & 0.3404\\
+DMSMg &	0.3493 & 0.3227\\
+NMIACE 	&0.2695 &0.2518 \\
+1M7CE &	0.3582 & 	0.3316\\
+BzCN &	0.4911 & 	0.4734\\
+Nai &	0.4025 &0.4025\\
+\end{tabular}
+\end {center}
+\caption{Distance of the predicted structure from the considered probing data.}
+\end{table}
+
+\begin{table}[H]
+\newcommand{\MyScale}{.18}
+\begin{tabular}{@{}ccccc@{}}
+
+%\toprule
+ \includegraphics[scale=\MyScale]{graphs/Supp_structures/native_structure}&
+ \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILU_NMIAMg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7MgCE_NMIAMgCE}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/NaiMg_CMCTMg}\\
+ \midrule
+Native&1M7ILU&1M7ILUMg&1M7MgCE&NaiMg\\
+structure &+NMIAMg&+1M7ILU3Mg&+NMIAMgCE&+CMCTMg\\
+\end{tabular}
+\caption{Predicted structures with conditions belonging to the same cluster}
+\end{table}
+
+\begin{table}[H]
 \begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
@@
 \begin{table}
-\newcommand{\MyScale}{.2}
-\begin{tabular}{@{}cccc@{}}
+\begin{center}
+\begin{tabular}{ll}
+Sample size & MCC\\
+\toprule
 
-%\toprule
- \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILU_NMIAMg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7MgCE_NMIAMgCE}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/NaiMg_CMCTMg}\\
- \midrule
-1M7ILU&1M7ILUMg&1M7MgCE&NaiMg\\
-+NMIAMg&+1M7ILU3Mg&+NMIAMgCE&+CMCTMg\\
+100 & 0.55 \\
+500 & 0.55  \\
+1000 &  0.81\\
+2000 & 	0.55  \\
+3000 &  0.77 \\
+5000 & 	0.55  \\ 
+10000& 0.8 \\
 \end{tabular}
-\caption{Predicted structures with conditions belonging to the same cluster}
+\end {center}
+\caption{ MCC of predicted structures in absence of probig data when varyingthe sample size }
 \end{table}

### 2019-06-19 IPANEMAP-2019.tex

\documentclass[a4,center,fleqn]{NAR}

%\usepackage{NAR-natbib}
\usepackage{graphics}
%\usepackage[latin1]{inputenc}
\usepackage{subcaption}
\usepackage{verbatim}%for comments
\usepackage[table,x11names,dvipsnames]{xcolor}
\usepackage{adjustbox}% to resize figures
%\usepackage[skip=4pt]{caption}  % to keep dstance between caption and table/figure
%\usepackage[sort&compress]{natbib}
\usepackage{xspace}
%\usepackage{algorithm}
%\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{todonotes}
\usepackage{booktabs}
\usepackage{relsize}
\usepackage{lipsum}
\usepackage{collcell}
\usepackage{pgfkeys}
\usepackage{amsmath}
\usepackage{varwidth}
\usepackage{tikz}
\usepackage{makecell}
\usepackage{gensymb}
\usepackage{inconsolata}

\usetikzlibrary{shapes.geometric, arrows,calc,fit,positioning}

\usetikzlibrary{shapes.geometric, arrows,calc}
%\tikzset { *|/.style={to path={	(perpendicular cs : horizontal line through ={(\tikztostart)},	vertical line through{(\tikztotarget)})	--(\tikztotarget) \tikztonodes	}	}	}

\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\input{macros.tex}

%%%%%% Article Customization %%%%%%%%%%%%%%%%%%%%%%%%

\setlength{\parskip}{.3em}
% for multicolumn text centering 
\hfuzz=5pt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Enter dates of publication
\copyrightyear{2019}
\pubdate{31 July 2019}
\pubyear{2019}
\jvolume{XX}
\jissue{YY}

%\articlesubtype{This is the article type (optional)}
\setcounter{secnumdepth}{3} % for subsection reference

\begin{document}

\title{\OurTool{}:  Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles}
\author{%
Afaf Saaidi\,$^{1}$\footnote{Both first authors contributed equally to the study},
Delphine Allouche\,$^{2\,*}$,
Mireille Regnier\,$^{1}$,
Bruno Sargueil\,$^{2}$
and Yann Ponty\,$^1$% 
\footnote{To whom correspondence should be addressed.
Tel: + 33 1 77578095; Email: yann.ponty@lix.polytechnique.fr}}

\address{%
$^{1}$\,CNRS UMR 7161, LIX, Ecole Polytechnique, France
and
$^{2}$\,CitCoM, Cibles thérapeutiques et conception de Médicaments, CNRS, University Paris Descartes, 4 avenue de l'observatoire 75006 PARIS France}
% Affiliation must include:
% Department name, institution name, full road and district address,
% state, Zip or postal code, country

\history{%
Received January 1, 2019;
Revised February 1, 2019;
Accepted March 1, 2019}

\maketitle

\begin{abstract}
The manual production of reliable structural models from chemical probing experiments, such as SHAPE probing, benefits from the integration of information derived from multiple protocols and reagents. However, the interpretation of multiple probing profiles remains a complex task, hindering the quality and reproducibility of modeling efforts.

\noindent We introduce \OurTool{}, an automated and robust approach for the computational modeling of RNA structure from multiple probing reactivity profiles. Input profiles can results from experiments based on diverse protocols, reagents, or even over a collection of mutants, and are jointly analyzed to predict the dominant conformations of an RNA.

\noindent \OurTool{} combines statistical sampling, unsupervised machine learning, and multi-objective optimization, to produce secondary structure models that are both stable and well-supported by experimental evidences. An analysis of several datasets demonstrates the performance of \OurTool, even in a mono probing setting, and confirms the potential of integrating multiple sources of probing data, leading to several recommendations for the experimental design of informative probing assays.

\noindent Availability: \OurTool{} is freely downloadable at

{\centering \url{\ipanemapurl}\\[.5em]}

\noindent \textbf{Contact:} \href{yann.ponty@lix.polytechnique.fr}{yann.ponty@lix.polytechnique.fr}
%\noindent \textbf{Supplementary information:} Supplementary data available at NAR online.
\end{abstract}

%%%%%%%%%%%%%% Introduction %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Introduction}

Historically used as a validation assays~\cite{Knapp1989}, enzymatic and chemical probing  is increasingly used in combination with computational methods to inform a rational prediction of secondary structure models for RNA~\cite{Mathews2004}. Such an integrated approach to structure modeling has led to sizable improvements in prediction accuracy~\cite{Washietl2012} and is currently at the core of successful modeling strategies~\cite{Miao2017}. However, the interpretation of probing data, to inform structure prediction, is challenged by a number of factors, including structural heterogeneity, experimental errors, structural dynamics and the natural variability of reactivity measurements across replicates.

SHAPE reagents represent a popular class of small molecules that react with the 2’ hydroxyl of flexible ribose~\cite{McGinnis2012}, although the exact properties observed by SHAPE remain the object of ongoing investigations~\cite{McGinnis2012,Hurst2018,Mlynsky2018,Frezza2019,Busan2019}.
As Ribose flexibility is proportional to the degree of freedom of the nucleotide, it is assumed that SHAPE reagents to discriminate nucleotides in stable interactions from others. Different SHAPE reagents are endowed with different dynamics, that can help differentiate Watson-Crick base pairs from more dynamics tertiary contacts~\cite{Gherghe2008,Steen2012,Rice2014,Busan2019}. Other chemical probes, harbouring different chemical reactivities, have been developed. Amongst the most popular, DiMethyl Sulfate (DMS) is a small molecule that methylates N1 of adenines and N3 of cytosine if they are not involved in a hydrogen bond. Similarly, 1-cyclohexyl-3-(2-morpholinoethyl)carbodiimide metho-p-toluene sulfonate (CMCT) reacts on the Watson-Crick face of guanosine (N1) and uracyl (N3) when not engaged in a hydrogen bond~\cite{Ehresmann1987,Brunel2000}. These two reagents therefore mostly reveal Watson-Crick base pairing, but also any other type of contact involving the same edge. This diversity of probes not only increases  coverage of the different positions and structural contexts in an RNA, but it also provides qualitatively different qualitative information, as been widely used since the very early days of RNA structure probing~\cite{Moazed1986, Ehresmann1987, Romaniuk1988, Butcher1994, Brunel2000,  Cordero2015, Somarowthu2015, Gross2017}.

% **************************************************************
% Keep this command to avoid text of first page running into the
% first page footnotes
\enlargethispage{-50.1pt}
% **************************************************************

Computationally, the reactivity of a nucleotide is typically used as a proxy for its unpairedness. The last couple of decades have nevertheless seen a series of paradigm shifts in the ways probing information is integrated, somehow mirroring the evolution of {\em ab initio} methods for secondary structure prediction. The seminal work of Mathews~\cite{Mathews2004} used cutoffs to transform reactivity values into hard constraints. Depending on the reagent/enzyme, significantly unreactive or reactive positions were forced to either paired or remain unpaired respectively within predicted structures. However, such \emph{hard} constraints turned out to be very sensitive to the choice of the cutoff, easily leading to artificially unpaired predictions or unsatisfiable sets of constraints. With the advent of SHAPE probing, subsequent methods lifted the requirement of thresholds and integrated reactivities directly as pseudo-energies, supplementing the Turner nearest-neighbor energy model~\cite{Turner2010} within an (pseudo-) energy minimization scheme~\cite{Deigan2009,Zarringhalam2012,Washietl2012}. Such methods can be seen as optimizing a tradeoff between the thermodynamic stability of a model, assessed by the Turner energy model, and its compatibility with the reactivity profile. As a consequence, popular packages for secondary structure prediction, such as \Software{RNAStructure}~\cite{Mathews2004} or the \Software{Vienna} package~\cite{Lorenz2011}, now routinely accept SHAPE-like reactivity profiles as input of their main structure prediction methods. These include the joint folding of aligned RNAs~\cite{Lavender2015}, partition function, statistical sampling and Maximal Expected Accuracy predictions~\cite{Spasic2017}\ldots

However, independent computational predictions for the same RNA using probing data obtained with different probes often yield largely different models, none of which are fully consistent with all the probing data. It follows that, while theoretically informative, a multiprobing strategy often leaves a user with different models, from which one cannot objectively decide. In order to identify the best fit with all the data, researchers resort to very intuitive and manual methods such as projecting the results obtained with one probe onto the different models obtained using the constraints from another probe~\cite{Herbreteau2005,James2008,Weill2004}. At the end of the modeling process, the modeler is often left with several alternatives, all of which may appear equally consistent with the probing data, and the choice of a "best" model becomes somewhat subjective. Here we have developed a new modeling procedure that takes into account multiple probing data and ultimately yields only one or two models.

%As a consequence, integrating the probing data as "hard" constraints, prohibiting any pairing for highly reactive nucleotides or forcing pairing of the unreactive, would preclude the prediction of valuable structure. For this reason it appeared more relevant and efficient to take probing data into account in the prediction pipeline as soft constraints, i.e as penalties in the score when elaborating the models~\cite{Deigan2009,Washietl2012,Zarringhalam2012}. 
% **************************************************************
% Keep this command to avoid text of first page running into the
% first page footnotes
%\enlargethispage{-65.1pt}
% **************************************************************

% **************************************************************
% Keep this command to avoid text of first page running into the
% first page footnotes
%\enlargethispage{-65.1pt}
% **************************************************************

%{\noindent\bf Contribution je ne mettrais pas de "contribution", c'est une intro pas un résumé ni une conclusion.}

\begin{comment}
\begin{itemize}
\item Complementarity of different probing data (protocols, reagents\ldots)
\item Auto-correlation of probing signal is important (partial validation of this claim due to partial data)
\item Why MEA can be bad... and the sensitivity of SHAPE experiments (dot-plot+ detection of outliers, PCA $\to$ Case study). Sampling from SHAPE modified ensembles induces higher robustness with respect to ``outlier'' (rotten) experiments
\item (wishful thinking) Despite relying on a stochastic sampling, results are reproducible
\item (wishful thinking) Performances of our methods are impacted by the choice of the chosen clustering/$\#$clusters
\item (wishful thinking) Increase of performances upon repeated experiments
\item (wishful thinking) Recommendations for experimentalists  (choosing the right combination of probes)
\end{itemize}
\end{comment}

\section*{Material and methods}

\subsection*{The \OurTool{} method}
\bsi{Ce chapitre est vraiment très réussi, il explique vraiment bien ce qu'on fait et c'est accessible meme pour des polios de biologistes. je trouve dommage qu'il soit cantoné au Mat et Met. A mon avis il faudrait soit le mettre tel quel comme premier chapitre des résultats, soit en faire une version light sans les détails de calculs le mettre comme premier chapitre des résultats et ne laisser que les calculs dans le Mat et Met}  
We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signal produced by various chemical probes. It performs clustering based on multiple sets of structures sampled in different experimental conditions, to ultimately return a set of structures that are representative of the dominant clusters. Its underlying rationale is that the prominent presence of a stable secondary structure within the  (pseudo-)Boltzmann ensembles induced by multiple experimental conditions should increase its likelihood to be (one of) the native structure(s) for a given RNA. It is thus hoped that integrating several reactivity profiles may be used to promote the native structure as one of the dominant structures within the multi-ensemble, and help circumvent the limitation of pseudo-energies derived from single reactivity profile, which are generally not sufficient to elect the native structure as its minimum (pseudo)-free energy candidate. In other words, combining ensembles of structures generated using multiple probing experiments is likely to denoise the Boltzmann (multi-)ensemble, and thus mitigate systematic biases induced by experimental conditions and reagents.

\begin{figure}
	{\centering\resizebox{.9\columnwidth}{!}{
			\input{diagram_NAR.tex}
		}\\}

	\caption{\OurTool{} workflow: \OurTool{} takes as input an RNA sequence with profiling data, denoted by reactivities, from various experimental conditions. \OurTool{} proceeds, first,  with a stochastic sampling that results into samples of predicted secondary structure. The data-driven predicted structures are then gathered in one sample, serving as input for the clustering step. \OurTool{} proceeds, then, with an iterative clustering that ends once the stopping criterion are reached. This step allows to identify the adequate number of clusters $k$ to be considered. The $k$ resulting clusters are then identified by their centroid structures. Clusters figuring on the 2D-Pareto frontier are considered to be optimal and subsequently their corresponding centroids are reported as the predicted structure through \OurTool.}\label{fig:approach}
\end{figure}

Our method, summarized in Figure~\ref{fig:approach}, takes as input a set $\mathcal{D}$ of probing experiments, each materialized by a reactivity profile. It starts by producing sets of representatives structures for each of the reactivity profiles using a SHAPE-directed variant of the classic Ding-Lawrence algorithm~\citep{Ding2003}. Following Deigan\etal\cite{Deigan2009}, soft constraints are used to complement the free energy contributions of the classic Turner energy model with pseudo energy contributions resulting from reactivity profiles derived from SHAPE probing experiments.  Given a reactivity $R_i$ for a position $i$, we associate a free-energy bonus to unpaired positions, defined as
$$\Delta G(i) = m \log(R_i +1 )+b$$ 
using $m=1.3$ and  $b=-0.4$. Those values are compared to those used by Deigan\etal\cite{Deigan2009}, based on the rationale that lower absolute values for pseudo-energy bonuses increase the chance of structures being simultaneously supported by multiple conditions. Those pseudo energy contributions effectively direct our predictions towards a subset of structures that are in good agreement with probing data.

In order to infer recurrent candidates across different samples, sampled structure sets are agglomerated while keeping track of the reagent/condition of origin for each structure, and clustered using the Mini Batch k-Means algorithm~\citep{Sculley2010} (\CL{}) based on a number of clusters determined by an iterative heuristics described in further details below. 
It takes as input a dissimilarity matrix, whose entries consist of the base pair distance between two sampled structures.
\CL{} was chosen as it requires less computational resources than the classic k-means algorithm, yet performs equally as well as an extensive collection including  both agglomerative (affinity propagation) and hierarchical (Ward, Diana, McQuitty) clustering algorithms (data not shown).

The number of clusters is a critical parameter of the \CL{} algorithm. It should, at the same time, remain small enough to ensure reproducibility, while being sufficiently large to discriminate outliers and ensure consistency within each cluster. We iteratively increase the number of clusters, deeming it appropriate if increasing its value leads to:
\begin{enumerate} 
	\item Splitting a significant cluster into two highly similar clusters, or; 
	\item Creating a poorly populated (outlier) cluster, while keeping highly similar clusters for each significant cluster from the previous iteration.
\end{enumerate} 
In practice, clusters are seldom exactly preserved across successive iterations, so we consider two clusters to be \Def{highly similar} if their Maximum Expected Accuracy (MEA) centroid structures~\citep{Lu2009}, differ by less than $\delta$ base pairs.
The $\delta$ parameter defaults to $1$, allowing the identification of clusters across successive generation in the presence of a minor variation.
Moreover, define the (pseudo-)Boltzmann probability of a structure $S$, generated for an experimental condition $d$ as part of a sampled set $\mathcal{S}_d$, as
\[\mathbb{P}_d(S) = \frac{e^{-E_d(S)/kT}}{\mathcal{Z}^*_d} \text{, with } \mathcal{Z}^*_d := \sum_{S'\in \mathcal{S}_d} e^{-E_d(S')/kT}\]
where $E_d(S)$ is the pseudo free-energy assigned to the structure $S$ within the experimental condition $d$.
We consider a cluster to be \Def{significantly populated} if the cumulative probability of its structures exceeds a predefined threshold $\epsilon$. 
The value of $\epsilon$ was set to $\#{\sf Datasets}/3$, ensuring that up to three clusters are deemed significantly populated, and used as our primary candidates.

Based on these definitions, our iterative heuristic consists in running \CL over an increasing number $\NumClust$ of clusters, starting with $\NumClust=2$, until a \Def{stopping criterion} is satisfied, that is either: 1) Two significantly populated clusters have associated centroids which are highly similar; or 2) Centroid structures of significantly populated clusters from the previous iteration are highly similar to those of the current iteration.

The last step of our approach consists in electing the most promising clusters, and returning their centroid. While the final number of clusters $\NumClust$ may potentially be large, only a handful of clusters are expected to represent structures that both stable, and supported by a large number of conditions. The remaining clusters are indeed probably artifacts of the clustering method, but nevertheless useful to filter \emph{noisy} structures. Thus we define the \Def{stability} of a cluster $C$ to denote its accumulated pseudo-Boltzmann probability accross conditions. It is computed as 
\[\text{Stability}(C) = \sum_{d\in \mathcal{D}} \mathbb{P}_d(S).\]
Moreover, we postulate that a perfect cluster should be representative of several conditions. In the presence of a set of experimental conditions $\mathcal{D}$, we consider that a cluster $C$ supports a given condition $d$ when its probability for $d$ exceeds a given threshold $\tau$, where $\NumClust$ is the number of clusters. The number of conditions supporting a cluster $C$ is defined as
\[\text{Support}(C) = |\{d\in \mathcal{D} \mid \sum_{S\in C\cap \mathcal{S}_d} \mathbb{P}_d(S)\ge \tau \}|.\]
The value of $\tau$ is set to $1/(\NumClust+1)$, ensuring at least one supporting cluster for each condition.

\OurTool evaluates the two above metrics on each cluster, and eliminates any cluster that is dominated by another with respect to both metrics. The remaining ones are \Def{Pareto optimal}, a classic concept in multi-objective optimization~\cite{Mattson2005}. Finally, we compute and return the MEA centroid~\citep{Lu2009} of the Pareto-optimal clusters as our final prediction(s).

\subsection*{Pairwise comparison of structural ensembles induced by reactivity profiles}\label{sec:dotplots}

We want to compare the structural ensembles induced by reactivity profiles, produced across diverse experimental conditions. To that purpose, we simply consider the base pair probability matrices, or dot-plots, resulting from supplementing the Turner energy model with pseudo energy terms. %This can be seen as a projection of experimentally constrained conformational spaces. 
Dot plots can be computed efficiently in the presence of pseudo-energy terms using a variant of the McCaskill algorithm~\citep{McCaskill1990}.

%To find the most compatible structural models across miscellaneous probing data, 
%As a first assessment of the structural compatibility across miscellaneous probing data, we compared their respective Boltzmann probability distributions where the pseudo-Boltzmann probability to observe a given structure from the ensemble is deduced from the base-pair probabilities computed through a recursive algorithm ~\cite{McCaskill1990}. 
%We remind that RNA folding process is a stochastic mechanism. For a given RNA a certain number of possible stable conformations is present in the ensemble.
%Assuming a Boltzmann thermodynamic equilibrium and the stability of  conditions, a Boltzmann probability to observe each secondary structure from the ensemble could be calculated~\cite{McCaskill1990}.

As a measure of the \Def{ensemble distance} $\Edist$ induced by probing data, we consider the dot-plots associated with experimental conditions $d$ and $d'$, and compute the squared Euclidean distance, such that
\[\Edist (d,d')=\sum\limits_{i=1}^\RL\sum\limits_{j=x+1}^\RL \left(\mathbb{P}(i,j\mid d)-\mathbb{P}(i,j\mid d')\right)^2.\]
with $\RL$ the length of the RNA sequence,  $\mathbb{P}(i,j\mid d)$ the Boltzmann probability of forming a base pair $(i,j)$ in the pseudo-Boltzmann ensemble associated with condition $d$.

Individual dot-plots were computed using the \Software{RNAfold} software in the \Software{Vienna Package 2.2.5}, using the \Software{-p} option in combination with the pseudo-energy terms introduced by Deigan\etal\cite{Deigan2009}.

% Yann:
%\begin{itemize}
% \item Sampling representative sets of structures. Pseudo potentials et distribution de Boltzmann (Utiliser refs, justifier et positionner)
%\item Clustering and extraction of representative structures (inc. optimal \#clusters)
%\item Selecting a subset of dominant clusters
%\item Comparing pseudo-Boltzmann ensembles
%+illust + On peut faire ça uniquement parceque les ARN sont de meme taille, ie l'alignement est sans ambiguite.
%Citer Halvorsen et al (+Laederach Plos CB 2013?)%
%\item Generating artificial SHAPE datasets
%Citer Suskozd (on generalise et etend au dela des grands rRNAs) et simulateur SHAPE (ref???)
%\end{itemize}

% \subsection*{Generative model for probing data}

%Many models to simulate SHAPE probing data starting from the analysis of the SHAPE reactivity distribution performed by ~\cite{Sukosd2013}, have been suggested.
%ull Access% c'est mal dit!===
%In order to generalize the simulation of probing data to other probes than SHAPE and to suggest more deterministic models, we build the normalized distribution of reactivities in function of three structural categories: {\em Helix, Helix-End} and {\em unpaired} using reactivity profiles from 6 RNAs with known structures and three available probing data ~\cite{Cordero2012}.  
%We constructed probabilistic models to simulate probing data from the different distributions; a number of 500 bins was set, the mid-value for each bin with the corresponding probability was considered as parameter for the generative model. Consequently, a random generator was implemented to simulate probing data for each single structural category and each specific probing method where the generator input is the mid-value vector w probability vector.

\subsection*{Datasets} 
\label{sec:datasets}
To validate our computational method quantitatively, we consider several datasets, depending on the availability of probing data for one or several reagents, restricted to the wild type or produced for several point-wise mutants. Each dataset consists of sequences and individual reactivities to one or several probes, at each position in the RNA, completed with one or several functionally-relevant secondary structures.

\paragraph{Cordero \emph{et al} dataset.} 
In order to test the integration of different probing sources, we consider a dataset introduced in Cordero\etal\cite{Cordero2012}, consisting of $6$ RNAs with known structures, for which reactivities are available for three reagents: {\tt NMIA-\SH, DMS} and {\tt CMCT}. 
Probing data were downloaded from the RMDB~\citep{Cordero2012a} on July 2017. In the RMDB, reactivity scores are reported for all nucleotides, including those that are not expected to react with a given reagent. Thus, for the DMS (resp. CMCT) probing, we restricted reactivities to positions featuring nucleotides {\sf A} and {\sf C} (resp. {\sf G} and {\sf U}), setting the reactivity to 0 for other positions. This assumption allowed to decrease the noise generated by reactivities associated with non-targeted nucleotides, leading to more accurate predictions (data not shown).

\paragraph{Hadjin \emph{et al} dataset.} 
A dataset was gathered by Hadjin\etal\cite{Hajdin2013} to validate the predictive capacities of probing data-driven predictions. It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (SHAPE -- 1M7) is available.
This dataset includes sequences originating from a variety of organisms, and spans lengths ranging from $34$ nts to $530$ nts, with a focus on riboswitches and complex RNA architectures.

\paragraph{Didymium structural model and probing data.} 
We considered the 188 nucleotides Lariat capping ribozyme from {\itshape Didymium iridis}, resolved 3.85 \AA{} resolution using X-ray cristallography (PDB: 4P8Z)~\citep{Meyer2014}.  We annotated the secondary structure elements using the \Software{DSSR} software from the 3DNA suite~\cite{Lu2015}. Non-canonical base pairs were removed~\citep{Smit2008}, by considering as the secondary structure the maximum subset of non-pseudoknotted base pairs.

Probing data were experimentally generated, as described in the next section, using a comprehensive set of conditions covering most of the popular probing reagents and SHAPE technologies for stop-inducing and mutation-inducing adducts. We also considered the presence/absence of Mg$^{2+}$, both to assess the capacity of \OurTool{} to recover tertiary interactions, and to assess the induced discrepancy on probing profiles and pseudo-Boltzmann ensembles.

\paragraph{Cheng  \emph{et al} dataset.} 
Starting from the assumption that a functional structure should be preserved during evolution, we wanted to assess the agreement that might exist between probing data profiles for a set of RNA mutants. 
We considered DMS probing data, generated by~\cite{Cheng2017} through systematic point-wise mutations, for the Lariat-capping ribozyme. We renormalized each reactivity profile following the method introduced by Deigan\etal\cite{Deigan2009}, restricted to the primer-free sequence: values greater than $1.5 $ times the interquartile range were discarded, and remaining values were divided by the mean of the top $10\%$ reactivities.  Overall, this constitutes a collection of 188 sequences, each having its associated reactivity profile.

\subsection*{Experimental probing protocols}
The structure of the Lariat capping ribozyme from \textit{Didymium iridis }(DiLCrz) was probed with different SHAPE reagents: 1M7 (1-methyl-7 nitrosatoic anhydre), NMIA (N-Methyl Isatoic Anhydre), BzCN (Benzoyl cyanide) and NAI (2-methylnicotinic acid imidazolide) in presence and absence of magnesium ions, and with DMS (DiMethylSulfate) and CMCT only in presence of magnesium ions.
As a preliminary experiment, we verified that DiLCrz generally adopts a single global architecture. To this end, we subjected DiLCrz to a standard denaturation protocol (80\degree{}C for 2 min in H2O, addition of 40 mM of HEPES at 7.5 pH, 100 mM of KCl, 5 mM of MgCl2, followed by 10 min at room temperature, and 10 min at 37\degree{}C), and observed the production of a single band on a non-denaturing PAGE, strongly suggesting the adoption of a single conformation.

\paragraph{Stops-inducing probing protocol (SHAPE-CE).} 
6 pmol of RNA were resuspended in 18 $\mu$l of water, denatured at 80\degree{}C for 2 minutes and cooled down at room temperature during 10 minutes in the probing buffer (40 mM HEPES at 7.5 pH, 100 mM KCl, in presence or absence of 5 mM MgCl2). After a 10 minutes incubation at 37\degree{}C, RNAs were treated with 2 mM of SHAPE reagent or DMSO (negative control) and incubated for 2 (BzCN), 5 (1M7), 30 (NMIA) or 60 (NAI) minutes at 37\degree{}C. Modified or unmodified RNAs were purified by ethanol precipitation and pellets were resuspended in 10 $\mu$l of water.
Modifications were revealed by reverse transcription using 5’ fluorescently labelled primers (D2 or D4 WellRED, Sigma Aldrich 5’ CTG-TGA-ACT-AAT-GCT-GTC-CTT-TAA 3’) and M-MLV RNAse H- reverse transcriptase (Promega) as previously described~\cite{Deforges2017}, with only minor modification of the originally described SHAPE protocol~\cite{Wilkinson2006}. cDNAs were separated by capillary electrophoresis (Beckman Coulter, Ceq8000). Data were analyzed using the  QuSHAPE~\cite{Karabiber2013} software. RNA probing was performed in triplicate with distinct RNA preparations.

\paragraph{Mutations-inducing probing (SHAPE-MaP).}
Probing was conducted as described in Smola\etal\cite{Smola2015}, except we used SuperScript III reverse transcriptase (ThermoFisher) at 50\degree{}C for 3 hours using the following specific primer:

{\centering \relsize{-1}%
5' CTT-CAT-AGC-CTT-ATG-CAG-TTG-CTT-TTT-TTT-TTT
   TTT-TTT-TTT-GAT-TGT-CTT-GGG-ATA-CCG-GAT 3'\\}

For 1M7ILU3, the denaturation step was repeated three times : After a 1 minute incubation at 95\degree{}C 10 mM of 1M7 was added and the mix was incubated 1 minute at 95\degree{}C. This step was repeated 3 times. For 1M7, 1M7Mg, NMIA and NMIAMg, experiments were conducted as described above, except that the NGS library preparation was adapted to Ion Torrent sequencing. Sequences were mapped and analyzed with \Software{ShapeMapper2} for 1M7ILU and 1M7ILU3. For the other conditions they are mapped with a script based on \Software{ShapeMapper}, but adapted to Ion Torrent output files.

\paragraph{DiMethylSulfate (DMS) and CMCT probing.}
DMS and CMCT probing were conducted essentially as described in~\cite{Weill2004,James2008}. Succinctly, 6 pmol of RNA were resuspended in 18 $\mu$l of water, denatured at 80\degree{}C for 2 minutes and cooled down at room temperature for 10 minutes in a probing buffer for DMS probing (40 mM HEPES pH 7.5, 100 mM KCl, 5 mM MgCl2) or in CMCT (50 mM of potassium borate pH 8, 100 mM KCl and 5 mM MgCl2). For DMS, RNA was then treated with 1 $\mu$l of DMS (1:12 in ethanol) or 1 $\mu$l of ethanol for 5 minutes at 37\degree{}C (mock reaction), the reaction was stopped by addition of 400 mM of Tris at 7.5 pH and immediately put on ice. For CMCT,  RNA was treated with 10 $\mu$l of CMCT (42mg/mL) or 10 $\mu$l H2O for 10 minutes at 37\degree{}C (mock reaction). Modified RNA were ethanol-precipitated and resuspended in 10 $\mu$l of water. Modifications were mapped as described above for SHAPE-CE experiments.

\subsection*{Benchmarking methodology}

The Matthews Correlation Coefficient (MCC) is a classic metrics for assessing the quality of a predicted structure $S$, identified by a set of base pairing positions, based on its agreement with an accepted reference structure $\Ref$. It can be interpreted as a compromise between the classic sensitivity and specificity metrics, and is defined as 
$$\text{MCC}(S\mid \Ref)=\frac{\text{TP} \times \text{TN} -\text{FP} \times \text{FN}}{\sqrt[]{(\text{TP}+\text{FP})(\text{TP}+\text{FN})(\text{TN}+\text{FP})(\text{TN}+\text{FN})}},$$
where TP, FP, TN, FN classically represent the correctly/erroneously predicted and correctly/erroneously omitted base pairs in $S$ with respect to $\Ref$.

For the sake of direct comparison with some competing methods~\cite{Spasic2017}, we also report the geometric mean metrics, defined as:
$$ \GMean(S\mid \Ref)=\sqrt{\text{Sens}(S\mid \Ref)\times \text{PPV}(S\mid\Ref)} $$
where $\text{Sens}(S\mid \Ref)$ represents the proportion of base pairs in $\Ref$ that are in $S$, and $\text{PPV}(S\mid \Ref)$ is the proportion of base pairs in $S$ that are also in $\Ref$. This quality metrics has very high correlation with the MCC, to which it is asymptotically equivalent as shown by~Gorodkin\etal\cite{Gorodkin2001}.

\section*{Results}

\subsection*{Implementation}

\OurTool{} was implemented in \Software{Python 2.7+}, and mainly depends on the \Software{RNAsubopt} software in the Vienna package~\cite{Lorenz2011}, and \Software{scikit-learn}~\cite{Pedregosa2012}. \OurTool{} is free software distributed under the terms of the MIT license, and can be freely downloaded, along with a collection of helper utilities, from:

{\centering \url{\ipanemapurl}\\}

\subsection*{Considering multiple probing conditions improves the quality of predictions}\label{sec:cordero}

%In order to validate our assumption of an enrichment in the presence of multiple probing profiles, we analyzed the Cordero \emph{et al} dataset, which contains CMCT, DMS and NMIA probing data for 6 RNAs. 
%

\colorlet{BadCol}{Burlywood1!70!red}

\begin{table*}
	\begin{adjustbox}{max width=\linewidth}%
		\setlength{\fboxsep}{0pt}
		%		\newcommand{\Base}[1]{\colorbox{blue!30}{\strut#1}}
		\newcommand{\Base}[1]{\colorbox{Aquamarine3!60}{\strut#1}}%
		\newcommand{\G}[1]{\colorbox{Aquamarine3!60}{\strut#1}}%
		\newcommand{\B}[1]{\colorbox{BadCol}{\strut#1}}%
		\begin{tabular}{@{}llr@{}} \toprule
			Sequence& \tt {GAUAUGAGGAGAGAUUUCAUUUUAAUGAAACACCGAAGAAGUAAAUCUUUCAGGUAAAAAGGACUCAUAUUGGACGAACCUCUGGAGAGCUUAUCUAAGAGAUAACACCGAAGGAGCAAAGCUAAUUUUAGCCUAAACUCUCAGGUAAAAGGACGGAG}
			& GM\\
			\midrule
			$\varnothing$& {\tt \G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\B{(}\B{(}\B{(}\B{(}\B{(}\B{(}\B{(}\B{(}\B{(}\B{.}\B{(}\B{(}\B{(}\B{(}\B{(}\G{.}\B{.}\B{.}\B{(}\B{.}\B{.}\B{.}\G{.}\B{.}\B{.}\B{)}\G{.}\G{.}\G{.}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\G{.}\B{.}\B{.}\B{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{(}\G{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\B{.}\B{.}\B{(}\G{(}\G{(}\B{(}\G{.}\G{.}\G{.}\B{)}\G{)}\G{)}\B{.}\B{.}\B{.}\G{.}\B{.}\G{(}\G{(}\G{.}\G{.}\G{.}\B{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{.}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\B{)}\G{.}\G{.}\G{)}\G{)}\B{)}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{)}}&.568\\
			CMCT & {\tt \G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{.}\G{(}\G{(}\G{(}\B{(}\B{(}\B{(}\B{(}\B{(}\B{.}\B{.}\G{.}\G{.}\G{.}\G{.}\B{)}\B{)}\B{)}\B{)}\B{)}\G{.}\G{.}\G{)}\G{)}\G{)}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\B{(}\G{.}\G{.}\B{(}\G{.}\G{.}\B{(}\B{(}\B{(}\B{(}\B{.}\B{.}\B{.}\B{(}\B{(}\B{(}\B{(}\B{(}\G{.}\G{(}\G{(}\G{(}\G{(}\G{(}\B{(}\G{.}\G{.}\G{.}\B{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{.}\B{.}\B{.}\B{.}\G{.}\G{.}\G{.}\G{.}\B{.}\B{.}\B{.}\G{.}\G{.}\G{.}\G{.}\G{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{.}\G{.}\G{.}\G{.}\G{.}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\B{)}\G{.}\G{.}\G{.}\G{.}\B{)}\G{.}\G{.}\B{)}\B{.}\B{.}\B{.}\B{.}}&.627\\
			DMS+CMCT & {\tt \G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{.}\G{(}\G{(}\G{(}\B{(}\B{(}\B{(}\B{(}\B{(}\B{.}\B{.}\G{.}\G{.}\G{.}\G{.}\B{)}\B{)}\B{)}\B{)}\B{)}\G{.}\G{.}\G{)}\G{)}\G{)}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\B{(}\B{(}\B{.}\B{.}\B{.}\B{.}\B{(}\B{(}\B{(}\B{(}\B{(}\G{.}\G{(}\G{(}\G{(}\G{(}\G{(}\B{(}\G{.}\G{.}\G{.}\B{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{.}\B{.}\B{.}\B{.}\G{.}\G{.}\G{.}\G{.}\B{.}\B{.}\B{.}\G{.}\G{.}\G{.}\G{.}\G{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{.}\G{.}\G{.}\G{.}\G{.}\B{)}\B{)}\B{)}\B{)}\B{)}\G{.}\B{)}\B{)}\B{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\B{.}\B{.}\B{.}\B{.}\B{.}}&.658\\
			DMS+CMCT+NMIA & {\tt \G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{(}\G{(}\G{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{.}\G{(}\G{(}\G{(}\B{(}\B{(}\B{(}\B{(}\B{(}\B{.}\B{.}\G{.}\G{.}\G{.}\G{.}\B{)}\B{)}\B{)}\B{)}\B{)}\G{.}\G{.}\G{)}\G{)}\G{)}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{(}\G{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{(}\G{(}\G{(}\G{(}\G{(}\B{(}\G{.}\G{.}\G{.}\B{)}\G{)}\G{)}\G{)}\G{)}\G{)}\G{.}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\B{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{(}\G{(}\G{(}\G{(}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{.}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\B{)}\G{.}\G{.}\G{)}\G{)}\G{)}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{.}\G{)}\G{)}\G{)}\G{)}\G{)}} & \textbf{.868}\\ \midrule
			Reference& {\tt \Base{((((((((......((((((....)))))).(((....(((.....)))...)))........))))))))........(((((......(((((.....))))).(((....(((....((((....)))).....)))...))).......)))))}}& 1\\
			\bottomrule
		\end{tabular}
	\end{adjustbox}\\
	\caption{Dot-bracket representations of the native structure and \OurTool{} predictions for the glycine riboswitch sequence of the Cordero\etal\cite{Cordero2012} data set, supplemented with an increasing collection of probing conditions. Positions with green and red backgrounds indicate correctly and incorrectly predicted base pairs.}\label{fig:glycine_example}
\end{table*}

Since \OurTool{} supports any number of probing profiles, we consider the Cordero~\emph{et al} dataset, further described. It consists of 6 RNAs of known structures, for which reactivity profiles are available for the CMCT, NMIA and DMS reagents. We executed \OurTool{} with default parameters on each sequence and any subset of the three available conditions. The centroid secondary structure  associated with the largest probability cluster was considered as the final prediction. Predictions were evaluated in term of the geometric mean (GM) metrics used by previous works~\cite{Spasic2017}.
\bsi{pourquoi utilise-t-on ici la Gm des bp distance et ensuite le MCC? à expliquer?}\ypi{Pour des raisons historiques, surtout dans le benchmark suivant, j'ai rajouté une phrase dans cette direction}As a control experiment, we also report ${\tt RNAfold}$ predictions in presence/absence of probing data, both in energy minimization (MFE) and maximum expected accuracy (MEA) modes. Figure~\ref{cordero1} shows the averaged GM values for all combinations of tools and probing data.

%* Ajouter des conditions de probing -> meilleurs conditions
\begin{figure}
	{\centering{\includegraphics[trim=.5cm .3cm .2cm .2cm, clip, width=\linewidth]{graphs/cordero_mean}}\\}

	\caption{Averaged GM values for structures  predicted using \OurTool{} for subset combination of probing data. MFE and MEA structures, as computed by RNAfold, are included as controls in the absence of probing data, and for mono-probing.}
	\label{cordero1}
\end{figure}

Interestingly, in the absence of probing data, MFE predictions are generally dominant on this dataset, trailed by the MEA, and followed by \OurTool which, in this setting, devolves into the Ding-Lawrence algorithm~\citep{Ding2003}. However, whenever probing reactivities are available, the single centroid returned by \OurTool always achieves higher GM values (Avg: $70\%$) than both MEA (Avg: $62\%$) and MFE  (Avg: $58\%$), whose relative performances depend on the probing reagent.%la moyenne sur les 6 ARNS et sur les 3 reactifs
Interestingly, the quality of MFE and MEA predictions does not systematically benefit from additional probing data. Indeed, for half of the reagents and methods, the average GM obtained in the presence of a single reagent is lower than in the absence of probing data. Also, the impact of single probing data varies greatly across the three reagents, and the GE values of predictions respectively informed by CMCT, DMS and NMIA are ordered increasingly for all approaches, except for a minor reversal of DMS and NMIA in MEA mode.

The joint analysis of pairs of probing conditions seems to average the quality of predictions. Any combination of probing data yields a GM value that is always greater than the worst-performing condition in the pair, yet worse than the best-performing alone. Interestingly, the addition of the worst performing condition (CMCT), does not equally affect the performances of DMS ($76\% \to 70.5\%$) %DMS \to DMS+CMCT
and NMIA  ($77.1\% \to 76.9\%$), %NMIA \to NMIA+CMCT
despite the latter conditions inducing similar GM values when considered alone. Indeed, CMCT+DMS yields GM values that are only remotely better than the worst-performing CMCT alone  ($70\% \to 70.5\%$), while CMCT+NMIA greatly outperforms CMCT alone  ($70\% \to 76.9\%$), almost matching the performance of the NMIA alone ($77.1\%$). It is also worth mentioning that NMIA+CMCT, combining the best and worst conditions, achieves a better combined performance than DMS+NMIA, the two best mono-probing conditions. This hints towards some level of complementarity between probing conditions, as already suggested by recent analyses \citep{Yu2018} (steen 2012, Rice 2014, Brunel 2000) and supported by further analyses in this paper.

Remarkably, the combination of the three conditions leads to the best overall predictions, averaging $78.5\%$ GM.
This improves by $8.5\%$, $2.7\%$ and $1.4\%$ the average predictions achieved using CMCT, DMS and NMIA respectively.
%CMCT/DMS/NMIA :0.6988333333/0.7588333333/0.771
Table~\ref{fig:glycine_example} illustrates the incremental refinement of \OurTool predictions for a glycine riboswitch upon increasing the number of considered probing conditions.

%\clearpage

\subsection*{\OurTool outperforms state-of-the-art predictive methods when informed by a single profile}
%Mono-probing Mathews (Hadjim)
%* On est meilleur sur la version courante de RSample
%* Multi conformation
We assess the predictive capacities of \OurTool in a classic setting where a single probing condition is available, and compare its performances against \Software{RSample} software recently introduced by Spacic\etal\cite{Spasic2017}, \Software{RSample} relies on a sampling/clustering method developed independently from the current work. It was shown to perform favorably against a comprehensive collection of state-of-the-art methods in probing-guided structure prediction, including \Software{RME}~\citep{Wu2015}, \Software{RNAprob}~\citep{Deng2016}, \Software{RNAprobing}~\citep{Washietl2012}, \Software{RNAsc}~\citep{Zarringhalam2012} and the \Software{fold} utility from the \Software{RNAstructure} suite~\citep{Reuter2010}.
We consider the dataset of hajdin~\emph{et al}, which consists of 24 RNAs of lengths ranging from 47 to 500 nucleotides, all believed to fold into a unique documented conformation.

\begin{figure}
	\includegraphics[width=\linewidth]{graphs/RsampleVsIPANEMAP/Accuracy}
	\caption{Geometric mean of positive predictive value and sensitivity achieved by the predictions of \OurTool{} and \Software{Rsample} over the Hadjin\etal\cite{Hajdin2013} data set, consisting of 24 RNAs (RNA length indicated within square brackets). \label{fig:Vsrsample}}
\end{figure}

We executed \OurTool with default parameters, producing 1~000 structures per sample for both software. The Pareto-optimality always implies a single returned cluster/structure in the case of a single probing condition. 
We also executed \Software{RSample} with default parameters, also returning a single structure for each sequence. 
We computed and report in Figure~\ref{fig:Vsrsample} the individual Geometric Means (GM) associated with predicted structures.

%\paragraph{Results description}
A first observation is that \OurTool generally outperforms \Software{RSample}. It averages $80.1\%$ GM as opposed to $68\%$ for its competitor.  Out of the 24 considered RNAs, \OurTool outperforms its competitor for 16 out of the 24 RNAs, achieves similar performances for 2 of them, and is dominated for the remaining 6 cases. Whenever dominant, \OurTool achieves an average GM score of $82\%$, dominating by $32.8\%$ its competitor, while being dominated by $3.5\%$ over the remaining examples (including ties). 
%Difference<-c(IPANEMAP-Rsample)
%> sum(Difference[1:18])
%[1] 3.17
%> sum(Difference[19:24])
%[1] -0.382
Remarkably, \OurTool predicts secondary structures much more consistently than its competitor. 
It ensures GM that are greater than 50\% in all cases, and even exceeds 60\% over all except for the yeast tRNA, encoding the Aspartic acid and HIV1 5' pseudoknot domain. The standard deviation of GM for \OurTool is $12.5\%$, compared to the larger value of $21\%$ for \Software{RSample} predictions.

%\paragraph{Discussion}

%Remarques These :
%\begin{itemize}
%  \item Fabrice : Y a t il une raison pour laquelle on est les meilleurs ?
%\end{itemize}
This observed discrepancy is somewhat surprising, given that \OurTool and \Software{RSample} share the same sampling/clustering based modeling. However, the two methods differ substantially whenever a single conformation is detected, in which case \Software{RSample} switches to a simple MEA prediction, whereas \OurTool{} always proceeds with an iterative clustering. This allows \OurTool{} to consider a larger number of clusters, which we believe helps eliminating structural outliers, leading to the election of a better final centroid structure.

%\clearpage

\subsection*{Comparing and exploiting multiple probing conditions: A case study on the Lariat capping ribozyme}

\newcommand{\Bull}[1]{{\sffamily #1}~\raisebox{1pt}{\tikzcircle[black, fill=cluster#1]{3pt}}}

\newcommand{\BullLab}[1]{Cluster \Bull{#1}}

\colorlet{clusterA}{SeaGreen}
\colorlet{clusterB}{Yellow}
\colorlet{clusterC}{gray}
\definecolor{clusterD}{HTML}{AFAFE9}
\colorlet{clusterE}{OliveGreen}
\colorlet{clusterF}{blue!90!black}
\colorlet{clusterG}{Orange}
\colorlet{clusterH}{SeaGreen!40}

\begin{figure*}
	{\centering {\includegraphics[width=.9\linewidth]{graphs/didy/didy}}\\}

	\caption{Model of the Lariat capping ribozyme from {\itshape Didymium iridis} (A -- Meyer\etal\cite{Meyer2014}), pseudoknot-free secondary structure as annotated by \Software{DSSR} (B -- pseudoknot ommitted) and \OurTool first-ranking prediction (C.) from 8 maximally-diverging probing conditions (NMIAMg, 1M7ILU3Mg, NAIMg, 1M7, CMCTMg,1M7ILU3,1M7Mg). \label{fig:didy}}
\end{figure*}
As benchmark RNA we used the lariat capping ribozyme from \textit{Didymium Iridis} (DiLCrz) which 3D structure has been modeled at 3,5 A from X-ray diffraction data~\cite{Meyer2014}. This RNA contains a large diversity of interactions, including helical domains, tertiary pairings, a kissing loop and a pseudoknot (Figure~\ref{fig:didy}).
It was probed as described in material and methods in 16 different conditions using different SHAPE and classical reagents, in presence/absence of Mg$^\text{2+}$ ions, using either the standard premature stop or the newly developed mutation induction technologies to map the modification sites. These experimental results were then used as input of \OurTool{}, either individually or combined,  for secondary structure prediction.

\paragraph{Comparison and clustering of conditions.}
Faced with diverse probing conditions, we first assess the compatibility of the conclusions drawn from different probing data, including the probing-free run of \OurTool as a control. We use the methodology described in the material and methods ~\ref{sec:dotplots} and, for each condition, compute the base pair probability distribution (a.k.a. dot-plot) in the  pseudo-Boltzmann ensemble, induced by the reactivity profile. We then computed the ensemble distance, the squared Euclidean distance between dot plots, for each pair of conditions. Figure~\ref{fig:PCA} summarizes the pairwise distance, projected onto a 2D surface by a principal component analysis.

\begin{figure}
	{\centering \includegraphics[width=\linewidth]{graphs/didy/PCA}\\}%, trim= 3.5cm .5cm 1cm 0,clip]

	\caption{2D spatial representation of conditions, computed by Principal Component Analysis (PCA) to optimally reflect pairwise ensemble distances between conditions. Colors and grayed areas indicate clusters of conditions.}\label{fig:PCA}
\end{figure}

A visual inspection of Figure~\ref{fig:PCA} suggests the presence of 8 clusters. In order to objectively build groups of compatible conditions, we performed a k-mean clustering using \Software{scikit-learn}, setting k to 8, and obtained the clusters highlighted in Figures~\ref{fig:biclustering} and \ref{fig:PCA}.  We obtain the following clusters:
\begin{itemize}\setlength{\itemsep}{0pt}\setlength{\parskip}{.1em}	
\item \BullLab{A}: \OneMSevILU, \NMIA, and \NMIAMg;
\item \BullLab{B}: \OneMSevCE, \NMIACE, \BzCN, and \NAICE;
\item \BullLab{C}: \OneMSevILUThreeMg and \OneMSevILUMg;
\item \BullLab{D}: \OneMSevILUThree;
\item \BullLab{E}: \NMIAMgCE and \OneMSevMgCE;
\item \BullLab{F}: \DMSMg and probing-free;
\item \BullLab{G}: \CMCTMg and \NAIMg;
\item \BullLab{H}: \BzCNMg.
\end{itemize}
The predicted clusters are generally consistent with the ordering resulting from a spectral biclustering, implemented in \Software{scikit-learn} and executed with default parameters, as illustrated by Figure~\ref{fig:glycine_example}.

The average ensemble distance and PCA visualization support a status of outliers for conditions in the cluster \Bull{B} (\OneMSevCE, \NMIACE, \BzCN, and \NAICE).  We hypothesize that such conditions may be either representative of alternative conformations, or be altogether erroneous. Consequently, we will single out these two clusters within our detailed analysis of the multi-probing performances.

\begin{figure}
	{\centering \includegraphics[width=\linewidth]{graphs/didy/bi_clustering}\\}%, trim= 3.5cm .5cm 1cm 0,clip]

	\caption{Discretized ensemble distance induced by conditions. Conditions ordered by spectral biclustering. Conditions regrouped within clusters by k-mean clustering appear as blocks on the x and y axis.}\label{fig:biclustering}
\end{figure}

%\clearpage

\paragraph{Assessment of predictions informed by individual probing conditions (mono).}
%First, we assess the individual performance of each probing condition considered as the sole source of structural constraints.  Then, analyze outperforming experimental conditions that ensure more accurate predictions, and spot \Def{outlier} conditions which, on the contrary, seem to increase the noise level, consequently leading to lower prediction performances for \OurTool{}.

For each probing condition, we run \OurTool{} on a single source of probing data, using a sample size of $1~000$ structures. For the sake of reproducibility, we executed \OurTool{} $10$ times for each condition. We report in Table~\ref{tab:thermo-ipan} the MCC values achieved by our prediction, averaged over the $10$ runs, along with the MCCs obtained by running \Software{RNAfold} with default parameters in energy minimization (MFE) and Maximum Expected Accuracy (MEA) modes.

The predictive performances of \OurTool averages a 70.5\% MCC, compared to 68.5\% and 68\% average MCC for MEA and MFE predictions respectively. The stability of predictions across conditions are comparable for all \OurTool, MEA and MFE-driven predictions (std dev. of 8\%). Excluding the outliers of cluster \Bull{B} improves the average MCC to 74\%.

Over the fourteen conditions, we observe a large discrepancy in the capacity of the experimental setup to inform predictions, with MCCs ranging from above 80\% (\NMIA, \NMIAMg, \OneMSevILU and \OneMSevILUMg) to around 60\% (\NAICE, \BzCN, \OneMSevCE and \NMIACE). These values are much better than those achieved in the absence of probing data (51\% MCC), and can be interpreted as indicative of good quality predictions. Indeed, it should be reminded that the MCC is a stringent metrics, taking values between -100\% and 100\%, 0\% being the expected MCC of a "coin tossing" random predictor. The lowest MCC value of 60\% is, in particular, equally consistent with Sensitivity/PPV values of 60\%/60\%, 80\%/46\% or 45\%/82\%.

Unsurprisingly, the presence/absence of Magnesium during the probing impacts the predictions, with an observed drop from an average 73\% MCC in the presence of Mg$^\text{2+}$ to 67\% MCC in the absence of Mg$^\text{2+}$. However, while the poorly performing conditions of cluster \Bull{B} are probed in the absence of Mg$^\text{2+}$, some of the most informative profiles (\OneMSevILU and \NMIA) were obtained in absence of Mg$^\text{2+}$.
An even more drastic change of performances can be observed when comparing the MCC of mutation-based protocols (avg 78\%) with stop-based ones (avg 66\%).

\begin{table}
	\Heatset{min=60,   
		max=83,   
		max colour=Aquamarine3, % colour at maximum
		min colour=BadCol,      % colour at minimum
		Min colour=BadCol, % colour for values below min
		Max colour=Aquamarine3   % colour for values above max
	}
	\newcommand{\Mut}{Mut}
	\newcommand{\Stop}{Stop}
	\newcommand{\Mg}{$\bullet$}
	\newcommand{\NoMg}{$\circ$}

	{\centering
			\begin{adjustbox}{max width=\linewidth}
		\begin{tabular}{@{}lcccHHH}
			\toprule
			                            &          &         &           & \multicolumn{3}{@{}c@{}}{MCC\%}                                                  \\
			Condition                   &  Clust.  &  Tech.  & Mg$^{2+}$ & \multicolumn1l{\relsize{-1}\OurTool} & \multicolumn1l{MFE} & \multicolumn1l{MEA} \\ \midrule
			\OneMSevILUMg                    & \Bull{C} &  \Mut   &    \Mg    & 85                                   & 82                  & 82                \\
			\OneMSevILU                      & \Bull{A} &  \Mut   &   \NoMg   & 84                                   & 82                  & 84                \\
			\NMIAMg                      & \Bull{A} &  \Mut   &    \Mg    & 81                                   & 80                  & 80                \\
			\NMIA                        & \Bull{A} &  \Mut   &   \NoMg   & 80                                   & 69                  & 69                \\
			\OneMSevILUThreeMg                   & \Bull{C} &  \Mut   &    \Mg    & 74                                   & 75                  & 75                \\
			\NMIAMgCE                    & \Bull{E} &  \Stop  &    \Mg    & 73                                   & 72                  & 73                \\
			\NAIMg                       & \Bull{G} &  \Stop  &    \Mg    & 73                                   & 68                  & 70                \\
			\BzCNMg                      & \Bull{H} &  \Stop  &    \Mg    & 71                                   & 68                  & 54                \\
			\OneMSevMgCE                     & \Bull{E} &  \Stop  &    \Mg    & 70                                   & 69                  & 70                \\
			\CMCTMg                      & \Bull{G} &  \Stop  &    \Mg    & 70                                   & 68                  & 73                \\
			\OneMSevILUThree                     & \Bull{D} &  \Mut   &   \NoMg   & 64                                   & 56                  & 61                \\
			\DMSMg                       & \Bull{F} &  \Stop  &    \Mg    & 63                                   & 66                  & 67                \\
			\NMIACE                      & \Bull{B} &  \Stop  &   \NoMg   & 61                                   & 60                  & 60                \\
			\OneMSevCE                       & \Bull{B} &  \Stop  &   \NoMg   & 60                                   & 61                  & 61                \\
			\BzCN                        & \Bull{B} &  \Stop  &   \NoMg   & 60                                   & 59                  & 59                \\
			\NAICE                         & \Bull{B} &  \Stop  &   \NoMg   & 60                                   & 59                  & 59                \\ \midrule
			Avg Technology              &          &  \Mut   &  $\star$  & 78                                   & 74                  & 75                \\
			                            &          &  \Stop  &  $\star$  & 66                                   & 65                  & 65                \\
			Avg -/+ Mg$^{2+}$           &          & $\star$ &   \NoMg   & 67                                   & 64                  & 65                \\
			                            &          & $\star$ &    \Mg    & 73                                   & 72                  & 71.5               \\ \midrule
			{\bfseries Average overall} &          & $\star$ &  $\star$  & 70.5                                 & 68                  & 68.5               \\ \bottomrule
			                            &          &         &           &
		\end{tabular}
		\end{adjustbox}\\[1em]}

	%data are from multiprobing7 
	\caption{Comparison of predicted secondary structures for the \didy{} obtained with \OurTool{} and computational methods from a mono-probing experiment. Averaged MCC of structures predicted with \OurTool{} and with classic deterministic alternatives. The structure was probed in presence (\Mg) or absence (\NoMg) of Magnesium, using either stops (\Stop) or mutations (\Mut) inducing technologies.
	}
	\label{tab:thermo-ipan}
\end{table}

\begin{figure}
	\newcommand{\W}{.725}
	{\centering\begin{tabular}{@{}c@{}c@{}c@{}}
			{\sf {\bfseries A} -- \relsize{-1}Bi-probing MCC \emph{vs} Min MCC of mono-probing}\\
			\includegraphics[width=\W\linewidth,trim=1cm 1cm .8cm .3cm,clip]{graphs/didy/MCC-vs-Min}\\[.5em]
			{\sf {\bfseries B} -- \relsize{-1}Bi-probing MCC \emph{vs} Average MCC of mono-probing}\\
			\includegraphics[width=\W\linewidth,trim=1cm 1cm .8cm .3cm,clip]{graphs/didy/MCC-vs-Avg}\\[.5em]
			{\sf {\bfseries C} -- \relsize{-1}Bi-probing MCC \emph{vs} Max MCC of mono-probing}\\
			\includegraphics[width=\W\linewidth,trim=1cm 1cm .8cm .3cm,clip]{graphs/didy/MCC-vs-Max}\\
		\end{tabular}\\}
	\caption{Shift in \OurTool{} predictive capacity (MCC) observed upon considering two conditions, compared to the worst ({\sf\bfseries A}),  average ({\sf\bfseries B}) 
		and best ({\sf\bfseries C}) 
		predictions achieved for mono-probing conditions. Green and red edges indicate enhanced and degraded predictions respectively compared to the reference (Min% 
		, Max 
		or Average of conditions). Thickness indicates absolute shift value (only values above 5\% reported for readability).\label{fig:pairwise}}
\end{figure}

\paragraph{Bi-probing mitigates the influence of outliers, but generally fails to improve predictions quality.}
We turn to a systematic exploration of the predictive capacities of bi-probing analyses, based on pairs of probing profiles, and attempt to quantify their impact on \OurTool predictions. For each pair and triplets of conditions, we executed \OurTool using 1~000 samples per condition, and considered the first returned structure. We then computed the associated MCC, and compare it with the MCC of the worst performing condition (Min), best-performing condition (Max) and average MCC over the single conditions experiments. A summary of the results over pairs is shown in Figure~\ref{fig:pairwise}.

A superficial inspection of the resulting MCCs appears to confirm the conclusions of Section~\ref{sec:cordero}. Indeed, considering two conditions leads to predictions whose quality fall between the worst and the best one, while being generally close to the average. More precisely, the average MCC value of pair-informed predictions remains at 70.7\%, improving only by 0.2\% over the average MCC of mono-probing predictions. Pair-informed predictions are marginally more reliable, with the median MCC increases to 72.5\% MCC from 70.5\% for mono-probing predictions. However, outliers strongly impact this picture, and ignoring the conditions in cluster \Bull{B} increases the average MCC to 75.1\%, while unsurprisingly decreasing the standard deviation (8.6\% $\to$ 6.5\%).

Compared to the minimum MCC of the pair, the MCC of bi-probing predictions is improved by 5.3\% on average over the MCC of the worst-performing conditions. Pair MCCs exceed their associated min. MCC by at least 5\% in 47 pairs out of the 120 possible pairs, while never being dominated by 5\%.

Bi-probing tends to perform similarly as the average of the two mono-probing conditions, with an average MCC improvement of 0.2\%. However, for 56 out of the 120 possible pairs, the bi-probing conditions exhibits an improvement of at least 1\% over the average, while a decay of at least 1\% is observed for 46 out of the remaining 64 pairs. Such a decay can partly be attributed to the disruptive effect of the outliers from cluster \Bull{B}, involved in 35 of the 46 observed loss of predictive performance. Moreover, removing this cluster induces an average MCC improvement of 1.1\%.

Predictions informed by two conditions remain, however, generally dominated by the best performing condition of the pair. On average, the MCC of the best condition is 4.9\% higher than the one achieved by bi-probing analysis. 
Outliers in cluster \Bull{B} are partly responsible for this situation, and their removal reduces the average MCC decay to 3.3\%. Moreover, for 9 pairs of conditions, the bi-probing analysis produces MCC that are at least 1\% better than the best of its two conditions, but these encouraging examples are generally dominated by 66 examples where the pair performs more than 1\% worse than the best mono-probing condition.

\paragraph{Considering three conditions yields improvement over the average performance of the triplet.}
Increasing the number of conditions to three boosts the average MCC to 72.3\% and even reaches 76.5\% over triplets that do not contain conditions from cluster \Bull{B}. The MCC gain over the worst condition in the triplet reaches 9.1\% on average.  Predictions are also more consistently good, with a median MCC of \%73.4. Disregarding conditions in cluster \Bull{B} does stabilizes the quality of prediction (stdev 8.4\% $\to$ 5.5\%), while increasing the average MCC to 77.4\%.

Compared to bi-probing, MCC values achieved by integrating triplets of conditions now indicate a clear gain over the average performances of individual conditions, supporting a notion of complementarity between experiments. Indeed tri-probing experiments lead to an average +1.7\% MCC improvement (median +2.6) over the average of mono-probing prediction in the triplet. The gain of quality is widespread, and observed for 65\% of the triplets. Finally, 39 triplets were found to induce MCC values greater than 83\%.

The enhanced performances of tri-probing over bi-probing are consistent with the \emph{voting} principle underlying the clustering used in \OurTool. Indeed, in a bi-probing setting, a single outlier condition, such as \NAICE, may entirely determine the final structure, due to its predictions being very tightly concentrated around a, presumably erroneous, structure. In the presence of three or more conditions, however, clusters resulting from an outlier condition typically end up being dominated, in the Pareto front, by compatible sets of structures originating from the pseudo-Boltzmann ensembles of alternative conditions. It follows that the influence of outliers over the final prediction remains limited.

%\clearpage

\begin{table}[ht]
	\Heatset{min=60,   
		max=83,   
		max colour=Aquamarine3, % colour at maximum
		min colour=BadCol,      % colour at minimum
		Min colour=BadCol, % colour for values below min
		Max colour=Aquamarine3   % colour for values above max
	}
    \newcommand{\Q}{\relsize{-1}}

	{\centering
		\begin{adjustbox}{max width=\linewidth}
			\begin{tabular}{@{}llHH@{}}
				\toprule
				&                                           & \multicolumn{2}{@{}c@{}}{MCC\%}                     \\
				& Conditions                                & \multicolumn{1}{@{}c@{}}{\Q{}Joint} &  \multicolumn{1}{@{}c@{}}{\Q{}Mono} \\ \midrule
				{Cluster \Bull{A}} & \NMIAMg, \OneMSevILU, \NMIA                       & 80                  & 82                      \\
				& \NMIAMg, \OneMSevILU                             & 82                  & 82                      \\
				& \OneMSevILU, \NMIA                               & 82                  & 82                      \\
				& \NMIAMg,  \NMIA                              & 80                  & 80                      \\ \midrule
				{Cluster \Bull{B}}                         & \NAICE, \BzCN, \OneMSevCE, \NMIACE                   & 60                  & 60                      \\
				& \NAICE, \BzCN, \OneMSevCE                           & 60                  & 60                      \\
				& \NAICE, \BzCN, \NMIACE                          & 61                  & 60                      \\
				& \NAICE, \OneMSevCE, \NMIACE                         & 61                  & 60                      \\
				& \BzCN, \OneMSevCE, \NMIACE                        & 61                  & 60                      \\
				& \NAICE, \BzCN                                  & 60                  & 60                      \\
				& \NAICE, \OneMSevCE                                 & 59                  & 60                      \\
				& \NAICE, \NMIACE                                & 60                  & 60                      \\
				& \BzCN, \OneMSevCE                                & 61                  & 60                      \\
				& \BzCN, \NMIACE                                & 60                  & 60                      \\
				& \OneMSevCE, \NMIACE                              & 62                  & 61                      \\ \midrule
				{Cluster \Bull{C}}                         & \OneMSevILUMg, \OneMSevILUThreeMg                        & 82                  & 80                      \\ \midrule
				{Cluster \Bull{E}} & \OneMSevMgCE, \NMIAMgCE                          & 73                  & 72                      \\ \midrule
				{Cluster \Bull{G}}                         & \NAIMg, \CMCTMg                              & 73                  & 72                      \\ \bottomrule
				%				\midrule
				%				{All 16 conditions} & .80                   & .70                       \\ \bottomrule
			\end{tabular}
		\end{adjustbox}
		\\}
	\caption{Predictions do not benefits from the joint consideration of similar conditions. For each cluster, all subsets of conditions in the cluster are considered, and the MCC of the resulting prediction ('Joint') is compared to the average MCC of predictions performed with individual conditions independently ('Mono'), revealing little improvement.}\label{fig:similarconds}
\end{table}

\paragraph{Similar conditions do not contribute to the quality of predictions.} 
Our distance assessment and clustering on conditions reveals groups of conditions that are highly compatible in their conclusions, while others seem to include highly diverging structural information. While some outlier conditions, such as those of cluster \Bull{B}, easily stand out as erroneous, the results of our systematic analysis of pairs and triplets supports a notion of \Def{complementarity between conditions}. Under this assumption, downstream analyses are more likely to benefit from the presence of highly diverging probing experiments, than the accumulation of similar, arguably redundant, profiles.

A first test of this hypothesis consists in executing \OurTool{} on every subset of conditions within clusters, and consider the quality of predictions (MCC). 
Note that our similarity notion and clustering of conditions is purely based on the ensemble features, and does not explicitly consider the MCC of individual conditions. 
As expected, compared to the average MCC over associated mono-probing analyses, a joint consideration of similar conditions only induces limited progress over single probing analyses (between -1\% and +1\%), leading to an modest average improvement (+0.2\% average MCC). This is consistent with the notion that supplementing a probing-based modeling with compatible conditions provides very little additional information, leading to a limited contribution of similar conditions.

\paragraph{Considering diverse conditions improves and stabilizes the quality of predictions.} 
Having established the redundant nature of similar conditions, and their limited contribution to downstream modeling, we turn to an analysis of conditions across clusters.  We run \OurTool{} the quality of predictions obtained by choosing a single condition in each of the 8 clusters. We consider any possible combination, except for the outlier cluster~\Bull{B}, for which we only consider \NAICE as a representative.

The results, summarized in Table~\ref{fig:differentconds}, demonstrate a clear advantage of including diverse conditions, leading to an average MCC value of 77\% in the absence of condition from cluster \Bull{B}. Those performances compare favorably against the 72\% MCC achieved on average by individual conditions, and reveal a remarkable stability, including a 5\% standard deviation, and a degradation in only two out the 24 tested cases ($\sim$1\% decay in both cases).  A similar trend can be observed in the, arguably disruptive presence of \NAICE, inducing an 72\% average MCC, albeit with an increased standard deviation of 9\%.

Finally, the consideration of four triplets of diverse conditions, reported in Table~\ref{fig:best}, allow to obtain the best prediction, achieving 85.3\% MCC. Meanwhile, considering the 16 conditions together yields a 80\% MCC value, increasing to 83\% in the absence of the four outlier conditions of cluster \Bull{B}. These results confirm that the strategy used by \OurTool{} is able to exploit the complementary information provided by multiple conditions, contributing to better predictions.

	\begin{table}[h]
	\Heatset{min=60,   
		max=83,   
		max colour=Aquamarine3, % colour at maximum
		min colour=BadCol,      % colour at minimum
		Min colour=BadCol, % colour for values below min
		Max colour=Aquamarine3   % colour for values above max
	}
    \newcommand{\MC}[1]{\multicolumn{1}{@{}c@{}}{#1}}
    \newcommand{\MCB}[1]{\multicolumn{2}{@{}c@{}}{#1}}
    \newcommand{\Q}{\relsize{-1}}

	\begin{adjustbox}{max width=\linewidth}
		\begin{tabular}{@{}llllHHHH@{}}
			\toprule
			\multicolumn{4}{@{}l@{}}{}                                    & \multicolumn{4}{@{}c@{}}{MCC\%}\\
			\multicolumn{4}{@{}l@{}}{Conditions in individual clusters}                                    & \MCB{-NAI} & \MCB{+NAI} \\
			\Bull{A} & \Bull{C} & \Bull{E} & \Bull{G}   & \MC{\Q Joint} & \MC{\Q Mono}  & \MC{\Q Joint}       & \MC{\Q Mono}      \\ \midrule
			\Q \NMIAMg          & \Q \OneMSevILUMg        & \Q \OneMSevMgCE          & \Q \NAIMg             & 84       & 72       & 83             & 71           \\
			\Q \NMIAMg          & \Q \OneMSevILUMg        & \Q \OneMSevMgCE          & \Q \CMCTMg            & 75       & 72       & 82             & 71           \\
			\Q \NMIAMg          & \Q \OneMSevILUMg        & \Q \NMIAMgCE         & \Q \NAIMg             & 79       & 73       & 82             & 71           \\
			\Q \NMIAMg          & \Q \OneMSevILUMg        & \Q \NMIAMgCE         & \Q \CMCTMg            & 82       & 72       & 60             & 71           \\
			\Q \NMIAMg          & \Q \OneMSevILUThreeMg       & \Q \OneMSevMgCE          & \Q \NAIMg             & 70       & 71       & 60             & 70           \\
			\Q \NMIAMg          & \Q \OneMSevILUThreeMg       & \Q \OneMSevMgCE          & \Q \CMCTMg            & 70       & 70       & 72             & 69           \\
			\Q \NMIAMg          & \Q \OneMSevILUThreeMg       & \Q \NMIAMgCE         & \Q \NAIMg             & 72       & 71       & 74             & 70           \\
			\Q \NMIAMg          & \Q \OneMSevILUThreeMg       & \Q \NMIAMgCE         & \Q \CMCTMg            & 75       & 71       & 60             & 70           \\
			\Q \OneMSevILU          & \Q \OneMSevILUMg        & \Q \OneMSevMgCE          & \Q \NAIMg             & 82       & 73       & 75             & 71           \\
			\Q \OneMSevILU          & \Q \OneMSevILUMg        & \Q \OneMSevMgCE          & \Q \CMCTMg            & 84       & 72       & 60             & 71           \\
			\Q \OneMSevILU          & \Q \OneMSevILUMg        & \Q \NMIAMgCE         & \Q \NAIMg             & 73       & 73       & 60             & 72           \\
			\Q \OneMSevILU          & \Q \OneMSevILUMg        & \Q \NMIAMgCE         & \Q \CMCTMg            & 82       & 73       & 82             & 71           \\
			\Q \OneMSevILU          & \Q \OneMSevILUThreeMg       & \Q \OneMSevMgCE          & \Q \NAIMg             & 84       & 71       & 60             & 70           \\
			\Q \OneMSevILU          & \Q \OneMSevILUThreeMg       & \Q \OneMSevMgCE          & \Q \CMCTMg            & 70       & 71       & 84             & 70           \\
			\Q \OneMSevILU          & \Q \OneMSevILUThreeMg       & \Q \NMIAMgCE         & \Q \NAIMg             & 74       & 72       & 60             & 70           \\
			\Q \OneMSevILU          & \Q \OneMSevILUThreeMg       & \Q \NMIAMgCE         & \Q \CMCTMg            & 74       & 71       & 75             & 70           \\
			\Q \NMIA            & \Q \OneMSevILUMg        & \Q \OneMSevMgCE          & \Q \NAIMg             & 82       & 72       & 75             & 71           \\
			\Q \NMIA            & \Q \OneMSevILUMg        & \Q \OneMSevMgCE          & \Q \CMCTMg            & 83       & 72       & 82             & 70           \\
			\Q \NMIA            & \Q \OneMSevILUMg        & \Q \NMIAMgCE         & \Q \NAIMg             & 73       & 73       & 59             & 71           \\
			\Q \NMIA            & \Q \OneMSevILUMg        & \Q \NMIAMgCE         & \Q \CMCTMg            & 82       & 72       & 82             & 71           \\
			\Q \NMIA            & \Q \OneMSevILUThreeMg       & \Q \OneMSevMgCE          & \Q \NAIMg             & 73       & 71       & 82             & 69           \\
			\Q \NMIA            & \Q \OneMSevILUThreeMg       & \Q \OneMSevMgCE          & \Q \CMCTMg            & 70       & 70       & 73             & 69           \\
			\Q \NMIA            & \Q \OneMSevILUThreeMg       & \Q \NMIAMgCE         & \Q \NAIMg             & 74       & 71       & 68             & 70           \\
			\Q \NMIA            & \Q \OneMSevILUThreeMg       & \Q \NMIAMgCE         & \Q \CMCTMg            & 74       & 71       & 73             & 69           \\ \midrule
			                &                 &                  & Average           & 77       & 72       & 72             & 70           \\ \bottomrule
		\end{tabular}
	\end{adjustbox}\\
	\caption{Predictive performances of conditions across clusters, coupled with impact assessment of NAI.  For each experiment, a single condition is chosen from each cluster except for cluster \Bull{B}. Clusters \Bull{D}, \Bull{F} and \Bull{H}, each consisting of a single condition, contribute their sole condition, leading to the list of conditions  $\mathcal{M}$:=\{\OneMSevILUThree, \BzCNMg and \DMSMg\}. 
		For each combination, the MCC of the predicted structural model is reported, as predicted from the conditions in the presence/absence of \NAICE ('Joint' columns). For each prediction, the average MCC value of mono conditions is reported as a baseline ('Mono' columns).
	}\label{fig:differentconds}
\end{table}

%
	\begin{table}
	\newcommand{\A}{45}
	\newcommand{\B}{1.2em}
	\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

	\centering
	\begin{adjustbox}{max width=\linewidth}
		\begin{tabular}{@{}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}cc@{}}
			\toprule
			\rotatebox{\A}{\Bull{C} \OneMSevILUMg} & \rotatebox{\A}{\Bull{D} \OneMSevILUThree} & \rotatebox{\A}{\Bull{A} \OneMSevILU} & \rotatebox{\A}{\Bull{F} \DMSMg} & \rotatebox{\A}{\Bull{E} \NMIAMgCE} & \rotatebox{\A}{\Bull{G} \CMCTMg} & \rotatebox{\A}{\Bull{H} \BzCNMg} &      MCC      & Mono-Avg \\ \midrule
			$\bullet$             &                                  &            $\bullet$            &           $\bullet$            &                                   &                                 &                                 & \textbf{.853} &   .77    \\
			$\bullet$             &                                  &            $\bullet$            &                                &             $\bullet$             &                                 &                                 & \textbf{.853} &   .81    \\
			$\bullet$             &                                  &                                 &           $\bullet$            &                                   &            $\bullet$            &                                 & \textbf{.853} &   .73    \\
			$\bullet$             &            $\bullet$             &                                 &                                &                                   &                                 &            $\bullet$            & \textbf{.853} &   .73    \\ \bottomrule
		\end{tabular}
	\end{adjustbox}\\

	\caption{Best triplets of conditions, achieving maximal obtained MCC value.}\label{fig:best}
\end{table}
%\begin{table}
%	\centering
%	\begin{adjustbox}{max width=.5\textwidth}
%		\begin{tabular}{@{}llllll@{}}
%			\toprule
%			Description & MCC & Mono-Avg\\
%			\midrule
%			12 conditions& {\bfseries.83}&.74 \\
%			{\em All 16 conditions} &	\textbf{.8}& .705\\
%			&	\textbf{.6}& \\
%			\bottomrule
%		\end{tabular}
%	\end{adjustbox}\\
%
%	\caption{Report the best possible score; + all conditions}
%\end{table}
%%

%\begin{figure*}
%\begin{adjustbox}{max width=\linewidth}
%  \begin{tabular}{ll}
% Seq & \tt {GGUUGGGUUGGGAAGUAUCAUGGCUAAUCACCAUGAUGCAAUCGGGUUGAACACUUAAUUGGGUUAAAACGGUGGGGGACGAUCCCGUAACAUCCGUCCUAACGGCGACAGACUGCACGGCCCUGCCUCUUAGGUGUGUUCAAUGAACAGUCGUUCCGAAAGGAAGCAUCCGGUAUCCCAAGACAAUC
%} \\
% 
% NATIVE &  \tt {(((((..(((((..(((((((((.......)))))))))..((((.(((.((((......(((((...((((..(((......))).......))))((....)).............))))).(((.....))))))).)))..........((((....))))....))))...)))))..)))))} \\
%
%.865 &\tt{(((((..((((((.(((((((((.......))))))))).((((((((((((((......((((....((((...((.......)).......))))......(.(........)..).)))).(((.....)))))))))))..........((((....))))...)))))).))))))..)))))}\\
% \\
%\end{tabular}
%\end{adjustbox}
%\caption{Content: 3D didymium + native structure + predicted base pairs}
%\end{figure*}
%\begin{figure}
%\includegraphics[width=\linewidth]{graphs/Arcs}
%\end{figure}
%\subsection*{\OurTool{} yields reproducible predictions}
%\as{Which structure?}

\subsection*{Supplementing wild-type profiles with random single mutants increases prediction quality}
In this final analysis, we decided to test the capacity of \OurTool to extract and exploit complementary information produced by the systematic probing of single-point mutants.
This approach may appear misguided, since single point mutants are expected to adopt different structures than the wild-type sequence. In fact, the conformational changes induced by mutations precisely constitute the signal that is exploited by downstream modeling pipelines~\cite{Cordero2015}. However, those changes are typically local, and the limited influence of outliers on \OurTool{} raises a hope that the benefit of including independently-produced conditions will outweigh the impact of noise introduced by local changes.

We generated 100 uniformly-distributed subsets consisting of 1, 2, 3 and 10 mutants extracted from the Cheng dataset~\cite{Cheng2017}, which we supplemented with the original wild-type sequence (WT).
For each set of probing profiles, we executed \OurTool{} with default parameters, using a sample size of 1~000 structures. 
We also analyzed the WT sequence alone, reproducing the analysis 100 times to compare the variability across sets of mutants to the one induced by stochastic aspects of our method. 
Finally, we performed a restricted analysis focusing on the WT, supplemented with 2 probing profiles selected from the 20 most similar conditions to the WT , as assessed by the ensemble distance metrics. By comparing our results with those obtained for unrestricted pairs of mutants, we assess the influence of structure variability on the quality of our predictions. We report in Figure~\ref{fig:variantanalysis} the distribution of MCC values associated with the first prediction returned by \OurTool.

For the WT, the average MCC achieved by our predictions is of 61.83\%, comparable to the worst condition of our previous dataset.
Our predictions are generally stable and, in 94\% out of the 100 runs, yield a structure having MCC between 60.5\% and 61\%. The remaining 6\% of the runs present improved MCC values, ranging from 76\% to 78\%, suggesting the existence of an alternative conformer in the pseudo-Boltzmann ensemble.
%%\as{ This value can not be fixed since we have  0.6134: 12\%, 0.7597: 2\%, 0.7649: 1\%, 0.7783: 1\%, 0.7665: 1\%, 0.7531: 1\% , is it ok Yann not to specify the improvemnt percentage ?}
%% Counter({0.6084: 82, 0.6134: 12, 0.7597: 2, 0.7649: 1, 0.7783: 1, 0.7665: 1, 0.7531: 1})
.%mean(Multiprobing_100WT)

When the WT and a single random mutant are jointly considered, the dispersion of the results increases, and MCC values ranging from 38.9\% to 85\% can be observed reached.%min(Multiprobing_1M)
These values, however, are not typical, and the average MCC in the presence of a single mutant increases to 64.9\%,%mean(Multiprobing_1M)
suggesting a positive contribution from the additional mutant profile.

This trend is confirmed when two random mutants are considered in addition to the WT. The average MCC across runs now increases to 66.6\%. %mean(Multiprobing_2M).
Interestingly, the dispersion of the MCC is lower for two mutants than for a single one, with low values that are now confined to 57\%.  This results suggests  a role of \emph{tie-breaker} played by the third condition, allowing to mitigate the deleterious effect of some erroneous outliers.
However, such outliers still play some disruptive role, and restricting our choice of mutants within the 20 conditions most similar to the WT further improves the MCC value by 2.8\% to 67.7\%, with MCC values that now almost reach 60\%.

The average MCC again increases in the presence of 3 mutants, to reach an average of 69.3\% MCC. For 10 mutants, the average MCC and overall distribution remains highly similar to that obtained by considering three mutants, suggesting peeking performances. These results support a notion of complementarity between the probing data produced across reasonably similar mutants, leading to a gradual decrease of the noise in the data. However, the bimodal nature of the distribution also suggests the existence of at least two conformations that are both supported by the WT and the mutant profiles.

\begin{figure}
	{\centering
		\includegraphics[width=\linewidth]{graphs/combinationscale.png}\\}

	\caption{Prediction accuracy (MCC) from Mutate-and-Map data. We consider the WT, supplemented with a set of $k=$1, 2, 3, or 10 single-point mutants, selected uniformly at random ({\sf WT+$k$M}). For 2 mutants, we also considered selecting mutants from the 20 most similar to the WT ({\sf WT+2M close}).}\label{fig:variantanalysis}
\end{figure}

%\begin{figure}[h]
%\begin{center}
%\includegraphics[scale=.4]{graphs/WT+combin}
%\label{Vsrsample}
%\caption{Accuracy of prediction reported as the MCC when considering only the WT, considering one mutants along side the WT and considering 10 mutants in addition to the WT.}

%\end{center}
%\end{figure}

%------------------------------------------------------------------------------

%======================

\section*{Conclusion and discussion}

We introduce \OurTool{}, an integrative method for the prediction of secondary structure models from multiple probing profiles. Based on the simple premise that good structural models should be stable and supported by most probing experiments, it uses a combination of stochastic backtrack in the pseudo-Boltzmann ensemble, coupled with a structural clustering to elect dominant conformations. This strategy is fast, stable and leads to predictions that clearly benefit from the availability of multiple probing profiles, as demonstrated by detailed analyses performed of four datasets.

To the best of our knowledge, \OurTool currently represents the only available method for a joint automated consideration of multiple probing profiles. Surprisingly, it also appears to produce the most accurate secondary structure models when informed by a single condition, as demonstrated by our reanalysis on the Hadjin\etal\cite{Hajdin2013} benchmark. We attribute this unexpected result to the adoption of a sampling/clustering approach, which seems to eliminate spurious outliers in the conformational space. In particular, the dominance of \OurTool is surprising since, while independently developed, our method shares the sampling/clustering philosophy of the recently introduced \Software{RSample}~\cite{Spasic2017}. We attribute these increased performances in part to a less stringent contribution of the pseudo-potentials, and to our adaptive choice of the number of clusters. Indeed, even in cases where a single dominant conformation is expected, considering multiple clusters seems to create more opportunities for discarding noisy random unstable conformations. Such conformations naturally occur in a Boltzmann-distributed subset of structures, and may end up \emph{polluting} the centroid structures by artificially inflating the cluster diversity. 
%The impact on \Software{RSample} predictions of such a \emph{noise} seems worsened by the fact that it returns the MEA, the centroid of the whole ensemble, whenever a dominant conformation is detected.

Despite its stochastic foundations, \OurTool{} is typically stable in its predictions. We report in Figure~\ref{fig:reproducibility} the MCC values achieved over  10 repetitions of our software for all probing datasets produced in the case of didymium. 
Interestingly, the worst most notable exception to the general observed stability is observed for the probing-free prediction. This confirms the widespread interpretation of SHAPE-induced pseudo-potentials as focusing predictions onto a subset of the Boltzmann ensemble. A similar behavior is observed for the other datasets considered in our study, as seen in Supplementary Material. %In the interest of reproducibility, ipanemap reports, and accepts as an option, the random seed used both within the main software and its non-deterministic dependencies, leading to identical repeats of the experiments.

Our analyses reveal that integrating multiple profiles of at least three conditions allows to mitigate the effect of outlier conditions and, at times, to produce better secondary structure models than the one inferred from the most accurate profile. A comparison of the joint performances to the average appears fair since no single reagent appears to systematically dominate in term of predictive capacity. It follows that a modeler cannot objectively favor \emph{a priori} a probing condition/reagent~\cite{Yu2018}. In other words, multiple experimental probing data, even combined in a pairwise fashion, were typically observed to mitigate the empirical risk of a misprediction.

A future work include solving the dilemma of whether or not to include Mg$^{\text{2+}}$ ions during the probing, in the perspective of a downstream computational prediction. On the one hand, RNA reaches its native 3D structure in the presence of Mg$^{\text{2+}}$ ions~\cite{Knapp1989}. However, such native structures include non WC pairings, and complex topological motifs such as pseudoknots, kissing loops, tetraloop interactions, that classic structure prediction algorithms do not explicitly capture. In contrast, in the presence of only monovalent ions, RNA is expected to form its helical domains, but leave its  tertiary contacts unstable. Therefore, probing the RNA without Mg$^{\text{2+}}$ could provide a signal that is less deceptive, and thus more informative, to predicting algorithms. Unfortunately, destabilizing tertiary contacts increases the probability of misfolding, or the coexistence in solution of multiple conformations. In this work, we probed a Didymium ribozyme which contains many tertiary motifs, systematically in the presence and absence of Mg$^{\text{2+}}$ (except for CMCT and DMS probes). Remarkably, the most accurate prediction was obtained with 1M7 and NMIA, irrespectively of the presence/absence of Mg$^{\text{2+}}$, this despite the fact that the reactivity profiles substantially diverged between the two conditions. However, for other reagents/protocols, predictions appeared positively impacted (+$10\%$ average MCC) by the presence of Mg$^{\text{2+}}$, painting a mixed picture that would deserve future investigations.
%\item je pense qu'il serait intéressant au moins pour un uo deux réactif de voir ce que donne le couple + et - Mg. Peut etre par exemple avec le 1M7ilu et un autre mauvais genre le NAI
%\end{itemize}
% il faudrait regrder ce que fait weeks avec le 1M7, le NMIA et le 1M6, je n'ai pas bien compris quelle tambouille il fait, mais ca part du meme genre de constatation : il y a une spécificité dans chacune des sondes qu'il n'arrive pas vraiment à formaliser donc il fait une tambouille qui lui permet de tous les intégrer (mais c'est peut etre un peu pédestre comme démarche!). Apres ce sont probablement des conditions tres proches ce qui n'apporte pas d'apres nos résultats bcq d'infos supplémentaires .. bref ca vaudrait peut etre le coup de comparer.

\begin{figure}
	{\centering\includegraphics[width=1\linewidth]{graphs/boxplotreproducibility.png}\\}

	\caption{MCC of predicted structures in presence/absence of Mg$^{2+}$, using both RT stops-inducing and mutation-inducing protocols for a collection of reagents, over 10 runs of \OurTool{}. }\label{fig:reproducibility}
\end{figure}

%
%%discussion et conclusion sont fusionnés dans ce type d'article.}
%
%
%
%\begin{itemize}
%	\item Our integrative approach compared to other tools did exhibit a trade-off  between  number of optimal clusters and accuracy but remains fast to execute. 
%	\item While based on the assumption that the different probing condition reveal complementary aspects of the same structure, it can still be enriched by experiments where the dominant conformation is expected to differ from the native one.
%	
%	\item 
%	
%	\item Complementarity
%	\item Pseudo-Energy models
%\end{itemize}

\section*{Authors contributions}
AS, BS, and YP designed the computational method. 
AS and YP implemented the computational method.
DA and BS planned and carried out all probing experiments.
AS and YP planned and carried out the computational experiments. 
AS, DA, BS and YP contributed to the interpretation of the results. 
All authors provided critical feedback and helped shape the research, analysis and manuscript

\section*{Acknowledgements}
The authors wish to express their gratitude to Ronny Lorenz for developing a probing-informed version of \Software{RNAsubopt}, allowing to sample the pseudo-Boltzmann ensemble, and Nathalie Chamond for helpful suggestions.

\section*{Funding}
This work has been supported by La Fondation pour la Recherche Medicale ([FRM
DBI20141423337]).
%\bibliographystyle{bioinformatics}
\bibliographystyle{nar}
%\bibliographystyle{natbib}%compa
\bibliography{biblio}

\end{document}

### 2019-06-19 SuppMat.tex

--- 2019-06-12 SuppMat.tex
@@
  \usepackage{tikz}
+ \usepackage{relsize}
   \usepackage{pgfkeys}
  \usepackage{graphicx}
+ \usepackage{xspace}
  \usepackage{amsmath,amssymb}
-
- \pgfkeys{/heat/.is family, /heat,
-      Max colour/.initial = Green4,
-      Min colour/.initial = Red1,
-      max colour/.initial = SpringGreen3,
-      mid colour/.initial = white,
-      min colour/.initial = Yellow1,
-      text colour/.initial = black,
-      Min color/.style = {Min colour=#1},% for our friends who can't spell
-      Max color/.style = {Max colour=#1},
-      min color/.style = {min colour=#1},
-      mid color/.style = {mid colour=#1},
-      max color/.style = {max colour=#1},
-      text color/.style = {text colour=#1},
-      min/.initial = -1,
-      mid/.initial = 0,
-      max/.initial = 1,
-      slider/.code={%
-         \tikz{\shade[left color=\HVal{min colour},%
-                      right color=\HVal{max colour}]%
-            (current page.south west) rectangle ++(#1,12pt);
-         }%
-      }%
-  }
-  \newcommand\Heatset[1]{\pgfkeys{/heat, #1}}
-  \newcommand\HVal[1]{\pgfkeysvalueof{/heat/#1}}
-
-  \newcolumntype{H}{>{\collectcell\Heat}r<{\endcollectcell}}
-  \newcommand\Heat[1]{% \Heat{number in the interval [min, max] }
-    \if\relax\detokenize{#1}\relax% empty cell
-    \else%
-      \pgfmathparse{int(100*(#1-\HVal{min})/(\HVal{max}-\HVal{min}))}% map number to [0,100]
-      \ifnum\pgfmathresult>100% too big
-        \edef\HeatCell{\noexpand\cellcolor{\HVal{Max colour}}}%
-      \else\ifnum\pgfmathresult<0% too small
-          \edef\HeatCell{\noexpand\cellcolor{\HVal{Min colour}}}%
-      \else\ifnum\pgfmathresult<50% between min and mid
-	      \pgfmathparse{int(2*\pgfmathresult)}% map number to [0,100]
-		  \edef\HeatCell{\noexpand\cellcolor{\HVal{mid colour}!\pgfmathresult!\HVal{min colour}}}%
-		\else% between min and max
-	      \pgfmathparse{int(2*(\pgfmathresult-50))}% map number to [0,100]
-          \edef\HeatCell{\noexpand\cellcolor{\HVal{max colour}!\pgfmathresult!\HVal{mid colour}}}%
-        \fi%
-        \fi%
-      \fi%
-      \HeatCell\textcolor{\HVal{text colour}}{$#1$}%
-    \fi%
-  }
-
- \pgfkeys{/heatsec/.is family, /heatsec,
-	Max colour/.initial = Green4,
-	Min colour/.initial = Red1,
-	max colour/.initial = SpringGreen3,
-	mid colour/.initial = white,
-	min colour/.initial = Yellow1,
-	text colour/.initial = black,
-	Min color/.style = {Min colour=#1},% for our friends who can't spell
-	Max color/.style = {Max colour=#1},
-	min color/.style = {min colour=#1},
-	mid color/.style = {mid colour=#1},
-	max color/.style = {max colour=#1},
-	text color/.style = {text colour=#1},
-	min/.initial = -1,
-	mid/.initial = 0,
-	max/.initial = 1,
-	slider/.code={%
-		\tikz{\shade[left color=\HVal{min colour},%
-			right color=\HVal{max colour}]%
-			(current page.south west) rectangle ++(#1,12pt);
-		}%
-	}%
-}
-\newcommand\HeatSecset[1]{\pgfkeys{/heatsec, #1}}
-\newcommand\HSVal[1]{\pgfkeysvalueof{/heatsec/#1}}
-
-\colorlet{BadCol}{Burlywood1!70!red}
-
-\newcolumntype{S}{>{\collectcell\HeatSec}r<{\endcollectcell}}
-\newcommand\HeatSec[1]{% \Heat{number in the interval [min, max] }
-	\if\relax\detokenize{#1}\relax% empty cell
-	\else%
-	\pgfmathparse{int(100*(#1-\HSVal{min})/(\HSVal{max}-\HSVal{min}))}% map number to [0,100]
-	\ifnum\pgfmathresult>100% too big
-	\edef\HeatCell{\noexpand\cellcolor{\HSVal{Max colour}}}%
-	\else\ifnum\pgfmathresult<0% too small
-	\edef\HeatCell{\noexpand\cellcolor{\HSVal{Min colour}}}%
-	\else\ifnum\pgfmathresult<50% between min and mid
-	\pgfmathparse{int(2*\pgfmathresult)}% map number to [0,100]
-	\edef\HeatCell{\noexpand\cellcolor{\HSVal{mid colour}!\pgfmathresult!\HSVal{min colour}}}%
-	\else% between min and max
-	\pgfmathparse{int(2*(\pgfmathresult-50))}% map number to [0,100]
-	\edef\HeatCell{\noexpand\cellcolor{\HSVal{max colour}!\pgfmathresult!\HSVal{mid colour}}}%
-	\fi%
-	\fi%
-	\fi%
-	\HeatCell\textcolor{\HSVal{text colour}}{$#1$}%
-	\fi%
-}
 
@@
 \usepackage{todonotes}
-\title{Supplementary materials}
-%\author{}
-%\date{} % Activate to display a given date or no date (if empty),
+\usepackage{inconsolata}
+
+\input{macros.tex}
+
+\title{\OurTool{}:  Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles\\[1em]
+	Supplementary material
+}
+\author{}
+\date{} % Activate to display a given date or no date (if empty),
          % otherwise the current date is printed 
-\newcommand{\Software}[1]{\texttt{#1}}
-\newcommand{\OurTool}{\Software{IPANEMAP}}
 \newcommand{\Start}{s}
 \newcommand{\End}{e}
-
-\newcommand{\yp}[1]{{\todo[color=blue!30]{\sf Yann: #1}}}
 
@@
 }
-\begin{table}[h]
+\begin{table}[H]
 \centering
+\newcommand{\B}{\multicolumn{1}{c}{$\bullet$}}
+\newcommand{\BR}{\multicolumn{1}{c}{$\bullet$}}
+\newcommand{\D}{\multicolumn{1}{c}{\color{gray}$\circ$}}
+\newcommand{\DR}{\multicolumn{1}{c}{\color{gray}$\circ$}}
 \begin{adjustbox}{max width=1\textwidth}
- \begin{tabular}{lHHHHHHHHHHHHHHHHHHHHH}
- &\multicolumn{8}{l}{ \OurTool{}} & \multicolumn{4}{l}{ RNAfold MFE} & \multicolumn{4}{l}{RNAfold MEA} \\
-% & \multicolumn{3}{l}{ GM }  & \multicolumn{3}{l}{ GM } & \multicolumn{3}{l}{ GM }& \multicolumn{3}{l}{ GM }  & \multicolumn{3}{l}{ GM } & \multicolumn{3}{l}{ GM } &\multicolumn{3}{l}{ GM }\\
+ \begin{tabular}{@{}lHHHHHHHHHHHHHHHH@{}}
  	\toprule
-\textbf{RNA }& \multicolumn{1}{l}{(-)} & \multicolumn{1}{l}{CMCT} & \multicolumn{1}{l}{NMIA} & \multicolumn{1}{l}{DMS} & \multicolumn{1}{l}{DMS+NMIA} & \multicolumn{1}{l}{DMS+CMCT}& \multicolumn{1}{l}{NMIA+CMCT}	 & \multicolumn{1}{l}{DMS+NMIA+CMCT} & \multicolumn{1}{l}{(-)}& \multicolumn{1}{l}{DMS }& \multicolumn{1}{l}{CMCT} &\multicolumn{1}{l}{ NMIA}& \multicolumn{1}{l}{(-)} & \multicolumn{1}{l}{DMS} & \multicolumn{1}{l}{CMCT} & \multicolumn{1}{l}{NMIA}\\
-\midrule
+ 	                              & \multicolumn{8}{l}{ \OurTool{}}                                     & \multicolumn{4}{l}{\Software{RNAfold} -- MFE mode} & \multicolumn{4}{l}{\Software{RNAfold} -- MEA mode}  \\
+ 	\textbf{CMCT}                 & \D    & \B    & \D       & \D    & \D    & \B       & \B    & \BR   & \D       & \D    & \B    & \DR                     & \D    & \D    & \B       & \D                       \\
+ 	\textbf{NMIA}                 & \D    & \D    & \B       & \D    & \B    & \D       & \B    & \BR   & \D       & \D    & \D    & \BR                     & \D    & \D    & \D       & \B                       \\
+ 	\textbf{DMS}                  & \D    & \D    & \D       & \B    & \B    & \B       & \D    & \BR   & \D       & \B    & \D    & \DR                     & \D    & \B    & \D       & \D                       \\ \midrule
  	5SRNA,E.Coli                  & 0.25  & 0.238 & 0.247    & 0.244 & 0.247 & 0.247    & 0.254 & 0.241 & 0.241    & 0.686 & 0.254 & 0.686                   & 0.241 & 0.686 & 0.269    & 0.686                    \\
@@
  	adenineriboswitch,add         & 1     & 0.956 & 1        & 0.977 & 1     & 0.956    & 1     & 1     & 1        & 0.333 & 0.41  & 1                       & 1     & 0.333 & 0.356    & 0.306                    \\
-tRNAphenylalanineyeast		&	0.286	&	0.746	&	0.976	&	0.746	&	0.976	&	0.746	&	0.976	&	0.976	&	0.976	&	0.976	&	0.976	&	0.976	&	0.334	&	0.976	&	0.976	&	0.976	\\
-
-\midrule
-Average&0.604&0.7&0.77&0.76&0.77&0.705&0.77&0.785&0.69&0.65&0.58&0.74&0.63&0.69&0.62&0.68\\
-\bottomrule
+ 	tRNAphenylalanineyeast        & 0.286 & 0.746 & 0.976    & 0.746 & 0.976 & 0.746    & 0.976 & 0.976 & 0.976    & 0.976 & 0.976 & 0.976                   & 0.334 & 0.976 & 0.976    & 0.976                    \\ \midrule
+ 	Average                       & 0.604 & 0.7   & 0.77     & 0.76  & 0.77  & 0.705    & 0.77  & 0.785 & 0.69     & 0.65  & 0.58  & 0.74                    & 0.63  & 0.69  & 0.62     & 0.68                     \\ \bottomrule
  \end{tabular}
 \end{adjustbox}
-\caption{ GM of the predicted structures with \OurTool{} compared to RNAfold MFE and MEA resulting structures}
+\caption{GM of the structures predicted with \OurTool{}, compared to predictions of \Software{RNAfold} in energy mimization (MFE) and accuracy maximization (MEA) modes.}
 %( ref. $supplementary\_data/benchmark\_6RNAs$)}
@@
 \subsection{Reproducibility}
-\begin{table}[h]
+\begin{table}[H]
 \centering
 \begin{adjustbox}{max width=1\textwidth}
- \begin{tabular}{lHHHHHHHHHHHHHHHHHHHHH}
+ \begin{tabular}{lHHHHHHHHHHHHHHHHHHHHH}\toprule
  &\multicolumn{3}{l}{ NMIA} & \multicolumn{3}{l}{ DMS} & \multicolumn{3}{l}{CMCT} &\multicolumn{3}{l}{ NMIA+DMS} & \multicolumn{3}{l}{ NMIA+CMCT} & \multicolumn{3}{l}{DMS+CMCT}& \multicolumn{3}{l}{NMIA+DMS+CMCT} \\
 % & \multicolumn{3}{l}{ GM }  & \multicolumn{3}{l}{ GM } & \multicolumn{3}{l}{ GM }& \multicolumn{3}{l}{ GM }  & \multicolumn{3}{l}{ GM } & \multicolumn{3}{l}{ GM } &\multicolumn{3}{l}{ GM }\\
-\toprule
-\textbf{RNA }& \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3} & \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3}& \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3} & \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3}& \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3} & \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3}& \multicolumn{1}{l}{run 1} & \multicolumn{1}{l}{run 2} & \multicolumn{1}{l}{run 3}\\
+
+Replicates& \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3} & \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3}& \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3} & \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3}& \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3} & \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3}& \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3}\\
 \midrule
@@
 
+\newpage
+
 %\section{IPANEMAP Vs Rsample}
-\section{Hadjin ~\emph{et al} dataset}
+\section{Detailed results on Hadjin~\emph{et al} dataset}
 
-  \Heatset{min=0.4,  
-	max=1,   
-	max colour=Aquamarine3, % colour at maximum
-	min colour=BadCol,      % colour at minimum
-	Min colour=BadCol, % colour for values below min
-	Max colour=Aquamarine3   % colour for values above max
-}
-\begin{table}
-\begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
-\begin{tabular}{llllHllH}
+ \begin{table}[h]
+
+{\centering
+\begin{adjustbox}{max width=1\textwidth}
+ \begin{tabular}{@{}llll@{}}
  	\toprule
-\textbf{RNA} & Length &  \multicolumn{3}{l}{\Software{Rsample}} &  \multicolumn{3}{l}{\Software{IPANEMAP}} \\
-\midrule
-	&		&	Sensitivity	&	PPV	&	\multicolumn1c{GM}	&	Sensitivity	&	PPV	&	\multicolumn1c{GM}	\\
-A	&	34	&	0.625	&	1	&	0.791	&	0.625	&	1	&	0.791	\\
-B	&	530	&	0.6622	&	0.6203	&	0.641	&	0.8986	&	0.8062	&	0.851	\\
-C	&	301	&	0.03	&	0.0313	&	0.031	&	0.6	&	0.6	&	0.6	\\
-D	&	214	&	0.6349	&	0.6667	&	0.651	&	0.7302	&	0.8519	&	0.789	\\
-E	&	500	&	0.6382	&	0.7823	&	0.707	&	0.5263	&	0.5517	&	0.539	\\
-F	&	47	&	0.4	&	0.75	&	0.548	&	0.6	&	0.75	&	0.671	\\
-G	&	118	&	0.7179	&	0.8	&	0.758	&	0.7692	&	0.8333	&	0.801	\\
-H	&	76	&	0.4762	&	0.5556	&	0.514	&	0.7619	&	0.7619	&	0.762	\\
-I	&	155	&	0.6786	&	0.8261	&	0.749	&	0.9821	&	0.9821	&	0.982	\\
-J	&	79	&	0.7727	&	0.85	&	0.81	&	0.9545	&	0.875	&	0.914	\\
-K	&	71	&	1	&	1	&	1	&	1	&	0.9545	&	0.977	\\
-L	&	66	&	0.5625	&	0.6923	&	0.624	&	0.625	&	0.7143	&	0.668	\\
-M	&	79	&	0.6538	&	0.8095	&	0.727	&	0.6538	&	0.68	&	0.667	\\
-N	&	120	&	0.2857	&	0.2632	&	0.274	&	0.9714	&	0.9189	&	0.945	\\
-O	&	97	&	0.9286	&	0.9286	&	0.929	&	0.9286	&	0.8966	&	0.912	\\
-P	&	511	&	0.9328	&	0.7762	&	0.851	&	0.8908	&	0.7681	&	0.827	\\
-Q	&	401	&	0.5826	&	0.5583	&	0.57	&	0.7304	&	0.75	&	0.74	\\
-R	&	425	&	0.8397	&	0.8271	&	0.833	&	0.9084	&	0.8815	&	0.895	\\
-S	&	336	&	0.5481	&	0.6129	&	0.58	&	0.8077	&	0.866	&	0.836	\\
-T	&	473	&	0.7986	&	0.7143	&	0.755	&	0.8819	&	0.8194	&	0.85	\\
-U	&	412	&	0.5379	&	0.6174	&	0.576	&	0.75	&	0.8462	&	0.797	\\
-V	&	75	&	0.619	&	0.6842	&	0.651	&	0.619	&	0.5652	&	0.591	\\
-W	&	174	&	0.8095	&	0.8644	&	0.836	&	0.8095	&	0.9808	&	0.891	\\
-X	&	154	&	0.875	&	0.913	&	0.894	&	0.875	&	0.913	&	0.894	\\
-
+ 	RNA ID & Description                      \\ \midrule
+ 	A      & PreQ1riboswitchBsubtilis         & 	M      & SARScoronaviruspseudoknot        \\
+ 	B      & 5domainof16SrRNAEcoli            & 	N      & 5SrRNAEcoli                      \\
+ 	C      & SignalrecognitonparticleRNAhuman & 	O      & cyclicdiGMPriboswitchVcholerae   \\
+ 	D      & GroupIintronAzoarcussp           & 	P      & 5domainof23SrRNAEcoli            \\
+ 	E      & HIV15primepseudoknotdomain       & 	Q      & RNasePBsubtilis                  \\
+ 	F      & Telomerasepseudoknothuman        & 	R      & GroupIIntronTthermophila         \\
+ 	G      & SAMIriboswitchTtengcongensis     & 	S      & HepatitisCvirusIRESdomain        \\
+ 	H      & tRNA phe Ecoli                   & 	T      & 5domainof16SrRNAHvolcanii        \\
+ 	I      & P546domainbI3groupIintron        & 	U      & GroupIIintronOiheyensis          \\
+ 	J      & TPPriboswitchEcoli               & 	V      & tRNAaspyeast                     \\
+ 	K      & AdenineriboswitchVvulnificus     & 	W      & LysineriboswitchTmaritime        \\
+ 	L      & FluorideriboswitchPsyringae      & 	X      & MBoxriboswitchBsubtilis          \\
  	\bottomrule
  \end{tabular}
-\end{adjustbox}
-\caption{\OurTool~ prediction performance compared to \Software{Rsample} tool}
-\end{table}
+\end{adjustbox}\\}
 
- \begin{table}[h]
-\centering
-\begin{adjustbox}{max width=1\textwidth}
- \begin{tabular}{@{}ll@{}}
- \toprule
-Rna ID & Description\\
- \midrule
-A	&	PreQ1riboswitchBsubtilis\\
-B	&	5domainof16SrRNAEcoli\\
-C	&	SignalrecognitonparticleRNAhuman\\
-D	&	GroupIintronAzoarcussp\\
-E	&	HIV15primepseudoknotdomain\\
-F	&	Telomerasepseudoknothuman\\
-G	&	SAMIriboswitchTtengcongensis\\
-H	&	tRNApheEcoli\\
-I	&	P546domainbI3groupIintron\\
-J	&	TPPriboswitchEcoli\\
-K	&	AdenineriboswitchVvulnificus\\
-L	&	FluorideriboswitchPsyringae\\
-M	&	SARScoronaviruspseudoknot\\
-N	&	5SrRNAEcoli\\
-O	&	cyclicdiGMPriboswitchVcholerae\\
-P	&	5domainof23SrRNAEcoli\\
-Q	&	RNasePBsubtilis\\
-R	&	GroupIIntronTthermophila\\
-S	&	HepatitisCvirusIRESdomain\\
-T	&	5domainof16SrRNAHvolcanii\\
-U	&	GroupIIintronOiheyensis\\
-V	&	tRNAaspyeast\\
-W	&	LysineriboswitchTmaritime\\
-X	&	MBoxriboswitchBsubtilis\\
-\bottomrule
-\end{tabular}
-\end{adjustbox}
-\caption{24 RNAs from Hadjin \emph{ et al.} dataset.}
+\caption{Labels used for the 24 RNAs in the Hadjin \emph{ et al.} dataset~\cite{Hajdin2013}.}
 \label{clyu}
@@
 \end{table}
-%  \Heatset{min=0,  
-%	max=1,   
-%	max colour=Aquamarine3, % colour at maximum
-%	min colour=BadCol,      % colour at minimum
-%	Min colour=BadCol, % colour for values below min
-%	Max colour=Aquamarine3   % colour for values above max
-%}
 
-\begin{table}
-\begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
+\begin{table}[H]
+	\begin{adjustbox}{max width=1\textwidth,max totalheight=\textheight,keepaspectratio}
 		\begin{tabular}{llHHHHHHHHHHHlHH}
@@
 
+\Heatset{min=0.4,  
+	max=1,   
+	max colour=Aquamarine3, % colour at maximum
+	min colour=BadCol,      % colour at minimum
+	Min colour=BadCol, % colour for values below min
+	Max colour=Aquamarine3   % colour for values above max
+}
+\begin{table}[H]
+{\centering \begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
+\begin{tabular}{llllHllH}
+	\toprule
+	      &        & \multicolumn{3}{l}{\Software{Rsample}}    & \multicolumn{3}{l}{\Software{IPANEMAP}}   \\
+	{RNA} & Length & Sensitivity & PPV    & \multicolumn1c{GM} & Sensitivity & PPV    & \multicolumn1c{GM} \\ \midrule
+	A     & 34     & 0.625       & 1      & 0.791              & 0.625       & 1      & 0.791              \\
+	B     & 530    & 0.6622      & 0.6203 & 0.641              & 0.8986      & 0.8062 & 0.851              \\
+	C     & 301    & 0.03        & 0.0313 & 0.031              & 0.6         & 0.6    & 0.6                \\
+	D     & 214    & 0.6349      & 0.6667 & 0.651              & 0.7302      & 0.8519 & 0.789              \\
+	E     & 500    & 0.6382      & 0.7823 & 0.707              & 0.5263      & 0.5517 & 0.539              \\
+	F     & 47     & 0.4         & 0.75   & 0.548              & 0.6         & 0.75   & 0.671              \\
+	G     & 118    & 0.7179      & 0.8    & 0.758              & 0.7692      & 0.8333 & 0.801              \\
+	H     & 76     & 0.4762      & 0.5556 & 0.514              & 0.7619      & 0.7619 & 0.762              \\
+	I     & 155    & 0.6786      & 0.8261 & 0.749              & 0.9821      & 0.9821 & 0.982              \\
+	J     & 79     & 0.7727      & 0.85   & 0.81               & 0.9545      & 0.875  & 0.914              \\
+	K     & 71     & 1           & 1      & 1                  & 1           & 0.9545 & 0.977              \\
+	L     & 66     & 0.5625      & 0.6923 & 0.624              & 0.625       & 0.7143 & 0.668              \\
+	M     & 79     & 0.6538      & 0.8095 & 0.727              & 0.6538      & 0.68   & 0.667              \\
+	N     & 120    & 0.2857      & 0.2632 & 0.274              & 0.9714      & 0.9189 & 0.945              \\
+	O     & 97     & 0.9286      & 0.9286 & 0.929              & 0.9286      & 0.8966 & 0.912              \\
+	P     & 511    & 0.9328      & 0.7762 & 0.851              & 0.8908      & 0.7681 & 0.827              \\
+	Q     & 401    & 0.5826      & 0.5583 & 0.57               & 0.7304      & 0.75   & 0.74               \\
+	R     & 425    & 0.8397      & 0.8271 & 0.833              & 0.9084      & 0.8815 & 0.895              \\
+	S     & 336    & 0.5481      & 0.6129 & 0.58               & 0.8077      & 0.866  & 0.836              \\
+	T     & 473    & 0.7986      & 0.7143 & 0.755              & 0.8819      & 0.8194 & 0.85               \\
+	U     & 412    & 0.5379      & 0.6174 & 0.576              & 0.75        & 0.8462 & 0.797              \\
+	V     & 75     & 0.619       & 0.6842 & 0.651              & 0.619       & 0.5652 & 0.591              \\
+	W     & 174    & 0.8095      & 0.8644 & 0.836              & 0.8095      & 0.9808 & 0.891              \\
+	X     & 154    & 0.875       & 0.913  & 0.894              & 0.875       & 0.913  & 0.894              \\ \bottomrule
+\end{tabular}
+\end{adjustbox}\\}
+
+\caption{\OurTool~ prediction performance compared to \Software{Rsample} tool}
+\end{table}
+
+%  \Heatset{min=0,  
+%	max=1,   
+%	max colour=Aquamarine3, % colour at maximum
+%	min colour=BadCol,      % colour at minimum
+%	Min colour=BadCol, % colour for values below min
+%	Max colour=Aquamarine3   % colour for values above max
+%}
+
 %  \HeatSecset{min=0,   
@@
 
-\section{Lariat capping ribozyme in {\itshape Didymium iridis}}
-\begin{table}
+\newpage 
+\section{Detailed results of Lariat capping ribozyme from {\itshape Didymium iridis}}
+
+\begin{table}[H]
+ \Heatset{min=0.2,  
+	max=.5,   
+	max colour=BadCol, % colour at maximum
+	min colour=Aquamarine3,      % colour at minimum
+	Min colour=Aquamarine3, % colour for values below min
+	Max colour=BadCol   % colour for values above max
+}
+\begin{center}
+\begin{tabular}{lHH}\toprule
+ & \multicolumn{2}{c}{Distance of reactivity}\\
+Condition & \multicolumn1c{to prediction}&\multicolumn1c{to native}\\
+\toprule
+\OneMSevILUMg & 0.234 & 0.2872\\
+\OneMSevILU &	0.2784 & 0.2961\\
+\NMIAMg &	0.2606 & 0.2518\\
+\NMIA &	0.2961 & 0.3316\\
+\OneMSevILUThreeMg &	0.4291 & 0.3936\\
+\NMIAMgCE &	0.4291 & 0.3404\\
+\NAIMg 	& 0.3493 & 0.305\\
+\BzCNMg & 	0.3759 & 0.2429\\
+\OneMSevMgCE &	0.4379 & 0.3759\\
+\CMCTMg &	0.4911 & 0.4113\\
+\OneMSevILUThree &	0.2961 & 0.3404\\
+\DMSMg &	0.3493 & 0.3227\\
+\NMIACE 	&0.2695 &0.2518 \\
+\OneMSevCE &	0.3582 & 	0.3316\\
+\BzCN &	0.4911 & 	0.4734\\
+\NAICE &	0.4025 &0.4025\\
+\bottomrule
+\end{tabular}
+\end {center}
+\caption{Distance of the predicted structure from the considered probing data.}
+\end{table}
+
+\begin{table}[H]
+\newcommand{\MyScale}{.18}
+\begin{adjustbox}{max width=\textwidth,keepaspectratio}
+\begin{tabular}{@{}c|cccc@{}}
+
+%\toprule
+ \includegraphics[scale=\MyScale]{graphs/Supp_structures/native_structure}&
+ \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILU_NMIAMg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7MgCE_NMIAMgCE}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/NaiMg_CMCTMg}\\
+ \midrule
+Native&\OneMSevILU&\OneMSevILUMg&\OneMSevMgCE&\NAIMg\\
+structure &+\NMIAMg&+\OneMSevILUThreeMg&+\NMIAMgCE&+\CMCTMg\\
+\end{tabular}
+\end{adjustbox}
+\caption{Predicted structures with conditions belonging to the same cluster}
+\end{table}
+
+\begin{table}[H]
+\newcommand{\MA}[1]{\multicolumn{1}{c}{\Acc{#1}}}
+\newcommand{\Acc}[1]{{\bfseries #1}}
 \begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
@@
 	\toprule
-&&\multicolumn{1}{l}{1M7ILUMg}  &\multicolumn{1}{l}{1M7ILU } &\multicolumn{1}{l}{NMIAMg}  &\multicolumn{1}{l}{NMIA } &\multicolumn{1}{l}{1M7ILU3Mg}  &\multicolumn{1}{l}{NMIAMgCE}  &\multicolumn{1}{l}{NaiMg} &\multicolumn{1}{l}{BzCNMg } &\multicolumn{1}{l}{1M7MgCE}&\multicolumn{1}{l}{CMCTMg}  &\multicolumn{1}{l}{1M7ILU3}  &\multicolumn{1}{l}{DMSMg}&\multicolumn{1}{l}{NMIACE }&\multicolumn{1}{l}{1M7CE}&\multicolumn{1}{l}{BzCN}&\multicolumn{1}{l}{Nai}  \\
-\midrule
-&&0.85&0.84&0.81&0.8&0.74&0.73&0.73&0.71&0.7&0.7&0.64&0.63&0.61&0.6&0.6&0.6\\
-1M7ILUMg  &0.85&0.85&0.84&0.8&0.82&0.82&0.82&0.75&0.79&0.78&0.79&0.79&0.85&0.61&0.79&0.79&0.85\\
-1M7ILU  &0.84&&0.84&0.82&0.82&0.7&0.83&0.72&0.82&0.85&0.83&0.84&0.84&0.84&0.6&0.84&0.6\\
-NMIAMg  &0.81&&&0.81&0.8&0.74&0.82&0.82&0.81&0.81&0.79&0.8&0.8&0.61&0.81&0.81&0.6\\
-NMIA  &0.8&&&&0.8&0.74&0.73&0.81&0.78&0.7&0.79&0.63&0.72&0.61&0.7&0.7&0.59\\
-1M7ILU3Mg  &0.74&&&&&0.74&0.74&0.7&0.74&0.74&0.7&0.63&0.74&0.74&0.74&0.74&0.74\\
-NMIAMgCE  &0.73&&&&&&0.73&0.73&0.73&0.73&0.73&0.61&0.73&0.61&0.73&0.73&0.6\\
-NaiMg &0.73&&&&&&&0.73&0.72&0.7&0.73&0.64&0.68&0.61&0.73&0.73&0.59\\
-BzCNMg  &0.71&&&&&&&&0.71&0.65&0.73&0.63&0.72&0.6&0.61&0.6&0.59\\
-1M7MgCE&0.7&&&&&&&&&0.7&0.7&0.7&0.7&0.61&0.61&0.6&0.6\\
-CMCTMg  &0.7&&&&&&&&&&0.7&0.63&0.74&0.61&0.6&0.6&0.6\\
-1M7ILU3  &0.64&&&&&&&&&&&0.64&0.62&0.62&0.65&0.65&0.59\\
-DMSMg&0.63&&&&&&&&&&&&0.63&0.61&0.6&0.6&0.6\\
-NMIACE&0.61&&&&&&&&&&&&&0.61&0.62&0.6&0.6\\
-1M7CE&0.6&&&&&&&&&&&&&&0.6&0.61&0.59\\
-BzCN&0.6&&&&&&&&&&&&&&&0.6&0.6\\
-Nai  &0.6&&&&&&&&&&&&&&&&0.6\\
-\bottomrule
+	                              & \multicolumn{1}{l}{Mono} & \MA{1} & \MA{2} & \MA{3} & \MA{4} & \MA{5} & \MA{6} & \MA{7} & \MA{8} & \MA{9} & \MA{10} & \MA{11} & \MA{12} & \MA{13} & \MA{14} & \MA{15} & \MA{16} \\ \midrule
+	%Mono                         &                          & 0.85   & 0.84   & 0.81   & 0.8    & 0.74   & 0.73   & 0.73   & 0.71   & 0.7    & 0.7     & 0.64    & 0.63    & 0.61    & 0.6     & 0.6     & 0.6     \\
+	\Acc{1} -- \OneMSevILUMg      & 0.85                     & 0.85   & 0.84   & 0.8    & 0.82   & 0.82   & 0.82   & 0.75   & 0.79   & 0.78   & 0.79    & 0.79    & 0.85    & 0.61    & 0.79    & 0.79    & 0.85    \\
+	\Acc{2} -- \OneMSevILU        & 0.84                     &        & 0.84   & 0.82   & 0.82   & 0.7    & 0.83   & 0.72   & 0.82   & 0.85   & 0.83    & 0.84    & 0.84    & 0.84    & 0.6     & 0.84    & 0.6     \\
+	\Acc{3} -- \NMIAMg            & 0.81                     &        &        & 0.81   & 0.8    & 0.74   & 0.82   & 0.82   & 0.81   & 0.81   & 0.79    & 0.8     & 0.8     & 0.61    & 0.81    & 0.81    & 0.6     \\
+	\Acc{4} -- \NMIA              & 0.8                      &        &        &        & 0.8    & 0.74   & 0.73   & 0.81   & 0.78   & 0.7    & 0.79    & 0.63    & 0.72    & 0.61    & 0.7     & 0.7     & 0.59    \\
+	\Acc{5} -- \OneMSevILUThreeMg & 0.74                     &        &        &        &        & 0.74   & 0.74   & 0.7    & 0.74   & 0.74   & 0.7     & 0.63    & 0.74    & 0.74    & 0.74    & 0.74    & 0.74    \\
+	\Acc{6} -- \NMIAMgCE          & 0.73                     &        &        &        &        &        & 0.73   & 0.73   & 0.73   & 0.73   & 0.73    & 0.61    & 0.73    & 0.61    & 0.73    & 0.73    & 0.6     \\
+	\Acc{7} -- \NAIMg             & 0.73                     &        &        &        &        &        &        & 0.73   & 0.72   & 0.7    & 0.73    & 0.64    & 0.68    & 0.61    & 0.73    & 0.73    & 0.59    \\
+	\Acc{8} -- \BzCNMg            & 0.71                     &        &        &        &        &        &        &        & 0.71   & 0.65   & 0.73    & 0.63    & 0.72    & 0.6     & 0.61    & 0.6     & 0.59    \\
+	\Acc{9} -- \OneMSevMgCE       & 0.7                      &        &        &        &        &        &        &        &        & 0.7    & 0.7     & 0.7     & 0.7     & 0.61    & 0.61    & 0.6     & 0.6     \\
+	\Acc{10} -- \CMCTMg           & 0.7                      &        &        &        &        &        &        &        &        &        & 0.7     & 0.63    & 0.74    & 0.61    & 0.6     & 0.6     & 0.6     \\
+	\Acc{11} -- \OneMSevILUThree  & 0.64                     &        &        &        &        &        &        &        &        &        &         & 0.64    & 0.62    & 0.62    & 0.65    & 0.65    & 0.59    \\
+	\Acc{12} -- \DMSMg            & 0.63                     &        &        &        &        &        &        &        &        &        &         &         & 0.63    & 0.61    & 0.6     & 0.6     & 0.6     \\
+	\Acc{13} -- \NMIACE           & 0.61                     &        &        &        &        &        &        &        &        &        &         &         &         & 0.61    & 0.62    & 0.6     & 0.6     \\
+	\Acc{14} -- \OneMSevCE        & 0.6                      &        &        &        &        &        &        &        &        &        &         &         &         &         & 0.6     & 0.61    & 0.59    \\
+	\Acc{15} -- \BzCN             & 0.6                      &        &        &        &        &        &        &        &        &        &         &         &         &         &         & 0.6     & 0.6     \\
+	\Acc{16} -- \NAICE            & 0.6                      &        &        &        &        &        &        &        &        &        &         &         &         &         &         &         & 0.6     \\ \bottomrule
 \end{tabular}
 \end{adjustbox}
-\caption{MCC of predicted structures with \OurTool{} in the case of Bi-probing}.
+\caption{MCC of structures prediction with \OurTool{} from any pair of conditions}.
 \end{table}
 
-\begin{table}
-\newcommand{\MyScale}{.2}
-\begin{tabular}{@{}cccc@{}}
+%\begin{table}
+%\begin{center}
+%\begin{tabular}{ll}
+%Sample size & MCC\\
 %\toprule
- \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILU_NMIAMg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7MgCE_NMIAMgCE}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/NaiMg_CMCTMg}\\
- \midrule
-1M7ILU&1M7ILUMg&1M7MgCE&NaiMg\\
-+NMIAMg&+1M7ILU3Mg&+NMIAMgCE&+CMCTMg\\
-\end{tabular}
-\caption{Predicted structures with conditions belonging to the same cluster}
-\end{table}
+%
+%100 & 0.55 \\
+%500 & 0.55  \\
+%1000 &  0.81\\
+%2000 & 	0.55  \\
+%3000 &  0.77 \\
+%5000 & 	0.55  \\ 
+%10000& 0.8 \\
+%\end{tabular}
+%\end {center}
+%\caption{ MCC of predicted structures in absence of probig data when varyingthe sample size }
+%\end{table}
 
-\subsection{Distribution of MCC values over 100 runs}
-\includegraphics[width=\linewidth]{graphs/histog}
+\begin{figure}[H]
+%\includegraphics[width=\linewidth]{graphs/histog}
+\includegraphics[width=\linewidth]{graphs/histog1}
+\caption{Distribution of MCC values over 100 runs from Mutate-and-Map data}
+\end{figure}
+
+%\bibliographystyle{bioinformatics}
+\bibliographystyle{nar}
+%\bibliographystyle{natbib}%compa
+\bibliography{biblio}
+
 \end{document}

### 2019-06-19 SuppMat.tex

@@
 \geometry{a4paper} % or letterpaper (US) or a5paper or....
-% \geometry{margin=2in} % for example, change the margins to 2 inches all round
+ \geometry{margin=1in} % for example, change the margins to 2 inches all round
 % \geometry{landscape} % set up the page for landscape
@@
 \usepackage{todonotes}
+\usepackage{times}
 \usepackage{inconsolata}
+
+\usepackage{rotating}
 
@@
 
-\section{Detailed results on Cordero~\emph{et al} dataset}
-\subsection{Structure prediction ordered by RNA}
+\section*{Multiprobing benchmark (Cordero~\emph{et al}~\cite{Cordero2012} dataset)}
+\subsection*{Structure prediction ordered by RNA}
 
@@
 \end{table}
-\subsection{Reproducibility}
+\subsection*{Stablity analysis}
 \begin{table}[H]
@@
 \end{adjustbox}
-\caption{ GM of the predicted structures with \OurTool{} over 3 runs, in the presence of 1,2 and 3 sources of probing data}
+\caption{GM of \OurTool predictions over three consecutive runs from up 3 sources of probing data.}
 %( ref. $supplementary\_data/benchmark\_6RNAs$)
@@
 
-\newpage
+%\section{IPANEMAP Vs Rsample}
+\section*{Mono probing benchmark (Hadjin~\emph{et al}~\cite{Hajdin2013} dataset)}
 
-%\section{IPANEMAP Vs Rsample}
-\section{Detailed results on Hadjin~\emph{et al} dataset}
-
- \begin{table}[h]
+ \begin{table}[H]
 
@@
 
-\caption{Labels used for the 24 RNAs in the Hadjin \emph{ et al.} dataset~\cite{Hajdin2013}.}
+\caption{Labels used as shorthands for the 24 RNAs in the Hadjin \emph{ et al.} dataset~\cite{Hajdin2013}.}
 \label{clyu}
@@
 \end{table}
+
+\subsection*{Detailed results of \OurTool{} and \Software{Rsample}}
+\begin{table}[H]
+	{\centering \begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
+			\begin{tabular}{llllHllH}
+				\toprule
+				&        & \multicolumn{3}{l}{\Software{Rsample}}    & \multicolumn{3}{l}{\Software{IPANEMAP}}   \\
+				{RNA} & Length & Sensitivity & PPV    & \multicolumn1c{GM} & Sensitivity & PPV    & \multicolumn1c{GM} \\ \midrule
+				A     & 34     & 0.625       & 1      & 0.791              & 0.625       & 1      & 0.791              \\
+				B     & 530    & 0.6622      & 0.6203 & 0.641              & 0.8986      & 0.8062 & 0.851              \\
+				C     & 301    & 0.03        & 0.0313 & 0.031              & 0.6         & 0.6    & 0.6                \\
+				D     & 214    & 0.6349      & 0.6667 & 0.651              & 0.7302      & 0.8519 & 0.789              \\
+				E     & 500    & 0.6382      & 0.7823 & 0.707              & 0.5263      & 0.5517 & 0.539              \\
+				F     & 47     & 0.4         & 0.75   & 0.548              & 0.6         & 0.75   & 0.671              \\
+				G     & 118    & 0.7179      & 0.8    & 0.758              & 0.7692      & 0.8333 & 0.801              \\
+				H     & 76     & 0.4762      & 0.5556 & 0.514              & 0.7619      & 0.7619 & 0.762              \\
+				I     & 155    & 0.6786      & 0.8261 & 0.749              & 0.9821      & 0.9821 & 0.982              \\
+				J     & 79     & 0.7727      & 0.85   & 0.81               & 0.9545      & 0.875  & 0.914              \\
+				K     & 71     & 1           & 1      & 1                  & 1           & 0.9545 & 0.977              \\
+				L     & 66     & 0.5625      & 0.6923 & 0.624              & 0.625       & 0.7143 & 0.668              \\
+				M     & 79     & 0.6538      & 0.8095 & 0.727              & 0.6538      & 0.68   & 0.667              \\
+				N     & 120    & 0.2857      & 0.2632 & 0.274              & 0.9714      & 0.9189 & 0.945              \\
+				O     & 97     & 0.9286      & 0.9286 & 0.929              & 0.9286      & 0.8966 & 0.912              \\
+				P     & 511    & 0.9328      & 0.7762 & 0.851              & 0.8908      & 0.7681 & 0.827              \\
+				Q     & 401    & 0.5826      & 0.5583 & 0.57               & 0.7304      & 0.75   & 0.74               \\
+				R     & 425    & 0.8397      & 0.8271 & 0.833              & 0.9084      & 0.8815 & 0.895              \\
+				S     & 336    & 0.5481      & 0.6129 & 0.58               & 0.8077      & 0.866  & 0.836              \\
+				T     & 473    & 0.7986      & 0.7143 & 0.755              & 0.8819      & 0.8194 & 0.85               \\
+				U     & 412    & 0.5379      & 0.6174 & 0.576              & 0.75        & 0.8462 & 0.797              \\
+				V     & 75     & 0.619       & 0.6842 & 0.651              & 0.619       & 0.5652 & 0.591              \\
+				W     & 174    & 0.8095      & 0.8644 & 0.836              & 0.8095      & 0.9808 & 0.891              \\
+				X     & 154    & 0.875       & 0.913  & 0.894              & 0.875       & 0.913  & 0.894              \\ \bottomrule
+			\end{tabular}
+		\end{adjustbox}\\}
+\caption{\OurTool~ prediction performance compared to \Software{Rsample} tool}
+\end{table}
+
+\subsection*{Stability analysis}
 
@@
 }
-\begin{table}[H]
-{\centering \begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
-\begin{tabular}{llllHllH}
-	\toprule
-	      &        & \multicolumn{3}{l}{\Software{Rsample}}    & \multicolumn{3}{l}{\Software{IPANEMAP}}   \\
-	{RNA} & Length & Sensitivity & PPV    & \multicolumn1c{GM} & Sensitivity & PPV    & \multicolumn1c{GM} \\ \midrule
-	A     & 34     & 0.625       & 1      & 0.791              & 0.625       & 1      & 0.791              \\
-	B     & 530    & 0.6622      & 0.6203 & 0.641              & 0.8986      & 0.8062 & 0.851              \\
-	C     & 301    & 0.03        & 0.0313 & 0.031              & 0.6         & 0.6    & 0.6                \\
-	D     & 214    & 0.6349      & 0.6667 & 0.651              & 0.7302      & 0.8519 & 0.789              \\
-	E     & 500    & 0.6382      & 0.7823 & 0.707              & 0.5263      & 0.5517 & 0.539              \\
-	F     & 47     & 0.4         & 0.75   & 0.548              & 0.6         & 0.75   & 0.671              \\
-	G     & 118    & 0.7179      & 0.8    & 0.758              & 0.7692      & 0.8333 & 0.801              \\
-	H     & 76     & 0.4762      & 0.5556 & 0.514              & 0.7619      & 0.7619 & 0.762              \\
-	I     & 155    & 0.6786      & 0.8261 & 0.749              & 0.9821      & 0.9821 & 0.982              \\
-	J     & 79     & 0.7727      & 0.85   & 0.81               & 0.9545      & 0.875  & 0.914              \\
-	K     & 71     & 1           & 1      & 1                  & 1           & 0.9545 & 0.977              \\
-	L     & 66     & 0.5625      & 0.6923 & 0.624              & 0.625       & 0.7143 & 0.668              \\
-	M     & 79     & 0.6538      & 0.8095 & 0.727              & 0.6538      & 0.68   & 0.667              \\
-	N     & 120    & 0.2857      & 0.2632 & 0.274              & 0.9714      & 0.9189 & 0.945              \\
-	O     & 97     & 0.9286      & 0.9286 & 0.929              & 0.9286      & 0.8966 & 0.912              \\
-	P     & 511    & 0.9328      & 0.7762 & 0.851              & 0.8908      & 0.7681 & 0.827              \\
-	Q     & 401    & 0.5826      & 0.5583 & 0.57               & 0.7304      & 0.75   & 0.74               \\
-	R     & 425    & 0.8397      & 0.8271 & 0.833              & 0.9084      & 0.8815 & 0.895              \\
-	S     & 336    & 0.5481      & 0.6129 & 0.58               & 0.8077      & 0.866  & 0.836              \\
-	T     & 473    & 0.7986      & 0.7143 & 0.755              & 0.8819      & 0.8194 & 0.85               \\
-	U     & 412    & 0.5379      & 0.6174 & 0.576              & 0.75        & 0.8462 & 0.797              \\
-	V     & 75     & 0.619       & 0.6842 & 0.651              & 0.619       & 0.5652 & 0.591              \\
-	W     & 174    & 0.8095      & 0.8644 & 0.836              & 0.8095      & 0.9808 & 0.891              \\
-	X     & 154    & 0.875       & 0.913  & 0.894              & 0.875       & 0.913  & 0.894              \\ \bottomrule
-\end{tabular}
-\end{adjustbox}\\}
-
-\caption{\OurTool~ prediction performance compared to \Software{Rsample} tool}
-\end{table}
 
@@
 
-\newpage 
-\section{Detailed results of Lariat capping ribozyme from {\itshape Didymium iridis}}
-
+\section*{Case-study: Lariat capping ribozyme of {\itshape Didymium iridis}}
+\subsection*{Compatibility of accessibility profiles with native and predicted structures}
 \begin{table}[H]
@@
 
-\begin{table}[H]
-\newcommand{\MyScale}{.18}
-\begin{adjustbox}{max width=\textwidth,keepaspectratio}
+\begin{sidewaysfigure}
+\newcommand{\MyScale}{.26}
+
+{\centering
+	\begin{adjustbox}{max width=\textwidth,max height=\textheight,keepaspectratio}
 \begin{tabular}{@{}c|cccc@{}}
@@
 \end{tabular}
-\end{adjustbox}
+\end{adjustbox}\\}
+
 \caption{Predicted structures with conditions belonging to the same cluster}
-\end{table}
+\end{sidewaysfigure}
 

### 2019-06-20 IPANEMAP-2019.tex
Updates from Overleaf

@@
 
-\usetikzlibrary{shapes.geometric, arrows,calc,fit,positioning}
+\usetikzlibrary{shapes.geometric, arrows, calc, fit, positioning}
 
@@
 and
-$^{2}$\,CitCoM, Cibles thérapeutiques et conception de Médicaments, CNRS, University Paris Descartes, 4 avenue de l'observatoire 75006 PARIS France}
+$^{2}$\,CitCoM, Cibles thérapeutiques et conception de Médicaments, CNRS, Université de Paris, 4 avenue de l'observatoire 75006 PARIS France}
 % Affiliation must include:
@@
 A mon avis il faudrait soit le mettre tel quel comme premier chapitre des résultats, soit en faire une version light sans les détails de calculs le mettre comme premier chapitre des résultats et ne laisser que les calculs dans le Mat et Met}
+\ypi{Dans la mesure où l'intro est courte, et étant donné le titre, je ne pense pas que sa présence en Mat Met limite  sur visibilité.
+En plus, c'est typiquement là où tu le trouves d'habitude dans ce type de papier (cf Rsample).}
 We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signal produced by various chemical probes.
@@
 
-\paragraph{Hadjin \emph{et al} dataset.} 
-A dataset was gathered by Hadjin\etal\cite{Hajdin2013} to validate the predictive capacities of probing data-driven predictions.
+\paragraph{Hajdin \emph{et al} dataset.} 
+A dataset was gathered by Hajdin\etal\cite{Hajdin2013} to validate the predictive capacities of probing data-driven predictions.
 It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (SHAPE -- 1M7) is available.
@@
 \subsection*{Experimental probing protocols}
-The structure of the Lariat capping ribozyme from \textit{Didymium iridis }(DiLCrz) was probed with different SHAPE reagents: 1M7 (1-methyl-7 nitrosatoic anhydre), NMIA (N-Methyl Isatoic Anhydre), BzCN (Benzoyl cyanide) and NAI (2-methylnicotinic acid imidazolide) in presence and absence of magnesium ions, and with DMS (DiMethylSulfate) and CMCT only in presence of magnesium ions.
+The structure of the Lariat capping ribozyme from \textit{Didymium iridis }(DiLCrz) was probed with different SHAPE reagents: 1M7 (1-methyl-7 nitrosatoic anhydre), NMIA (N-Methyl Isatoic Anhydre), BzCN (Benzoyl cyanide) and NAI (2-methylnicotinic acid imidazolide) in presence and absence of Mg$^\text{2+}$, and with DMS (DiMethylSulfate) and CMCT only in presence of Mg$^\text{2+}$.
 As a preliminary experiment, we verified that DiLCrz generally adopts a single global architecture.
-To this end, we subjected DiLCrz to a standard denaturation protocol (80\degree{}C for 2 min in H2O, addition of 40 mM of HEPES at 7.5 pH, 100 mM of KCl, 5 mM of MgCl2, followed by 10 min at room temperature, and 10 min at 37\degree{}C), and observed the production of a single band on a non-denaturing PAGE, strongly suggesting the adoption of a single conformation.
+To this end, we subjected DiLCrz to a standard denaturation protocol (80\degree{}C for 2 min in H$_\text{2}$O, addition of 40 mM of HEPES at 7.5 pH, 100 mM of KCl, 5 mM of MgCl$_\text{2}$, followed by 10 min at room temperature, and 10 min at 37\degree{}C), and observed the production of a single band on a non-denaturing PAGE, strongly suggesting the adoption of a single conformation.
 
 \paragraph{Stops-inducing probing protocol (SHAPE-CE).} 
-6 pmol of RNA were resuspended in 18 $\mu$l of water, denatured at 80\degree{}C for 2 minutes and cooled down at room temperature during 10 minutes in the probing buffer (40 mM HEPES at 7.5 pH, 100 mM KCl, in presence or absence of 5 mM MgCl2).
+6 pmol of RNA were resuspended in 18 $\mu$l of water, denatured at 80\degree{}C for 2 minutes and cooled down at room temperature during 10 minutes in the probing buffer (40 mM HEPES at 7.5 pH, 100 mM KCl, in presence or absence of 5 mM MgCl$_\text{2}$.
 After a 10 minutes incubation at 37\degree{}C, RNAs were treated with 2 mM of SHAPE reagent or DMSO (negative control) and incubated for 2 (BzCN), 5 (1M7), 30 (NMIA) or 60 (NAI) minutes at 37\degree{}C.
@@
 DMS and CMCT probing were conducted essentially as described in~\cite{Weill2004,James2008}.
-Succinctly, 6 pmol of RNA were resuspended in 18 $\mu$l of water, denatured at 80\degree{}C for 2 minutes and cooled down at room temperature for 10 minutes in a probing buffer for DMS probing (40 mM HEPES pH 7.5, 100 mM KCl, 5 mM MgCl2) or in CMCT (50 mM of potassium borate pH 8, 100 mM KCl and 5 mM MgCl2).
+Succinctly, 6 pmol of RNA were resuspended in 18 $\mu$l of water, denatured at 80\degree{}C for 2 minutes and cooled down at room temperature for 10 minutes in a probing buffer for DMS probing (40 mM HEPES pH 7.5, 100 mM KCl, 5 mM MgCl$_\text{2}$) or in CMCT (50 mM of potassium borate pH 8, 100 mM KCl and 5 mM MgCl$_\text{2}$).
 For DMS, RNA was then treated with 1 $\mu$l of DMS (1:12 in ethanol) or 1 $\mu$l of ethanol for 5 minutes at 37\degree{}C (mock reaction), the reaction was stopped by addition of 400 mM of Tris at 7.5 pH and immediately put on ice.
-For CMCT,  RNA was treated with 10 $\mu$l of CMCT (42mg/mL) or 10 $\mu$l H2O for 10 minutes at 37\degree{}C (mock reaction).
+For CMCT,  RNA was treated with 10 $\mu$l of CMCT (42mg/mL) or 10 $\mu$l H$_\text{2}$O for 10 minutes at 37\degree{}C (mock reaction).
 Modified RNA were ethanol-precipitated and resuspended in 10 $\mu$l of water.
@@
 	\includegraphics[width=\linewidth]{graphs/RsampleVsIPANEMAP/Accuracy}
-	\caption{Geometric mean of positive predictive value and sensitivity achieved by the predictions of \OurTool{} and \Software{Rsample} over the Hadjin\etal\cite{Hajdin2013} data set, consisting of 24 RNAs (RNA length indicated within square brackets).
+	\caption{Geometric mean of positive predictive value and sensitivity achieved by the predictions of \OurTool{} and \Software{Rsample} over the Hajdin\etal\cite{Hajdin2013} data set, consisting of 24 RNAs (RNA length indicated within square brackets).
 \label{fig:Vsrsample}}
@@
 	\caption{Model of the Lariat capping ribozyme from {\itshape Didymium iridis} (A -- Meyer\etal\cite{Meyer2014}), pseudoknot-free secondary structure as annotated by \Software{DSSR} (B -- pseudoknot ommitted) and \OurTool first-ranking prediction (C.) from 8 maximally-diverging probing conditions (NMIAMg, 1M7ILU3Mg, NAIMg, 1M7, CMCTMg,1M7ILU3,1M7Mg).
-\label{fig:didy}}
+\bsi{figure à refaite.
+Faut il montrer les interactions tertiaires?
+} \label{fig:didy}}
 \end{figure*}
@@
 Conditions ordered by spectral biclustering.
-Conditions regrouped within clusters by k-mean clustering appear as blocks on the x and y axis.}\label{fig:biclustering}
+Conditions regrouped within clusters by k-mean clustering appear as blocks on the x and y axis.\bsi{pourquoi les clusters E et F ne sont ils pas regroupés comme les autres?}}\label{fig:biclustering}
 \end{figure}
@@
 
-Unsurprisingly, the presence/absence of Magnesium during the probing impacts the predictions, with an observed drop from an average 73\% MCC in the presence of Mg$^\text{2+}$ to 67\% MCC in the absence of Mg$^\text{2+}$.
+Unsurprisingly, the presence/absence of Magnesium ions during the probing impacts the predictions, with an observed drop from an average 73\% MCC in the presence of Mg$^\text{2+}$ to 67\% MCC in the absence of Mg$^\text{2+}$.
 However, while the poorly performing conditions of cluster \Bull{B} are probed in the absence of Mg$^\text{2+}$, some of the most informative profiles (\OneMSevILU and \NMIA) were obtained in absence of Mg$^\text{2+}$.
@@
 Averaged MCC of structures predicted with \OurTool{} and with classic deterministic alternatives.
-The structure was probed in presence (\Mg) or absence (\NoMg) of Magnesium, using either stops (\Stop) or mutations (\Mut) inducing technologies.
+The structure was probed in presence (\Mg) or absence (\NoMg) of Magnesium ions, using either stops (\Stop) or mutations (\Mut) inducing technologies.
 	}
@@
 We turn to a systematic exploration of the predictive capacities of bi-probing analyses, based on pairs of probing profiles, and attempt to quantify their impact on \OurTool predictions.
-For each pair and triplets of conditions, we executed \OurTool using 1~000 samples per condition, and considered the first returned structure.
+For each pair and triplets of conditions, we executed \OurTool using 1000 samples per condition, and considered the first returned structure.
 We then computed the associated MCC, and compare it with the MCC of the worst performing condition (Min), best-performing condition (Max) and average MCC over the single conditions experiments.
@@
 
-A superficial inspection of the resulting MCCs appears to confirm the conclusions of Section~\ref{sec:cordero}.
+A superficial inspection of the resulting MCCs appears to confirm the conclusions of Section Bruno~\ref{sec:cordero}.
 Indeed, considering two conditions leads to predictions whose quality fall between the worst and the best one, while being generally close to the average.
@@
 Having established the redundant nature of similar conditions, and their limited contribution to downstream modeling, we turn to an analysis of conditions across clusters.
-We run \OurTool{} the quality of predictions obtained by choosing a single condition in each of the 8 clusters.
-We consider any possible combination, except for the outlier cluster~\Bull{B}, for which we only consider \NAICE as a representative.
+We run \OurTool{} the quality of predictions obtained by choosing a single condition in each of the 8 clusters.\bsi{il manque des trucs dans cette phrase (celle qui commence par we run), non?} We consider any possible combination, except for the outlier cluster~\Bull{B}, for which we only consider \NAICE as a representative.
 
@@
 %% Counter({0.6084: 82, 0.6134: 12, 0.7597: 2, 0.7649: 1, 0.7783: 1, 0.7665: 1, 0.7531: 1})
-.%mean(Multiprobing_100WT)
+%mean(Multiprobing_100WT)
 
@@
 However, the bimodal nature of the distribution also suggests the existence of at least two conformations that are both supported by the WT and the mutant profiles.
+\bsi{il y a une conformation alternative connue. Tu pourrais me donner un model provenant des 2 clusters de MCC que je regarde si pas hasard c'est ca?}
 
@@
 To the best of our knowledge, \OurTool currently represents the only available method for a joint automated consideration of multiple probing profiles.
-Surprisingly, it also appears to produce the most accurate secondary structure models when informed by a single condition, as demonstrated by our reanalysis on the Hadjin\etal\cite{Hajdin2013} benchmark.
+Surprisingly, it also appears to produce the most accurate secondary structure models when informed by a single condition, as demonstrated by our reanalysis on the Hajdin\etal\cite{Hajdin2013} benchmark.
 We attribute this unexpected result to the adoption of a sampling/clustering approach, which seems to eliminate spurious outliers in the conformational space.
@@
 However, such native structures include non WC pairings, and complex topological motifs such as pseudoknots, kissing loops, tetraloop interactions, that classic structure prediction algorithms do not explicitly capture.
-In contrast, in the presence of only monovalent ions, RNA is expected to form its helical domains, but leave its  tertiary contacts unstable.
+In contrast, in the presence of only monovalent ions, RNA is expected to form its helical domains, but leave its  tertiary contacts unstable, with the notable exception of the G quadruplex.
 Therefore, probing the RNA without Mg$^{\text{2+}}$ could provide a signal that is less deceptive, and thus more informative, to predicting algorithms.
@@
 \begin{figure}
-	{\centering\includegraphics[width=1\linewidth]{graphs/boxplotreproducibility.png}\\}
+	{\centering\includegraphics[width=1.2\linewidth]{graphs/boxplotreproducibility.png}\\}
 
@@
 \section*{Acknowledgements}
-The authors wish to express their gratitude to Ronny Lorenz for developing a probing-informed version of \Software{RNAsubopt}, allowing to sample the pseudo-Boltzmann ensemble, and Nathalie Chamond for helpful suggestions.
+The authors wish to express their gratitude to Ronny Lorenz for developing a probing-informed version of \Software{RNAsubopt}, allowing to sample the pseudo-Boltzmann ensemble, Nathalie Chamond for helpful suggestions and Benoit Masquida for the gift of DiLCRz plasmid and helpful discussions about protocols and DiLCRz structure .
 

### 2019-06-20 SuppMat.tex
Updates from Overleaf

@@
 	\begin{adjustbox}{max width=\textwidth,max height=\textheight,keepaspectratio}
-\begin{tabular}{@{}c|cccc@{}}
+\begin{tabular}{@{}c|ccccc@{}}
 
@@
  \includegraphics[scale=\MyScale]{graphs/Supp_structures/native_structure}&
- \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILU_NMIAMg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7MgCE_NMIAMgCE}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/NaiMg_CMCTMg}\\
+ \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILU_NMIAMg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}&
+ \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7MgCE_NMIAMgCE}
+ &
+ \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7CE_NMIACE}
+ & \includegraphics[scale=\MyScale]{graphs/Supp_structures/NaiMg_CMCTMg}\\
  \midrule
-Native&\OneMSevILU&\OneMSevILUMg&\OneMSevMgCE&\NAIMg\\
-structure &+\NMIAMg&+\OneMSevILUThreeMg&+\NMIAMgCE&+\CMCTMg\\
+Native&\OneMSevILU&\OneMSevILUMg&\OneMSevMgCE& \1M7CE &\NAIMg\\
+structure &+\NMIAMg&+\OneMSevILUThreeMg&+\NMIAMgCE&+\NMIACE &+\CMCTMg\\
 \end{tabular}

### 2019-06-21 IPANEMAP-2019.tex
Updates from Overleaf

@@
 \ypi{Dans la mesure où l'intro est courte, et étant donné le titre, je ne pense pas que sa présence en Mat Met limite  sur visibilité.
-En plus, c'est typiquement là où tu le trouves d'habitude dans ce type de papier (cf Rsample).}
+En plus, c'est typiquement là où tu le trouves d'habitude dans ce type de papier (cf Rsample).} \bsi{OK je propose la chose suivante:on laisse tt comme ca le math et met, mais on commence les résultats par un mini-chapitre du genre "principles of IPANEMAP" IPANEMAP relies on ..
+prediction with RNAsubopt ...
+sampling ...
+clustering ..
+evaluation by front pareto..
+Its originality relies on ...
+See material and methods for details }
 We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signal produced by various chemical probes.
@@
 $$\Delta G(i) = m \log(R_i +1 )+b$$ 
-using $m=1.3$ and  $b=-0.4$.
-Those values are compared to those used by Deigan\etal\cite{Deigan2009}, based on the rationale that lower absolute values for pseudo-energy bonuses increase the chance of structures being simultaneously supported by multiple conditions.
+using $m=1.3$ and  $b=-0.4$.\bsi{j'avais cru comprendre que ces valeurs avaient été adapté pour IPANEMAP.
+Sont elles paramétrables lorsuq'ion lance IPANEMAP?
+Si oui il faudrait peut etre en toucher un mot qque part?} Those values are compared to those used by Deigan\etal\cite{Deigan2009}, based on the rationale that lower absolute values for pseudo-energy bonuses increase the chance of structures being simultaneously supported by multiple conditions.
 Those pseudo energy contributions effectively direct our predictions towards a subset of structures that are in good agreement with probing data.
@@
 
-\newcommand{\Bull}[1]{{\sffamily #1}~\raisebox{1pt}{\tikzcircle[black, fill=cluster#1]{3pt}}}
-
-\newcommand{\BullLab}[1]{Cluster \Bull{#1}}
-
-\colorlet{clusterA}{SeaGreen}
-\colorlet{clusterB}{Yellow}
-\colorlet{clusterC}{gray}
-\definecolor{clusterD}{HTML}{AFAFE9}
-\colorlet{clusterE}{OliveGreen}
-\colorlet{clusterF}{blue!90!black}
-\colorlet{clusterG}{Orange}
-\colorlet{clusterH}{SeaGreen!40}
-
 \begin{figure*}
@@
 Faced with diverse probing conditions, we first assess the compatibility of the conclusions drawn from different probing data, including the probing-free run of \OurTool as a control.
-We use the methodology described in the material and methods ~\ref{sec:dotplots} and, for each condition, compute the base pair probability distribution (a.k.a.
-dot-plot) in the  pseudo-Boltzmann ensemble, induced by the reactivity profile.
+We use the methodology described in the material and methods~\ref{sec:dotplots} and, for each condition, compute the base pair probability distribution (aka \emph{dot-plot}) in the  pseudo-Boltzmann ensemble, induced by the reactivity profile.
 We then computed the ensemble distance, the squared Euclidean distance between dot plots, for each pair of conditions.
@@
 
+\bsi{sauf erreur de ma part le cluster B des parias c'est tous les CE sans Mg .je me demande s'ils n'auraient pas tous un même biais expérimental...}
+
 \begin{figure}
@@
 Conditions ordered by spectral biclustering.
-Conditions regrouped within clusters by k-mean clustering appear as blocks on the x and y axis.\bsi{pourquoi les clusters E et F ne sont ils pas regroupés comme les autres?}}\label{fig:biclustering}
+Conditions regrouped within clusters by k-mean clustering appear as blocks on the x and y axis.}\label{fig:biclustering}
 \end{figure}

### 2019-06-21 SuppMat.tex
Updates from Overleaf

@@
 }
+
 \begin{table}[H]
@@
 %\section{IPANEMAP Vs Rsample}
-\section*{Mono probing benchmark (Hadjin~\emph{et al}~\cite{Hajdin2013} dataset)}
+\section*{Mono probing benchmark (Hajdin~\emph{et al}~\cite{Hajdin2013} dataset)}
 
  \begin{table}[H]
-
 {\centering
@@
  	RNA ID & Description                      \\ \midrule
- 	A      & PreQ1riboswitchBsubtilis         & 	M      & SARScoronaviruspseudoknot        \\
- 	B      & 5domainof16SrRNAEcoli            & 	N      & 5SrRNAEcoli                      \\
- 	C      & SignalrecognitonparticleRNAhuman & 	O      & cyclicdiGMPriboswitchVcholerae   \\
- 	D      & GroupIintronAzoarcussp           & 	P      & 5domainof23SrRNAEcoli            \\
- 	E      & HIV15primepseudoknotdomain       & 	Q      & RNasePBsubtilis                  \\
- 	F      & Telomerasepseudoknothuman        & 	R      & GroupIIntronTthermophila         \\
- 	G      & SAMIriboswitchTtengcongensis     & 	S      & HepatitisCvirusIRESdomain        \\
- 	H      & tRNA phe Ecoli                   & 	T      & 5domainof16SrRNAHvolcanii        \\
- 	I      & P546domainbI3groupIintron        & 	U      & GroupIIintronOiheyensis          \\
- 	J      & TPPriboswitchEcoli               & 	V      & tRNAaspyeast                     \\
- 	K      & AdenineriboswitchVvulnificus     & 	W      & LysineriboswitchTmaritime        \\
- 	L      & FluorideriboswitchPsyringae      & 	X      & MBoxriboswitchBsubtilis          \\
+ 	A      & Pre-Q1 riboswitch, B. subtilis      & 	M      &SARS corona virus pseudoknot       \\
+ 	B      & 5’ domain of 16S rRNA, E. coli            & 	N      & 	5S rRNA, E. coli                      \\
+ 	C      &	Signal recogniton particle RNA human & 	O      & cyclic-di-GMP riboswitch, V. cholerae   \\
+ 	D      & 	Group I intron, Azoarcus sp.          & 	P      & 5’ domain of 23S rRNA, E. coli            \\
+ 	E      & 	HIV 15 prime pseudoknot domain       & 	Q      & RNase PB. subtilis                \\
+ 	F      & Telomerase pseudoknot human       & 	R      & 	Group I Intron, T. thermophila        \\
+ 	G      & 	SAM I riboswitch, T. tengcongensis   & 	S      & 	Hepatitis C virus IRES domain       \\
+ 	H      & 	tRNA(phe), E. coli                & 	T      & 	5' domain of 16S rRNA, H. volcanii      \\
+ 	I      & P546 domain, bI3 group I intron       & 	U      & 	Group II intron, O. iheyensis         \\
+ 	J      & TPP riboswitch, E. coli              & 	V      & 	tRNA(asp), yeast                    \\
+ 	K      &	Adenine riboswitch, V. vulnificus    & 	W      & 	Lysine riboswitch, T. maritima        \\
+ 	L      & Fluoride riboswitch, P. syringae     & 	X      & M-Box riboswitch, B. subtilis          \\
  	\bottomrule
@@
 
-\caption{Labels used as shorthands for the 24 RNAs in the Hadjin \emph{ et al.} dataset~\cite{Hajdin2013}.}
+\caption{Labels used as shorthands for the 24 RNAs in the Hajdin \emph{ et al.} dataset~\cite{Hajdin2013}.}
 \label{clyu}
@@
 \toprule
-\OneMSevILUMg & 0.234 & 0.2872\\
-\OneMSevILU &	0.2784 & 0.2961\\
-\NMIAMg &	0.2606 & 0.2518\\
-\NMIA &	0.2961 & 0.3316\\
-\OneMSevILUThreeMg &	0.4291 & 0.3936\\
-\NMIAMgCE &	0.4291 & 0.3404\\
-\NAIMg 	& 0.3493 & 0.305\\
-\BzCNMg & 	0.3759 & 0.2429\\
-\OneMSevMgCE &	0.4379 & 0.3759\\
-\CMCTMg &	0.4911 & 0.4113\\
-\OneMSevILUThree &	0.2961 & 0.3404\\
-\DMSMg &	0.3493 & 0.3227\\
-\NMIACE 	&0.2695 &0.2518 \\
-\OneMSevCE &	0.3582 & 	0.3316\\
-\BzCN &	0.4911 & 	0.4734\\
-\NAICE &	0.4025 &0.4025\\
+\Bull{C}~\OneMSevILUMg & 0.234 & 0.2872\\
+\Bull{A}~\OneMSevILU &	0.2784 & 0.2961\\
+\Bull{A}~\NMIAMg &	0.2606 & 0.2518\\
+\Bull{A}~\NMIA &	0.2961 & 0.3316\\
+\Bull{C}~\OneMSevILUThreeMg &	0.4291 & 0.3936\\
+\Bull{E}~\NMIAMgCE &	0.4291 & 0.3404\\
+\Bull{G}~\NAIMg 	& 0.3493 & 0.305\\
+\Bull{H}~\BzCNMg & 	0.3759 & 0.2429\\
+\Bull{E}~\OneMSevMgCE &	0.4379 & 0.3759\\
+\Bull{G}~\CMCTMg &	0.4911 & 0.4113\\
+\Bull{D}~\OneMSevILUThree &	0.2961 & 0.3404\\
+\Bull{F}~\DMSMg &	0.3493 & 0.3227\\
+\Bull{B}~\NMIACE 	&0.2695 &0.2518 \\
+\Bull{B}~\OneMSevCE &	0.3582 & 	0.3316\\
+\Bull{B}~\BzCN &	0.4911 & 	0.4734\\
+\Bull{B}~\NAICE &	0.4025 &0.4025\\
 \bottomrule
@@
  \midrule
-Native&\OneMSevILU&\OneMSevILUMg&\OneMSevMgCE& \1M7CE &\NAIMg\\
-structure &+\NMIAMg&+\OneMSevILUThreeMg&+\NMIAMgCE&+\NMIACE &+\CMCTMg\\
+Native&\OneMSevILU&\OneMSevILUMg&\OneMSevMgCE& \OneMSevCE &\NAIMg\\
+structure &+ \NMIAMg&+ \OneMSevILUThreeMg&+ \NMIAMgCE&+ \NMIACE &+ \CMCTMg\\
 \end{tabular}

### 2019-06-21 IPANEMAP-2019.tex

@@
 The centroid secondary structure  associated with the largest probability cluster was considered as the final prediction.
-Predictions were evaluated in term of the geometric mean (GM) metrics used by previous works~\cite{Spasic2017}.
-\bsi{pourquoi utilise-t-on ici la Gm des bp distance et ensuite le MCC?
-à expliquer?}\ypi{Pour des raisons historiques, surtout dans le benchmark suivant, j'ai rajouté une phrase dans cette direction}As a control experiment, we also report ${\tt RNAfold}$ predictions in presence/absence of probing data, both in energy minimization (MFE) and maximum expected accuracy (MEA) modes.
+Predictions were evaluated in term of the geometric mean (GM) metrics, for the sake of comparison with previous studies~\cite{Spasic2017}.
+As a control experiment, we also report ${\tt RNAfold}$ predictions in presence/absence of probing data, both in energy minimization (MFE) and maximum expected accuracy (MEA) modes.
 Figure~\ref{cordero1} shows the averaged GM values for all combinations of tools and probing data.
@@
 
-%Remarques These :
-%\begin{itemize}
-%  \item Fabrice : Y a t il une raison pour laquelle on est les meilleurs ?
-%\end{itemize}
 This observed discrepancy is somewhat surprising, given that \OurTool and \Software{RSample} share the same sampling/clustering based modeling.
@@
 
-	\caption{Model of the Lariat capping ribozyme from {\itshape Didymium iridis} (A -- Meyer\etal\cite{Meyer2014}), pseudoknot-free secondary structure as annotated by \Software{DSSR} (B -- pseudoknot ommitted) and \OurTool first-ranking prediction (C.) from 8 maximally-diverging probing conditions (NMIAMg, 1M7ILU3Mg, NAIMg, 1M7, CMCTMg,1M7ILU3,1M7Mg).
-\bsi{figure à refaite.
-Faut il montrer les interactions tertiaires?
-} \label{fig:didy}}
+	\caption{Model of the Lariat capping ribozyme from {\itshape Didymium iridis} (A -- PDB 4P8Z:A by Meyer\etal\cite{Meyer2014}), secondary structure, as annotated by \Software{DSSR} (B -- pseudoknot ommitted) and \OurTool first-ranking prediction (C.), obtained from the integration of 8 maximally-diverging probing conditions (\NMIAMg, \OneMSevILUThreeMg, \NAIMg, \OneMSevCE, \CMCTMg,\OneMSevILUThree, and \OneMSevMgCE).\label{fig:didy}}
 \end{figure*}
@@
 We obtain the following clusters:
-\begin{itemize}\setlength{\itemsep}{0pt}\setlength{\parskip}{.1em}	
+\begin{bulletlist}%\setlength{\itemsep}{0pt}\setlength{\parskip}{.1em}	
 \item \BullLab{A}: \OneMSevILU, \NMIA, and \NMIAMg;
@@
 \item \BullLab{H}: \BzCNMg.
-\end{itemize}
+\end{bulletlist}
 The predicted clusters are generally consistent with the ordering resulting from a spectral biclustering, implemented in \Software{scikit-learn} and executed with default parameters, as illustrated by Figure~\ref{fig:glycine_example}.
@@
 	\newcommand{\NoMg}{$\circ$}
+	\newcommand{\M}[1]{\multicolumn{1}{l}{#1}}
 
@@
 			                            &          &         &           & \multicolumn{3}{@{}c@{}}{MCC\%}              \\
-			Condition                   &  Clust.  &  Tech.  & Mg$^{2+}$ & \multicolumn1l{\relsize{-1}\OurTool} & \multicolumn1l{MFE} & \multicolumn1l{MEA} \\ \midrule
+			Condition                   &  Clust.  &  Tech.  & Mg$^{2+}$ & \M{\relsize{-1}\OurTool} & \M{MFE} & \M{MEA} \\ \midrule
 			\OneMSevILUMg               & \Bull{C} &  \Mut   &    \Mg    & 85                       & 82      & 82      \\
@@
 			{\bfseries Average overall} &          & $\star$ &  $\star$  & 70.5                     & 68      & 68.5    \\ \bottomrule
-			                            &          &         &           &
+			                            &          &         &           &                          &
 		\end{tabular}
@@
 
-A superficial inspection of the resulting MCCs appears to confirm the conclusions of Section Bruno~\ref{sec:cordero}.
+A superficial inspection of the resulting MCCs appears to confirm the conclusions reached on the Cordero~\emph{et al}~\cite{Cordero2012} dataset.
 Indeed, considering two conditions leads to predictions whose quality fall between the worst and the best one, while being generally close to the average.
@@
 Having established the redundant nature of similar conditions, and their limited contribution to downstream modeling, we turn to an analysis of conditions across clusters.
-We run \OurTool{} the quality of predictions obtained by choosing a single condition in each of the 8 clusters.\bsi{il manque des trucs dans cette phrase (celle qui commence par we run), non?} We consider any possible combination, except for the outlier cluster~\Bull{B}, for which we only consider \NAICE as a representative.
+We run \OurTool{} to assess the quality of predictions informed by condition chosen across the 8 clusters.We consider any possible combination, except for the outlier cluster~\Bull{B}, for which we only consider \NAICE as a representative.
 
@@
 However, the bimodal nature of the distribution also suggests the existence of at least two conformations that are both supported by the WT and the mutant profiles.
-\bsi{il y a une conformation alternative connue. Tu pourrais me donner un model provenant des 2 clusters de MCC que je regarde si pas hasard c'est ca?}
 
@@
 \begin{figure}
-	{\centering\includegraphics[width=1.2\linewidth]{graphs/boxplotreproducibility.png}\\}
+	{\centering\includegraphics[width=1\linewidth]{graphs/boxplotreproducibility.png}\\}
 

### 2019-06-21 IPANEMAP-2019.tex
Updates from Overleaf

@@
 cDNAs were separated by capillary electrophoresis (Beckman Coulter, Ceq8000).
-Data were analyzed using the  QuSHAPE~\cite{Karabiber2013} software.
+Data were analyzed using the  \Software{QuSHAPE}~\cite{Karabiber2013} software.
 RNA probing was performed in triplicate with distinct RNA preparations.
+The corresponding datasets are named after the probe followed by "-CE"
 
@@
 For 1M7ILU3, the denaturation step was repeated three times : After a 1 minute incubation at 95\degree{}C 10 mM of 1M7 was added and the mix was incubated 1 minute at 95\degree{}C.
-This step was repeated 3 times.
+This step was repeated 3 times(3D).
 For 1M7, 1M7Mg, NMIA and NMIAMg, experiments were conducted as described above, except that the NGS library preparation was adapted to Ion Torrent sequencing.
@@
 
-Compared to bi-probing, MCC values achieved by integrating triplets of conditions now indicate a clear gain over the average performances of individual conditions, supporting a notion of complementarity between experiments.
+Compared to bi-probing, MCC values achieved by integrating triplets of conditions now indicate a clear gain over the average performances of individual conditions, supporting a notion of complementary between experiments.
 Indeed tri-probing experiments lead to an average +1.7\% MCC improvement (median +2.6) over the average of mono-probing prediction in the triplet.
@@
 We generated 100 uniformly-distributed subsets consisting of 1, 2, 3 and 10 mutants extracted from the Cheng dataset~\cite{Cheng2017}, which we supplemented with the original wild-type sequence (WT).
-For each set of probing profiles, we executed \OurTool{} with default parameters, using a sample size of 1~000 structures. 
+For each set of probing profiles \bsi{c'est ce qui correspondrait chez nous à du DMS Map mg ilu ce qu'il a fait.
+Tu peux l'insérer la avec notre nomenclature entre paranthèse step ?}, we executed \OurTool{} with default parameters, using a sample size of 1~000 structures.
 We also analyzed the WT sequence alone, reproducing the analysis 100 times to compare the variability across sets of mutants to the one induced by stochastic aspects of our method. 

### 2019-06-21 SuppMat.tex

@@
 \begin{table}[H]
- \Heatset{min=0.2,  
-	max=.5,   
+ \Heatset{min=20,  
+	max=60,   
 	max colour=BadCol, % colour at maximum
@@
 }
+ \HeatSecset{min=60,  
+	max=85,   
+	max colour=Aquamarine3, % colour at maximum
+	min colour=BadCol,      % colour at minimum
+	Min colour=BadCol, % colour for values below min
+	Max colour=Aquamarine3   % colour for values above max
+}
 \begin{center}
-\begin{tabular}{lHH}\toprule
- & \multicolumn{2}{c}{Distance of reactivity}\\
-Condition & \multicolumn1c{to prediction}&\multicolumn1c{to native}\\
+\begin{tabular}{llHHS}
 	\toprule
-\Bull{C}~\OneMSevILUMg & 0.234 & 0.2872\\
-\Bull{A}~\OneMSevILU &	0.2784 & 0.2961\\
-\Bull{A}~\NMIAMg &	0.2606 & 0.2518\\
-\Bull{A}~\NMIA &	0.2961 & 0.3316\\
-\Bull{C}~\OneMSevILUThreeMg &	0.4291 & 0.3936\\
-\Bull{E}~\NMIAMgCE &	0.4291 & 0.3404\\
-\Bull{G}~\NAIMg 	& 0.3493 & 0.305\\
-\Bull{H}~\BzCNMg & 	0.3759 & 0.2429\\
-\Bull{E}~\OneMSevMgCE &	0.4379 & 0.3759\\
-\Bull{G}~\CMCTMg &	0.4911 & 0.4113\\
-\Bull{D}~\OneMSevILUThree &	0.2961 & 0.3404\\
-\Bull{F}~\DMSMg &	0.3493 & 0.3227\\
-\Bull{B}~\NMIACE 	&0.2695 &0.2518 \\
-\Bull{B}~\OneMSevCE &	0.3582 & 	0.3316\\
-\Bull{B}~\BzCN &	0.4911 & 	0.4734\\
-\Bull{B}~\NAICE &	0.4025 &0.4025\\
-\bottomrule
+	         &                    &      \multicolumn{2}{c}{Distance of reactivity (\%)}      &                       \\
+	Cluster  & Condition          & \multicolumn1c{to prediction} & \multicolumn1c{to native} & \multicolumn1c{MCC\%} \\ \toprule
+	\Bull{C} & \OneMSevILUMg      & 23                            & 29                        & 85                    \\
+	\Bull{A} & \OneMSevILU        & 32                            & 40                        & 84                    \\
+	\Bull{A} & \NMIAMg            & 26                            & 27                        & 81                    \\
+	\Bull{A} & \NMIA              & 30                            & 38                        & 80                    \\
+	\Bull{C} & \OneMSevILUThreeMg & 42                            & 38                        & 74                    \\
+	\Bull{E} & \NMIAMgCE          & 46                            & 39                        & 73                    \\
+	\Bull{G} & \NAIMg             & 38                            & 39                        & 73                    \\
+	\Bull{H} & \BzCNMg            & 38                            & 29                        & 71                    \\
+	\Bull{E} & \OneMSevMgCE       & 48                            & 42                        & 70                    \\
+	\Bull{G} & \CMCTMg            & 49                            & 45                        & 70                    \\
+	\Bull{D} & \OneMSevILUThree   & 33                            & 39                        & 64                    \\
+	\Bull{F} & \DMSMg             & 38                            & 38                        & 63                    \\
+	\Bull{B} & \NMIACE            & 31                            & 38                        & 61                    \\
+	\Bull{B} & \OneMSevCE         & 38                            & 41                        & 60                    \\
+	\Bull{B} & \BzCN              & 51                            & 46                        & 60                    \\
+	\Bull{B} & \NAICE             & 41                            & 44                        & 60                    \\ \bottomrule
 \end{tabular}

### 2019-06-21 IPANEMAP-2019.tex
Updates from Overleaf

@@
 RNA probing was performed in triplicate with distinct RNA preparations.
-The corresponding datasets are named after the probe followed by "-CE"
+The corresponding data sets are named after the probe followed by "-CE" and Mg if probed in presence of MgCl$_\text{2}$.
 
@@
 Sequences were mapped and analyzed with \Software{ShapeMapper2} for 1M7ILU and 1M7ILU3.
-For the other conditions they are mapped with a script based on \Software{ShapeMapper}, but adapted to Ion Torrent output files.
+For the other conditions they are mapped with a script based on \Software{ShapeMapper}, but adapted to Ion Torrent output files.The corresponding data sets are named after the probe followed by "-MAP", Mg if probed in presence of MgCl$_\text{2}$, and IL  or IT for Illumina or Ion Torrent sequencing, respectively.
 
@@
 Modified RNA were ethanol-precipitated and resuspended in 10 $\mu$l of water.
-Modifications were mapped as described above for SHAPE-CE experiments.
+Modifications were mapped as described above for SHAPE-CE experiments.The corresponding data sets are named after the probe followed by "-CE" and Mg.
 
@@
 \begin{figure}
-	{\centering\includegraphics[width=1\linewidth]{graphs/boxplotreproducibility.png}\\}
+	{\centering\includegraphics[width=1\linewidth]{graphs/boxplotreproducibility}\\}
 

### 2019-06-21 IPANEMAP-2019.tex
Updates from Overleaf

@@
 SHAPE reagents represent a popular class of small molecules that react with the 2’ hydroxyl of flexible ribose~\cite{McGinnis2012}, although the exact properties observed by SHAPE remain the object of ongoing investigations~\cite{McGinnis2012,Hurst2018,Mlynsky2018,Frezza2019,Busan2019}.
-As Ribose flexibility is proportional to the degree of freedom of the nucleotide, it is assumed that SHAPE reagents to discriminate nucleotides in stable interactions from others.
+As ribose flexibility is proportional to the degree of freedom of the nucleotide, it is assumed that SHAPE reagents to discriminate nucleotides in stable interactions from others.
 Different SHAPE reagents are endowed with different dynamics, that can help differentiate Watson-Crick base pairs from more dynamics tertiary contacts~\cite{Gherghe2008,Steen2012,Rice2014,Busan2019}.
@@
 A dataset was gathered by Hajdin\etal\cite{Hajdin2013} to validate the predictive capacities of probing data-driven predictions.
-It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (SHAPE -- 1M7) is available.
+It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (1M7-\SH) is available.
 This dataset includes sequences originating from a variety of organisms, and spans lengths ranging from $34$ nts to $530$ nts, with a focus on riboswitches and complex RNA architectures.
@@
 Modified or unmodified RNAs were purified by ethanol precipitation and pellets were resuspended in 10 $\mu$l of water.
-Modifications were revealed by reverse transcription using 5’ fluorescently labelled primers (D2 or D4 WellRED, Sigma Aldrich 5’ CTG-TGA-ACT-AAT-GCT-GTC-CTT-TAA 3’) and M-MLV RNAse H- reverse transcriptase (Promega) as previously described~\cite{Deforges2017}, with only minor modification of the originally described SHAPE protocol~\cite{Wilkinson2006}.
+Modifications were revealed by reverse transcription using 5' fluorescently labelled primers (D2 or D4 WellRED, Sigma Aldrich 5' CTG-TGA-ACT-AAT-GCT-GTC-CTT-TAA 3') and M-MLV RNAse (H-) reverse transcriptase (Promega) as previously described~\cite{Deforges2017}, with only minor modification of the originally described SHAPE protocol~\cite{Wilkinson2006}.
 cDNAs were separated by capillary electrophoresis (Beckman Coulter, Ceq8000).
@@
 
-For 1M7ILU3, the denaturation step was repeated three times : After a 1 minute incubation at 95\degree{}C 10 mM of 1M7 was added and the mix was incubated 1 minute at 95\degree{}C.
-This step was repeated 3 times(3D).
+For \OneMSevILUThreeMg, the denaturation step was repeated three times : After a 1 minute incubation at 95\degree{}C, 10 mM of 1M7 was added and the mix was incubated 1 minute at 95\degree{}C.
+This step was repeated 3 times (3D).
 For 1M7, 1M7Mg, NMIA and NMIAMg, experiments were conducted as described above, except that the NGS library preparation was adapted to Ion Torrent sequencing.
-Sequences were mapped and analyzed with \Software{ShapeMapper2} for 1M7ILU and 1M7ILU3.
-For the other conditions they are mapped with a script based on \Software{ShapeMapper}, but adapted to Ion Torrent output files.The corresponding data sets are named after the probe followed by "-MAP", Mg if probed in presence of MgCl$_\text{2}$, and IL  or IT for Illumina or Ion Torrent sequencing, respectively.
+Sequences were mapped and analyzed with \Software{ShapeMapper2} for \OneMSevILUThreeMg and \OneMSevILUThreeMg.
+For the other conditions they are mapped with a script based on \Software{ShapeMapper}, but adapted to Ion Torrent output files.
+The corresponding data sets are named after the probe followed by "-MAP", Mg if probed in presence of MgCl$_\text{2}$, and IL  or IT for Illumina or Ion Torrent sequencing, respectively.
 
@@
 For DMS, RNA was then treated with 1 $\mu$l of DMS (1:12 in ethanol) or 1 $\mu$l of ethanol for 5 minutes at 37\degree{}C (mock reaction), the reaction was stopped by addition of 400 mM of Tris at 7.5 pH and immediately put on ice.
-For CMCT,  RNA was treated with 10 $\mu$l of CMCT (42mg/mL) or 10 $\mu$l H$_\text{2}$O for 10 minutes at 37\degree{}C (mock reaction).
+For CMCT,  RNA was treated with 10 $\mu$l of CMCT (42 mg/mL) or 10 $\mu$l H$_\text{2}$O for 10 minutes at 37\degree{}C (mock reaction).
 Modified RNA were ethanol-precipitated and resuspended in 10 $\mu$l of water.
@@
 Interestingly, in the absence of probing data, MFE predictions are generally dominant on this dataset, trailed by the MEA, and followed by \OurTool which, in this setting, devolves into the Ding-Lawrence algorithm~\citep{Ding2003}.
-However, whenever probing reactivities are available, the single centroid returned by \OurTool always achieves higher GM values (Avg: $70\%$) than both MEA (Avg: $62\%$) and MFE  (Avg: $58\%$), whose relative performances depend on the probing reagent.%la moyenne sur les 6 ARNS et sur les 3 reactifs
+However, whenever probing reactivities are available, the single centroid returned by \OurTool always achieves higher GM values (Avg: $70\%$) than both MEA (Avg: $62\%$) and MFE  (Avg: $58\%$), whose relative performances depend on the probing reagent.
+%la moyenne sur les 6 ARNS et sur les 3 reactifs
 Interestingly, the quality of MFE and MEA predictions does not systematically benefit from additional probing data.
@@
 We assess the predictive capacities of \OurTool in a classic setting where a single probing condition is available, and compare its performances against \Software{RSample} software recently introduced by Spacic\etal\cite{Spasic2017}, \Software{RSample} relies on a sampling/clustering method developed independently from the current work.
-It was shown to perform favorably against a comprehensive collection of state-of-the-art methods in probing-guided structure prediction, including \Software{RME}~\citep{Wu2015}, \Software{RNAprob}~\citep{Deng2016}, \Software{RNAprobing}~\citep{Washietl2012}, \Software{RNAsc}~\citep{Zarringhalam2012} and the \Software{fold} utility from the \Software{RNAstructure} suite~\citep{Reuter2010}.
+It was shown to perform favorably against a comprehensive collection of state-of-the-art methods in probing-guided structure prediction, including \Software{RME}~\citep{Wu2015}, \Software{RNAprob}~\citep{Deng2016}, \Software{RNAprobing}~\citep{Washietl2012}, \Software{RNAsc}~\citep{Zarringhalam2012} and the \Software{fold} utility from the \Software{RNAStructure} suite~\citep{Reuter2010}.
 We consider the dataset of hajdin~\emph{et al}, which consists of 24 RNAs of lengths ranging from 47 to 500 nucleotides, all believed to fold into a unique documented conformation.
@@
 \end{figure*}
-As benchmark RNA we used the lariat capping ribozyme from \textit{Didymium Iridis} (DiLCrz) which 3D structure has been modeled at 3,5 A from X-ray diffraction data~\cite{Meyer2014}.
+As benchmark RNA we used the Lariat capping ribozyme from \textit{Didymium Iridis} (DiLCrz) which 3D structure has been modeled at 3,85 A from X-ray diffraction data~\cite{Meyer2014}.
 This RNA contains a large diversity of interactions, including helical domains, tertiary pairings, a kissing loop and a pseudoknot (Figure~\ref{fig:didy}).
-It was probed as described in material and methods in 16 different conditions using different SHAPE and classical reagents, in presence/absence of Mg$^\text{2+}$ ions, using either the standard premature stop or the newly developed mutation induction technologies to map the modification sites.
+It was probed as described in Material and Methods in 16 different conditions using different SHAPE and classical reagents, in presence/absence of Mg$^\text{2+}$ ions, using either the standard premature stop or the newly developed mutation induction technologies to map the modification sites.
 These experimental results were then used as input of \OurTool{}, either individually or combined,  for secondary structure prediction.
@@
 Faced with diverse probing conditions, we first assess the compatibility of the conclusions drawn from different probing data, including the probing-free run of \OurTool as a control.
-We use the methodology described in the material and methods~\ref{sec:dotplots} and, for each condition, compute the base pair probability distribution (aka \emph{dot-plot}) in the  pseudo-Boltzmann ensemble, induced by the reactivity profile.
+We use the methodology described in the Material and Methods~\ref{sec:dotplots} and, for each condition, compute the base pair probability distribution (aka \emph{dot-plot}) in the  pseudo-Boltzmann ensemble, induced by the reactivity profile.
 We then computed the ensemble distance, the squared Euclidean distance between dot plots, for each pair of conditions.

### 2019-06-21 IPANEMAP-2019.tex

@@
 \begin{figure}
-	\newcommand{\W}{.725}
-	{\centering\begin{tabular}{@{}c@{}c@{}c@{}}
-			{\sf {\bfseries A} -- \relsize{-1}Bi-probing MCC \emph{vs} Min MCC of mono-probing}\\
-			\includegraphics[width=\W\linewidth,trim=1cm 1cm .8cm .3cm,clip]{graphs/didy/MCC-vs-Min}\\[.5em]
-			{\sf {\bfseries B} -- \relsize{-1}Bi-probing MCC \emph{vs} Average MCC of mono-probing}\\
-			\includegraphics[width=\W\linewidth,trim=1cm 1cm .8cm .3cm,clip]{graphs/didy/MCC-vs-Avg}\\[.5em]
-			{\sf {\bfseries C} -- \relsize{-1}Bi-probing MCC \emph{vs} Max MCC of mono-probing}\\
-			\includegraphics[width=\W\linewidth,trim=1cm 1cm .8cm .3cm,clip]{graphs/didy/MCC-vs-Max}\\
-		\end{tabular}\\}
+	\newcommand{\W}{.86}
+	{\centering\includegraphics[width=\W\linewidth]{graphs/didy/pairs-spiders}\\}
 	\caption{Shift in \OurTool{} predictive capacity (MCC) observed upon considering two conditions, compared to the worst ({\sf\bfseries A}),  average ({\sf\bfseries B}) 

### 2019-06-21 IPANEMAP-2019.tex
Updates from Overleaf

@@
 % Enter dates of publication
-\copyrightyear{2019}
-\pubdate{31 July 2019}
-\pubyear{2019}
-\jvolume{XX}
-\jissue{YY}
+\copyrightyear{??}
+\pubdate{???????}
+\pubyear{????}
+\jvolume{??}
+\jissue{??}
 
@@
 \history{%
-Received January 1, 2019;
-Revised February 1, 2019;
-Accepted March 1, 2019}
+Received July 1, 2019;
+Revised ??????;
+Accepted ??????}
 
@@
 \subsection*{The \OurTool{} method}
-\bsi{Ce chapitre est vraiment très réussi, il explique vraiment bien ce qu'on fait et c'est accessible meme pour des polios de biologistes.
-je trouve dommage qu'il soit cantoné au Mat et Met.
-A mon avis il faudrait soit le mettre tel quel comme premier chapitre des résultats, soit en faire une version light sans les détails de calculs le mettre comme premier chapitre des résultats et ne laisser que les calculs dans le Mat et Met}
-\ypi{Dans la mesure où l'intro est courte, et étant donné le titre, je ne pense pas que sa présence en Mat Met limite  sur visibilité.
-En plus, c'est typiquement là où tu le trouves d'habitude dans ce type de papier (cf Rsample).} \bsi{OK je propose la chose suivante:on laisse tt comme ca le math et met, mais on commence les résultats par un mini-chapitre du genre "principles of IPANEMAP" IPANEMAP relies on ..
-prediction with RNAsubopt ...
-sampling ...
-clustering ..
-evaluation by front pareto..
-Its originality relies on ...
-See material and methods for details }
 We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signal produced by various chemical probes.
@@
 $$\Delta G(i) = m \log(R_i +1 )+b$$ 
-using $m=1.3$ and  $b=-0.4$.\bsi{j'avais cru comprendre que ces valeurs avaient été adapté pour IPANEMAP.
-Sont elles paramétrables lorsuq'ion lance IPANEMAP?
-Si oui il faudrait peut etre en toucher un mot qque part?} Those values are compared to those used by Deigan\etal\cite{Deigan2009}, based on the rationale that lower absolute values for pseudo-energy bonuses increase the chance of structures being simultaneously supported by multiple conditions.
+using $m=1.3$ and  $b=-0.4$.
+Those values differ from those used by Deigan\etal\cite{Deigan2009}, and were adapted based on the rationale that lower absolute values for pseudo-energy bonuses increase the chance of structures being simultaneously supported by multiple conditions.
 Those pseudo energy contributions effectively direct our predictions towards a subset of structures that are in good agreement with probing data.
@@
 Starting from the assumption that a functional structure should be preserved during evolution, we wanted to assess the agreement that might exist between probing data profiles for a set of RNA mutants.
-We considered DMS probing data, generated by~\cite{Cheng2017} through systematic point-wise mutations, for the Lariat-capping ribozyme.
+We considered DMS probing data, generated by~\cite{Cheng2017} through systematic point-wise mutations, for the Lariat-capping ribozyme (equivalent to \Cond{DMS}{mg}{MaP}{ilu}{} in the nomenclature below).
 We renormalized each reactivity profile following the method introduced by Deigan\etal\cite{Deigan2009}, restricted to the primer-free sequence: values greater than $1.5 $ times the interquartile range were discarded, and remaining values were divided by the mean of the top $10\%$ reactivities.
@@
 RNA probing was performed in triplicate with distinct RNA preparations.
-The corresponding data sets are named after the probe followed by "-CE" and Mg if probed in presence of MgCl$_\text{2}$.
+The corresponding data sets are named after the probe followed by {\sc CE}, and {\sc mg} if probed in presence of MgCl$_\text{2}$.
 
@@
 For the other conditions they are mapped with a script based on \Software{ShapeMapper}, but adapted to Ion Torrent output files.
-The corresponding data sets are named after the probe followed by "-MAP", Mg if probed in presence of MgCl$_\text{2}$, and IL  or IT for Illumina or Ion Torrent sequencing, respectively.
+The corresponding data sets are named after the probe followed by {\sc MaP}.
+The presence of {\sc mg} as a superscript indicates the presence of MgCl$_\text{2}$, while the subscript {\sc il} (Illumina) or {\sc it} (Ion Torrent) specify the sequencing technology.
 
@@
 Modified RNA were ethanol-precipitated and resuspended in 10 $\mu$l of water.
-Modifications were mapped as described above for SHAPE-CE experiments.The corresponding data sets are named after the probe followed by "-CE" and Mg.
+Modifications were mapped as described above for SHAPE-CE experiments.The corresponding data sets are named after the probe, followed by {CE} and {\sc mg}.
 
@@
 \end{figure*}
-As benchmark RNA we used the Lariat capping ribozyme from \textit{Didymium Iridis} (DiLCrz) which 3D structure has been modeled at 3,85 A from X-ray diffraction data~\cite{Meyer2014}.
+As benchmark RNA we used the Lariat capping ribozyme from \textit{Didymium Iridis} (DiLCrz) which 3D structure has been modeled at 3.85 \AA{} from X-ray diffraction data~\cite{Meyer2014}.
 This RNA contains a large diversity of interactions, including helical domains, tertiary pairings, a kissing loop and a pseudoknot (Figure~\ref{fig:didy}).
@@
 We hypothesize that such conditions may be either representative of alternative conformations, or be altogether erroneous.
-Consequently, we will single out these two clusters within our detailed analysis of the multi-probing performances.
+Consequently, we will single out this cluster within our detailed analysis of the multi-probing performances.
 
-\bsi{sauf erreur de ma part le cluster B des parias c'est tous les CE sans Mg .je me demande s'ils n'auraient pas tous un même biais expérimental...}
+%\bsi{sauf erreur de ma part le cluster B des parias c'est tous les CE sans Mg .je me demande s'ils n'auraient pas tous un même biais expérimental...}
 
@@
 We generated 100 uniformly-distributed subsets consisting of 1, 2, 3 and 10 mutants extracted from the Cheng dataset~\cite{Cheng2017}, which we supplemented with the original wild-type sequence (WT).
-For each set of probing profiles \bsi{c'est ce qui correspondrait chez nous à du DMS Map mg ilu ce qu'il a fait.
-Tu peux l'insérer la avec notre nomenclature entre paranthèse step ?}, we executed \OurTool{} with default parameters, using a sample size of 1~000 structures.
+For each set of probing profiles (\Cond{DMS}{mg}{MaP}{ilu}{}), we executed \OurTool{} with default parameters, using a sample size of 1~000 structures. 
 We also analyzed the WT sequence alone, reproducing the analysis 100 times to compare the variability across sets of mutants to the one induced by stochastic aspects of our method. 

### 2019-06-21 SuppMat.tex
Updates from Overleaf

@@
 \section*{Multiprobing benchmark (Cordero~\emph{et al}~\cite{Cordero2012} dataset)}
-\subsection*{Structure prediction ordered by RNA}
+\subsection*{MCC values of individual predictions}
 
@@
 \subsection*{Compatibility of accessibility profiles with native and predicted structures}
+We assess and validate the overall quality of probing data using a basic notion of distance between discretized profiles and structure.
+More precisely, given a secondary structure $S$ of length $n$, we define the {\bf 1D projection} of $S$ as the vector $v_S$ such that
+$$ \forall i\in[1,n], v_S(i)=\begin{cases} 1 &\text{if $i$ unpaired in $S$}\\ 0 &\text{otherwise.} \end{cases} $$
+
+Given a vector $R$ of reactivities, we define a {\bf discrete reactivity} vector $d_R$ as
+$$ \forall i\in[1,n], d_R(i)=\begin{cases} 1 &\text{if $R[i]$ within the $30\%$ largest reactivity values}\\ 
+0 & \text{if $R[i]$ within the $30\%$ lowest reactivity values}\\
+1/2 &\text{otherwise.} \end{cases} $$
+By focusing on the largest and lowest values, we avoid the influence of outliers, and hope to partially mitigate the natural variability of probing reactivities.
+
+We now define the {\bf normalized distance} $||R,S||$ of a reactivity profile $R$ and a structure $S$ as
+$$ ||R,S|| = \frac{\sum_{i=1}^n |v_S(i)-d_R(i)|-40\%\times \frac{n}{2}}{n\times 60\%}.$$
+Note that exactly 40\% of the positions are assigned to $1/2$, leading to $|v_S(i)-d_R(i)|=1/2$ irrespectively of the value taken by $v_S(i)$.
+The overall contribution of $40\%\times \frac{n}{2}$ needs to be corrected for, thus the subtraction.
+The divisor term of $n\times 60\%$ accounts for the fact that, after compensating for the contribution of the $[30,70]$ percentile, only $60\%$ of the original values are effectively covered by the sum.
+We thus renormalize by $n\times 60\%$ to obtain a final value scaling between $0$ (all highly reactive pos.
+unpaired, all lowly reactive pos.
+paired) and $1$ (all highly reactive pos.
+paired, all lowly reactive pos.
+unpaired).
 \begin{table}[H]
@@
 }
-\begin{center}
+{\centering
 \begin{tabular}{llHHS}
@@
 	\Bull{B} & \NAICE             & 41                            & 44                        & 60                    \\ \bottomrule
-\end{tabular}
-\end {center}
+\end{tabular}\\}
 \caption{Distance of the predicted structure from the considered probing data.}

### 2019-06-24 IPANEMAP-2019.tex
Updates from Overleaf

@@
 This step was repeated 3 times (3D).
-For 1M7, 1M7Mg, NMIA and NMIAMg, experiments were conducted as described above, except that the NGS library preparation was adapted to Ion Torrent sequencing.
+For \NMIA and \NMIAMg, experiments were conducted as described above, except that the NGS library preparation was adapted to Ion Torrent sequencing.
 Sequences were mapped and analyzed with \Software{ShapeMapper2} for \OneMSevILUThreeMg and \OneMSevILUThreeMg.

### 2019-07-03 IPANEMAP-2019.tex
Updates from Overleaf

@@
 \usepackage{graphics}
-%\usepackage[latin1]{inputenc}
 \usepackage{subcaption}
@@
 % Enter dates of publication
-\copyrightyear{??}
-\pubdate{???????}
-\pubyear{????}
-\jvolume{??}
-\jissue{??}
+\copyrightyear{2019}
+\pubdate{$\star$}
+\pubyear{2019}
+\jvolume{$\star$}
+\jissue{$\star$}
 
@@
 \address{%
-$^{1}$\,CNRS UMR 7161, LIX, Ecole Polytechnique, France
+$^{1}$\,CNRS UMR 7161, LIX, Ecole Polytechnique, 1 rue Estienne d'Orves, 91120 Palaiseau, France
 and
-$^{2}$\,CitCoM, Cibles thérapeutiques et conception de Médicaments, CNRS, Université de Paris, 4 avenue de l'observatoire 75006 PARIS France}
+$^{2}$\,CitCoM, Cibles th\'erapeutiques et conception de M\'edicaments, CNRS, Universit\'e de Paris, 4 avenue de l'observatoire, 75006 PARIS, France}
 % Affiliation must include:
@@
 Received July 1, 2019;
-Revised ??????;
-Accepted ??????}
+Revised $\star$;
+Accepted $\star$}
 
 \maketitle
-
 \begin{abstract}
-The manual production of reliable structural models from chemical probing experiments, such as SHAPE probing, benefits from the integration of information derived from multiple protocols and reagents.
+The manual production of reliable RNA structure models from chemical probing experiments, such as SHAPE probing, benefits from the integration of information derived from multiple protocols and reagents.
 However, the interpretation of multiple probing profiles remains a complex task, hindering the quality and reproducibility of modeling efforts.
@@
 Such an integrated approach to structure modeling has led to sizable improvements in prediction accuracy~\cite{Washietl2012} and is currently at the core of successful modeling strategies~\cite{Miao2017}.
-However, the interpretation of probing data, to inform structure prediction, is challenged by a number of factors, including structural heterogeneity, experimental errors, structural dynamics and the natural variability of reactivity measurements across replicates.
+However, the interpretation of probing data, to inform structure prediction, is challenged by a number of factors, including structural heterogeneity, experimental errors, structural dynamics and the potential variability of reactivity measurements across replicates.
 
 SHAPE reagents represent a popular class of small molecules that react with the 2’ hydroxyl of flexible ribose~\cite{McGinnis2012}, although the exact properties observed by SHAPE remain the object of ongoing investigations~\cite{McGinnis2012,Hurst2018,Mlynsky2018,Frezza2019,Busan2019}.
-As ribose flexibility is proportional to the degree of freedom of the nucleotide, it is assumed that SHAPE reagents to discriminate nucleotides in stable interactions from others.
+As ribose flexibility is proportional to the degree of freedom of the nucleotide, it is assumed that SHAPE reagents discriminate nucleotides in stable interactions from others.
 Different SHAPE reagents are endowed with different dynamics, that can help differentiate Watson-Crick base pairs from more dynamics tertiary contacts~\cite{Gherghe2008,Steen2012,Rice2014,Busan2019}.
@@
 For \OneMSevILUThreeMg, the denaturation step was repeated three times : After a 1 minute incubation at 95\degree{}C, 10 mM of 1M7 was added and the mix was incubated 1 minute at 95\degree{}C.
-This step was repeated 3 times (3D).
+This step was repeated 3 times for the samples labeled 3D.
 For \NMIA and \NMIAMg, experiments were conducted as described above, except that the NGS library preparation was adapted to Ion Torrent sequencing.
@@
 The corresponding data sets are named after the probe followed by {\sc MaP}.
-The presence of {\sc mg} as a superscript indicates the presence of MgCl$_\text{2}$, while the subscript {\sc il} (Illumina) or {\sc it} (Ion Torrent) specify the sequencing technology.
+The presence of {\sc mg} as a superscript indicates the presence of MgCl$_\text{2}$, while the subscript {\sc il} (Illumina) or {\sc it} (Ion Torrent) specifies the sequencing technology.
 
@@
 
-The joint analysis of pairs of probing conditions seems to average the quality of predictions.
+The joint analysis of pairs of probing conditions appears to average the quality of predictions.
 Any combination of probing data yields a GM value that is always greater than the worst-performing condition in the pair, yet worse than the best-performing alone.
@@
 It is also worth mentioning that NMIA+CMCT, combining the best and worst conditions, achieves a better combined performance than DMS+NMIA, the two best mono-probing conditions.
-This hints towards some level of complementarity between probing conditions, as already suggested by recent analyses \citep{Yu2018} (steen 2012, Rice 2014, Brunel 2000) and supported by further analyses in this paper.
+This hints towards some level of complementarity between probing conditions, as already suggested by recent analyses \citep{Brunel2000,Steen2012,Rice2014,Yu2018}  and supported by further analyses in this paper.
 
@@
 
-We generated 100 uniformly-distributed subsets consisting of 1, 2, 3 and 10 mutants extracted from the Cheng dataset~\cite{Cheng2017}, which we supplemented with the original wild-type sequence (WT).
+We generated 100 uniformly-distributed subsets consisting of 1, 2, 3 and 10 mutants extracted from the Cheng~\emph{et al} dataset~\cite{Cheng2017}, which we supplemented with the original wild-type sequence (WT).
 For each set of probing profiles (\Cond{DMS}{mg}{MaP}{ilu}{}), we executed \OurTool{} with default parameters, using a sample size of 1~000 structures. 
@@
 
-Our analyses reveal that integrating multiple profiles of at least three conditions allows to mitigate the effect of outlier conditions and, at times, to produce better secondary structure models than the one inferred from the most accurate profile.
+Our analyses reveal that integrating multiple profiles of at least three conditions allows to mitigate the effect of outlier conditions and, at times, to produce better secondary structure models than the one inferred only from the most accurate profile.
 A comparison of the joint performances to the average appears fair since no single reagent appears to systematically dominate in term of predictive capacity.
@@
 \section*{Funding}
-This work has been supported by La Fondation pour la Recherche Medicale ([FRM
+This work, including AS and AD phD fellowships, has been supported by La Fondation pour la Recherche Medicale ([FRM
 DBI20141423337]).

### 2019-07-10 IPANEMAP-2019.tex
Updates from Overleaf

@@
 \section*{Funding}
-This work, including AS and AD phD fellowships, has been supported by La Fondation pour la Recherche Medicale ([FRM
+This work, including AS and DA phD fellowships, has been supported by La Fondation pour la Recherche Medicale ([FRM
 DBI20141423337]).

### 2019-07-10 IPANEMAP-2019.tex

@@
 \noindent \OurTool{} combines statistical sampling, unsupervised machine learning, and multi-objective optimization, to produce secondary structure models that are both stable and well-supported by experimental evidences.
-An analysis of several datasets demonstrates the performance of \OurTool, even in a mono probing setting, and confirms the potential of integrating multiple sources of probing data, leading to several recommendations for the experimental design of informative probing assays.
+The analysis of several reactivity, both publicly available and specifically produced experimentally, demonstrates the good performances of \OurTool, even in a mono probing setting.
+It confirms the potential of integrating multiple sources of probing data, leading to recommendations for the experimental design of informative probing assays.
 
@@
 At the end of the modeling process, the modeler is often left with several alternatives, all of which may appear equally consistent with the probing data, and the choice of a "best" model becomes somewhat subjective.
-Here we have developed a new modeling procedure that takes into account multiple probing data and ultimately yields only one or two models.
+Here we have developed a new modeling procedure that jointly takes into account multiple probing data, and ultimately yields a small collection of secondary structure models.
 
@@
 We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signal produced by various chemical probes.
-It performs clustering based on multiple sets of structures sampled in different experimental conditions, to ultimately return a set of structures that are representative of the dominant clusters.
+
+It performs a structural clustering across multiple sets of structures sampled in different experimental conditions, and ultimately returns a set of structures that are representative of the dominant conformations supported across conditions.
 Its underlying rationale is that the prominent presence of a stable secondary structure within the  (pseudo-)Boltzmann ensembles induced by multiple experimental conditions should increase its likelihood to be (one of) the native structure(s) for a given RNA.
@@
 
-Since \OurTool{} supports any number of probing profiles, we consider the Cordero~\emph{et al} dataset, further described.
+Since \OurTool{} supports any number of probing profiles, we consider the Cordero~\emph{et al}~\cite{Cordero2012a} dataset, further described in the previous section.
 It consists of 6 RNAs of known structures, for which reactivity profiles are available for the CMCT, NMIA and DMS reagents.
@@
 It follows that a modeler cannot objectively favor \emph{a priori} a probing condition/reagent~\cite{Yu2018}.
-In other words, multiple experimental probing data, even combined in a pairwise fashion, were typically observed to mitigate the empirical risk of a misprediction.
+In other words, multiple experimental probing profiles, even combined in a pairwise fashion, were typically observed to mitigate the empirical risk of a misprediction, and should be considered.
 

### 2019-07-17 IPANEMAP-2019.tex

@@
 \begin{abstract}
-The manual production of reliable RNA structure models from chemical probing experiments, such as SHAPE probing, benefits from the integration of information derived from multiple protocols and reagents.
+The manual production of reliable RNA structure models from chemical probing experiments, benefits from the integration of information derived from multiple protocols and reagents.
 However, the interpretation of multiple probing profiles remains a complex task, hindering the quality and reproducibility of modeling efforts.
 
-\noindent We introduce \OurTool{}, an automated and robust approach for the computational modeling of RNA structure from multiple probing reactivity profiles.
-Input profiles can results from experiments based on diverse protocols, reagents, or even over a collection of mutants, and are jointly analyzed to predict the dominant conformations of an RNA.
+\noindent We introduce \OurTool{}, an automated method for the computational modeling of RNA structure from multiple probing reactivity profiles.
+Input profiles can result from experiments based on diverse protocols, reagents, or even a collection of mutants, and are jointly analyzed to predict the dominant conformations of an RNA.
 
-\noindent \OurTool{} combines statistical sampling, unsupervised machine learning, and multi-objective optimization, to produce secondary structure models that are both stable and well-supported by experimental evidences.
+\noindent \OurTool{} combines sampling, unsupervised learning, and multi-optimization, to produce secondary structure models that are both stable and well-supported by experimental evidences.
 The analysis of several reactivity, both publicly available and specifically produced experimentally, demonstrates the good performances of \OurTool, even in a mono probing setting.
@@
 \noindent \textbf{Contact:} \href{yann.ponty@lix.polytechnique.fr}{yann.ponty@lix.polytechnique.fr}
-%\noindent \textbf{Supplementary information:} Supplementary data available at NAR online.
+
+\noindent Supplementary information available at NAR online.
 \end{abstract}
@@
 
-Computationally, the reactivity of a nucleotide is typically used as a proxy for its unpairedness.
+Computationally, the reactivity of a nucleotide is typically used as a proxy for the unpairedness of individual nucleotides.
 The last couple of decades have nevertheless seen a series of paradigm shifts in the ways probing information is integrated, somehow mirroring the evolution of {\em ab initio} methods for secondary structure prediction.
@@
 As a control experiment, we also report ${\tt RNAfold}$ predictions in presence/absence of probing data, both in energy minimization (MFE) and maximum expected accuracy (MEA) modes.
-Figure~\ref{cordero1} shows the averaged GM values for all combinations of tools and probing data.
+Figure~\ref{cordero1} shows the averaged GM values for all combinations of tools and probing data (Detailed results can be found in Supp.
+Tables S1 and S2).
 
@@
 
-Interestingly, in the absence of probing data, MFE predictions are generally dominant on this dataset, trailed by the MEA, and followed by \OurTool which, in this setting, devolves into the Ding-Lawrence algorithm~\citep{Ding2003}.
+In the absence of probing data, MFE predictions are generally dominant on this dataset, trailed by the MEA, and followed by \OurTool which, in this setting, devolves into the Ding-Lawrence algorithm~\citep{Ding2003}.
 However, whenever probing reactivities are available, the single centroid returned by \OurTool always achieves higher GM values (Avg: $70\%$) than both MEA (Avg: $62\%$) and MFE  (Avg: $58\%$), whose relative performances depend on the probing reagent.
@@
 We computed and report in Figure~\ref{fig:Vsrsample} the individual Geometric Means (GM) associated with predicted structures.
+Sensitivity and PPV reported in Supp.
+Table S4, while Supp.
+Table S5 reports the stability of predictions across 10 runs of \OurTool.
 
@@
 
-\paragraph{Assessment of predictions informed by individual probing conditions (mono).}
+\paragraph{Assessment of predictions informed by individual probing conditions (mono probing).}
 %First, we assess the individual performance of each probing condition considered as the sole source of structural constraints.
@@
 
-Over the fourteen conditions, we observe a large discrepancy in the capacity of the experimental setup to inform predictions, with MCCs ranging from above 80\% (\NMIA, \NMIAMg, \OneMSevILU and \OneMSevILUMg) to around 60\% (\NAICE, \BzCN, \OneMSevCE and \NMIACE).
-These values are much better than those achieved in the absence of probing data (51\% MCC), and can be interpreted as indicative of good quality predictions.
-Indeed, it should be reminded that the MCC is a stringent metrics, taking values between -100\% and 100\%, 0\% being the expected MCC of a "coin tossing" random predictor.
+Across the sixteen conditions, we observe a large discrepancy in the capacity of the experimental setup to inform predictions, with MCCs ranging from above 80\% (\NMIA, \NMIAMg, \OneMSevILU and \OneMSevILUMg) to around 60\% (\NAICE, \BzCN, \OneMSevCE and \NMIACE).
+Interestingly the MCC value cannot be trivially anticipated from an assessment of the compatibility of the reactivity scores with the predicted structure, or even with the native structure (see Supp.
+Table S6).
+ These predictive capacities clearly outperform those achieved in the absence of probing data (51\% MCC), and can be interpreted as indicative of good quality predictions.
+Indeed, remind that the MCC is a stringent metrics, taking values between -100\% and 100\%, 0\% being the expected MCC of a "coin tossing" random predictor.
 The lowest MCC value of 60\% is, in particular, equally consistent with Sensitivity/PPV values of 60\%/60\%, 80\%/46\% or 45\%/82\%.
@@
 Unsurprisingly, the presence/absence of Magnesium ions during the probing impacts the predictions, with an observed drop from an average 73\% MCC in the presence of Mg$^\text{2+}$ to 67\% MCC in the absence of Mg$^\text{2+}$.
-However, while the poorly performing conditions of cluster \Bull{B} are probed in the absence of Mg$^\text{2+}$, some of the most informative profiles (\OneMSevILU and \NMIA) were obtained in absence of Mg$^\text{2+}$.
+However, while the poorly performing conditions in Cluster \Bull{B} are probed in the absence of Mg$^\text{2+}$, some of the most informative profiles (\OneMSevILU and \NMIA) were obtained in absence of Mg$^\text{2+}$.
 An even more drastic change of performances can be observed when comparing the MCC of mutation-based protocols (avg 78\%) with stop-based ones (avg 66\%).
@@
 We then computed the associated MCC, and compare it with the MCC of the worst performing condition (Min), best-performing condition (Max) and average MCC over the single conditions experiments.
-A summary of the results over pairs is shown in Figure~\ref{fig:pairwise}.
+A summary of the results over pairs is shown in Figure~\ref{fig:pairwise} (see details in Supp.
+Table S7).
 
@@
 
-A first test of this hypothesis consists in executing \OurTool{} on every subset of conditions within clusters, and consider the quality of predictions (MCC). 
+As a test of this hypothesis, we executed \OurTool{} on every subset of conditions within clusters, and report in Figure~\ref{fig:similarconds} the quality (MCC) of predictions for any subset of similar conditions (see Supp.
+Figure~S1 for associated models).
 Note that our similarity notion and clustering of conditions is purely based on the ensemble features, and does not explicitly consider the MCC of individual conditions. 
@@
 By comparing our results with those obtained for unrestricted pairs of mutants, we assess the influence of structure variability on the quality of our predictions.
-We report in Figure~\ref{fig:variantanalysis} the distribution of MCC values associated with the first prediction returned by \OurTool.
+We report in Figure~\ref{fig:variantanalysis} the distribution of MCC values associated with the first prediction returned by \OurTool (see Supp.
+Figure S2 for details).
 
@@
 This confirms the widespread interpretation of SHAPE-induced pseudo-potentials as focusing predictions onto a subset of the Boltzmann ensemble.
-A similar behavior is observed for the other datasets considered in our study, as seen in Supplementary Material.
+A similar behavior is observed for the other datasets considered in our study, as seen in Supp.
+Tables S2 and S5.
 %In the interest of reproducibility, ipanemap reports, and accepts as an option, the random seed used both within the main software and its non-deterministic dependencies, leading to identical repeats of the experiments.

### 2019-07-17 SuppMat.tex

@@
 
-\usepackage{rotating}
+\usepackage[figurename=Supplementary~Figure, tablename=Supplementary~Table]{caption}
+\DeclareCaptionLabelFormat{cont}{Figure S\arabic{figure}}
+\captionsetup[figure]{labelformat=cont}
+\DeclareCaptionLabelFormat{contt}{Table S\arabic{table}}
+\captionsetup[table]{labelformat=contt}
+\captionsetup{labelfont={bf}}
 
@@
 
-\begin{sidewaysfigure}
-\newcommand{\MyScale}{.26}
+{\newcommand{\MyScale}{.26}
 
+\begin{figure}
 {\centering
-	\begin{adjustbox}{max width=\textwidth,max height=\textheight,keepaspectratio}
-\begin{tabular}{@{}c|ccccc@{}}
+  \sf	\begin{adjustbox}{max width=\textwidth,max height=\textheight,keepaspectratio}
+\begin{tabular}{@{}ccc@{}}
 
@@
  \includegraphics[scale=\MyScale]{graphs/Supp_structures/native_structure}&
- \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILU_NMIAMg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}&
+ \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILU_NMIAMg}& 
+ \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}\\
+Native&\OneMSevILU&\OneMSevILUMg\\ 
+structure &+ \NMIAMg&+ \OneMSevILUThreeMg\\[2em]
  \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7MgCE_NMIAMgCE}
@@
  & \includegraphics[scale=\MyScale]{graphs/Supp_structures/NaiMg_CMCTMg}\\
- \midrule
-Native&\OneMSevILU&\OneMSevILUMg&\OneMSevMgCE& \OneMSevCE &\NAIMg\\
-structure &+ \NMIAMg&+ \OneMSevILUThreeMg&+ \NMIAMgCE&+ \NMIACE &+ \CMCTMg\\
+\OneMSevMgCE& \OneMSevCE &\NAIMg\\
++ \NMIAMgCE&+ \NMIACE &+ \CMCTMg\\
 \end{tabular}
@@
 \caption{Predicted structures with conditions belonging to the same cluster}
-\end{sidewaysfigure}
+
+\end{figure}
 

### 2019-10-28 Response-NAR-Round1.tex

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[9pt,hyperref]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{amsmath} % for much better looking tables
\usepackage{xspace}
\usepackage{url}
\usepackage{soul}
\usepackage[table,x11names,dvipsnames]{xcolor}
\usepackage{amssymb} % for much better looking tables
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{ntheorem} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
\usepackage{framed}
\usepackage{cabin}
\usepackage{tgtermes}
%\usepackage{tgadventor}
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

\newcommand{\Answer}[1]{\noindent{\color{purple!80!black}{{\sffamily Response:} }{#1}}}
\newcommand{\Comment}[1]{\item{\color{black}{{\sffamily Comment:} }{#1}}}

\usepackage{todonotes}
\newcommand{\TODO}[2][All]{{\todo[color=blue!20,inline]{{TODO (#1)}: #2}}}

%%% END Article customizations

%%% The "real" document content comes below...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{macros.tex}

\title{Response to reviewers\\[.3em]\OurTool{}:  Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles}
\author{
Afaf Saaidi \and
Delphine Allouche \and
Mireille Regnier \and
Bruno Sargueil \and
Yann Ponty}

\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

\section{General comments}

Points géneraux :

On a probablement pas été super clairs, mais :
Pas de comparaison entre protocoles, plutôt un test, rendu possible par IPANEMAP de la complémentarité des protocoles
Pas optimisation du traitement des donnés de sondage, mais traitement uniforme
Approche multi probing nous parait susceptible d’être robuste à l’incorporation d’une réactivité moins informative.

\section{Detailed responses}

\subsection*{Reviewer \#1}

\begin{enumerate}
	\Comment{The authors do not provide any sense of how raw data sets correlate in their experiments. They only analyze the ensembles of structures predicted based on their data. Thus for example, in Figure 5 we notice that the method of detection (CE) seems to be a far greater differentiator in predictions than the probe used (e.g. not all 1M7 cluster). The authors should add a panel clustering the raw probing data to compare it with the IPANEMAP clustering. This will require some clever design of distance metrics and mapping 1M7 to DMS maybe complex, but they should do a heat map like Figure 6 but with the raw data minimally.}

	\Answer{	On a fait ça, c’est en supp mat. On observe :
		Corrélation inverse entre distance à la native et MCC\%;
		Difficulté à regarder le clustering car peu significatif
	}

	\Comment{I am also concerned by the fact that they only have one (-) No Prob condition. For both CE and MaP data, a background subtraction is required. It is not at all clear what the No Prob condition is and why it was not independently subtracted from all the data prior to analysis.}

	\Answer{There is a misunderstanding here, the “(-) No Prob” conditions refers to a modeling process that does not use probing data as constraints. We understand that this was confusing and we no refer to it as “(-) No Exp” for modeling without experimental data. “(-) No Cons” for modeling without experimental constraints. Of course we have carried out control experiments without any modification reageant for each of the experiments reported, as now explicitly specified in the material and methods section.}

	\Comment{I am concerned the MaP data they collected is not reporting on structure. For example, quality controls need to be established showing that DMS reactivities follow expected profiles, i.e. A and C having higher mean reactivity that G and U.}

	\Answer{We understand this legitimate concern, expressed by multiple referees. The probing data reported on the structure for each probing conditions and reagents is now provided as supplementary material.
	Please note that the DMS results were not obtained following a ShapeMap-like strategy but by a “classical” RT-stop technique and Capilary electrophoresis.}

	\Comment{The main reason I am concerned with their MaP data is that they used superscript III instead of II for the RT. In the Smola paper, superscript III was attempted, but it was shown it does not detect adduct formation by mutation. If the CE to MaP correlations are low for the data, which I suspect they are, the authors will have to repeat MaP profiling with Superscript II to show this change in the protocol does not affect the data. I suspect however it will and may change the conclusions of the paper.}

	\Answer{It is exact that the Weeks laboratory uses an RNAse H minus mutant of the MMLV RT sold as “Superscript II” and advises to use it. However as far as we can tell from reviewing Dr Weeks bibliography, they never reported the test of “superscript III” (another RNAse H- mutant of the MMLV RT sold by the same company), and therefore never shown that does not detect the adduct. Actually in Smola et al. they specify “Reaction conditions have been optimized for SuperScript II reverse transcriptase only. Other reverse transcriptase enzymes and derivatives have not been tested and should not be used”. We tried superscript III in the conditions described and analysed the histograms provided by ShapeMapper which are metrics for the success of a shapeMap experiment as defined by smola et al. 2015.
	Please find below some examples of the histograms obtained using Superscript III:

	\begin{itemize}
	\item NMIA
	\TODO{Add plot}
	\item NMIA mg
	\TODO{Add plot}	
	\item 1M7
	\TODO{Add plot}
	\end{itemize}

	As can be seen on the right  histograms the read depth is sufficient (>10k reads/nucleotides) for these experiments to be interpreted. The histograms on the left shows that there are significantly more mutations in the modified (denatured and SHAPE) samples than in the untreated ones. These histograms are very similar to those shown by smola et al. as exemplary of a successful experiment. 
	In addition the shapeMap results are consistent with the X-ray structure (see figure XX in the supp. material), and their use with any prediction software as shown to improve the accuracy of the obtained model(s).

	In conclusion there is no doubt about the relevance of these results and we do not see the necessity to repeat them using superscript II.}

	\Comment{One possibility if the authors indeed find the MaP data is not correct because of superscript III would be to not include it in the analysis. Unfortunately this may somewhat temper the general interest of the paper, as a majority of SHAPE and probing data is now collected with MaP and -stop protocols.}

	\Answer{As described above, we did not find any reason to doubt the quality of those results, and believe they should remain included in our analysis.}

	\Comment{The authors also need to make their raw data available, I would recommend using the SNRNASM standard (PMID 21610212) and adding the files in the supplement. In addition the raw read files and alignments should really be deposited into the sequence read archive (SRA) so others can assess the quality of the mutation rates.}

	\Answer{We agree that raw data should be made available to the reader for the sake of reproducibility. Unfortunately, the SNANASM were no longer available at the time of this revision. We have informed the creators of the resource. In the meanwhile, we have thus deposited the reactivity results on the GitHub page of IPANEMAP, where they also serve as illustrative examples.

	The raw data were deposited to the SRA under identifier \TODO[Bruno]{Deposit data in SRA}.}

	\end{enumerate}

	\subsection*{Reviewer \#2}

	\begin{enumerate}
\Comment{I find this tool and study very interesting, especially the systematic assessment of the impact and combined power of various probing protocols for DiLCrz. To be ready for publication, the manuscript needs some further discussion of the DiLCrz results and a revision of the method presentation as detailed below.}

	\Answer{\TODO{...}}

	\Comment{The authors picked with DiLCrz a hard example since most problems of prediction accuracy are caused by the confusion of the nested structure model in the presence of the reactivities of the pseudoknot. That is, the quite stable PK-subhelix of length 4 (ignored in the nested model and the evaluation) will have a strong impact on reactivities that can not (correctly) be captured. Thus, the authors mainly evaluate how much the prediction of local structures in subsequence 65-120 is correct or not (as also visible in the reported structures within the supplement). Eventually, this also impacts the generality of the results.

	Thus, I think the authors should add a respective discussion to the manuscript.}

	\Answer{\TODO[Yann]{Check if we get better results by blocking the pseudoknot.}}

	\Comment{The methods section needs some reordering to first formaly introduce clusters and respective measures (stability/support) (e.g. in line 3-1-57) before introducing and discussing MBkM.

	For this restructuring, you might want to introduce the cummulative sample multi-set "$\mathcal{S} = \sum_d \mathcal{S}_d$" from which the clusters (also multi-sets?!) form a partitioning.}

	\Answer{\TODO{...}}

	\Comment{You do not introduce how you are able to compute MEA centroids without explicitely computing base pair probabilities. Checking your source code, I found the bp-probs are computed analogously to the structure probabilities, which is fine but should be explicitely mentioned to be reproducable.}

	\Answer{\TODO{...}}

	\Comment{It is totally unclear to me (and I didnt took the time to track it down in your code), how you derive an MEA for a cluster with structures from different reactivity data sets, since the same structure can be part of the cluster multiple times (with different probe-constraint probabilites). Do you sum up probe-specific probabilities in that case? Do you derive new base-pair probabilities for the cluster (and if so how)?
	Please provide somewhere details of your methods concerning the whole bp-prob and MEA centroid computation (at least in the supplementary material). Otherwise, your method description leaves out a central point of your algorithmic pipeline and renders is non-reproducable!}

	\Answer{\TODO{...}}

	\Comment{Please replace "kT" by "\text{R}T" in the structure probability formula (and whereever needed)!}

	\Answer{\TODO{...}}

	\Comment{What is the difference betwee the "cummulative probability of [a cluster's] structures" (line 3-1-58) and the formally introduced "Stability(C)" notion? If there is none, please use accordingly, otherwise provide respective details.}

	\Answer{\TODO{...}}

	\Comment{Please provide (eg in the supplement) some further details how to derived the values for m abd b for the Deigan method. So far, the description is quite vague.. }

	\Answer{\TODO{...}}

	\Comment{Within this context, you might want to discuss that too low pseudo-energy boni will yield similar results as hard constraints (which you want to avoid).}

	\Answer{\TODO{...}}

	\Comment{The formula for "$Stability(C)$" misses a "$\sum_{S\in C \cap \mathcal{S}_d}$" between the sum over d and the probability.}

	\Answer{\TODO{...}}

	\Comment{Please put brackets around the probability sum within the "$Support(C)$" definition to make clear that you are comparing $\tau$ to the sum.}

	\Answer{\TODO{...}}

	\Comment{If I get it right, if you are facing two clusters, one dominant in stability and the other in support, you will report two centroids, right?
	So reactivities that are inducing very high stability might dominate your prediction? Please check this hypothesis and if true discuss within the manuscript!}

	\Answer{\TODO{...}}

	\Comment{If you are NOT reporting multiple structures (as indicated in your results) how do you deal with cases illustrated above?}

	\Answer{\TODO{...}}

	\Comment{On page 7 you write "Fig. 5 suggests the presence of 8 clusters", which I do not agree... for me the PCA depiction would yield 4 clusters, i.e. B, C, D and the merge of all others. Thus, please rephrase the sentence or provide some other reason why you came up with 8 clusters from the beginning..}
	\Answer{\TODO{...}}

	\Comment{Concerning data normalization (Cordero et al.): when resetting all non-reacting position to 0, dont you loose the "background noise" of the experiment? 
	Did you test other options like subtracting the minimum/average/maximum of all non-reacting positions from all reactivities (before setting non-reacting to zero)?}

	\Comment{I would be very interested to see the performance of multi-probe predictions via RNAfold (MFE+MEA) when merging their reactivity data.
	For instance: just average the reactivities (per position) and provide them as new pseudo-reactivity data to RNAfold.
	While this might extremely mess up predictions, it might also yield similar effects as IPANEMAP, i.e. importance of contradicting reactivities is reduced the more experiments are considered..}

	\Answer{\TODO{...}}

	\Comment{Given your tool's documentation, a couple of questions arise that should be answered within the README.

	\begin{itemize}\item How are the reported free energies evaluated since your predictions are based on (multiple) reactivity-biased predictions? (your code hints at RNAeval calls, but please be verbose in the documentation to enable a clear intepretation of the results and comparability with other tools' output)\end{itemize}}

	\Answer{\TODO{...}}

	\Comment{\begin{itemize}\item How are the reported Boltzmann probabilities computed? based on the overall sample set? the respective cluster? (in both cases: what energies are used?) or based on the true (unconstraint) partition function computed e.g. by RNAfold?\end{itemize}}

	\Answer{\TODO{...}}

	\Comment{Given Table 1 (CMCT) and the general unawareness of the Nussinov-like MEA-optimization of the importance of base pair stacking, you might be able to further improve your results when implementing an MEA variant that excludes lonely base pairs.}

	\Answer{\TODO{...}}

	\Comment{I highly recommend to integrate IPANEMAP into the package management system conda via the bioconda channel. This would:
		\begin{itemize}
	\item be easy since IPANEMAP is eventually a set of python scripts (only a simple build and dependency script needs to be integrated into bioconda's recipes github repo)
	\item  handle all dependencies for the user
	\item ensure a simple installation of the tool and all dependencies 
	\item increase visibility of the tool
			\end{itemize}}

	\Answer{\TODO{...}}

	\Comment{When reading your work, it reminds me of another application, ie. the computation of unpaired probabilities for RNA-RNA interaction prediction in the presence of multiple probing data. For instance, see 
	https://doi.org/10.1093/bioinformatics/bty1029
	where single-experiment-reactivities were integrated (via Vienna package routines). 
	Given your pipeline, you should be able to derive suitable probabilities either directly from Pareto-optimal cluster(s) or from the overall sample set.}

	\Answer{\TODO{...}}

	\Comment{Fix the following typos/unfortunate choices of words:\begin{itemize}
	\item general: et al $\to$ et al.
	\item 1-2-30 : please provide the abbreviation details for SHAPE
	\item 1-1-21 : remove comma after "experiments"
	\item 1-2-37 : "more dynamics"
	\item 1-2-49 : "qualitative"
	\item 1-2-53 : "of individual nucleotides"
	\item 2-1-23 : "These include " beside structure optimization " the joint folding.."
	\item 2-1-49 : signal $\to$ signals
	\item 2-2-38 : criterion "are" $\to$ "is"
	\item 2-2-39 : identified $\to$ represented
	\item 3-1-5 and 8 : better replace "$R_i$" with "$d_i$", since you are using R for reference later on
	\item 3-1-60 : maybe replace "\#Datasets" with "$|\mathcal{D}|$" ?!
	\item 3-2-17 : missing "are" in "that both"
	\item 3-2-32 : move "where $k$ is the number of clusters" to line 40, where it is needed.
	\item 4-1-8 : Within the formula "$Dist(d,d')$" you have to replace "$x$" with "$i$"
	\item 4-1-52 : "Non-canonical" $\to$ "Crossing" or "Pseudoknot-forming"
	\item 5-1-40 : "metrics" $\to$ "metric"
	\item 6-2-44 : "hajdin" $\to$ "Hajdin"
	\item 8-1-31 : "Figure 1" $\to$ "6"
	\item 10-1-56 : "Figure" $\to$ "Table"
		\end{itemize}}	
	\end{enumerate}

	\subsection*{Reviewer \#3}

	\begin{enumerate}
	\Comment{My primary concern is that the authors use a pseudo-free energy formulation (page 3) with parameters that are the same for all conditions and probing reagents.  I am very surprised that this works.  The Diegan et al. work fit the parameters empirically for SHAPE data.  Sean Eddy [Annu. Rev. Biophys. 2014. 43:433] more explicitly drew the connection between these parameters and the log odds ratio of a nucleotide being paired, following earlier work of Sukosd et al. [Nucleic Acids Research. 2013. 41: 2807].  The Das lab [Biochemistry. 2012. 51: 7037] also explicitly fit the log odds for DMS data using gamma distributions.  I would expect that each probing reagent and each probing condition would need different parameters.}

		\Answer{We also expect that additional conversion formulae, and a specific parameter calibration, for each condition, would be beneficial in order to obtain better predictive performances. In fact, it is one of our project to produce more extensive data for a subset of the conditions explored in this study, and to attempt to systematically learn reactivities-to-pseudo-energy conversion models.

		However, the partial understanding of what is precisely being measured through different types of probing, in conjunction with the relative scarcity of available data for certain reagents/technologies, would probably induce very heterogeneous quality for such pseudo-energies across conditions. In particular, training for those parameters would probably lead to overfitting in cases where only a couple of reactivity profiles are currently available.

		Rather than address this heterogeneity as a possible confounding factor in our analyses, we have preferred to adopt a homogeneous treatment for all sources of probing data. It seems to us that the benefits of considering comparable reactivity profiles (produced in an homogeneous manner, and similarly normalized) far outweigh the loss in predicting accuracy.

		\TODO{Discuss the fact that the normalization used in various studies are deemed equivalent}

		}

	\Comment{The manuscript needs to be clearer about how the two parameters (m and b) determined for this work.  What optimizations were performed?  It needs to be clear that proper cross-validations were performed if the parameters were optimized.  Conversely, if the optimizations were not performed, it seems like an opportunity was then lost to have the best accuracy.}

		\Answer{\TODO{Clarify}}

	\Comment{I am also concerned that the manuscript is not as clear as it should be about the three types of multiple data that are used.  It can use multiple reagents, probing data for multiple solution conditions, or probing data across multiple sequence variants (a.k.a. mutate and map data).  It is a real achievement that one program handles all three cases.  The abstract does state this (briefly).  The Introduction is not clear on this points, just calling the data “multiple probing data.”  Also, the text itself is sometimes unclear about what is meant.  On page 6, for example, the sentence “It is also worth mentioning that NMIA+CMCT, combining the best and worst conditions, achieves a better combined performance than DMS+NMIA, the two best mono-probing conditions.” calls these “multiple conditions”, but these data are from multiple probing reagents.  I would like to see a revised manuscript be clearer about the three types of information in the Introduction.  In the text, multiple reagents, multiple solution conditions, and multiple sequence variants should be clearly denoted as different multiple data types. }

	\Answer{\TODO{Clarify in manuscript what conditions means in the context of the study}}

	\Comment{My third concern is that IPANEMAP is not entirely stable, in that there are variations in the results from run to run.  Users of the software will want the overall results to be stable; it is hard for most users to use a program multiple times and integrate the results.  It seems to me that the stability would be improved by using larger sample sizes, and that this should be explored for this manuscript.  If larger sample sizes don’t improve the stability, this should be addressed. }

	\Answer{\TODO{...}}

	\Comment{On page 3, the text states that “Non-canonical base pairs were removed (40), by considering as the secondary structure the maximum subset of non-pseudoknotted base pairs.” I think the manuscript should state that non-canonical pairs were removed and that pseudoknots were also removed.  Just removing the pseudoknots would not remove the non-canonical pairs. }

	\Answer{\TODO{...}}

	\Comment{On page 5, the MCC should be stated as the compromise between sensitivity and PPV (not sensitivity and specificity).}

	\Answer{\TODO{...}}

	\Comment{Figure 2 shows the accuracy of structure prediction using NMIA, DMS, and/or CMCT data.  As noted above, the pseudo free energy parameters for SHAPE (NMIA) will not be optimal for DMS and CMCT.  The benchmarks should therefore also be run using the pseudo free energies fit by the Das lab for DMS and CMCT [Biochemistry. 2012. 51: 7037].  My hypothesis is that the structure prediction accuracy will be higher with the properly fit parameters.}

	\Answer{\TODO{...}}

	\Comment{In the Conclusion and Discussion, it is an oversimplification to state that Rsample also uses stochastic sampling.  Rsample attempts to adaptively define the pseudo free energies to not perturb probabilities for nucleotides in an initial stochastic sample that are consistent with the SHAPE mapping data.  Perhaps this manuscript should also mention that detail and state this adaptive approach is possibly not as accurate (when producing a single structure) as the Diegan approach.}

	\Answer{\TODO{...}}

	\Comment{Typos, etc.:
		\begin{itemize}
		\item The quality of writing is excellent.  There are few typos and the text is clear.  
		\item I noted in the second column on page 3 that k is defined before the equation that uses it (1/(1+k)).  It would be more conventional to define k after the equation.  (I thought I had missed something as I read through).  
		\item In the Funding section, “phD” should be “PhD”.
		\item There are many problems with the references, however.  Many references are incomplete (missing page and volume, e.g. 22, 23, …).  A number of references also have superfluous months with the years (e.g. 5, 6, 7,…). 
	\end{itemize}}

	\Answer{\TODO{...}}

	\Comment{It would be helpful if the caption to figure 2 named the benchmark dataset.  This is clear in the main text, but will make it easier to quickly digest the manuscript if the captions are all self-contained.}

	\Answer{\TODO{...}}

	\end{enumerate}

	\subsection{Reviewer \#4}

	\begin{enumerate}

	\Comment{IPANEMAP is benchmarked against Rsample in a mono-probe setting. It is shown to outperform Rsample, where the authors attribute this to the sample denoising IPANEMAP performs via clustering (page 7). I am not sure if this is indeed the case and also think this benchmark could be improved in several ways.

	First, Rsample’s sampling routine is quite different than IPANEMAP’s and could be a relevant factor. It entails sampling from SHAPE reactivity distributions followed by partition function optimization. The clustering’s contribution to performance gains can be assessed by applying such clustering to Rsample’s output sample, right before the MEA step. Such analysis is useful because if this routine is key to improvements, it may be adopted by numerous other methods that analyze Boltzmann samples.}

	\Answer{\TODO{...}}

	\Comment{Second, my understanding is that a sample size of 1,000 was used in all benchmarks. However, unlike the 1,000 size-convention, Spasic et al. used 10,000 structures in their benchmarks. It is possible that this deeper sampling is necessary to Rsample’s performance. I think it would be fair to compare IPANEMAP and Rsample performances using 10,000-structure samples. }

	\Answer{\TODO{...}}

	\Comment{Third, the authors mention that Rsample was “shown to perform favorably against a comprehensive collection of state-of-the-art methods…” This is incorrect. In fact, based on a t-test, Spastic et al. concluded that “all methods that use experimental restraints perform similarly.” A similar conclusion, also based on a statistical test, was reported by Deng et al. (2016). This matters because it points out a major weakness of this work, namely, it does not consider the statistical significance of its findings despite reporting substantial variability between predictions. Additionally, it highlights the possibility that a comparison to another method may result in different findings since there are a few sequences with dramatic performance gaps, which were not considered by Spasic et al.}

	\Comment{Fourth, I think the stability of predictions over 10 independent samples should have been compared between the two algorithms. I could find stability results for IPANEMAP only (Table S5). In my experience, reproducibility of predictions is the most challenging issue with Boltzmann-based algorithms, especially when relying on clustering techniques. }

	\Answer{\TODO{...}}

	\Comment{Fifth, the Methods section indicates IPANEMAP does not use default “m” and “b” values for Deigan’s algorithm because different values seem to work better. It is unclear to me which of the analyzed datasets was used to determine these values or how the “m” and “b” parameters were optimized. These details should be included in the text. Moreover, a comparison between the algorithms should train them on the same dataset. If the Hajdin dataset was used in IPANEMAP’s parameter optimization, I think it is fair to re-optimize Rsample’s parameters too. This is because of the few sequences that were not considered by Spastic et al., for which performance gains are huge. The different training sets could be a factor in the reported performance gains. }

	\Answer{\TODO{...}}

	\Comment{Sixth, according to the abstract, IPANEMAP includes an unsupervised learning component. However, if the Deigan’s “m” and “b” parameters were optimized against known structures, then I don’t see which parts are unsupervised. This is another reason why I request to clarify the parameter optimization step. }

	\Answer{\TODO{...}}

	\Comment{The bulk of the manuscript describes various benchmarks, typically comparing average accuracies over multiple IPANEMAP runs, over multiple RNAs, or over multiple conditions. However, a comparison of two sets of data points should consider their means in the context of their std’s, otherwise it is often not too informative or conclusive. What I’m missing here are classical tests that quantify the statistical significance associated with each comparison. This is missing in all tests conducted in this work, and I suspect that a few of the reported mean differences might not be very significant. One example of where a test is needed is the comparison to Rsample -  see my comment above. Another example is the comparison to MFE and MEA (Table 2). There are several more. Note that if multiple tests are jointly considered/reported then p-value adjustment (e.g., B-H) might be warranted.}

	\Answer{\TODO{...}}

	\Comment{The authors generated 16 new datasets that enabled them to consider various ways of combining data from multiple conditions. This is impressive, with two caveats. First, only a single RNA of medium length was probed. Second, it is my understanding  that no replicate data was generated or analyzed, with the exception of SHAPE-CE. Typically, methods are tested on at least a few RNAs of varying lengths and whenever possible on datasets like Hajdin et al.’s. From such studies, we know that data-directed gains vary substantially between RNAs. For this reason, I would be hesitant to draw any conclusions from probing data for a single RNA. That said, I do not expect the authors to generate more data. I do think it should be emphasized that this is merely a case study. The purpose of this study and what one could conclude from it should also be clarified. While the subsection’s title (page 7) indicates this is a case study, I believe that some statements in the Abstract and Discussion require further substantiation and should be toned down.}

	\Answer{\TODO{...}}

	\Comment{The other reason I feel these results should be framed as a case study is that replicate data was not considered in the analysis. The manuscript attempts to differentiate between conditions or between algorithms, but the variation within each condition should be assessed in order to confirm that observed differences are in fact meaningful. }

	\Answer{\TODO{...}}

	\Comment{The assessment of mono-probing predictions on page 8 does not seem to result in a clear and significant take-home message. Accuracies seem to be comparable between IPANEMAP, MEA, and MFE, but I don’t see what we learn from this. Note also that the data-directed MEA and MFE algorithms were optimized for SHAPE data only, hence they might not do as well on non-SHAPE data (Deng et al., RNA, 2016).}

	\Answer{\TODO{...}}

	\Comment{Additionally, there is strong A/C bias in DMS data obtained by stop-based protocols. I don’t know if it was accounted for during normalization step, if normalization was indeed applied. I also don’t know which datasets were used to optimize IPANEMAP’s parameters, but as I commented above, a fair comparison should train all methods on the same data.}

	\Answer{\TODO{...}}

	\Comment{Finally, I don’t see why IPANEMAP was compared to MEA and MFE in this subsection but to Rsample in the previous one. Any special reason why Rsample is not included? The authors may also want to consider moving this part to the Supp.}

	\Answer{\TODO{...}}

	\Comment{My comments above re: IPANEMAP, MEA, and MFE also apply to their comparison for the Cordero et al. dataset (page 5).}

	\Answer{\TODO{...}}

	\Comment{Please edit for typos and grammar. There are also frequent transitions between present and past tense within paragraphs, which don’t read well.}

	\Answer{\TODO{...}}

	\Comment{I think an explicit discussion of IPANEMAP’s limitations is warranted in Discussion or as an independent subsection.}

	\Answer{\TODO{...}}

	\Comment{Many details could be moved to the Supp. to improve readability, specifically descriptions of results that essentially parse a relevant table for the reader. Often, the text gets so focused on the trees that one can’t see the forest. Briefly stating the main trend or conclusion with reference to a table should suffice. Examples include the last 2 paragraphs on page 9, first paragraph on page 10, and third paragraph on right column on page 11 (“For the WT…”).}

	\Answer{\TODO{...}}

	\Comment{Much space is allocated to outlier detection (cluster B) and then to describing analyses with and without cluster B. I feel that the authors can safely designate cluster B as an outlier and exclude it from the analyses described in the main text. Analyses with cluster B can be described in the Supp.}

	\Answer{\TODO{...}}

	\Comment{Were all the analyzed datasets normalized by any standard method (e.g., 2\%-8\% or box-plot)? I noticed that the mutate-and-map dataset was renormalized, and to the best of my knowledge, the Hajdin dataset is also normalized. However, I didn’t see any information re: Cordero dataset, and more importantly re: the newly generated data analyzed here. Normalization is especially important in the context of differential analysis between conditions.}

	\Answer{\TODO{...}}

	\Comment{Fig. 5: Why was k set to 8 as opposed to, say, 4 or 5? As the authors note, determining the number of clusters is tricky and can significantly impact results. }

	\Answer{\TODO{...}}

	\Comment{The authors compare stop-based to mutation-based experiments. However, the term “stop-based” is somewhat misleading because there is another variable here - the modification assay platform. MaP protocols must be assayed by NGS whereas the stop-based protocols used in this work rely on CE for modification detection. It is possible that the differences observed between MaP and CE data (see e.g., Fig. 5) have little to do with the stop strategy and more with the sequencing platform (CE vs NGS). The use of “stop-based” might create an impression that this feature is necessarily responsible for these differences. The ideal comparison should be between experiments such as SHAPE-MaP and SHAPE-seq, but given that experiments were already performed, I recommend changing the term and clarifying the technical differences between these subsets of conditions.}

	\Answer{\TODO{...}}

	\Comment{Does IPANEMAP perform better when sampling deeper? More generally, was sample size of 1,000 selected following performance optimization or simply because there is a convention in the field that 1,000 suffices for reproducibility at the base-pair level?}

	\Answer{\TODO{...}}

	\Comment{Table 1 could be moved to the Supp.}

	\Answer{\TODO{...}}

	\Comment{Table 2 summarizes performance differences at the Mut/Stop and -/+Mg levels. I wonder if such differences could have been detected already at the reactivity level and without the need for Boltzmann sampling. For example, one could lump all “Mut” profiles together as replicates for condition A and all “Stop” profiles as replicates for condition B and then perform differential reactivity analysis to spot differentially reactive regions and their statistical significance.}

	\Answer{\TODO{...}}

	\Comment{In Table 2, the two conditions in cluster C perform quite differently: 74\% and 85\%. Can the authors suggest why this happens? Can the performance differences be picked up by a differential reactivity analysis? For example, there may be a certain differentially reactive region, which the ensemble distance measure does not pick up. }

	\Answer{\TODO{...}}

	\Comment{Does bi-probing improve precision (std) over mono-probing? Similarly, does probing with multiple similar conditions improve precision compared to mono- or bi-probing (Table 3)?}

	\Answer{\TODO{...}}

	\Comment{Performance assessments in this work focus primarily on averages although precision is an important factor, especially when statistical sampling is involved and large variation in prediction accuracy is observed for different RNAs. The conclusion that one might not gain much or at all from increasing the no. of probes or using similar ones overlooks the notion that redundancy potentially increases our confidence in the results. For example, replicates are inherently redundant and ideally not too diverse, yet they are generated to decrease std around the mean although the mean is always lower than some of the point estimates. A related question is as follows.}

	\Answer{\TODO{...}}

	\Comment{For bi-probing, if we repeat predictions with 10 independent Boltzmann samples, which strategy is more precise: averaging predictions between the two probes or the corresponding bi-probing prediction?}

	\Answer{\TODO{...}}

	\Comment{Table 5: I do not think it demonstrates a significant finding. All subsets include $1M7-MaP_{IL}^{MG}$, whose mono performance is already 0.85, as compared to 0.853. According to Table 2, it is the best-performing condition. My guess is that the 0.003 gain is an artifact rather than a manifestation of complementary information embedded in the additional conditions. If the point made is that it’s advantageous to jointly analyze conditions rather than average them out, then I think it could be made in the context of the other results. However, as shown here, using too many conditions might negatively affect performance, so not sure if robust guidelines can be derived from these results.}

	\Answer{\TODO{...}}

	\Comment{Fig. 9: $1M7-MaP_{IL}^{MG}$ is a top performer on the Lariat data per Table 2, but its MCC variance over 10 runs is high per Fig. 9. If I interpret Fig. 9 correctly, the 85\% MCC from Table 2 is not shown for this condition and maybe resulted from another IPANEMAP run. In any case, I wonder if Table 2 lists the results of single runs or averages over 10 runs. If single runs are listed, how representative are they for the two conditions with high variances (Fig. 9)? Maybe replace them with average runs? Same questions for results with multiple probes. Note that the case for $NMIA-MaP_{IT}$ is similar as 80\% seems to be max performance.}

	\Answer{\TODO{...}}

	\Comment{Page 8, line 55: 1000 samples or 1000 structures per condition?}

	\Answer{\TODO{...}}

	\Comment{Methods, page 3: The definition of highly similar clusters uses a default value $\delta=1$. My understanding is that this means that highly similar clusters must have the exact same centroids, such that minor variation is not allowed. Is that correct?}

	\Answer{\TODO{...}}

	\Comment{Methods, page 3: Is $E_{d}(S)$ the pseudo-energy term of the total score or is it the entire score, including the NNTM free-energy? }

	\Answer{\TODO{...}}

	\Comment{My impression is that the main purpose of the clustering was to detect the outliers in cluster B. Since I find this to be a technical piece of the entire analysis, and not a very interesting one, I suggest moving most of the text + Fig. 5 \& 6 to the Supp.}

	\Answer{\TODO{...}}

	\Comment{Page 5, benchmarks on Cordero et al.’s dataset: It is noted that the prediction is the centroid of the largest probability cluster. My understanding, based on the method’s section, is that the prediction should be the centroid of the pareto-optimal clusters. It is unclear whether this is a modification of the method, and if yes, if it applies to all other benchmarks or not. Please clarify. }

	\Answer{\TODO{...}}

	\Comment{Page 10, line 56: Figure 3 should be Table 3.}

\end{enumerate}

%\bibliographystyle{amsplain}
%\bibliography{response}
\end{document}

### 2019-11-16 Response-NAR-Round1.tex
Updates from Overleaf

@@
 \documentclass[9pt,hyperref]{article} % use larger type; default would be 10pt
+\setcounter{secnumdepth}{1}
 
@@
 %%% PACKAGES
+\usepackage{hyperref}
+\hypersetup{
+	colorlinks, linkcolor={red!80!black},
+	citecolor={blue}, urlcolor={red!50!black}
+}
+
 \usepackage{amsmath} % for much better looking tables
@@
 \usepackage{cabin}
+\usepackage{enumitem}
 \usepackage{tgtermes}
@@
 \usepackage{todonotes}
-\newcommand{\TODO}[2][All]{{\todo[color=blue!20,inline]{{TODO (#1)}: #2}}}
+\newcommand{\TODO}[2][All]{{\todo[color=blue!10,inline]{\color{black}{TODO (#1)}: #2}}}
 
@@
 
+\tableofcontents
+
 \section{General comments}
+
+First of all, the authors would like to express their deepest gratitude for the quality and thoroughness of the reports written by the reviewers, covering a remarkably comprehensive range of aspects of the submitted manuscript.
+We have individually addressed each of the \ref{comment:last} comments to the best of our ability, but would like to start this note with higher-level answers to recurrent questions:
 
@@
 
-\subsection*{Reviewer \#1}
+\subsection{Reviewer \#1}
 
@@
 The authors should add a panel clustering the raw probing data to compare it with the IPANEMAP clustering.
-This will require some clever design of distance metrics and mapping 1M7 to DMS maybe complex, but they should do a heat map like Figure 6 but with the raw data minimally.}
+This will require some clever design of distance metrics and mapping 1M7 to DMS may be complex, but they should do a heat map like Figure 6 but with the raw data minimally.}
 
-	\Answer{	On a fait ça, c’est en supp mat. On observe :
+	\Answer{
+		\TODO{On a fait ça, c’est en supp mat. On observe :
 			Corrélation inverse entre distance à la native et MCC\%;
-		Difficulté à regarder le clustering car peu significatif
+			Difficulté à regarder le clustering car peu significatif}
 	}
@@
 	\Answer{We agree that raw data should be made available to the reader for the sake of reproducibility.
-Unfortunately, the SNANASM were no longer available at the time of this revision.
+Unfortunately, the SNANASM was no longer available at the time of this revision.
 We have informed the creators of the resource.
@@
 
-	\subsection*{Reviewer \#2}
+	\subsection{Reviewer \#2}
 
-	\begin{enumerate}
+	\begin{enumerate}[resume]
 \Comment{I find this tool and study very interesting, especially the systematic assessment of the impact and combined power of various probing protocols for DiLCrz.
@@
 
-	\subsection*{Reviewer \#3}
+	\subsection{Reviewer \#3}
 
-	\begin{enumerate}
+	\begin{enumerate}[resume]
 	\Comment{My primary concern is that the authors use a pseudo-free energy formulation (page 3) with parameters that are the same for all conditions and probing reagents.
@@
 
-	\begin{enumerate}
+	\begin{enumerate}[resume]
 
@@
 
-	\Comment{Page 10, line 56: Figure 3 should be Table 3.}
+	\Comment{Page 10, line 56: Figure 3 should be Table 3.\label{comment:last}}
+
+	\Answer{\TODO{...}}
 

### 2019-11-16 SuppMat.tex

@@
 
-\section*{Multiprobing benchmark (Cordero~\emph{et al}~\cite{Cordero2012} dataset)}
-\subsection*{MCC values of individual predictions}
+\tableofcontents
+
+\section{Multiprobing benchmark (Cordero~\emph{et al}~\cite{Cordero2012} dataset)}
+\subsection{MCC values of individual predictions}
 
@@
 \end{table}
-\subsection*{Stablity analysis}
+\subsection{Stablity analysis}
 \begin{table}[H]
@@
 %\section{IPANEMAP Vs Rsample}
-\section*{Mono probing benchmark (Hajdin~\emph{et al}~\cite{Hajdin2013} dataset)}
+\section{Mono probing benchmark (Hajdin~\emph{et al}~\cite{Hajdin2013} dataset)}
 
@@
 
-\subsection*{Detailed results of \OurTool{} and \Software{Rsample}}
+\subsection{Detailed results of \OurTool{} and \Software{Rsample}}
 \begin{table}[H]
@@
 
-\subsection*{Stability analysis}
+\subsection{Stability analysis}
 
@@
 
-\section*{Case-study: Lariat capping ribozyme of {\itshape Didymium iridis}}
-\subsection*{Compatibility of accessibility profiles with native and predicted structures}
+\section{Case-study: Lariat capping ribozyme of {\itshape Didymium iridis}}
+\subsection{Correlation analysis of reactivities}
+\subsection{Compatibility of accessibility profiles with native and predicted structures}
 We assess and validate the overall quality of probing data using a basic notion of distance between discretized profiles and structure.

### 2019-12-22 Response-NAR-Round1.tex
Updates from Overleaf

@@
 %%% PAGE DIMENSIONS
-\usepackage{geometry} % to change the page dimensions
+\usepackage[margin=1.3in]{geometry} % to change the page dimensions
 \geometry{a4paper} % or letterpaper (US) or a5paper or....
@@
 
+\usepackage[authoryear]{natbib}
 \usepackage{amsmath} % for much better looking tables
@@
 \usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
-\usepackage{framed}
+\usepackage{mdframed}
+\usepackage{relsize}
+\usepackage{pgf}
 \usepackage{cabin}
@@
 \usepackage{tgtermes}
+\usepackage{collcell}
+\usepackage{adjustbox}% to resize figures
+
 %\usepackage{tgadventor}
@@
 
-\newcommand{\Answer}[1]{\noindent{\color{purple!80!black}{{\sffamily Response:} }{#1}}}
-\newcommand{\Comment}[1]{\item{\color{black}{{\sffamily Comment:} }{#1}}}
+\newcommand{\FAnswer}[1]{{\color{black!95}#1}}
+\newcommand{\IAnswer}[1]{\FAnswer{\hfill $\to$ #1}}
+\newcommand{\PAnswer}[1]{\FAnswer{\\#1}}
 
-\usepackage{todonotes}
+\newcommand{\Answer}[1]{\noindent{{{\sffamily Response:} }{\FAnswer{#1}}}}
+\newcommand{\Comment}[1]{\item{\color{red!40!black}{{\sffamily Comment:} }{#1}}}
+\newcommand{\Quote}[1]{\begin{mdframed}[backgroundcolor=black!2,linecolor=gray!50, linewidth=.6mm]#1\end{mdframed}}
+
+\usepackage[colorinlistoftodos]{todonotes}
 \newcommand{\TODO}[2][All]{{\todo[color=blue!10,inline]{\color{black}{TODO (#1)}: #2}}}
+\newcommand{\TODOBruno}[2][Bruno]{{\todo[color=red!10,inline]{\color{black}{TODO (#1)}: #2}}}
+\newcommand{\TODOAfaf}[2][Afaf]{{\todo[color=Aquamarine3,inline]{\color{black}{TODO (#1)}: #2}}}
 
@@
 \tableofcontents
+\listoftodos
 
@@
 
+\TODO[Yann]{Faire résumé}
 Points géneraux :
-
-On a probablement pas été super clairs, mais :
+\begin{itemize}
+\item On a probablement pas été super clairs, mais :
 Pas de comparaison entre protocoles, plutôt un test, rendu possible par IPANEMAP de la complémentarité des protocoles
-Pas optimisation du traitement des donnés de sondage, mais traitement uniforme
-Approche multi probing nous parait susceptible d’être robuste à l’incorporation d’une réactivité moins informative.
+\item Pas optimisation du traitement des donnés de sondage, mais traitement uniforme
+\item Approche multi probing nous parait susceptible d’être robuste à l’incorporation d’une réactivité moins informative.
+\end{itemize}
 
@@
 	\Answer{
-		\TODO{On a fait ça, c’est en supp mat. On observe :
-			Corrélation inverse entre distance à la native et MCC\%;
-			Difficulté à regarder le clustering car peu significatif}
+		First of all, we need to stress that it is not the objective of our manuscript to establish the predictive potential of various probing technologies and reagents.
+Rather, we wish to embrace the diversity of available protocols, and harness the resulting data to improve its integration and exploitation within structure prediction pipelines.
+
+		That being said, we agree that a cross-comparison of the raw reactivity profiles is potentially informative, and useful to better understand the data.
+Accordingly, we now dedicate a whole section of the supplementary data to a correlation analysis, whose results and conclusions are reproduced below.
+
+		\Quote{
+			\subsection*{Cross-analysis of raw reactivities profiles}
+
+			We compared raw reactivity data produced across various conditions, in order to test their overall compatibility, investigate their main determinants and compare their impacts on predictions.
+
+						\includegraphics[width=.54\textwidth]{graphs/didy/reactivity_correlation.pdf}\hfill
+			\includegraphics[width=.45\textwidth]{graphs/didy/bi_clustering.pdf}\\
+			\noindent Figure~A: Correlation of reactivity profiles (left) and induced (pseudo-)ensemble distance (right).
+Conditions clustered in our analysis due to their impact of the pseudo-Boltzmann ensemble (black boxes, right) do not necessarily correlate very well in term of raw reactivity (white boxes, left).
+
+			\paragraph{Correlation metrics and rationale.}To that purpose, we consider the Pearson product-moment \Def{correlation coefficients} $C_{R,R'}$ of two reactivity profile $R$ and $R'$, defined as
+			$$ C_{R,R'} = \frac{\sum^n_{i=1}\left(R_i - \mu_R\right)\left(R'_i - \mu_{R'}\right)}{\sqrt{\sum ^n _{i=1}\left(R_i - \mu_{R}\right)^2} \sqrt{\sum ^n _{i=1}\left(R'_i - \mu_{R'}\right)^2}},$$ 
+			where $\mu_{R}$ and $\mu_{R'}$  denote the mean reactivity within $R$ and $R'$ respectively.
+
+			The choice of the \Def{correlation coefficient as a measurement of pairwise similarity} for probing profiles is motivated by differences in the distributions of reactivities across experimental conditions and protocols.
+			Indeed, $C_{R,R'}$ remains unchanged by the application of any rescaling, through an affine transformation with positive slope, to either $R$, $R'$ (or both).
+Moreover, this metrics does not scale with the number of positions.
+			Thus, we can restrict the sums and mean values to positions for which reactivities are available in both experiments, and still compare the coefficient with another pair of reactivity profiles.
+This property is important in order to tolerate \Def{missing values} introduced by experimental artifacts and reagent properties, while still being able to compare the coefficients.
+
+			\paragraph{Results.} We computed the correlation coefficient for all pairs of reactivity profiles, and report the results in the left matrix of Figure~A.
+For the sake of comparison, we also remind the Ensemble Distance matrix, indicating the difference in impact on the computational predictions of each pair of conditions.
+
+			We first observe that almost all of the pairs of reactivity profiles show a positive correlation, as expected from their observation of a common structure.
+A notable exception is \OneMSevMgCE, having average negative correlation with all its peers (except for \NMIAMgCE).
+Its outlier status is confirmed by its large ensemble distance to other conditions, yet does not prevent it to achieve a respectable MCC of 70\%.
+Finally, the perfectly negative correlation between \CMCTMg{} and \DMSMg{} can be dismissed as a computational artifact, since CMCT and DMS react with different nucleotides, and the correlation cannot be computed.
+
+			Moreover, we observe that conditions clustered from the Ensemble Distance metrics (due to their impact of the pseudo-Boltzmann ensemble) do not always correlate in term of raw reactivity, with very poor correlations occurring within clustered conditions.
+This essentially reflects the fact that similar profiles may induce very similar pseudo-Boltzmann ensembles, \emph{ie} different premises sometimes imply the same conclusions.
+
+			Conversely, one observes very few strong correlations outside of clusters.
+In fact, the only two exceptions are \OneMSevILU and \OneMSevILUThree which, despite belonging to different clusters and achieving different qualities of prediction, correlate almost perfectly with each other and with other reactivity profiles produced using 1M7/SHAPE-MaP.
+Other natural candidates for explaining the reactivity correlations (Map vs CE, +Mg vs -Mg) do not appear to be supported by this analysis.
+
+			Overall, this analysis confirms the consistency of the produced probing profiles (positive correlation), but also suggests a very delicate effect of reactivity data on the downstream predictions, with very different profiles leading to similar predictions (\OneMSevILU), and similar profiles inducing very different ones (\OneMSevILU vs \OneMSevILUThree).}
+
+				In particular, while reactivity profiles correlate positively in a pairwise fashion, as can be expected, these results do not support the hypothesis that the method of detection (CE vs Map) would represent a strong differentiator at the level of reactivity profiles.
+In fact, the only clear clusters are:
+		\begin{itemize}
+			\item  $\{\NMIAMg,\NMIA\}$;
+			\item  $\{\NMIACE,\NAICE,\OneMSevCE\}$;
+			\item $\{\OneMSevILU, \OneMSevILUThree,\OneMSevILUThreeMg,\OneMSevILUMg\}$,
+		\end{itemize} having respectively a similar reagent, technology, or both, as a common trait.
+
 	}
@@
 
-	\Answer{There is a misunderstanding here, the “(-) No Prob” conditions refers to a modeling process that does not use probing data as constraints.
-We understand that this was confusing and we no refer to it as “(-) No Exp” for modeling without experimental data.
-“(-) No Cons” for modeling without experimental constraints.
-Of course we have carried out control experiments without any modification reageant for each of the experiments reported, as now explicitly specified in the material and methods section.}
+	\Answer{There is a misunderstanding here, the “(-) No Prob” conditions refers to a baseline computational prediction that does not use any probing data as constraints (Turner energy model only).
+We understand that this may be confusing to the reader, and we now refer to it as “(-) No Exp” for modeling without experimental data.
+
+	 Of course we have carried out control experiments without any modification reageant for each of the reported experiments, as now explicitly specified in the material and methods section.
+	 \TODOBruno{J'écris ça, mais je suis plus trop sûr, tu peux vérifier: \\ \didy{} was also probed with DMS (DiMethylSulfate) and CMCT, this time only in presence of Mg$^\text{2+}$, resulting in a total of 16 \Def{probing conditions}, a term we use in the following to denote a combination of probing technology, reagent and presence/absence of Mg$^\text{2+}$ (+ sequencing technology).
+For each probing condition, three experiments were performed in presence/absence of the reagent and in a denatured setting, following classic protocols~\cite{Wilkinson2006,Smola2015}.}}
 
@@
 	\Answer{We understand this legitimate concern, expressed by multiple referees.
-The probing data reported on the structure for each probing conditions and reagents is now provided as supplementary material.
-	Please note that the DMS results were not obtained following a ShapeMap-like strategy but by a “classical” RT-stop technique and Capilary electrophoresis.}
+The probing data reported on the structure for each probing conditions and reagents is now provided in Supplementary Figures S3 and S4.
+Also, please note that the DMS results were not obtained following a SHAPE-MaP-like strategy but by a “classical” RT-stop technique and Capilary electrophoresis.}
 
@@
 	\Answer{It is exact that the Weeks laboratory uses an RNAse H minus mutant of the MMLV RT sold as “Superscript II” and advises to use it.
-However as far as we can tell from reviewing Dr Weeks bibliography, they never reported the test of “superscript III” (another RNAse H- mutant of the MMLV RT sold by the same company), and therefore never shown that does not detect the adduct.
-Actually in Smola et al.
-they specify “Reaction conditions have been optimized for SuperScript II reverse transcriptase only.
+However as far as we can tell from reviewing Dr Weeks bibliography, they never reported the test of “superscript III” (another RNAse H- mutant of the MMLV RT sold by the same company), and therefore never showed that this technology does not detect the adduct.
+
+		More precisely, in \cite{Smola2015}, the authors state that “Reaction conditions have been optimized for SuperScript II reverse transcriptase only.
 Other reverse transcriptase enzymes and derivatives have not been tested and should not be used”.
-We tried superscript III in the conditions described and analysed the histograms provided by ShapeMapper which are metrics for the success of a shapeMap experiment as defined by smola et al.
-2015.
+In this work, we tried superscript III in the  described conditions, and inspected the coverage/mutation distribution graphs produced by ShapeMapper.
+Those graphs are suggested by the Weeks group to assess the success of a SHAPE-MaP experiment, as described in \cite{Smola2015}.
+
 	Please find below some examples of the histograms obtained using Superscript III:
@@
 	\begin{itemize}
-	\item NMIA
-	\TODO{Add plot}
-	\item NMIA mg
-	\TODO{Add plot}	
-	\item 1M7
-	\TODO{Add plot}
+	\item \NMIA:
+
+	{\centering\includegraphics[width=.8\linewidth]{graphs/didy/NMIA}\\}
+
+	\item \NMIAMg:
+
+	{\centering\includegraphics[width=.8\linewidth]{graphs/didy/NMIA-Mg}\\}
+
+	\item \OneMSevILUThree:
+
+	{\centering\includegraphics[width=.8\linewidth]{graphs/didy/1M7}\\}
+
 	\end{itemize}
 
-	As can be seen on the right  histograms the read depth is sufficient (>10k reads/nucleotides) for these experiments to be interpreted.
+	As can be seen on the right  histograms the read depth is sufficient ($>$10k reads/nucleotides) for these experiments to be interpreted (\emph{ie} for the frequency of mutations to be a good estimate of the probability).
 The histograms on the left shows that there are significantly more mutations in the modified (denatured and SHAPE) samples than in the untreated ones.
-These histograms are very similar to those shown by smola et al.
-as exemplary of a successful experiment.
-	In addition the shapeMap results are consistent with the X-ray structure (see figure XX in the supp.
-material), and their use with any prediction software as shown to improve the accuracy of the obtained model(s).
 
-	In conclusion there is no doubt about the relevance of these results and we do not see the necessity to repeat them using superscript II.}
+	These histograms are very similar to those deemed as exemplary of a successful experiment by \cite{Smola2015},. 
+	In addition, those SHAPE-MaP results are consistent with the X-ray structure (see Supplementary Figure S2 and S3), and their use with our prediction software was shown to improve the accuracy of the obtained model(s).
+
+	In conclusion, a thorough re-inspection of the data does not lead us to question the relevance of these results, and we do not see the necessity to repeat them using superscript II.}
 
@@
 
-	\Answer{As described above, we did not find any reason to doubt the quality of those results, and believe they should remain included in our analysis.}
+	\Answer{As discussed above, we did not find any reason to doubt the quality of our results, and believe they should remain included in our analysis.}
 
@@
 	\Answer{We agree that raw data should be made available to the reader for the sake of reproducibility.
-Unfortunately, the SNANASM was no longer available at the time of this revision.
-We have informed the creators of the resource.
-In the meanwhile, we have thus deposited the reactivity results on the GitHub page of IPANEMAP, where they also serve as illustrative examples.
+Unfortunately, the SNRNASM resource was no longer available at the time of this revision.
+We have informed the creators of the resource, but have not witnessed accessibility improvement over the course of our revisions.
+	In the meanwhile, we have deposited the reactivity files on the GitHub page of IPANEMAP, where they also serve as illustrative examples.
 
-	The raw data were deposited to the SRA under identifier \TODO[Bruno]{Deposit data in SRA}.}
+	We experienced other issues while attempting to upload the raw data onto the SRA.
+Namely, this database requires a structured definition of the experiment, including several mandatory fields (growth culture...) that we could not reasonably fill given the nature of our experiments.
+In fact, a cursory search of the database showed that primary data produced by the Weeks lab is very rarely deposited in the SRA, indicating that similar difficulties are also encountered by other groups.
+
+	We have permanently uploaded the data at:
+
+	{\centering \url{http://www.lix.polytechnique.fr/~ponty/suppdata/ipanemap/}\\}
+
+	and commit to its continued accessibility (but will also pursue public venues for better guarantees of a continued availability). 
+	\TODO{Add minimal page and reactivities, including scripts/instructions to compute reactivities.}
+
+}
 
 	\end{enumerate}
+
+\newpage
 
@@
 	\begin{enumerate}[resume]
-\Comment{I find this tool and study very interesting, especially the systematic assessment of the impact and combined power of various probing protocols for DiLCrz.
-To be ready for publication, the manuscript needs some further discussion of the DiLCrz results and a revision of the method presentation as detailed below.}
-
-	\Answer{\TODO{...}}
 
@@
 
-	\Answer{\TODO[Yann]{Check if we get better results by blocking the pseudoknot.}}
+	\Answer{It is true that at least four helices of DiLCrz (DP2, DP2.1, P10 and P9) are somewhat easy to predict, and are indeed consistently found in all predictions, including in the probing-free \Software{RNAfold} MFE prediction.
+The changes in performances are thus mostly due to the other 6 helices (omitting the pseudoknot).
+We emphasized this fact in the description of our result section, to provide the reader additional context for the interpretation of our results.
+However, we disagree that this impacts the generality of our results.
+In particular, badly produced/poorly combined reactivities may easily disrupt the four above helices, and the fact that our method is able to preserve those helices, yet predict other structure elements, should be considered an asset of our method.
 
-	\Comment{The methods section needs some reordering to first formaly introduce clusters and respective measures (stability/support) (e.g. in line 3-1-57) before introducing and discussing MBkM.
+	The remark regarding the impact of the pseudoknot subhelix (P7) on the reactivity is very perceptive, and certainly noteworthy.
+Indeed, positions 111 to 115 (the 5' region of P7) are typically associated with lower reactivities, providing incentives to algorithms for pairing them within their proposed models.
+Accordingly, one finds those positions to be paired to erroneous partners within most of our models.
+Such is not the case, however, for the 3' region of P7, which is typically left unpaired in our models.
+}
 
+	\Comment{The methods section needs some reordering to first formaly introduce clusters and respective measures (stability/support) (e.g.
+in line 3-1-57) before introducing and discussing MBkM.
 For this restructuring, you might want to introduce the cummulative sample multi-set "$\mathcal{S} = \sum_d \mathcal{S}_d$" from which the clusters (also multi-sets?!) form a partitioning.}
 
-	\Answer{\TODO{...}}
+	\Answer{We fully agree that reproducibility is essential, and have realized that our previous manuscript was lacking in details.
+Accordingly, we have largely revised our description of the method to clarify the nature of the clusters (indeed multisets), and the exact computation of centroids (including base-pair/unpaired position probabilities within a cluster).
+
+	However, we have chosen to avoid an overly formal description of the method, as we believe such a writing style would not represent a good fit with the expectations of the typical NAR audience.
+We have adopted a linear narration of the method, which we completed in the hope to now ensure full reproducibility of the method.}
 
 	\Comment{You do not introduce how you are able to compute MEA centroids without explicitely computing base pair probabilities.
-Checking your source code, I found the bp-probs are computed analogously to the structure probabilities, which is fine but should be explicitely mentioned to be reproducable.}
+Checking your source code, I found the bp-probs are computed analogously to the structure probabilities, which is fine but should be explicitely mentioned to be reproducable.
 
-	\Answer{\TODO{...}}
-
-	\Comment{It is totally unclear to me (and I didnt took the time to track it down in your code), how you derive an MEA for a cluster with structures from different reactivity data sets, since the same structure can be part of the cluster multiple times (with different probe-constraint probabilites).
+		It is totally unclear to me (and I didnt took the time to track it down in your code), how you derive an MEA for a cluster with structures from different reactivity data sets, since the same structure can be part of the cluster multiple times (with different probe-constraint probabilites).
 Do you sum up probe-specific probabilities in that case?
 Do you derive new base-pair probabilities for the cluster (and if so how)?
+
 	Please provide somewhere details of your methods concerning the whole bp-prob and MEA centroid computation (at least in the supplementary material).
@@
 
-	\Answer{\TODO{...}}
+	\Answer{As mentioned in our response to the previous point, we revised our method description to include more details on the probability distribution on which we base our centroid/MEA structure.
+The computation of the MEA is performed using a strict reimplementation of the dynamic programming algorithm introduced by \cite{Lu2009}, so we just refer the reader to this article rather than reproducing its content.
 
-	\Comment{Please replace "kT" by "\text{R}T" in the structure probability formula (and whereever needed)!}
+	Please find below the description of the clustering method, as now appears in our revised document.
+		\Quote{\paragraph{Clustering across conditions.} In order to infer recurrent conformations across sampled sets, \OurTool{} agglomerate structure (multi)sets while keeping track of their condition of origin, and \Def{clusters} with respect to the \Def{base-pair distance}, the number of base pairs differing between two structures.
+A clustering algorithm then partitions the (multi)set of sampled structures into \emph{clusters}, (multi)sets of structures such that the accumulated sum of distances over clusters is minimized.
 
-	\Answer{\TODO{...}}
+				Among the many available options, we chose the Mini Batch k-Means algorithm~\citep{Sculley2010} (MBkM), implemented in the \Software{scikit-learn} pacakage~\cite{Pedregosa2012}, which requires less computational resources than the classic k-means algorithm, yet performed similarly in preliminary studies as an extensive collection of both agglomerative (affinity propagation) and hierarchical (Ward, Diana, McQuitty) clustering algorithms.
+A dissimilarity matrix, presenting the pairwise base-pair distance between structures, is precomputed and fed to the clustering algorithm.
+
+				Any cluster $C$ output by the clustering is a multiset of structures, each labeled with its origin condition of $\mathcal{D}$.
+				The \Def{cluster probability of a structure feature} $f$ (base pair or unpaired base) within a cluster $C$ is then defined as 
+				$$ \mathbb{P}_C(f) = \frac{\sum_{\substack{S\in \mathcal{C}\\\text{s.t. } f\in S}} e^{-E(S)/RT}}{\sum_{\substack{S'\in \mathcal{C}}} e^{-E(S')/RT}} $$ 
+				where $R$ represents the Boltzmann constant, $E(S)$ is the Turner free-energy, and $\mathcal{C}$ is the (non-redundant) set of structures in $C$.
+From those probabilities we define the \Def{centroid structure} of a cluster as its Maximum Expected Accuracy (MEA) structure, computed efficiently following Lu\etal~\citep{Lu2009}.
+
+				Moreover, define the (pseudo-)\Def{Boltzmann condition probability} of a structure $S$, generated for a probing condition $d$ as part of a sampled  set $\mathcal{S}_d$, as
+				$$\mathbb{P}_d(S) = \frac{e^{-E_d(S)/RT}}{\mathcal{Z}^*_d} \text{, with } \mathcal{Z}^*_d := \sum_{S'\in \mathcal{S}_d} e^{-E_d(S')/RT}$$
+				where $E_d(S)$ is the pseudo free-energy assigned to the structure $S$ within the probing condition $d$.
+				The \Def{stability} of a cluster $C$ denoted its accumulated pseudo-Boltzmann probability across conditions, computed as
+				$$\text{Stability}(C) = \sum_{d\in \mathcal{D}} \sum_{S\in C\cap \mathcal{S}_d}  \mathbb{P}_d(S).$$
+				A cluster is deemed \Def{significantly populated} if its stability exceeds a predefined threshold $\epsilon$. 
+				We set  $\epsilon =\#{\sf Datasets}/3$ by default, such that at most three clusters are deemed significantly populated, and used as our primary candidates.
+				Finally, we consider two clusters to be \Def{highly similar} if their centroid structures differ by at most $\delta$ base pairs ($\delta=1$ by default), allowing the identification of clusters in the presence of minor variations.
+
+				The targeted number of clusters is a critical parameter of the \CL{} algorithm.
+It should, at the same time, remain small enough to ensure reproducibility, while being sufficiently large to discriminate outliers and ensure consistency within each cluster.
+We determine an \Def{optimal number of clusters} $\NumClust$ using an iterative heuristics, gradually increasing the number of clusters until a significantly-populated cluster is split into two similar clusters, or a poorly populated (outlier) cluster is created.
+				Namely, our iterative heuristic consists in running \CL over an increasing number $\NumClust$ of clusters, starting with $\NumClust=2$, until the following \Def{stopping criterion} is met: 1) Two significantly populated clusters have associated centroids which are highly similar; or 2) Centroid structures of significantly populated clusters from the previous iteration are highly similar to those of the current iteration.}
+}
+
+	\Comment{Please replace "kT" by "\text{R}T" in the structure probability formula (and wherever needed)!}
+
+	\Answer{We agree that using $k$ was unfortunate in this context, leading to an unnecessarily notational overload, and have used $R$ instead as suggested.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{Those two notions indeed coincide and we had initially used a sentence to avoid a forward definition.
+We have now moved the stability definition earlier in the method description, and use it to define the notion of \Def{significantly populated cluster}.}
 
-	\Comment{Please provide (eg in the supplement) some further details how to derived the values for m abd b for the Deigan method. So far, the description is quite vague.. }
+	\Comment{Please provide (eg in the supplement) some further details how to derived the values for m abd b for the Deigan method.
+So far, the description is quite vague.
+Within this context, you might want to discuss that too low pseudo-energy boni will yield similar results as hard constraints (which you want to avoid).}
 
-	\Answer{\TODO{...}}
-
-	\Comment{Within this context, you might want to discuss that too low pseudo-energy boni will yield similar results as hard constraints (which you want to avoid).}
-
-	\Answer{\TODO{...}}
+	\Answer{The values $m$ and $b$ were loosely calibrated by grid search based on the \cite{Cordero2012} dataset, revealing a limited influence of those factors on the predictive performances. }
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{Absolutely, this is now fixed.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{Thanks for the suggestion, which greatly improves the readability of the formula.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{
+		The interpretation of the reviewer is accurate, and it is true that the configuration of reactivities may give rise to \Def{tight} (focused on a few structures), or \Def{wide} (compatible will multiple conformers) pseudo Boltzmann Ensembles.
+Combining a \emph{tight} and \emph{wide} pseudo-Boltzmann ensembles will grant more influence to the former over the final prediction.
+		In the worst case scenario, the combination of a poorly-informed tight Ensemble, which we informally call a \emph{bully}, with an accurate wide (\emph{open-minded}) Ensemble, may lead the bully to almost solely drive the ultimate prediction.
+However, we believe such cases to be rare, as indicated by the good performances of our approach on single probing experiments and that, in general, tight ensembles are indicative of successful experiments.
+
+		Moreover, in the presence of multiple probing conditions, bullies can be overruled by a population of informed, even if open-minded, individuals.
+Our approach can indeed be seen, in first order approximation, as a weighted voting mechanism where each voter splits his vote between her/his preferred candidates.
+Since the probability of independently ill-informed bullies to agree by chance decreases exponentially with their number\footnote{In the absence of a systematic underlying bias, synchronizing the bullies, which in itself would be informative, or at the very least worthy of investigation.}, the voting process will ultimately favor a population of open-minded individuals over a, possibly majority, subpopulation of bullies.
+
+		Finally, it is worth reminding that reactivity values are normalized, so that the average reactivity (over nucleotides touched by the probe) remains the same across conditions, so that the probability of a tight Ensemble being an artifact due to a given probing condition is expected to be limited.
+	}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{
+		Over a normal execution, \OurTool{} {\bf does report} all the dominant clusters (and even dominated ones), along with their centroids.
+		In the context of benchmarking, unless explicitly specified, we systematically picked a cluster uniformly at random from the output to assess the predictive performances.
+
+		However, in an overwhelming proportion of our experiments, a single cluster is reported, due to a generally good correlation  of stability with support.
+For instance, it is mathematically impossible for a single condition to feature more than one Pareto cluster, since the cluster $C^\star$ having max stability $s^\star$ is such that: either $C$ has support $t^\star=1$, in which case it dominates any other cluster with signature $(s<s^\star,t\in\{0,1\})$; or $C$ has support $t^\star=0$, ruling out the possibility of $t=1$ for any other cluster (since it would imply $s>s^\star$).
+
+}
 
@@
 Thus, please rephrase the sentence or provide some other reason why you came up with 8 clusters from the beginning..}
-	\Answer{\TODO{...}}
+
+	\Answer{We respectfully disagree, and still see 8 clusters in the PCA (and distance matrix, which we reformatted and reproduced below).
+Of course, this view is quite subjective and we carefully chose our wording (\emph{suggest}) to reflect the inherent limitations of a visual inspection.
+	\Quote{
+			{\centering \includegraphics[width=.75\linewidth]{graphs/didy/bi_clustering}\\}%
+		}
+	On a more objective level, an examination of the distance matrix above indicates that merging conditions in the clusters A, E, F, G and H as suggested would be ill-advised, since some of the pairs of clusters ((E,F,G) vs H, A vs E\ldots) are actually quite distant.
+This  fact is somewhat obfuscated by the low-dimensionality of the PCA.
+}
 
@@
 	Did you test other options like subtracting the minimum/average/maximum of all non-reacting positions from all reactivities (before setting non-reacting to zero)?}
+
+	\Answer{
+		First, we need to stress that this cancellation of non-reactive nucleotides only applies to DMS and CMCT probing, and reflects the fact that those reagents are not supposed to react with certain nucleotides.
+Namely, DMS only reacts with \Ab{} and \Cb{}, while CMCT only reacts with \Gb{} and \Ub{}.
+We did verify that the reactivities associated with non-reacting nucleotides, as assessed using the relevant formula, are overwhelmingly close to zero so normalizing by the minimum/average/maximum would have no effect.
+Overall, these observations strongly support the idea of non-informative reactivity values at those positions.
+For instance, using them within the computation of the normalization (based on a percentile) would systematically increase the average reactivity, leading the pseudo-energy bonuses to act as \emph{hard constraints} for all practical purposes.
+
+		Moreover, the reactivity normalization is based on percentiles over reacting positions, and there is no reason to suspect a systematic bias between the nature of nucleotides and the structural context.
+It follows that the effect of the normalization should be very similar when starting from a fully informed reactivity profile, or ignoring reactivities associated with \Gb{}/\Ub{} nucleotides.
+
+		Finally, we must point out that we do not set the reactivities of non-reacting nucleotides to $0$, but rather to \emph{non-reactive}, usually encoded as a large negative value.
+This value is then interpreted as a missing data within \Software{RNAfold}, so that no pseudo-energy (rather than the bonus induced by a reactivity of $0$ by the \cite{Deigan2009} conversion) is used for those positions.
+
+}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{
+		In principle, this idea of averaging reactivities is natural, but has limits. 
+		In particular, it does not sample from the same distribution as our approach, as can be seen in the following toy example:
+
+		{\centering
+			\begin{tabular}{ll}
+				Structure $S^\star$ (free-energy $E^\star$)& \texttt{(((..)))(((..)))} \\
+				Structure $S^\varnothing$ ($E^\varnothing=0$)& \texttt{................} \\
+				Reactivity profile $r^\alpha$                          & \texttt{0011100000000000} \\
+				Reactivity profile $r^\beta$                          & \texttt{0000000000011100}
+			\end{tabular}\\}
+
+		Now, consider a simplified ensemble of structures, consisting only of $S^\star$ and $S^\varnothing$, the open chain having free-energy $E^\varnothing = 0$. 
+		Given a structure $S$ and a reactivity profile $r$, the overall pseudo-energy bonus is then 
+		$$ \Delta G_r(S) = \sum_i m \log(r_i+1) + b.$$
+		The energy bonuses for the combinations of structure/(averaged)reactivity profiles are
+
+			$$
+			\begin{array}{c|cc}
+				\Delta G_r(S) & S:=S^\star & S:=S^\varnothing\\ \hline
+				r:=r^\alpha & 2m\log 2 + 4\, b & 3m\log 2 + 16\, b\\
+				r:=r^\beta & 2m\log 2 + 4\, b & 3m\log 2 + 16\, b\\
+				r:=\frac{r^\alpha+r^\beta}{2} & 4m\log 3/2 + 6\, b & 6m\log 3/2 + 16\, b			
+				\end{array}
+			$$
+
+		For instance, the emission probability of $S^\star$ given a reactivity profile $r$ is
+		\begin{align*}\mathbb{P}(S^\star\mid r) &= \frac{\displaystyle e^{-\frac{\Delta G_{r}(S^\star) + E^{\star}}{RT}}}{\displaystyle e^{-\frac{\Delta G_{r}(S^\star) + E^{\star}}{RT}} + e^{-\frac{\Delta G_{r}(S^\varnothing)}{RT}} }\\
+		&= \frac{1}{1 + e^{\frac{\Psi(R) + E^{\star}}{RT}}}
+		\end{align*}
+		with  $\Psi(r) := \Delta G_{r}(S^\star) -\Delta G_{r}(S^\varnothing)$.
+
+		It follows that the expected frequency of occurrences of $S^\star$ within an overall multiset of $M$ structures sampled by \OurTool{} ($M/2$ structures for each condition) is given by 
+		\begin{align*}
+		\mathbb{E}\left(\frac{\#\text{Occ}(S^\star)}{M}\mid r^\alpha \wedge r^\beta\right) &= \frac{\mathbb{P}(S^\star\mid r^\alpha) + \mathbb{P}(S^\star\mid r^\beta)}{2}\\
+		 &= \frac{1}{1 + e^{\frac{-m \log(2) -12\,b + E^{\star}}{RT}}} \\
+		 &= \frac{1}{1 + \frac{e^{\frac{-12\,b + E^{\star}}{RT}}}{2^m}}
+		\end{align*}
+		while the same quantity, measured within $M$ structures sampled from an averaged reactivity profile $\frac{r^\alpha+r^\beta}{2}$, becomes
+		\begin{align*}
+			\mathbb{E}\left(\frac{\#\text{Occ}(S^\star)}{M}\mid \frac{r^\alpha+r^\beta}{2}\right) &= \frac{1}{1 + e^{\frac{\Psi\left(\frac{r^\alpha+r^\beta}{2}\right) + E^{\star}}{RT}}}\\
+			  &=\frac{1}{1 + \frac{e^{\frac{-12\,b + E^{\star}}{RT}}}{(3/2)^{2m}}}.
+		\end{align*}
+		The two quantities are strictly different for any value of $m$ (except $m=0$), and the ratio of probability can go in different directions depending on the reactivities profiles and (exponentially large) ensembles of conformations.
+
+		More empirically, we believe such a strategy would impact the robustness and versatility of our whole approach. 
+		Indeed, in the presence of two sources of reactivities showing strong disagreement, indicating either the probing of two conformations, or an erroneous condition, averaged reactivity terms will end up favoring unpaired conformations, while our approach will probably produce two clusters (leaving aside the final Pareto optimization).
+	}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{Free-energies are re-computed using the \Software{RNAeval} tool, we have completed the documentation to make this point more explicit (and refer to a refined definition of these terms in the manuscript).}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{The reported Boltzmann probabilities, also called Stability in the manuscript, are strictly computed as described in the manuscript.
+We have made this point explicit in our revised documentation.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{We agree that this would be a potentially fruitful development.
+This would require a modification of the dynamic programming equations presented by \cite{Lu2009}, for instance by leveraging the Waterman grammar into something similar to:
+	\begin{align*}
+	  S &\to \texttt{(}\texttt{(} \,T\, \texttt{)}\texttt{)}\, S \mid \texttt{.}\, S \mid \varepsilon \\
+	  T &\to \texttt{(}\, T\, \texttt{)} \mid \texttt{(}\texttt{(} \,T\, \texttt{)}\texttt{)}\, U\mid   \texttt{.}\, S \\
+	  U &\to \texttt{(}\texttt{(} \,T\, \texttt{)}\texttt{)}\, S \mid \texttt{.}\, S 
+	\end{align*}
+	Such a modification would triple the memory requirement, but would remain polynomial in time and should not constitute the bottleneck of the whole method.
+However, it would also warrant an extensive validation effort, and lead to the description of further material within an, already consequent, Methods section.
+For these reasons, we prefer to leave this promising extension to a future version of \OurTool.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{Again, we agree that integrating the package within \Software{conda} would be ideal.
+However, since the \Software{ViennaRNA} package is not currently available on \Software{bioconda} for all platforms (only Mac OS and Linux), this would still require a manual installation and would somewhat mitigate the benefits of such an integration.
+
+		In the new version of the documentation, we at least partially ease the installation process, and suggest using \Software{pip} (available in any \Software{Python} release) to install all the necessary dependencies through the command:
+
+		{\centering \Software{pip install cython scipy numpy sklearn}\\[.5em]}
+
+		The \Software{ViennaRNA} package, however,  still needs to be independently installed.
+We have contacted the maintainers of the \Software{ViennaRNA} package to (help) make the package more universaly accessible, and will certainly pursue the  integration of future versions of \OurTool{} within \Software{bioconda} (a port to \Software{Python3} is currently underway).
+	}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{This is a great suggestion for a future application of our method, and will rank high within the list of our priorities (along with the, arguably natural analysis of multi-states RNAs).
+However, we feel that such a study would warrant a validation effort that would go beyond the scope of the current paper.}
 
 	\Comment{Fix the following typos/unfortunate choices of words:\begin{itemize}
-	\item general: et al $\to$ et al.
-	\item 1-2-30 : please provide the abbreviation details for SHAPE
-	\item 1-1-21 : remove comma after "experiments"
-	\item 1-2-37 : "more dynamics"
-	\item 1-2-49 : "qualitative"
-	\item 1-2-53 : "of individual nucleotides"
-	\item 2-1-23 : "These include " beside structure optimization " the joint folding.."
-	\item 2-1-49 : signal $\to$ signals
-	\item 2-2-38 : criterion "are" $\to$ "is"
-	\item 2-2-39 : identified $\to$ represented
-	\item 3-1-5 and 8 : better replace "$R_i$" with "$d_i$", since you are using R for reference later on
-	\item 3-1-60 : maybe replace "\#Datasets" with "$|\mathcal{D}|$" ?!
-	\item 3-2-17 : missing "are" in "that both"
-	\item 3-2-32 : move "where $k$ is the number of clusters" to line 40, where it is needed.
-	\item 4-1-8 : Within the formula "$Dist(d,d')$" you have to replace "$x$" with "$i$"
-	\item 4-1-52 : "Non-canonical" $\to$ "Crossing" or "Pseudoknot-forming"
-	\item 5-1-40 : "metrics" $\to$ "metric"
-	\item 6-2-44 : "hajdin" $\to$ "Hajdin"
-	\item 8-1-31 : "Figure 1" $\to$ "6"
-	\item 10-1-56 : "Figure" $\to$ "Table"
+	\item general: et al $\to$ et al. \IAnswer{Done}
+	\item 1-2-30 : please provide the abbreviation details for SHAPE \IAnswer{Done}
+	\item 1-1-21 : remove comma after "experiments" \IAnswer{Done}
+	\item 1-2-37 : "more dynamics"\IAnswer{Done}
+	\item 1-2-49 : "qualitative"\IAnswer{Done}
+	\item 1-2-53 : "of individual nucleotides"\IAnswer{Done}
+	\item 2-1-23 : "These include " beside structure optimization " the joint folding.."\IAnswer{Done}
+	\item 2-1-49 : signal $\to$ signals\IAnswer{Done}
+	\item 2-2-38 : criterion "are" $\to$ "is"\IAnswer{Done}
+	\item 2-2-39 : identified $\to$ represented\IAnswer{Done}
+	\item 3-1-5 and 8 : better replace "$R_i$" with "$d_i$", since you are using R for reference later on \PAnswer{We replaced $R_i$ with $r_i$ to avoid clashing with the new notation for the Boltzmann constant $R$.
+We also denote the reference structure as $\Ref$ in the definitions of the evaluation metrics.}
+	\item 3-1-60 : maybe replace "\#Datasets" with "$|\mathcal{D}|$" ? \IAnswer{Done}
+	\item 3-2-17 : missing "are" in "that both" \IAnswer{Done}
+	\item 3-2-32 : move "where $k$ is the number of clusters" to line 40, where it is needed. \IAnswer{Done}
+	\item 4-1-8 : Within the formula "$Dist(d,d')$" you have to replace "$x$" with "$i$" \IAnswer{Done}
+	\item 4-1-52 : "Non-canonical" $\to$ "Crossing" or "Pseudoknot-forming" \PAnswer{We clarified this description, which originally mixed the notions of non-canonical BPs and pseudoknots in a confusing manner.
+The new sentence reads \Quote{Non-canonical base pairs were removed, and a non-pseudoknotted secondary structure was extracted as a maximum subset of non-crossing base pairs~\citep{Smit2008}.}}
+	\item 5-1-40 : "metrics" $\to$ "metric" \IAnswer{Done}
+	\item 6-2-44 : "hajdin" $\to$ "Hajdin" \IAnswer{Done}
+	\item 8-1-31 : "Figure 1" $\to$ "6" \IAnswer{Done}
+	\item 10-1-56 : "Figure" $\to$ "Table" \IAnswer{Done}
 		\end{itemize}}	
@@
 
+	\newpage
 	\subsection{Reviewer \#3}
@@
 I am very surprised that this works.
-The Diegan et al.
-work fit the parameters empirically for SHAPE data.
-Sean Eddy [Annu.
-Rev.
-Biophys.
-2014.
-43:433] more explicitly drew the connection between these parameters and the log odds ratio of a nucleotide being paired, following earlier work of Sukosd et al.
-[Nucleic Acids Research.
-2013.
-41: 2807].
-The Das lab [Biochemistry.
-2012.
-51: 7037] also explicitly fit the log odds for DMS data using gamma distributions.
-I would expect that each probing reagent and each probing condition would need different parameters.}
+The \cite{Deigan2009} work fit the parameters empirically for SHAPE data.
+\cite{Eddy2014} more explicitly drew the connection between these parameters and the log odds ratio of a nucleotide being paired, following earlier work of \cite{Suekoesd2013}.
+\cite{Cordero2012} also explicitly fit the log odds for DMS data using gamma distributions.
+I would expect that each probing reagent and each probing condition would need different parameters.\label{arg:homogenous}}
 
-		\Answer{We also expect that additional conversion formulae, and a specific parameter calibration, for each condition, would be beneficial in order to obtain better predictive performances.
-In fact, it is one of our project to produce more extensive data for a subset of the conditions explored in this study, and to attempt to systematically learn reactivities-to-pseudo-energy conversion models.
+		\Answer{We also expect that additional conversion formulae, and a specific parameter calibration, for each condition, would be beneficial to the downstream predictive performances.
+In fact, it is one of our ongoing (recently funded) projects  to produce more extensive data for a subset of the conditions explored in this study, and to attempt to systematically learn reactivities-to-pseudo-energy conversion models.
 
-		However, the partial understanding of what is precisely being measured through different types of probing, in conjunction with the relative scarcity of available data for certain reagents/technologies, would probably induce very heterogeneous quality for such pseudo-energies across conditions.
-In particular, training for those parameters would probably lead to overfitting in cases where only a couple of reactivity profiles are currently available.
+		However, the conjunction of:
+		\begin{itemize}
+			\item Distinguishable, yet large-variance and highly-overlapping, reactivity distributions being observed for unpaired and paired nucleotides;
+			\item High reproducibility for  reactivity profiles across repeats (and, albeit less frequently, across probes as shown in Figure A above);
+		\end{itemize}
+		supports the notion that the determinants of SHAPE reactivity extends far the current binary paired/unpaired model. 
+		This hypothesis is already explored to some extent in \cite{Suekoesd2013}, and confirmed by recent mechanistic studies~\citep{McGinnis2012,Sexton2017,Hurst2018,Mlynsky2018,Frezza2019,Busan2019}.
 
-		Rather than address this heterogeneity as a possible confounding factor in our analyses, we have preferred to adopt a homogeneous treatment for all sources of probing data.
-It seems to us that the benefits of considering comparable reactivity profiles (produced in an homogeneous manner, and similarly normalized) far outweigh the loss in predicting accuracy.
+		This leads us to a strong suspicion that future pseudo-energy models will have to consider more structural contexts than the current binary one.
+Those contexts will need to be captured by the dynamic programming schemes.
+The training of such extended models would probably induce a highly heterogeneous quality for pseudo-energies models trained across conditions, given  the relative scarcity of available data for certain reagents/technologies.
+In particular, training extended parameters would probably lead to overfitting in contexts where only a couple of reactivity profiles are currently available.
 
-		\TODO{Discuss the fact that the normalization used in various studies are deemed equivalent}
+		Moreover, going back to the suggested use of the gamma distribution in the case of DMS, it is worth pointing out that \cite{Cordero2012} state in their supplementary material that
+		\Quote{
+		The performance of the algorithm using this probabilistic potential for SHAPE, DMS, and CMCT reactivities is given in Table S6 and is identical to
+		results obtained for the standard potential with slope ( m ) and intercept b ) optimi zed through grid
+		search .}
+		We also note the relative stagnation of performances of state-of-the-art tools integrating SHAPE data, despite an increasing sophistication of the pseudo-free-energy models.
+So, while we believe progress can be made in the integration of reactivities, we do not expect them to radically change the quality of predicted models, nor to invalidate the findings of this manuscript.
+
+		To summarize, it seems too early in this manuscript to pursue a specialization of energy models to each reagent/probing technology, and the benefits of considering comparable reactivity profiles (produced in an homogeneous manner, and similarly normalized) far outweigh the expected loss in predicting accuracy.
+Consequently, rather than specialize our treatment of individual probes and address this heterogeneity as a possible confounding factor in our analyses (in a way that would remain to be determined), we prefer to preserve a homogeneous treatment for all sources of probing data.
 
@@
 
-		\Answer{\TODO{Clarify}}
+		\Answer{
+			The (hyper)parameters $m$ and $b$ were initially calibrated on the \cite{Cordero2012} dataset, revealing a very limited impact of the parameters on the prediction accuracy.
+We kept those values for multi probing, as the scarcity of available data did not allow us to afford an extensive cross-validation.
+We updated the description as shown below to be more explicit.
+			\Quote{
+			Those values were halved in comparison to those recommended by Deigan\etal\citep{Deigan2009}, following both a grid search optimization on the Cordero dataset, and based on the rationale that lower absolute values for pseudo-energy bonuses increase the expected overlap between pseudo-Boltzmann ensembles.}
+		}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{
+		We agree that stability of the predictions is potentially problematic, especially in the context of reproducing earlier results.
+
+		Larger sample sizes represent a natural option for stabilizing the predictions, and indeed seem to decrease the variability.
+However, a substantial increase of the number $x$ of samples induces extreme computational demands, due to the current necessity to compute the dissimilarity matrix in $\Theta(x^2.n)$ time and $\Theta(x^2)$ memory.
+In this context, simply multiplying by 10 the number of samples would induce a 100 fold increase of the time requirement.
+In the future, we will revisit the clustering component to learn suitable embeddings for secondary structures, granting access to linear-time EM algorithms (eg based on Gaussian Mixture Models).
+
+		\Quote{
+	{\centering \includegraphics[width=.7\linewidth]{graphs/boxplotreproducibility}\\}
+}					
+		Moreover, as noted by this reviewer, an expert user would ensure a greater stability by running \OurTool several times (say 10), and electing the most recurrent structure.
+This process would achieve asymptotic determinism, since the frequency of a structure would converge towards real emission probabilities.
+Moreover, most frequent structures seem to be the best quality ones, and are emitted by \OurTool with fairly large probability, ensuring convergence in a reasonable number of iterations.
+		However, such a process would be cumbersome for a basic user.
+
+		In the future, for the sake of reproducibility, we also wish to report, and letting the user specify, a \Def{random seed}, which will  allow the exact reproduction of an analysis.
+However, such a feature is currently not supported by \Software{RNAsubopt}, so we are working towards its inclusion to the \Software{ViennaRNA} package.
+
+	}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{This is entirely correct.
+We clarified this description, which originally mixed the notions of non-canonical BPs and pseudoknots in a confusing manner.
+The sentence now reads: \Quote{Non-canonical base pairs were removed, and a non-pseudoknotted secondary structure was extracted as a maximum subset of non-crossing base pairs~\citep{Smit2008}.}}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{While the statement was meant to be informal, we agree (and it is quite established in the field) that the sensitivity is not a relevant metric for assessing the predictive capacities of structure prediction methods, and that the MCC correlates very well with the product of the Sensitivity and PPV.
+We have decided to preserve the informal tone, and now state: \Quote{It can be interpreted as a compromise between the main metrics derived from the confusion matrix, and is defined as [..] }}
 
@@
 As noted above, the pseudo free energy parameters for SHAPE (NMIA) will not be optimal for DMS and CMCT.
-The benchmarks should therefore also be run using the pseudo free energies fit by the Das lab for DMS and CMCT [Biochemistry.
-2012.
-51: 7037].
+The benchmarks should therefore also be run using the pseudo free energies fit by the Das lab for DMS and CMCT~\citep{Cordero2012}.
 My hypothesis is that the structure prediction accuracy will be higher with the properly fit parameters.}
 
-	\Answer{\TODO{...}}
+	\Answer{
+		Unfortunately, due to our current reliance on \Software{RNAfold}, which takes reactivities as input rather than direct pseudo-energy terms, we could not test this hypothesis.
+However, we do not expect such a fitting procedure to produce substantially better results.
+Indeed, as stated in the supplementary material of the \cite{Cordero2012} manuscript:
+		\Quote{The performance of the algorithm using this probabilistic potential for SHAPE, DMS, and CMCT reactivities is given in Table S6 and is identical to results obtained for the standard potential with slope (m) and intercept (b) optimized through grid-search.}
+		While this leaves open the possibility of improvements upon re-optimization of the $m$ and $b$ parameters, this statement indicates that the nature of the conversion itself does not play a large role in the final performances.
+	}
 
@@
 Rsample attempts to adaptively define the pseudo free energies to not perturb probabilities for nucleotides in an initial stochastic sample that are consistent with the SHAPE mapping data.
-Perhaps this manuscript should also mention that detail and state this adaptive approach is possibly not as accurate (when producing a single structure) as the Diegan approach.}
+Perhaps this manuscript should also mention that detail and state this adaptive approach is possibly not as accurate (when producing a single structure) as the \cite{Deigan2009} approach.}
 
-	\Answer{\TODO{...}}
+	\Answer{
+	This refined free-energies could indeed be impacting the predictive capabilities of \Software{RSample} (although their underlying rationale is convincing), we came to realize upon further inspection that \Software{RSample} actually never uses sampling/clustering on the \cite{Hajdin2013} dataset.
+Rather, the method concludes from the corrective factors that a single conformation is most likely, and simply returns the MEA of the whole pseudo-ensemble.
+We have updated our discussion to include this alternative explanation, which seems to explain the strong discrepancy, and supports the notion that clustering can be used to \emph{clean up} a noisy Boltzmann ensemble.
+	\Quote{
+In this context, the good performance of \OurTool in comparison with \Software{RSample}~\citep{Spasic2017} may appear surprising.
+Indeed, while independently developed and used~\citep{Deforges2017}, both methods share a common sampling/clustering scheme.
+However, \Software{RSample}~\citep{Spasic2017} reverts to a MEA~\citep{Lu2009} computation whenever a single conformation is suspected.
+The MEA structure coincides with the centroid of the whole Boltzmann ensemble, suggesting that the increased performances of \OurTool could be at least partially attributed to the use of sampling, and to our adaptive choice of the number of clusters.
+Moreover, even in cases where a single dominant conformation is expected, considering multiple clusters seems to create more opportunities for discarding noisy random unstable conformations.
+Such conformations naturally occur in a Boltzmann-distributed subset of structures, and may end up \emph{polluting} the centroid structures by artificially inflating the cluster diversity.
+}}
 
@@
 (I thought I had missed something as I read through).
-		\item In the Funding section, “phD” should be “PhD”.
+\IAnswer{Done}
+		\item In the Funding section, “phD” should be “PhD”. \IAnswer{Done}
 		\item There are many problems with the references, however.
@@
 5, 6, 7,…).
-	\end{itemize}}
-
-	\Answer{\TODO{...}}
-
-	\Comment{It would be helpful if the caption to figure 2 named the benchmark dataset.
-This is clear in the main text, but will make it easier to quickly digest the manuscript if the captions are all self-contained.}
-
-	\Answer{\TODO{...}}
+\PAnswer{We have extensively sanitized our references, and we believe our bibliography now adheres to accepted standards.}
+		\item It would be helpful if the caption to figure 2 named the benchmark dataset.
+This is clear in the main text, but will make it easier to quickly digest the manuscript if the captions are all self-contained.\PAnswer{We have systematically revised the captions to make the as self-contained as possible, and in particular always name the dataset on which the data was produced.}
+	\end{itemize}
+}
 
 	\end{enumerate}
+\newpage
 
@@
 It is shown to outperform Rsample, where the authors attribute this to the sample denoising IPANEMAP performs via clustering (page 7).
-I am not sure if this is indeed the case and also think this benchmark could be improved in several ways.
+I am not sure if this is indeed the case and also think this benchmark could be improved in several ways.}
 
-	First, Rsample’s sampling routine is quite different than IPANEMAP’s and could be a relevant factor.
+	\Answer{First of all, we would like to point out that the performances of \OurTool{} in a mono-probing setting were explored as a validation of the basic predictive capacities of \OurTool{} in a context where reference performances are available.
+The very good behavior of \OurTool{} on this benchmark came as a pleasant surprise which we deemed worthy of a mention in the paper.
+
+	In particular, this benchmark should not be mistaken for a central result in our paper, and we do not claim that \OurTool{} supersedes \Software{RSample}.
+Indeed, the benchmark on single-state RNAs is not the most salient point of the \cite{Spasic2017} paper, since the main application of \Software{RSample} is to identify multiple conformations, a use-case that we did not specifically tried to address while designing \OurTool{}.}
+
+	\Comment{First, Rsample’s sampling routine is quite different than IPANEMAP’s and could be a relevant factor.
 It entails sampling from SHAPE reactivity distributions followed by partition function optimization.
@@
 
-	\Answer{\TODO{...}}
+	\Answer{In the case of single-state RNAs, highlighting the shared sampling/clustering paradigm with \Software{RSample} turned out to be misguided on our part, since \Software{RSample} simply computes the MEA of the whole ensemble rather than to rely on a sampling/clustering combination.
+We have modified our discussion on the comparison with \Software{RSample} to reflect the fact that clustering is simply not used by \Software{RSample} in this setting.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{We agree that our description may have been imprecise in that respect, but we did use the recommended numbers of samples in our execution of \Software{RSample}.
+More precisely, as described in \cite{Spasic2017}, two different rounds of sampling were used in the method: a) to estimate the corrective terms; b) to identify alternative conformations, if several conformation are suspected.
+The recommendations of \cite{Spasic2017} are to use 10,000 samples for (a), and 1,000 samples for (b), which was done for the reported predictions.
+In the submitted manuscript we only referred to (b) for the number of samples, which was not very useful to the reader since step b) is never performed by \Software{RSample} on the benchmark.}
 
 	\Comment{Third, the authors mention that Rsample was “shown to perform favorably against a comprehensive collection of state-of-the-art methods…” This is incorrect.
-In fact, based on a t-test, Spastic et al.
-concluded that “all methods that use experimental restraints perform similarly.” A similar conclusion, also based on a statistical test, was reported by Deng et al.
-(2016).
-This matters because it points out a major weakness of this work, namely, it does not consider the statistical significance of its findings despite reporting substantial variability between predictions.
-Additionally, it highlights the possibility that a comparison to another method may result in different findings since there are a few sequences with dramatic performance gaps, which were not considered by Spasic et al.}
+In fact, based on a t-test, \cite{Spasic2017} concluded that “all methods that use experimental restraints perform similarly.” A similar conclusion, also based on a statistical test, was reported by \cite{Deng2016}.}
+
+	\Answer{We thank the reviewer for pointing out this remark in the \cite{Spasic2017} paper which we overlooked in our initial reading.
+We corrected the corresponding sentence in our revised manuscript to
+		\Quote{It was shown to perform \Rev{comparably as} a comprehensive collection of state-of-the-art methods in probing-guided structure prediction, including \Software{RME}~\citep{Wu2015}, \Software{RNAprob}~\citep{Deng2016}, \Software{RNAprobing}~\citep{Washietl2012}, \Software{RNAsc}~\citep{Zarringhalam2012} and the \Software{fold} utility from the \Software{RNAStructure} suite~\citep{Reuter2010}.}
+		 This does not however challenge the conclusions drawn from this section, as \Software{RSample} remains as good a representative of the current state-of-the-art for mono-probing structure prediction.}
+
+	\Comment{This matters because it points out a major weakness of this work, namely, it does not consider the statistical significance of its findings despite reporting substantial variability between predictions.
+Additionally, it highlights the possibility that a comparison to another method may result in different findings since there are a few sequences with dramatic performance gaps, which were not considered by \cite{Spasic2017}}
+
+	\Answer{The significance (or lack thereof) of claims in the context of RNA benchmarks is, of course, a legitimate concern, and we thank the reviewer for pointing it out.
+The way it is used in \cite{Spasic2017} should probably be taken with a grain of salt due to the dubious obedience of produced data to classic assumptions (\emph{eg} normality of distribution), but it nevertheless represents the best option available at the time.
+
+	In the context of comparing the performances of \OurTool{} to \Software{RSample}, we performed a two-tailed Student $t$-test on paired data with unequal variance.
+The test rejects the $H_0$ hypothesis \emph{Both methods have similar average performance}, with P-value of $6\times 10^{-3}$, much lower than the usual significance threshold $\alpha=5 \times 10^{-2}$  (computed statistics $t=2.98 > t_{\text{crit}}=2.07$).
+	}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{As observed above, on the \cite{Hajdin2013} dataset, \Software{RSample} always ends up running an MEA algorithm, an approach which is deterministic for a given set of base pair probabilities.
+Moreover, the base pair probabilities are based on correcting terms that are themselves estimated from 10~000 sampled structures (a number chosen to ensure stable estimates).
+Therefore, in theory, \Software{RSample} is virtually deterministic in its predictions, and quite unsurprisingly we could not detect a single variation of its predictions within 10 executions over the \cite{Hajdin2013} dataset.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{The \cite{Hajdin2013} was not used in the initial training.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{
+		The distance-based clustering of structures, arguably the most crucial component of \OurTool, is fully unsupervised.
+We do not believe that a loose calibration of two numerical parameters should raise a reasonable concerns of overfitting.
+	}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{
+	Despite our previously-stated objections, we understand that such metrics are expected for new methods.
+We assessed the statistical significance of our main assertions using the two-tailed Student $t$-test on paired data with unequal variance, as recommended by \cite{Xu2012}:
+	\begin{itemize}
+		\item \cite{Hajdin2013} dataset.\\
+		Hypothesis $H_0$: \OurTool and \Software{RSample} have similar expected GM\\
+		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=6\times 10^{-3}$ $\to$ Rejected at $\alpha=5\times10^{-2}$ 
+		\item \cite{Cordero2012} dataset.\\
+		  Due to the limited size of the dataset (6 RNAs), none of our observations regarding the average behavior are sufficient to reach statistical significance.
+We have toned down the title of the corresponding section, and added a remark to the conclusion of the associated results section.
+		\item \didy{}  dataset.\\
+		Hypothesis $H_0$: \OurTool and MFE have similar expected MCC\\
+		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=2.2\times 10^{-2}$ $\to$ Borderline rejected at $\alpha=5\times10^{-2}$\\
+		Hypothesis $H_0$: \OurTool and MEA have similar expected MCC\\
+		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=14.4\times 10^{-2}$ $\to$ Not significant at $\alpha=5\times10^{-2}$ \\
+		Hypothesis $H_0$: Joint analysis of triplet have same MCC as average of three member conditions\\
+		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=1.4\times 10^{-9}$ $\to$ Rejected at $\alpha=5\times10^{-2}$ \\
+		Hypothesis $H_0$: Joint analysis of triplets and pairs of conditions have same MCC\\
+		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=2.1\times 10^{-2}$ $\to$ Borderline rejected at $\alpha=5\times10^{-2}$ \\
+		\item \cite{Cheng2017} dataset.\\
+		Hypothesis $H_0$: WT and WT+1M have similar expected MCC\\
+		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=9.7\times10^{-4}$ $\to$ Rejected at $\alpha=5\times10^{-2}$\\ 
+		Hypothesis $H_0$: WT+1M and WT+2MC have similar expected MCC\\
+		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=2.4\times10^{-2}$ $\to$ Rejected at $\alpha=5\times10^{-2}$
+	\end{itemize}
+	}
 
@@
 This is impressive, with two caveats.
-First, only a single RNA of medium length was probed.
-Second, it is my understanding  that no replicate data was generated or analyzed, with the exception of SHAPE-CE.
+First, only a single RNA of medium length was probed.}
+
+	\Answer{
+		It must be noted that we do not base our main conclusion (benefit from the joint analysis of multiple reactivity) solely on the \didy{} case study, but also on the Cordero dataset, where the best performances are also observed upon considering a combination of reactivity profiles associated with DMS, CMCT and NMIA probes.
+A similar trend is also observed on the \cite{Cheng2017} dataset.
+	}
+
+	\Comment{Second, it is my understanding  that no replicate data was generated or analyzed, with the exception of SHAPE-CE.
 Typically, methods are tested on at least a few RNAs of varying lengths and whenever possible on datasets like Hajdin et al.’s.
@@
 For this reason, I would be hesitant to draw any conclusions from probing data for a single RNA.
+
 	That said, I do not expect the authors to generate more data.
@@
 
-	\Answer{\TODO{...}}
+	\Answer{We already clarified that this is merely a case study.
+We wish we could do more (e.g.
+test on a wider dataset) but this is simply unrealistic, and way beyond the standards enforced for the dissemination of a new {\bf method} in structural bioinformatics.
+
+		 However, it must be noted that we do not base our main conclusion (complementarity of probing data, benefit of their integration) solely on the \didy{} case study, but also on the Cordero dataset, where the best performances are also observed upon considering a combination of reactivity profiles associated with DMS, CMCT and NMIA probes.
+	}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{
+		\TODOBruno{Did we produce replicates?}
+	}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{We agree, and in fact do not draw any conclusion from these results.
+We only report those values as references, to support the hypothesis that the MCCs achieved by \OurTool are probably more representative of the reactivity profiles/condition than of a specific predictive model.}
 
-	\Comment{Additionally, there is strong A/C bias in DMS data obtained by stop-based protocols.
-I don’t know if it was accounted for during normalization step, if normalization was indeed applied.
-I also don’t know which datasets were used to optimize IPANEMAP’s parameters, but as I commented above, a fair comparison should train all methods on the same data.}
+	\Comment{Additionally, there is strong A/C bias in DMS data obtained by stop-based protocols. I don’t know if it was accounted for during normalization step, if normalization was indeed applied. }
 
-	\Answer{\TODO{...}}
+	\Answer{
+		We were unsure how to interpret this remark: \Ab{} and \Cb{} are indeed the only reactive nucleotides, so the normalization step is indeed restricted to positions featuring those two nucleotides.
+If the reviewer is referring to the observations of \cite{Sexton2017} of opposite {\em \Ab{} vs \Cb} biases for RT-based and mutations-based probing respectively, we did not account for it.
+Indeed, while including corrective terms would seem to represent an opportunity for improved predictions, we would rather treat of reactivity data homogeneously than achieve the best MCCs possible.
+More details in our response to Comment \ref{arg:homogenous} of Reviewer \#3.
+	}
+
+	\Comment{I also don’t know which datasets were used to optimize IPANEMAP’s parameters, but as I commented above, a fair comparison should train all methods on the same data.}
+
+	\Answer{Again, we object to such an adversarial view of this particular analysis (and, more generally, our work).
+\OurTool offers new capabilities (joint integration of multiple probing data), which were not supported by previous tools.
+Our extensive comparison to earlier efforts on mono-probing data is mainly meant to validate our specific implementation of the \cite{Ding2003} sampling/clustering in the context of single probing data.
+	}
 
@@
 
-	\Answer{\TODO{...}}
-
-	\Comment{My comments above re: IPANEMAP, MEA, and MFE also apply to their comparison for the Cordero et al. dataset (page 5).}
-
-	\Answer{\TODO{...}}
+	\Answer{
+		Again, we include those MCCs for reference only, and leave their interpretation to the reader. We do not feel the need to include a huge collection of tools, as this is not a benchmark.
+	}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{In this revised version, we tried to enforce a consistent use of the past tense in the results section.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{Although we already described some of those limitations in the method, we now include such a discussion more explicitly in our conclusion/discussion.
+	\Quote{Current limitations of \OurTool include the lack of an explicit support for pseudoknots, which will be addressed in a future version, using  polynomial-time dynamic programming~\citep{Janssen2015}, or a iterative heuristic alternative~\citep{Hajdin2013}.
+The inherently stochastic nature of the sampling scheme may also be challenging for certain types of analysis.
+However, despite its stochastic foundations, \OurTool{} is typically stable in its predictions, as can be seen in Figure~\ref{fig:reproducibility} for all probing datasets produced over our \didy case study.
+			Interestingly, the most notable exception to the general stability is observed for the probing-free prediction, consistent with the widespread interpretation of SHAPE-induced pseudo-potentials as focusing predictions onto a subset of the Boltzmann ensemble.
+A similar behavior is observed for the other datasets considered in our study (see Supp.
+Tables S2 and S5).
+%In the interest of reproducibility, ipanemap reports, and accepts as an option, the random seed used both within the main software and its non-deterministic dependencies, leading to identical repeats of the experiments.
+			Finally, the current clustering method induce a growth of time and memory requirements that scales quadratically with the number of samples.
+While this allows a joint analysis of a dozen of reactivity profiles in less than an hour on a personal computer, we expect the ability to push the number of samples will lead to better, even more reproducible, predictions.
+To that purpose, we will explore embedding techniques and linear-time community detection algorithms as alternatives to MBkM~\citep{Sculley2010}.
+	}	
+}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{We respectfully disagree with this proposal of an alternative writing style, and prefer to include sufficient factual statements to let the readers form an opinion regarding our reasoning and conclusion (at the risk of being tediously descriptive at times).
+}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{Again, we respectfully disagree with this vision.
+What is important about cluster B is not only the fact that it can be identified from our new metrics (more conclusively than than from the raw reactivities), but also that it can be used to assess the impact of contamination by an erroneous/outlier condition.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{\TODOAfaf{Can you confirm that a similar normalization was used?}}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{We came to this conclusion from a visual inspection of the distance matrix in Figure 5, reproduced below. 
+		\Quote{
+	\includegraphics[width=.45\textwidth]{graphs/didy/PCA} \hfill
+	\includegraphics[width=.45\textwidth]{graphs/didy/bi_clustering}\\
+	}
+
+		This number may not seem to be supported by a inspection Figure 5 (PCA, left above), but it must be reminded that a PCA is simply a projection.
+Points that are distant in the PCA are typically distant in the original higher-dimensional space, but close points in the PCA may also be quite distant due to the missing depth perception.
+
+		On Figure 5, we already see a clear separation between B, D, C, A, and H.
+Moreover, merging clusters E, F and G would lead to the aggregation of substantially different conditions, as can be seen in Figure 6 (right above), so we considered 8 conditions.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{We agree that SHAPE CE should not be mischaracterized as a representative for all SHAPE protocols based on RT stops (notoriously including SHAPE-Seq).
+Our description was misleading in that respect, so we modified and removed the corresponding column in Table 2.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{
+		We were happy with the overall reproducibility of our results with 1~000 structures, which has been the standard practice in the field since the original \cite{Ding2003} paper.
+Moreover, our computational requirements scale quadratically with the number of structure, so we recommend to stay with this number, which represents in our opinion a good tradeoff between reproducibility and efficiency.
+
+		To address this question more precisely, we reexecuted the \didy benchmark using a sample size of 2~000, and observed a minor non-significant variation in the MCC (ie the results did not allow to refute the hypothesis of equal expected MCCs for both experiments).}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{We disagree and, while anecdotal, we believe this figure illustrates nicely a case where the addition of data leads to an improvement of the model.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{For this revision, we added a principled comparison of the raw reactivity profiles based on correlation of reactivity.
+Its results are now included in the supplementary material and reproduced below
+	\Quote{\centering\includegraphics[width=.64\textwidth]{graphs/didy/reactivity_correlation.pdf}}
+	As can be seen in this matrix, "MaP" conditions essentially lump in two classes at the reactivity level (that do not fully align with the "MaP" clusters A/C/D).
+However, "CE" conditions do not form a clear cluster, and in many case correlate better with a "MaP" conditions than with "CE" conditions in the same cluster.
+}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{Indeed, the difference can be traced back, for the conditions in C, to a reactivity in the 5' terminal region that is greater for  \OneMSevILUThreeMg than for \OneMSevILUMg, resulting in a loss of the basal helix in the model predicted by \OneMSevILUThreeMg.
+This variation, coupled with the shift of a small helix, is alone responsible for the change of performances (the MCC does not take prisoners...).
+}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{As stated in the manuscript, bi-probing does not improve the standard deviation of the MCC (8\% for mono-probing, 8.4\% for bi-probing).
+Considering the sets of similar conditions in Table 3 increases the stdev to around 10\%.
+However, this observation merely stems from a statistical artifact, as the large number of conditions in Cluster B leads to a multiplication of MCC around 60\%, resulting in what would appear to be a bimodal distribution ($\to$ large variance/stdev) without reflecting any underlying reality.}
 
@@
 For example, replicates are inherently redundant and ideally not too diverse, yet they are generated to decrease std around the mean although the mean is always lower than some of the point estimates.
-A related question is as follows.}
+A related question is as follows.
 
-	\Answer{\TODO{...}}
+	For bi-probing, if we repeat predictions with 10 independent Boltzmann samples, which strategy is more precise: averaging predictions between the two probes or the corresponding bi-probing prediction?}
 
-	\Comment{For bi-probing, if we repeat predictions with 10 independent Boltzmann samples, which strategy is more precise: averaging predictions between the two probes or the corresponding bi-probing prediction?}
+	\Answer{
+		The meaning of \emph{averaging the predictions} is unclear to us in this context (our method produces 2D model(s), how does one average models?), and could be interpreted in several ways that we try to cover in the following sentences.
 
-	\Answer{\TODO{...}}
+		If the intent is to compare the MCC of a bi-probing prediction with the average of the two mono predictions, this is precisely the content of Figure 7.B, and our analysis does not reveal a substantial reduction of the stdev.
+
+		The intent may also be to compare, for a given pair of conditions, the stdev of the MCC to the stdev of the mean MCC over two independent mono executions (by running several experiments).
+In this setting, we either analyze a single pair, and obtain results that are probably not significant, or we need to report summary statistics over stdevs of a dataset of pairs of conditions, something that would not only be heavily time-consuming, but also hard to interpret.
+We will welcome clarification over this point over the next round of reviews.
+
+		However, if the objective metric of quality is the reproducibility of predictions using \OurTool (\emph{i.e.} to decrease the stdev as low as possible), a very simple solution is to perform 10+ independent runs of \OurTool, and elect the dominant conformation.
+Indeed, in most cases, \OurTool produces a well-defined structure with high probability, as suggested by Figure 8 (WT) and 9.
+By running \OurTool repeatedly, the frequency of the dominant structure becomes an estimator, with variance decreasing on the number of iterations, for the emission probability of the structure.
+This process ultimately converges towards a deterministic election of the most probable structure, and should already be fairly close to determinism after 10 iterations.
+Moreover, the time requirement scales linearly with the number of iterations, as opposed to quadratically if one were to multiply the number of samples.
+
+		In the future, we also wish to report, and letting the user specify, a \Def{random seed}, which will  allow the exact reproduction of an analysis. 
+		However, such a feature is currently not supported by \Software{RNAsubopt}, so we must first work towards its inclusion to the \Software{ViennaRNA} package.
+		}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{Upon further inspection of the prediction, it appears that the reviewer's intuition is correct and the improvement in comparison to the best of the group is marginal.
+We have removed the Table, and any mention to those specific triplets in the text.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{
+		We would like to sincerely thank this reviewer for noticing and voicing this apparent inconsistency, which led us to realize a severe mistake in Figure 9.
+For reasons to be determined, the labeling of the y-axis (MCC) ended up being dilated, showing MCCs in the 70\%/82\% range.
+The revised version of Figure 9, reproduced below, now features corrected axis and is now consistent with the averages MCC reported in Table 2.
+		\Quote{
+				{\centering
+					\begin{minipage}{.58\linewidth}
+					\includegraphics[width=\linewidth]{graphs/boxplotreproducibility}
+					\end{minipage}\hfill
+					\begin{minipage}{.4\linewidth}
+	\Heatset{min=60,   
+		max=83,   
+		max colour=Aquamarine3, % colour at maximum
+		min colour=BadCol,      % colour at minimum
+		Min colour=BadCol, % colour for values below min
+		Max colour=Aquamarine3   % colour for values above max
+	}
+					\newcommand{\M}[1]{\multicolumn{1}{l}{#1}}
+						\begin{adjustbox}{max width=\linewidth}
+							\begin{tabular}{@{}lccHHH}
+								\toprule
+								&          &           & \multicolumn{3}{@{}c@{}}{MCC\%}              \\
+								Condition                   &  Clust.  & Mg$^{2+}$ & \M{\relsize{-1}\OurTool} & \M{MFE} & \M{MEA} \\ \midrule
+								\OneMSevILUMg               & \Bull{C} &    \Mg    & 85                       & 82      & 82      \\
+								\OneMSevILU                 & \Bull{A} &   \NoMg   & 84                       & 82      & 84      \\
+								\NMIAMg                     & \Bull{A} &    \Mg    & 81                       & 80      & 80      \\
+								\NMIA                       & \Bull{A} &   \NoMg   & 80                       & 69      & 69      \\
+								\OneMSevILUThreeMg          & \Bull{C} &    \Mg    & 74                       & 75      & 75      \\
+								\NMIAMgCE                   & \Bull{E} &    \Mg    & 73                       & 72      & 73      \\
+								\NAIMg                      & \Bull{G} &    \Mg    & 73                       & 68      & 70      \\
+								\BzCNMg                     & \Bull{H} &    \Mg    & 71                       & 68      & 54      \\
+								\OneMSevMgCE                & \Bull{E} &    \Mg    & 70                       & 69      & 70      \\
+								\CMCTMg                     & \Bull{G} &    \Mg    & 70                       & 68      & 73      \\
+								\OneMSevILUThree            & \Bull{D} &   \NoMg   & 64                       & 56      & 61      \\
+								\DMSMg                      & \Bull{F} &    \Mg    & 63                       & 66      & 67      \\
+								\NMIACE                     & \Bull{B} &   \NoMg   & 61                       & 60      & 60      \\
+								\OneMSevCE                  & \Bull{B} &   \NoMg   & 60                       & 61      & 61      \\
+								\BzCN                       & \Bull{B} &   \NoMg   & 60                       & 59      & 59      \\
+								\NAICE                      & \Bull{B} &   \NoMg   & 60                       & 59      & 59      \\ \midrule
+								Avg Technology              &          &  $-$  & 78                       & 74      & 75      \\
+								&          &  $-$  & 66                       & 65      & 65      \\
+								Avg -/+ Mg$^{2+}$           &          &   \NoMg   & 67                       & 64      & 65      \\
+								&          &    \Mg    & 73                       & 72      & 71.5    \\ \midrule
+								{\bfseries Average overall} &          &  $-$  & 70.5                     & 68      & 68.5    \\ \bottomrule
+								&          &           &                          &
+							\end{tabular}
+					\end{adjustbox}
+					\end{minipage}
+					\\}
+		}
+
+		Average MCC values over 10 runs are indeed reported in Table 2, and an independent dataset was generated to create Figure 9, leading to minor discrepancies.
+For multiple probes, only a single result was generated.
+	}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{Fixed.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{Two centroids distant by at most one base-pair (included) will be deemed highly similar, so a (very) minor variation will be tolerated.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{$E_{d}(S)$ denotes the entire score, including NNTM free energy plus the energy bonus.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{We already addressed this comment above, and respectfully disagree with this assessment.}
 
@@
 
-	\Answer{\TODO{...}}
+	\Answer{In a mono-probing setting, due to properties of the two criteria used for the multi-objective Pareto optimization, it is rigorously impossible for \OurTool{} to return more than one cluster.
+Indeed, the cluster $C^\star$ having max stability $s^\star$ is such that: either $C$ has support $t^\star=1$, in which case it dominates any other cluster with signature $(s<s^\star,t\in\{0,1\})$; or $C$ has support $t^\star=0$, ruling out the possibility of $t=1$ for any other cluster (since it would imply $s>s^\star$).
+Note that \OurTool also reports dominated clusters for furter inspection.
+	}
 
@@
 
-%\bibliographystyle{amsplain}
-%\bibliography{response}
+\bibliographystyle{abbrvnat}
+\bibliography{biblio}
 \end{document}

### 2019-12-22 IPANEMAP-2019.tex

@@
 \begin{abstract}
-The manual production of reliable RNA structure models from chemical probing experiments, benefits from the integration of information derived from multiple protocols and reagents.
+The manual production of reliable RNA structure models from chemical probing experiments benefits from the integration of information derived from multiple protocols and reagents.
 However, the interpretation of multiple probing profiles remains a complex task, hindering the quality and reproducibility of modeling efforts.
@@
 
-SHAPE reagents represent a popular class of small molecules that react with the 2’ hydroxyl of flexible ribose~\cite{McGinnis2012}, although the exact properties observed by SHAPE remain the object of ongoing investigations~\cite{McGinnis2012,Hurst2018,Mlynsky2018,Frezza2019,Busan2019}.
+Reagents used within \Rev{selective 2 -hydroxyl acylation analyzed by primer extension (SHAPE)~\cite{Wilkinson2006} protocols} represent a popular class of small molecules.
+They react with the 2’-hydroxyl of flexible ribose~\cite{McGinnis2012}, although the exact properties observed by SHAPE remain the object of ongoing investigations~\cite{McGinnis2012,Sexton2017,Hurst2018,Mlynsky2018,Frezza2019,Busan2019}.
 As ribose flexibility is proportional to the degree of freedom of the nucleotide, it is assumed that SHAPE reagents discriminate nucleotides in stable interactions from others.
-Different SHAPE reagents are endowed with different dynamics, that can help differentiate Watson-Crick base pairs from more dynamics tertiary contacts~\cite{Gherghe2008,Steen2012,Rice2014,Busan2019}.
+Different SHAPE reagents are endowed with different dynamics, that can help differentiate Watson-Crick base pairs from more dynamic tertiary contacts~\cite{Gherghe2008,Steen2012,Rice2014,Busan2019}.
 Other chemical probes, harbouring different chemical reactivities, have been developed.
@@
 These two reagents therefore mostly reveal Watson-Crick base pairing, but also any other type of contact involving the same edge.
-This diversity of probes not only increases  coverage of the different positions and structural contexts in an RNA, but it also provides qualitatively different qualitative information, as been widely used since the very early days of RNA structure probing~\cite{Moazed1986, Ehresmann1987, Romaniuk1988, Butcher1994, Brunel2000,  Cordero2015, Somarowthu2015, Gross2017}.
+This diversity of probes not only increases coverage of the different positions and structural contexts in an RNA, but it also provides different qualitative information\Rev{, including in an \emph{in vivo} setting~\cite{Wang2019}}.
+It has been widely used since the very early days of RNA structure probing~\cite{Moazed1986, Ehresmann1987, Romaniuk1988, Butcher1994, Brunel2000,  Cordero2015, Somarowthu2015, Gross2017}.
 
@@
 
-Computationally, the reactivity of a nucleotide is typically used as a proxy for the unpairedness of individual nucleotides.
+Computationally, the reactivity of a nucleotide is typically used as a proxy \Rev{to assess the unpaired nature of individual nucleotides}.
 The last couple of decades have nevertheless seen a series of paradigm shifts in the ways probing information is integrated, somehow mirroring the evolution of {\em ab initio} methods for secondary structure prediction.
@@
 However, such \emph{hard} constraints turned out to be very sensitive to the choice of the cutoff, easily leading to artificially unpaired predictions or unsatisfiable sets of constraints.
-With the advent of SHAPE probing, subsequent methods lifted the requirement of thresholds and integrated reactivities directly as pseudo-energies, supplementing the Turner nearest-neighbor energy model~\cite{Turner2010} within an (pseudo-) energy minimization scheme~\cite{Deigan2009,Zarringhalam2012,Washietl2012}.
+With the advent of SHAPE probing, subsequent methods lifted the requirement of thresholds and integrated reactivities directly as pseudo-energies, supplementing the Turner nearest-neighbor energy model~\cite{Turner2010} within an (pseudo) energy minimization scheme~\cite{Deigan2009,Zarringhalam2012,Washietl2012}.
 Such methods can be seen as optimizing a tradeoff between the thermodynamic stability of a model, assessed by the Turner energy model, and its compatibility with the reactivity profile.
 As a consequence, popular packages for secondary structure prediction, such as \Software{RNAStructure}~\cite{Mathews2004} or the \Software{Vienna} package~\cite{Lorenz2011}, now routinely accept SHAPE-like reactivity profiles as input of their main structure prediction methods.
-These include the joint folding of aligned RNAs~\cite{Lavender2015}, partition function, statistical sampling and Maximal Expected Accuracy predictions~\cite{Spasic2017}\ldots
-
-However, independent computational predictions for the same RNA using probing data obtained with different probes often yield largely different models, none of which are fully consistent with all the probing data.
+\Rev{Besides free-energy optimization, such methods notably} include the joint folding of aligned RNAs~\cite{Lavender2015}, partition function, statistical sampling and Maximal Expected Accuracy predictions~\cite{Spasic2017}\ldots
+
+However, independent computational predictions for the same RNA using probing data obtained with different probes often yield substantially different models, none of which are fully consistent with all the probing data.
 It follows that, while theoretically informative, a multiprobing strategy often leaves a user with different models, from which one cannot objectively decide.
 In order to identify the best fit with all the data, researchers resort to very intuitive and manual methods such as projecting the results obtained with one probe onto the different models obtained using the constraints from another probe~\cite{Herbreteau2005,James2008,Weill2004}.
-At the end of the modeling process, the modeler is often left with several alternatives, all of which may appear equally consistent with the probing data, and the choice of a "best" model becomes somewhat subjective.
-Here we have developed a new modeling procedure that jointly takes into account multiple probing data, and ultimately yields a small collection of secondary structure models.
+At the end of the modeling process, the modeler is often left with several alternatives, all of which may appear equally consistent with the probing data.
+Thus, we have developed a new modeling procedure that jointly takes into account multiple probing data, and ultimately yields a small collection of secondary structure models.
 
@@
 \subsection*{The \OurTool{} method}
-We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signal produced by various chemical probes.
+We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signals produced by various chemical probes \Rev{to predict one or several secondary structure models for a given RNA.}
 
@@
 The data-driven predicted structures are then gathered in one sample, serving as input for the clustering step.
-\OurTool{} proceeds, then, with an iterative clustering that ends once the stopping criterion are reached.
+\OurTool{} proceeds, then, with an iterative clustering that ends once \Rev{a stopping criterion is} reached.
 This step allows to identify the adequate number of clusters $k$ to be considered.
-The $k$ resulting clusters are then identified by their centroid structures.
+The $k$ resulting clusters are then \Rev{represented} by their centroid structures.
 Clusters figuring on the 2D-Pareto frontier are considered to be optimal and subsequently their corresponding centroids are reported as the predicted structure through \OurTool.}\label{fig:approach}
@@
 
-Our method, summarized in Figure~\ref{fig:approach}, takes as input a set $\mathcal{D}$ of probing experiments, each materialized by a reactivity profile.
-It starts by producing sets of representatives structures for each of the reactivity profiles using a SHAPE-directed variant of the classic Ding-Lawrence algorithm~\citep{Ding2003}.
-Following Deigan\etal\cite{Deigan2009}, soft constraints are used to complement the free energy contributions of the classic Turner energy model with pseudo energy contributions resulting from reactivity profiles derived from SHAPE probing experiments.
-Given a reactivity $R_i$ for a position $i$, we associate a free-energy bonus to unpaired positions, defined as
-$$\Delta G(i) = m \log(R_i +1 )+b$$ 
+\paragraph{Sampling the pseudo-Boltzmann ensembles.} Our method, summarized in Figure~\ref{fig:approach}, takes as input a set $\mathcal{D}$ of probing experiments, each materialized by a reactivity profile.
+It starts by producing (multi)sets of representative structures for each of the reactivity profiles using a SHAPE-directed variant of the classic Ding-Lawrence algorithm~\citep{Ding2003}.
+Following Deigan\etal\cite{Deigan2009}, \Def{soft constraints} are used to complement the free energy contributions of the classic Turner energy model with pseudo energy contributions resulting from the reactivity derived from a probing experiment.
+Given a reactivity $r_i$ for a position $i$ in a probing $d$, we associate a free-energy bonus to unpaired positions, defined as
+$$\Delta G_d(i) = m \log(r_i +1 )+b$$ 
 using $m=1.3$ and  $b=-0.4$.
-Those values differ from those used by Deigan\etal\cite{Deigan2009}, and were adapted based on the rationale that lower absolute values for pseudo-energy bonuses increase the chance of structures being simultaneously supported by multiple conditions.
-Those pseudo energy contributions effectively direct our predictions towards a subset of structures that are in good agreement with probing data.
-
-In order to infer recurrent candidates across different samples, sampled structure sets are agglomerated while keeping track of the reagent/condition of origin for each structure, and clustered using the Mini Batch k-Means algorithm~\citep{Sculley2010} (\CL{}) based on a number of clusters determined by an iterative heuristics described in further details below.
-It takes as input a dissimilarity matrix, whose entries consist of the base pair distance between two sampled structures.
-\CL{} was chosen as it requires less computational resources than the classic k-means algorithm, yet performs equally as well as an extensive collection including  both agglomerative (affinity propagation) and hierarchical (Ward, Diana, McQuitty) clustering algorithms (data not shown).
-
-The number of clusters is a critical parameter of the \CL{} algorithm.
+\Rev{Those values were halved in comparison to those recommended by Deigan\etal\cite{Deigan2009}, following a grid search optimization on the Cordero dataset, and based on the rationale that lower absolute values for pseudo-energy bonuses increase the expected overlap between pseudo-Boltzmann ensembles.} Those pseudo energy contributions effectively guide our predictions towards a subset of structures that are in good agreement with probing data.
+For each condition $d$ in $\mathcal{D}$, we use the soft constraints framework~\cite{Lorenz2016} of the \Software{RNAfold} software~\cite{Lorenz2011} to produce a random (multi)set $\mathcal{S}_d$ of $\NumStruct$ structures in the pseudo-Boltzmann ensemble.
+
+\Rev{\paragraph{Clustering across conditions.} In order to infer recurrent conformations across sampled sets, \OurTool{} agglomerate structure (multi)sets while keeping track of their condition of origin, and \Def{clusters} with respect to the \Def{base-pair distance}, the number of base pairs differing between two structures.
+A clustering algorithm then partitions the (multi)set of sampled structures into \emph{clusters}, (multi)sets of structures such that the accumulated sum of distances over clusters is minimized.
+
+Among the many available options, we chose the Mini Batch $\NumClust$-Means algorithm~\citep{Sculley2010} (MBkM), implemented in the \Software{scikit-learn} pacakage~\cite{Pedregosa2012}, which requires less computational resources than the classic $\NumClust$-means algorithm, yet performed similarly in preliminary studies as an extensive collection of both agglomerative (affinity propagation) and hierarchical (Ward, Diana, McQuitty) clustering algorithms.
+A dissimilarity matrix, presenting the pairwise base-pair distance between structures, is precomputed and fed to the clustering algorithm.
+
+Any cluster $C$ output by the clustering is a multiset of structures, each labeled with its origin condition of $\mathcal{D}$.
+The \Def{cluster probability of a structure feature} $f$ (base pair or unpaired base) within a cluster $C$ is then defined as 
+$$ \mathbb{P}_C(f) = \frac{\sum_{\substack{S\in \mathcal{C}\\\text{s.t. } f\in S}} e^{-E(S)/RT}}{\sum_{\substack{S'\in \mathcal{C}}} e^{-E(S')/RT}} $$ 
+where $R$ represents the Boltzmann constant, $E(S)$ is the Turner free-energy, and $\mathcal{C}$ is the (non-redundant) set of structures in $C$.
+From those probabilities we define the \Def{centroid structure} of a cluster as its Maximum Expected Accuracy (MEA) structure, computed efficiently following Lu\etal~\citep{Lu2009}.
+
+Moreover, define the (pseudo-)\Def{Boltzmann condition probability} of a structure $S$, generated for a probing condition $d$ as part of a sampled  set $\mathcal{S}_d$, as
+$$\mathbb{P}_d(S) = \frac{e^{-E_d(S)/RT}}{\mathcal{Z}^*_d} \text{, with } \mathcal{Z}^*_d := \sum_{S'\in \mathcal{S}_d} e^{-E_d(S')/RT}$$
+where $E_d(S)$ is the pseudo free-energy, including the Turner free-energy, assigned to the structure $S$ within the probing condition $d$.
+The \Def{stability} of a cluster $C$ denoted its accumulated pseudo-Boltzmann probability across conditions, computed as
+$$\text{Stability}(C) = \sum_{d\in \mathcal{D}} \sum_{S\in C\cap \mathcal{S}_d}  \mathbb{P}_d(S).$$
+A cluster is deemed \Def{significantly populated} if its stability exceeds a predefined threshold $\epsilon$. 
+We set  $\epsilon =\Rev{|\mathcal{D}|}/3$ by default, such that at most three clusters are deemed significantly populated, and used as our primary candidates.
+Finally, we consider two clusters to be \Def{highly similar} if their centroid structures differ by at most $\delta$ base pairs ($\delta=1$ by default), allowing the identification of clusters in the presence of minor variations.
+
+The targeted number of clusters is a critical parameter of the \CL{} algorithm.
 It should, at the same time, remain small enough to ensure reproducibility, while being sufficiently large to discriminate outliers and ensure consistency within each cluster.
-We iteratively increase the number of clusters, deeming it appropriate if increasing its value leads to:
-\begin{enumerate} 
-	\item Splitting a significant cluster into two highly similar clusters, or; 
-	\item Creating a poorly populated (outlier) cluster, while keeping highly similar clusters for each significant cluster from the previous iteration.
-\end{enumerate} 
-In practice, clusters are seldom exactly preserved across successive iterations, so we consider two clusters to be \Def{highly similar} if their Maximum Expected Accuracy (MEA) centroid structures~\citep{Lu2009}, differ by less than $\delta$ base pairs.
-The $\delta$ parameter defaults to $1$, allowing the identification of clusters across successive generation in the presence of a minor variation.
-Moreover, define the (pseudo-)Boltzmann probability of a structure $S$, generated for an experimental condition $d$ as part of a sampled set $\mathcal{S}_d$, as
-\[\mathbb{P}_d(S) = \frac{e^{-E_d(S)/kT}}{\mathcal{Z}^*_d} \text{, with } \mathcal{Z}^*_d := \sum_{S'\in \mathcal{S}_d} e^{-E_d(S')/kT}\]
-where $E_d(S)$ is the pseudo free-energy assigned to the structure $S$ within the experimental condition $d$.
-We consider a cluster to be \Def{significantly populated} if the cumulative probability of its structures exceeds a predefined threshold $\epsilon$. 
-The value of $\epsilon$ was set to $\#{\sf Datasets}/3$, ensuring that up to three clusters are deemed significantly populated, and used as our primary candidates.
-
-Based on these definitions, our iterative heuristic consists in running \CL over an increasing number $\NumClust$ of clusters, starting with $\NumClust=2$, until a \Def{stopping criterion} is satisfied, that is either: 1) Two significantly populated clusters have associated centroids which are highly similar; or 2) Centroid structures of significantly populated clusters from the previous iteration are highly similar to those of the current iteration.
-
-The last step of our approach consists in electing the most promising clusters, and returning their centroid.
-While the final number of clusters $\NumClust$ may potentially be large, only a handful of clusters are expected to represent structures that both stable, and supported by a large number of conditions.
-The remaining clusters are indeed probably artifacts of the clustering method, but nevertheless useful to filter \emph{noisy} structures.
-Thus we define the \Def{stability} of a cluster $C$ to denote its accumulated pseudo-Boltzmann probability accross conditions.
-It is computed as
-\[\text{Stability}(C) = \sum_{d\in \mathcal{D}} \mathbb{P}_d(S).\]
-Moreover, we postulate that a perfect cluster should be representative of several conditions.
-In the presence of a set of experimental conditions $\mathcal{D}$, we consider that a cluster $C$ supports a given condition $d$ when its probability for $d$ exceeds a given threshold $\tau$, where $\NumClust$ is the number of clusters.
+We determine an \Def{optimal number of clusters} $\NumClust$ using an iterative heuristics, gradually increasing the number of clusters until a significantly-populated cluster is split into two similar clusters, or a poorly populated (outlier) cluster is created.
+Namely, our iterative heuristic consists in running \CL over an increasing number $\NumClust$ of clusters, starting with $\NumClust=2$, until the following \Def{stopping criterion} is met: 1) Two significantly populated clusters have associated centroids which are highly similar; or 2) Centroid structures of significantly populated clusters from the previous iteration are highly similar to those of the current iteration.}
+
+\paragraph{Filtering the promising conformer(s).} Finally, we \Def{elect the most promising cluster(s)}, and return their centroid(s).
+While the final number of clusters may potentially be large, only a handful of clusters are expected to represent structures that \Rev{are} both stable, and supported by a large number of conditions.
+The remaining clusters are indeed probably artifacts of the clustering method, but nevertheless useful to filter out \emph{noisy} structures.
+
+\Rev{We postulate that a perfect cluster should have large stability, as defined above, and be representative of several conditions.} In the presence of a set of experimental conditions $\mathcal{D}$, we consider that a cluster $C$ supports a given condition $d$ when its probability within $d$ exceeds a given threshold $\tau$.
 The number of conditions supporting a cluster $C$ is defined as
-\[\text{Support}(C) = |\{d\in \mathcal{D} \mid \sum_{S\in C\cap \mathcal{S}_d} \mathbb{P}_d(S)\ge \tau \}|.\]
-The value of $\tau$ is set to $1/(\NumClust+1)$, ensuring at least one supporting cluster for each condition.
-
-\OurTool evaluates the two above metrics on each cluster, and eliminates any cluster that is dominated by another with respect to both metrics.
+$$\text{Support}(C) = \left|\left\{d\in \mathcal{D} \mid \left(\textstyle\sum_{S\in C\cap \mathcal{S}_d} \mathbb{P}_d(S)\right)\ge \tau \right\}\right|.$$
+The value of $\tau$ is set to $1/(\NumClust+1)$ \Rev{with $\NumClust$ the final number of clusters},  ensuring at least one supporting cluster for each condition.
+
+\OurTool evaluates the Stability and Support metrics for each cluster, and filters out any cluster that is dominated by some other with respect to both metrics.
 The remaining ones are \Def{Pareto optimal}, a classic concept in multi-objective optimization~\cite{Mattson2005}.
-Finally, we compute and return the MEA centroid~\citep{Lu2009} of the Pareto-optimal clusters as our final prediction(s).
+\OurTool computes and returns the MEA centroid~\cite{Lu2009} of the Pareto-optimal clusters as its final prediction(s).
 
@@
 As a measure of the \Def{ensemble distance} $\Edist$ induced by probing data, we consider the dot-plots associated with experimental conditions $d$ and $d'$, and compute the squared Euclidean distance, such that
-\[\Edist (d,d')=\sum\limits_{i=1}^\RL\sum\limits_{j=x+1}^\RL \left(\mathbb{P}(i,j\mid d)-\mathbb{P}(i,j\mid d')\right)^2.\]
+$$\Edist (d,d')=\sum\limits_{i=1}^\RL\sum\limits_{j=i+1}^\RL \left(\mathbb{P}(i,j\mid d)-\mathbb{P}(i,j\mid d')\right)^2.$$
 with $\RL$ the length of the RNA sequence,  $\mathbb{P}(i,j\mid d)$ the Boltzmann probability of forming a base pair $(i,j)$ in the pseudo-Boltzmann ensemble associated with condition $d$.
@@
 
-\paragraph{Cordero \emph{et al} dataset.} 
+\paragraph{Hajdin dataset.} 
+A dataset was gathered by Hajdin\etal\cite{Hajdin2013} to validate the predictive capacities of probing data-driven predictions.
+It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (1M7-\SH) is available.
+This dataset includes sequences originating from a variety of organisms, and spans lengths ranging from $34$ nts to $530$ nts, with a focus on riboswitches and complex RNA architectures.
+
+\paragraph{Cordero dataset.} 
 In order to test the integration of different probing sources, we consider a dataset introduced in Cordero\etal\cite{Cordero2012}, consisting of $6$ RNAs with known structures, for which reactivities are available for three reagents: {\tt NMIA-\SH, DMS} and {\tt CMCT}.
@@
 
-\paragraph{Hajdin \emph{et al} dataset.} 
-A dataset was gathered by Hajdin\etal\cite{Hajdin2013} to validate the predictive capacities of probing data-driven predictions.
-It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (1M7-\SH) is available.
-This dataset includes sequences originating from a variety of organisms, and spans lengths ranging from $34$ nts to $530$ nts, with a focus on riboswitches and complex RNA architectures.
-
-\paragraph{Didymium structural model and probing data.} 
+\paragraph{Didymium structural model and probing data (\didy{} dataset).} 
 We considered the 188 nucleotides Lariat capping ribozyme from {\itshape Didymium iridis}, resolved 3.85 \AA{} resolution using X-ray cristallography (PDB: 4P8Z)~\citep{Meyer2014}.
 We annotated the secondary structure elements using the \Software{DSSR} software from the 3DNA suite~\cite{Lu2015}.
-Non-canonical base pairs were removed~\citep{Smit2008}, by considering as the secondary structure the maximum subset of non-pseudoknotted base pairs.
-
-Probing data were experimentally generated, as described in the next section, using a comprehensive set of conditions covering most of the popular probing reagents and SHAPE technologies for stop-inducing and mutation-inducing adducts.
+\Rev{Non-canonical base pairs were removed, and a non-pseudoknotted secondary structure was extracted as the maximum subset of non-crossing base pairs~\citep{Smit2008}.}
+
+Probing data were experimentally generated, as described in the next section, using a comprehensive set of conditions covering some of the popular probing reagents and SHAPE technologies. 
 We also considered the presence/absence of Mg$^{2+}$, both to assess the capacity of \OurTool{} to recover tertiary interactions, and to assess the induced discrepancy on probing profiles and pseudo-Boltzmann ensembles.
 
-\paragraph{Cheng  \emph{et al} dataset.} 
+\paragraph{Cheng dataset.} 
 Starting from the assumption that a functional structure should be preserved during evolution, we wanted to assess the agreement that might exist between probing data profiles for a set of RNA mutants.
@@
 \subsection*{Experimental probing protocols}
-The structure of the Lariat capping ribozyme from \textit{Didymium iridis }(DiLCrz) was probed with different SHAPE reagents: 1M7 (1-methyl-7 nitrosatoic anhydre), NMIA (N-Methyl Isatoic Anhydre), BzCN (Benzoyl cyanide) and NAI (2-methylnicotinic acid imidazolide) in presence and absence of Mg$^\text{2+}$, and with DMS (DiMethylSulfate) and CMCT only in presence of Mg$^\text{2+}$.
-As a preliminary experiment, we verified that DiLCrz generally adopts a single global architecture.
-To this end, we subjected DiLCrz to a standard denaturation protocol (80\degree{}C for 2 min in H$_\text{2}$O, addition of 40 mM of HEPES at 7.5 pH, 100 mM of KCl, 5 mM of MgCl$_\text{2}$, followed by 10 min at room temperature, and 10 min at 37\degree{}C), and observed the production of a single band on a non-denaturing PAGE, strongly suggesting the adoption of a single conformation.
+
+\Rev{To systematically assess the potential of multiple sources of probing, we considered a difficult example, the \textit{Didymium iridis} Lariat Capping ribozyme (\didy{}).
+The native structure of \didy{}, shown in Figure~\ref{fig:didy} is highly complex, and features two pseudoknots which cannot be explicitly modeled by most computational methods, making \didy{} a challenging target for secondary structure prediction.}
+
+\didy{} was probed with different SHAPE reagents: 1M7 (1-methyl-7 nitrosatoic anhydre), NMIA (N-Methyl Isatoic Anhydre), BzCN (Benzoyl cyanide) and NAI (2-methylnicotinic acid imidazolide) in presence and absence of Mg$^\text{2+}$.
+\Rev{\didy{} was also probed with DMS (DiMethylSulfate) and CMCT, this time only in presence of Mg$^\text{2+}$, resulting in a total of 16 \Def{probing conditions}, a term we use in the following to denote a combination of probing technology, reagent and presence/absence of Mg$^\text{2+}$ (+ sequencing technology).
+For each probing condition, three experiments were performed in presence/absence of the reagent, and in a denatured context, following classic SHAPE protocols~\cite{Wilkinson2006,Smola2015}.}
+
+As a preliminary experiment, we verified that \didy{} generally adopts a single global architecture.
+To this end, we subjected \didy{} to a standard denaturation protocol (80\degree{}C for 2 min in H$_\text{2}$O, addition of 40 mM of HEPES at 7.5 pH, 100 mM of KCl, 5 mM of MgCl$_\text{2}$, followed by 10 min at room temperature, and 10 min at 37\degree{}C), and observed the production of a single band on a non-denaturing PAGE, strongly suggesting the adoption of a single conformation.
 
@@
 For \NMIA and \NMIAMg, experiments were conducted as described above, except that the NGS library preparation was adapted to Ion Torrent sequencing.
-Sequences were mapped and analyzed with \Software{ShapeMapper2} for \OneMSevILUThreeMg and \OneMSevILUThreeMg.
-For the other conditions they are mapped with a script based on \Software{ShapeMapper}, but adapted to Ion Torrent output files.
+Sequences were mapped and analyzed with \Software{ShapeMapper2}~\cite{Busan2018} for \OneMSevILUThreeMg and \OneMSevILUThreeMg.
+For the other conditions they are mapped with a script based on \Software{ShapeMapper}~\cite{Smola2015}, but adapted to Ion Torrent output files.
 The corresponding data sets are named after the probe followed by {\sc MaP}.
@@
 
-The Matthews Correlation Coefficient (MCC) is a classic metrics for assessing the quality of a predicted structure $S$, identified by a set of base pairing positions, based on its agreement with an accepted reference structure $\Ref$.
-It can be interpreted as a compromise between the classic sensitivity and specificity metrics, and is defined as
+The Matthews Correlation Coefficient (MCC) is a classic metric for assessing the quality of a predicted structure $S$, identified by a set of base pairing positions, based on its agreement with an accepted reference (native) structure $\Ref$.
+It can be interpreted as a compromise \Rev{between the main metrics derived from the confusion matrix}, and is defined as
 $$\text{MCC}(S\mid \Ref)=\frac{\text{TP} \times \text{TN} -\text{FP} \times \text{FN}}{\sqrt[]{(\text{TP}+\text{FP})(\text{TP}+\text{FN})(\text{TN}+\text{FP})(\text{TN}+\text{FN})}},$$
@@
 
-For the sake of direct comparison with some competing methods~\cite{Spasic2017}, we also report the geometric mean metrics, defined as:
+For the sake of direct comparison with some competing methods~\cite{Spasic2017}, we also report the Geometric Mean (GM) metric, defined as:
 $$ \GMean(S\mid \Ref)=\sqrt{\text{Sens}(S\mid \Ref)\times \text{PPV}(S\mid\Ref)} $$
 where $\text{Sens}(S\mid \Ref)$ represents the proportion of base pairs in $\Ref$ that are in $S$, and $\text{PPV}(S\mid \Ref)$ is the proportion of base pairs in $S$ that are also in $\Ref$.
-This quality metrics has very high correlation with the MCC, to which it is asymptotically equivalent as shown by~Gorodkin\etal\cite{Gorodkin2001}.
+This quality metric has very high correlation with the MCC, to which it is equivalent for large values of $n$, as shown by~Gorodkin\etal\cite{Gorodkin2001}.
+
+\Rev{
+\subsection*{Assessment of statistical significance}
+Following Xu\etal\cite{Xu2012}, we assessed the statistical significance of observed changes in predictive performances, using a two-tailed Student $t$-test with unequal variance, using a classic significance threshold $\alpha=5\%$.
+When comparing performances within a given dataset, we used the paired version of the test.
+
+Intuitively, the $t$-test estimates the probability that an observed change in performance can be attributed to chance alone.
+It considers the hypothesis $H_0$ that the two distributions have equal mean and computes a \Def{P-value}, \emph{i.e.} the likelihood of the observed values under $H_0$.
+If the P-value is lower than $\alpha$, then $H_0$ is deemed \Def{refuted}, and the observed difference in average performance is considered as significant.
+
+}
 
@@
 
-\subsection*{Considering multiple probing conditions improves the quality of predictions}\label{sec:cordero}
-
-%In order to validate our assumption of an enrichment in the presence of multiple probing profiles, we analyzed the Cordero \emph{et al} dataset, which contains CMCT, DMS and NMIA probing data for 6 RNAs.
+\subsection*{Considering multiple probing conditions \Rev{appears to} improve the quality of predictions}\label{sec:cordero}
+
+%In order to validate our assumption of an enrichment in the presence of multiple probing profiles, we analyzed the Cordero \emph al} dataset, which contains CMCT, DMS and NMIA probing data for 6 RNAs.
 %
@@
 
-Since \OurTool{} supports any number of probing profiles, we consider the Cordero~\emph{et al}~\cite{Cordero2012a} dataset, further described in the previous section.
+Since \OurTool{} supports any number of probing profiles, we considered the Cordero\etal\cite{Cordero2012a} dataset, further described in the previous section.
 It consists of 6 RNAs of known structures, for which reactivity profiles are available for the CMCT, NMIA and DMS reagents.
@@
 The centroid secondary structure  associated with the largest probability cluster was considered as the final prediction.
-Predictions were evaluated in term of the geometric mean (GM) metrics, for the sake of comparison with previous studies~\cite{Spasic2017}.
+Predictions were evaluated in term of the geometric mean (GM) metric, for the sake of comparison with previous studies~\cite{Spasic2017}.
 As a control experiment, we also report ${\tt RNAfold}$ predictions in presence/absence of probing data, both in energy minimization (MFE) and maximum expected accuracy (MEA) modes.
@@
 
-	\caption{Averaged GM values for structures  predicted using \OurTool{} for subset combination of probing data.
-MFE and MEA structures, as computed by RNAfold, are included as controls in the absence of probing data, and for mono-probing.}
+	\caption{\Rev{Predictive performances (GM) of \OurTool{} on the Cordero~\cite{Cordero2012a} dataset consisting of reactivities associated with the CMCT, NMIA and DMS probing for 6 reference RNAs.
+(Pseudo) MFE and MEA structures, predicted using \Software{RNAfold}~\cite{Lorenz2011}, are included as references in the case of a single condition, and in the absence of probing.}}
 	\label{cordero1}
@@
 
-In the absence of probing data, MFE predictions are generally dominant on this dataset, trailed by the MEA, and followed by \OurTool which, in this setting, devolves into the Ding-Lawrence algorithm~\citep{Ding2003}.
+In the absence of probing data, MFE predictions are generally dominant on this dataset, trailed by the MEA, and followed by \OurTool which, in this setting, devolves into the classic Ding-Lawrence algorithm~\citep{Ding2003}.
 However, whenever probing reactivities are available, the single centroid returned by \OurTool always achieves higher GM values (Avg: $70\%$) than both MEA (Avg: $62\%$) and MFE  (Avg: $58\%$), whose relative performances depend on the probing reagent.
@@
 It is also worth mentioning that NMIA+CMCT, combining the best and worst conditions, achieves a better combined performance than DMS+NMIA, the two best mono-probing conditions.
-This hints towards some level of complementarity between probing conditions, as already suggested by recent analyses \citep{Brunel2000,Steen2012,Rice2014,Yu2018}  and supported by further analyses in this paper.
 
 Remarkably, the combination of the three conditions leads to the best overall predictions, averaging $78.5\%$ GM.
-This improves by $8.5\%$, $2.7\%$ and $1.4\%$ the average predictions achieved using CMCT, DMS and NMIA respectively.
+This \Rev{seemingly} improves by $8.5\%$, $2.7\%$ and $1.4\%$ over the average predictions achieved using CMCT, DMS and NMIA respectively.
 %CMCT/DMS/NMIA :0.6988333333/0.7588333333/0.771
 Table~\ref{fig:glycine_example} illustrates the incremental refinement of \OurTool predictions for a glycine riboswitch upon increasing the number of considered probing conditions.
+\Rev{However, these observations merely suggest an improvement, as statistical significance cannot be established on such a small dataset.}
+Still, they suggests some level of complementarity between probing conditions, as already suggested by recent analyses~\citep{Brunel2000,Steen2012,Rice2014,Yu2018}  and supported by further analyses in this paper.
 
@@
 %* Multi conformation
-We assess the predictive capacities of \OurTool in a classic setting where a single probing condition is available, and compare its performances against \Software{RSample} software recently introduced by Spacic\etal\cite{Spasic2017}, \Software{RSample} relies on a sampling/clustering method developed independently from the current work.
-It was shown to perform favorably against a comprehensive collection of state-of-the-art methods in probing-guided structure prediction, including \Software{RME}~\citep{Wu2015}, \Software{RNAprob}~\citep{Deng2016}, \Software{RNAprobing}~\citep{Washietl2012}, \Software{RNAsc}~\citep{Zarringhalam2012} and the \Software{fold} utility from the \Software{RNAStructure} suite~\citep{Reuter2010}.
-We consider the dataset of hajdin~\emph{et al}, which consists of 24 RNAs of lengths ranging from 47 to 500 nucleotides, all believed to fold into a unique documented conformation.
+We assessed the predictive capacities of \OurTool in a classic setting where a single probing condition is available, and compared its performances against \Software{RSample} software recently introduced by Spacic\etal\cite{Spasic2017}, \Software{RSample} relies on a sampling/clustering method developed independently from the current work.
+It was shown to perform \Rev{comparably as} a comprehensive collection of state-of-the-art methods in probing-guided structure prediction, including \Software{RME}~\citep{Wu2015}, \Software{RNAprob}~\citep{Deng2016}, \Software{RNAprobing}~\citep{Washietl2012}, \Software{RNAsc}~\citep{Zarringhalam2012} and the \Software{fold} utility from the \Software{RNAStructure} suite~\citep{Reuter2010}.
+We considered the Hajdin dataset~\cite{Hajdin2013}, which consists of 24 RNAs of lengths ranging from 47 to 500 nts, all believed to fold into a unique documented conformation.
 
@@
 	\includegraphics[width=\linewidth]{graphs/RsampleVsIPANEMAP/Accuracy}
-	\caption{Geometric mean of positive predictive value and sensitivity achieved by the predictions of \OurTool{} and \Software{Rsample} over the Hajdin\etal\cite{Hajdin2013} data set, consisting of 24 RNAs (RNA length indicated within square brackets).
-\label{fig:Vsrsample}}
+	\caption{\Rev{Comparison of predictive performances (GM) achieved by  \OurTool{} and \Software{Rsample} over the Hajdin\etal\cite{Hajdin2013} data set, consisting of 24 RNAs which associated SHAPE reactivities.
+The length of each individual RNA is indicated within square brackets.} \label{fig:Vsrsample}}
 \end{figure}
 
-We executed \OurTool with default parameters, producing 1~000 structures per sample for both software.
-The Pareto-optimality always implies a single returned cluster/structure in the case of a single probing condition.
-We also executed \Software{RSample} with default parameters, also returning a single structure for each sequence. 
+\Rev{We used the default number of sampled structures for \Software{RSample}, namely 10~000 to compute correction factors and 1~000 for the clustering phase, and \OurTool (1~000 structures).
+For all RNAs, a single structure was returned by both software.
+Note that the Pareto-optimality always implies a single returned cluster/structure for \OurTool{} in a mono probing setting.}
 We computed and report in Figure~\ref{fig:Vsrsample} the individual Geometric Means (GM) associated with predicted structures.
@@
 %\paragraph{Results description}
-A first observation is that \OurTool generally outperforms \Software{RSample}.
-It averages $80.1\%$ GM as opposed to $68\%$ for its competitor.
+A first observation is that \OurTool generally outperforms \Software{RSample}, averaging $80\%$ GM as opposed to $68\%$ for its competitor.
+\Rev{The superiority on \OurTool on this benchmark appears statistically significant according to a two-tailed paired $t$-test (P-value$=6\times10^{-3}$) with observed data.}
 Out of the 24 considered RNAs, \OurTool outperforms its competitor for 16 out of the 24 RNAs, achieves similar performances for 2 of them, and is dominated for the remaining 6 cases.
@@
 It ensures GM that are greater than 50\% in all cases, and even exceeds 60\% over all except for the yeast tRNA, encoding the Aspartic acid and HIV1 5' pseudoknot domain.
-The standard deviation of GM for \OurTool is $12.5\%$, compared to the larger value of $21\%$ for \Software{RSample} predictions.
+The standard deviation of GM for \OurTool is $12.4\%$, compared to the larger value of $21\%$ for \Software{RSample} predictions.
 
@@
 
-	\caption{Model of the Lariat capping ribozyme from {\itshape Didymium iridis} (A -- PDB 4P8Z:A by Meyer\etal\cite{Meyer2014}), secondary structure, as annotated by \Software{DSSR} (B -- pseudoknot ommitted) and \OurTool first-ranking prediction (C.), obtained from the integration of 8 maximally-diverging probing conditions (\NMIAMg, \OneMSevILUThreeMg, \NAIMg, \OneMSevCE, \CMCTMg,\OneMSevILUThree, and \OneMSevMgCE).\label{fig:didy}}
+	\caption{\Rev{Model of the Lariat capping ribozyme from {\itshape Didymium iridis} (A -- PDB 4P8Z:A by Meyer\etal\cite{Meyer2014}), secondary structure, as annotated by \Software{DSSR}~\cite{Lu2015} (B) and \OurTool first-ranking prediction (C.), obtained from the integration of 8 maximally-diverging probing conditions (\NMIAMg, \OneMSevILUThreeMg, \NAIMg, \OneMSevCE, \CMCTMg,\OneMSevILUThree, and \OneMSevMgCE).}\label{fig:didy}}
 \end{figure*}
-As benchmark RNA we used the Lariat capping ribozyme from \textit{Didymium Iridis} (DiLCrz) which 3D structure has been modeled at 3.85 \AA{} from X-ray diffraction data~\cite{Meyer2014}.
-This RNA contains a large diversity of interactions, including helical domains, tertiary pairings, a kissing loop and a pseudoknot (Figure~\ref{fig:didy}).
-It was probed as described in Material and Methods in 16 different conditions using different SHAPE and classical reagents, in presence/absence of Mg$^\text{2+}$ ions, using either the standard premature stop or the newly developed mutation induction technologies to map the modification sites.
+To assess the potential of our method, we considered the Lariat capping ribozyme from \textit{Didymium Iridis} (\didy{}) whose 3D structure has been modeled at 3.85 \AA{} from X-ray diffraction data~\cite{Meyer2014}.
+\Rev{This RNA contains a large diversity of interactions, including helical domains, tertiary pairings, a kissing loop and a pseudoknot (Figure~\ref{fig:didy}).
+Its automated modeling is therefore a challenging case for computational  structure prediction methods.
+For instance, running \Software{RNAfold} with default parameters only recovers four (DP2, DP2.1, P10 and P9) of the ten helices.
+Those are universally recovered by energy-based prediction tools, so differences in performances will mainly be observed between nucleotides 35 and 163.}
+
+As described in the Material and Methods section, we probed \didy{}  in 16 different conditions using different SHAPE and classical reagents, in presence/absence of Mg$^\text{2+}$ ions, using either the standard premature stop or the newly developed mutation induction technologies to map the modification sites.
 These experimental results were then used as input of \OurTool{}, either individually or combined,  for secondary structure prediction.
@@
 \paragraph{Comparison and clustering of conditions.}
-Faced with diverse probing conditions, we first assess the compatibility of the conclusions drawn from different probing data, including the probing-free run of \OurTool as a control.
-We use the methodology described in the Material and Methods~\ref{sec:dotplots} and, for each condition, compute the base pair probability distribution (aka \emph{dot-plot}) in the  pseudo-Boltzmann ensemble, induced by the reactivity profile.
-We then computed the ensemble distance, the squared Euclidean distance between dot plots, for each pair of conditions.
+Faced with diverse probing conditions, we first assessed the compatibility of the conclusions drawn from different probing data, including the probing-free run of \OurTool as a control.
+We used the methodology described in the Material and Methods~\ref{sec:dotplots} and, for each condition, computed the base pair probability distribution (aka \emph{dot-plot}) in the  pseudo-Boltzmann ensemble induced by the reactivity profile.
+We then computed the Ensemble Distance, the squared Euclidean distance between dot plots, for each pair of conditions.
 Figure~\ref{fig:PCA} summarizes the pairwise distance, projected onto a 2D surface by a principal component analysis.
@@
 \end{bulletlist}
-The predicted clusters are generally consistent with the ordering resulting from a spectral biclustering, implemented in \Software{scikit-learn} and executed with default parameters, as illustrated by Figure~\ref{fig:glycine_example}.
+The predicted clusters are generally consistent with the ordering resulting from a spectral biclustering, implemented in \Software{scikit-learn} and executed with default parameters, as illustrated by \Rev{Figure~\ref{fig:biclustering}}.
 
@@
 
-For each probing condition, we run \OurTool{} on a single source of probing data, using a sample size of $1~000$ structures.
-For the sake of reproducibility, we executed \OurTool{} $10$ times for each condition.
-We report in Table~\ref{tab:thermo-ipan} the MCC values achieved by our prediction, averaged over the $10$ runs, along with the MCCs obtained by running \Software{RNAfold} with default parameters in energy minimization (MFE) and Maximum Expected Accuracy (MEA) modes.
+For each probing condition, we executed \OurTool{} on a single reactivity profile, using a sample size of $1~000$ structures.
+For the sake of reproducibility, we executed \OurTool{} $10$ times for each condition, and report in Table~\ref{tab:thermo-ipan} the average MCC values.
+We also report for reference the MCCs obtained by running \Software{RNAfold} with default parameters in energy minimization (MFE) and Maximum Expected Accuracy (MEA) modes.
 
 The predictive performances of \OurTool averages a 70.5\% MCC, compared to 68.5\% and 68\% average MCC for MEA and MFE predictions respectively.
-The stability of predictions across conditions are comparable for all \OurTool, MEA and MFE-driven predictions (std dev.
+\Rev{This suggests modest, barely significant, improvements compared to classic prediction paradigms (P-values of 2\% and 14\% respectively).}
+The stability of predictions across conditions seem comparable for all \OurTool, MEA and MFE-driven predictions (std dev.
 of 8\%).
-Excluding the outliers of cluster \Bull{B} improves the average MCC to 74\%.
-
-Across the sixteen conditions, we observe a large discrepancy in the capacity of the experimental setup to inform predictions, with MCCs ranging from above 80\% (\NMIA, \NMIAMg, \OneMSevILU and \OneMSevILUMg) to around 60\% (\NAICE, \BzCN, \OneMSevCE and \NMIACE).
-Interestingly the MCC value cannot be trivially anticipated from an assessment of the compatibility of the reactivity scores with the predicted structure, or even with the native structure (see Supp.
-Table S6).
- These predictive capacities clearly outperform those achieved in the absence of probing data (51\% MCC), and can be interpreted as indicative of good quality predictions.
-Indeed, remind that the MCC is a stringent metrics, taking values between -100\% and 100\%, 0\% being the expected MCC of a "coin tossing" random predictor.
+Excluding the outliers of cluster \Bull{B} improved the average MCC to 74\%.
+
+Across the sixteen conditions, we observed a large discrepancy in the capacity of the experimental setup to inform predictions, with MCCs ranging from above 80\% (\NMIA, \NMIAMg, \OneMSevILU and \OneMSevILUMg) to around 60\% (\NAICE, \BzCN, \OneMSevCE and \NMIACE).
+\Rev{Interestingly the MCC value cannot be trivially anticipated from the compatibility of the reactivity scores with the predicted structure, or even with the native structure (see Supp.
+Table S6).}
+ These predictive capacities clearly outperformed those achieved in the absence of probing data (51\% MCC), and can be interpreted as indicative of predictions of good overall quality.
+Indeed, it is worth reminding that the MCC is a stringent metric, taking values between -100\% and 100\%, 0\% being the expected MCC of a "coin tossing" random predictor.
 The lowest MCC value of 60\% is, in particular, equally consistent with Sensitivity/PPV values of 60\%/60\%, 80\%/46\% or 45\%/82\%.
 
-Unsurprisingly, the presence/absence of Magnesium ions during the probing impacts the predictions, with an observed drop from an average 73\% MCC in the presence of Mg$^\text{2+}$ to 67\% MCC in the absence of Mg$^\text{2+}$.
-However, while the poorly performing conditions in Cluster \Bull{B} are probed in the absence of Mg$^\text{2+}$, some of the most informative profiles (\OneMSevILU and \NMIA) were obtained in absence of Mg$^\text{2+}$.
-An even more drastic change of performances can be observed when comparing the MCC of mutation-based protocols (avg 78\%) with stop-based ones (avg 66\%).
+Unsurprisingly, the presence/absence of Magnesium ions during the probing was observed to impact the predictions, with an observed drop from an average 73\% MCC in the presence of Mg$^\text{2+}$ to 67\% MCC in the absence of Mg$^\text{2+}$.
+However, while the poorly performing conditions in Cluster \Bull{B} \Rev{were} probed in the absence of Mg$^\text{2+}$, some of the most informative profiles (\OneMSevILU and \NMIA) were obtained in absence of Mg$^\text{2+}$.
+\Rev{Even more drastic changes of performances were observed when comparing the MCCs of SHAPE-MaP experiments (avg 78\%) with the historical SHAPE CE technology (avg 66\%).}
 
@@
 	}
-	\newcommand{\Mut}{Mut}
-	\newcommand{\Stop}{Stop}
-	\newcommand{\Mg}{$\bullet$}
-	\newcommand{\NoMg}{$\circ$}
 	\newcommand{\M}[1]{\multicolumn{1}{l}{#1}}
@@
 			\begin{adjustbox}{max width=\linewidth}
-		\begin{tabular}{@{}lcccHHH}
+		\begin{tabular}{@{}lccHHH}
 			\toprule
-			                            &          &         &           & \multicolumn{3}{@{}c@{}}{MCC\%}              \\
-			Condition                   &  Clust.  &  Tech.  & Mg$^{2+}$ & \M{\relsize{-1}\OurTool} & \M{MFE} & \M{MEA} \\ \midrule
-			\OneMSevILUMg               & \Bull{C} &  \Mut   &    \Mg    & 85                       & 82      & 82      \\
-			\OneMSevILU                 & \Bull{A} &  \Mut   &   \NoMg   & 84                       & 82      & 84      \\
-			\NMIAMg                     & \Bull{A} &  \Mut   &    \Mg    & 81                       & 80      & 80      \\
-			\NMIA                       & \Bull{A} &  \Mut   &   \NoMg   & 80                       & 69      & 69      \\
-			\OneMSevILUThreeMg          & \Bull{C} &  \Mut   &    \Mg    & 74                       & 75      & 75      \\
-			\NMIAMgCE                   & \Bull{E} &  \Stop  &    \Mg    & 73                       & 72      & 73      \\
-			\NAIMg                      & \Bull{G} &  \Stop  &    \Mg    & 73                       & 68      & 70      \\
-			\BzCNMg                     & \Bull{H} &  \Stop  &    \Mg    & 71                       & 68      & 54      \\
-			\OneMSevMgCE                & \Bull{E} &  \Stop  &    \Mg    & 70                       & 69      & 70      \\
-			\CMCTMg                     & \Bull{G} &  \Stop  &    \Mg    & 70                       & 68      & 73      \\
-			\OneMSevILUThree            & \Bull{D} &  \Mut   &   \NoMg   & 64                       & 56      & 61      \\
-			\DMSMg                      & \Bull{F} &  \Stop  &    \Mg    & 63                       & 66      & 67      \\
-			\NMIACE                     & \Bull{B} &  \Stop  &   \NoMg   & 61                       & 60      & 60      \\
-			\OneMSevCE                  & \Bull{B} &  \Stop  &   \NoMg   & 60                       & 61      & 61      \\
-			\BzCN                       & \Bull{B} &  \Stop  &   \NoMg   & 60                       & 59      & 59      \\
-			\NAICE                      & \Bull{B} &  \Stop  &   \NoMg   & 60                       & 59      & 59      \\ \midrule
-			Avg Technology              &          &  \Mut   &  $\star$  & 78                       & 74      & 75      \\
-			                            &          &  \Stop  &  $\star$  & 66                       & 65      & 65      \\
-			Avg -/+ Mg$^{2+}$           &          & $\star$ &   \NoMg   & 67                       & 64      & 65      \\
-			                            &          & $\star$ &    \Mg    & 73                       & 72      & 71.5    \\ \midrule
-			{\bfseries Average overall} &          & $\star$ &  $\star$  & 70.5                     & 68      & 68.5    \\ \bottomrule
-			                            &          &         &           &                          &
+			                            &          &           & \multicolumn{3}{@{}c@{}}{MCC\%}              \\
+			Condition                   &  Clust.  & Mg$^{2+}$ & \M{\relsize{-1}\OurTool} & \M{MFE} & \M{MEA} \\ \midrule
+			\OneMSevILUMg               & \Bull{C} &    \Mg    & 85                       & 82      & 82      \\
+			\OneMSevILU                 & \Bull{A} &   \NoMg   & 84                       & 82      & 84      \\
+			\NMIAMg                     & \Bull{A} &    \Mg    & 81                       & 80      & 80      \\
+			\NMIA                       & \Bull{A} &   \NoMg   & 80                       & 69      & 69      \\
+			\OneMSevILUThreeMg          & \Bull{C} &    \Mg    & 74                       & 75      & 75      \\
+			\NMIAMgCE                   & \Bull{E} &    \Mg    & 73                       & 72      & 73      \\
+			\NAIMg                      & \Bull{G} &    \Mg    & 73                       & 68      & 70      \\
+			\BzCNMg                     & \Bull{H} &    \Mg    & 71                       & 68      & 54      \\
+			\OneMSevMgCE                & \Bull{E} &    \Mg    & 70                       & 69      & 70      \\
+			\CMCTMg                     & \Bull{G} &    \Mg    & 70                       & 68      & 73      \\
+			\OneMSevILUThree            & \Bull{D} &   \NoMg   & 64                       & 56      & 61      \\
+			\DMSMg                      & \Bull{F} &    \Mg    & 63                       & 66      & 67      \\
+			\NMIACE                     & \Bull{B} &   \NoMg   & 61                       & 60      & 60      \\
+			\OneMSevCE                  & \Bull{B} &   \NoMg   & 60                       & 61      & 61      \\
+			\BzCN                       & \Bull{B} &   \NoMg   & 60                       & 59      & 59      \\
+			\NAICE                      & \Bull{B} &   \NoMg   & 60                       & 59      & 59      \\ \midrule
+			Avg Technology              &          &  $-$  & 78                       & 74      & 75      \\
+			                            &          &  $-$  & 66                       & 65      & 65      \\
+			Avg -/+ Mg$^{2+}$           &          &   \NoMg   & 67                       & 64      & 65      \\
+			                            &          &    \Mg    & 73                       & 72      & 71.5    \\ \midrule
+			{\bfseries Average overall} &          &  $-$  & 70.5                     & 68      & 68.5    \\ \bottomrule
+			                            &          &           &                          &
 		\end{tabular}
-		\end{adjustbox}\\[1em]}
+		\end{adjustbox}}
 
 	%data are from multiprobing7 
-	\caption{Comparison of predicted secondary structures for the \didy{} obtained with \OurTool{} and computational methods from a mono-probing experiment.
-Averaged MCC of structures predicted with \OurTool{} and with classic deterministic alternatives.
-The structure was probed in presence (\Mg) or absence (\NoMg) of Magnesium ions, using either stops (\Stop) or mutations (\Mut) inducing technologies.
+	\caption{\Rev{Predictive performances (MCC) of \OurTool{}, averaged over 10 runs, for the \didy{} dataset, consisting of 16 reactivity profiles.
+(Pseudo) MFE and MEA structures, predicted using \Software{RNAfold}~\cite{Lorenz2011}, are included for reference.
+The structure was probed in presence (\Mg) or absence (\NoMg) of Mg$^{2+}$.}
 	}
@@
 \begin{figure}
-	\newcommand{\W}{.86}
+	\newcommand{\W}{.7}
 	{\centering\includegraphics[width=\W\linewidth]{graphs/didy/pairs-spiders}\\}
@@
 
-\paragraph{Bi-probing mitigates the influence of outliers, but generally fails to improve predictions quality.}
-We turn to a systematic exploration of the predictive capacities of bi-probing analyses, based on pairs of probing profiles, and attempt to quantify their impact on \OurTool predictions.
-For each pair and triplets of conditions, we executed \OurTool using 1000 samples per condition, and considered the first returned structure.
-We then computed the associated MCC, and compare it with the MCC of the worst performing condition (Min), best-performing condition (Max) and average MCC over the single conditions experiments.
+\paragraph{Bi-probing mitigates the influence of outliers, but does not significantly improve prediction quality.}
+Next, we turned to a systematic exploration of the predictive capacities of bi-probing analyses, based on pairs of probing profiles, and attempted to quantify their impact on \OurTool predictions.
+For each pair and triplets of conditions, we executed \OurTool using 1~000 structures per condition, and considered the first returned structure.
+We then computed the associated MCC, and compared it with the MCC of the worst performing condition (Min), best-performing condition (Max) and average MCC over the single conditions experiments.
 A summary of the results over pairs is shown in Figure~\ref{fig:pairwise} (see details in Supp.
@@
 
-A superficial inspection of the resulting MCCs appears to confirm the conclusions reached on the Cordero~\emph{et al}~\cite{Cordero2012} dataset.
-Indeed, considering two conditions leads to predictions whose quality fall between the worst and the best one, while being generally close to the average.
-More precisely, the average MCC value of pair-informed predictions remains at 70.7\%, improving only by 0.2\% over the average MCC of mono-probing predictions.
-Pair-informed predictions are marginally more reliable, with the median MCC increases to 72.5\% MCC from 70.5\% for mono-probing predictions.
-However, outliers strongly impact this picture, and ignoring the conditions in cluster \Bull{B} increases the average MCC to 75.1\%, while unsurprisingly decreasing the standard deviation (8.6\% $\to$ 6.5\%).
-
-Compared to the minimum MCC of the pair, the MCC of bi-probing predictions is improved by 5.3\% on average over the MCC of the worst-performing conditions.
-Pair MCCs exceed their associated min.
-MCC by at least 5\% in 47 pairs out of the 120 possible pairs, while never being dominated by 5\%.
-
-Bi-probing tends to perform similarly as the average of the two mono-probing conditions, with an average MCC improvement of 0.2\%.
-However, for 56 out of the 120 possible pairs, the bi-probing conditions exhibits an improvement of at least 1\% over the average, while a decay of at least 1\% is observed for 46 out of the remaining 64 pairs.
+A superficial inspection of the resulting MCCs appears to confirm the conclusions reached on the Cordero\etal\cite{Cordero2012} dataset.
+Indeed, considering two conditions led to predictions whose quality fall between the worst and the best one, while being generally close to the average.
+More precisely, the average MCC value of pair-informed predictions remained at 70.7\%, improving only by 0.2\% over the average MCC of mono-probing predictions.
+Pair-informed predictions were marginally more reliable, with the median MCC \Rev{increasing} to 72.5\% MCC from 70.5\% for mono-probing predictions.
+However, outliers strongly impacted the overall picture, and ignoring the conditions in cluster \Bull{B} increased the average MCC to 75.1\%, while unsurprisingly decreasing the standard deviation (8.6\% $\to$ 6.5\%).
+
+Compared to the minimum MCC of the pair, the MCC of bi-probing predictions was improved by 5.3\% on average over the MCC of the worst-performing conditions.
+Pair MCCs exceeded their associated min.
+MCC by at least 5\% in 47 pairs out of the 120 possible pairs, while never being dominated by more than  5\% MCC.
+
+Bi-probing seemed to perform similarly as the average of the two mono-probing conditions, with an average MCC improvement of 0.2\%.
+However, for 56 out of the 120 possible pairs, the bi-probing conditions exhibited an improvement of at least 1\% over the average, while a decay of at least 1\% was observed for 46 out of the remaining 64 pairs.
 Such a decay can partly be attributed to the disruptive effect of the outliers from cluster \Bull{B}, involved in 35 of the 46 observed loss of predictive performance.
-Moreover, removing this cluster induces an average MCC improvement of 1.1\%.
-
-Predictions informed by two conditions remain, however, generally dominated by the best performing condition of the pair.
+Moreover, removing this cluster induced an average MCC improvement of 1.1\%.
+
+Predictions informed by two conditions remained, however, generally dominated by the best performing condition of the pair.
 On average, the MCC of the best condition is 4.9\% higher than the one achieved by bi-probing analysis.
 Outliers in cluster \Bull{B} are partly responsible for this situation, and their removal reduces the average MCC decay to 3.3\%.
-Moreover, for 9 pairs of conditions, the bi-probing analysis produces MCC that are at least 1\% better than the best of its two conditions, but these encouraging examples are generally dominated by 66 examples where the pair performs more than 1\% worse than the best mono-probing condition.
+Moreover, for 9 pairs of conditions, the bi-probing analysis produced MCCs that are at least 1\% better than the best of its two conditions, but these encouraging examples were generally dominated by 66 examples where the pair performs more than 1\% worse than the best mono-probing condition.
 
 \paragraph{Considering three conditions yields improvement over the average performance of the triplet.}
-Increasing the number of conditions to three boosts the average MCC to 72.3\% and even reaches 76.5\% over triplets that do not contain conditions from cluster \Bull{B}.
-The MCC gain over the worst condition in the triplet reaches 9.1\% on average.
-Predictions are also more consistently good, with a median MCC of \%73.4.
-Disregarding conditions in cluster \Bull{B} does stabilizes the quality of prediction (stdev 8.4\% $\to$ 5.5\%), while increasing the average MCC to 77.4\%.
-
-Compared to bi-probing, MCC values achieved by integrating triplets of conditions now indicate a clear gain over the average performances of individual conditions, supporting a notion of complementary between experiments.
-Indeed tri-probing experiments lead to an average +1.7\% MCC improvement (median +2.6) over the average of mono-probing prediction in the triplet.
-The gain of quality is widespread, and observed for 65\% of the triplets.
+Increasing the number of conditions to three boosted the average MCC to 72.3\% and even reached 76.5\% over triplets that do not contain conditions from cluster \Bull{B}.
+The MCC gain over the worst condition in the triplet reached 9.1\% on average.
+Predictions were also more consistently good, with a median MCC of \%73.4.
+Disregarding conditions in cluster \Bull{B} did stabilize the quality of prediction (stdev 8.4\% $\to$ 5.5\%), while increasing the average MCC to 77.4\%.
+
+Compared to bi-probing, MCC values achieved by integrating triplets of conditions indicated a clear gain over the average performances of individual conditions, supporting a notion of complementary between experiments.
+Indeed tri-probing experiments induced an average +1.7\% MCC improvement (median +2.6) over the average of mono-probing prediction in the triplet \Rev{(P-value$=1.4\times 10^{-9}$)}.
+The gain of quality was widespread, and observed for 65\% of the triplets.
 Finally, 39 triplets were found to induce MCC values greater than 83\%.
 
-The enhanced performances of tri-probing over bi-probing are consistent with the \emph{voting} principle underlying the clustering used in \OurTool.
-Indeed, in a bi-probing setting, a single outlier condition, such as \NAICE, may entirely determine the final structure, due to its predictions being very tightly concentrated around a, presumably erroneous, structure.
-In the presence of three or more conditions, however, clusters resulting from an outlier condition typically end up being dominated, in the Pareto front, by compatible sets of structures originating from the pseudo-Boltzmann ensembles of alternative conditions.
-It follows that the influence of outliers over the final prediction remains limited.
+The enhanced performances of tri-probing over bi-probing \Rev{was deemed statistically significant (P-value$=2\times 10^{-2}$), and} consistent with the \emph{voting} principle underlying the clustering used in \OurTool.
+\Rev{Indeed, in a bi-probing setting, a single outlier (\emph{e.g.} \NAICE) may entirely determine the final structure due to its Boltzmann ensemble being very tightly concentrated around a, presumably erroneous, centroid structure.
+In the presence of three or more conditions, however, clusters resulting from an outlier condition are typically expected to be dominated in the Pareto front, by compatible structures originating from alternative conditions.
+It follows that the influence of outliers over the final prediction is mitigated.}
 
@@
 \paragraph{Similar conditions do not contribute to the quality of predictions.} 
-Our distance assessment and clustering on conditions reveals groups of conditions that are highly compatible in their conclusions, while others seem to include highly diverging structural information.
-While some outlier conditions, such as those of cluster \Bull{B}, easily stand out as erroneous, the results of our systematic analysis of pairs and triplets supports a notion of \Def{complementarity between conditions}.
-Under this assumption, downstream analyses are more likely to benefit from the presence of highly diverging probing experiments, than the accumulation of similar, arguably redundant, profiles.
-
-As a test of this hypothesis, we executed \OurTool{} on every subset of conditions within clusters, and report in Figure~\ref{fig:similarconds} the quality (MCC) of predictions for any subset of similar conditions (see Supp.
+Our distance assessment and clustering on conditions revealed groups of conditions that were highly compatible in their conclusions, while others seemed to include highly diverging structural information.
+While some outlier conditions, such as those of cluster \Bull{B}, clearly stood out as erroneous, the results of our systematic analysis of pairs and triplets supported a notion of \Def{complementarity between conditions}.
+Under this assumption, downstream analyses would be more likely to benefit from the presence of highly diverging probing experiments, than the accumulation of similar (arguably redundant) profiles.
+
+To test this hypothesis, we executed \OurTool{} on every subset of conditions within clusters, and report in Table~\ref{fig:similarconds} the quality (MCC) of predictions for any subset of similar conditions (see Supp.
 Figure~S1 for associated models).
-Note that our similarity notion and clustering of conditions is purely based on the ensemble features, and does not explicitly consider the MCC of individual conditions. 
-As expected, compared to the average MCC over associated mono-probing analyses, a joint consideration of similar conditions only induces limited progress over single probing analyses (between -1\% and +1\%), leading to an modest average improvement (+0.2\% average MCC).
+\Rev{Note that such an analysis does not use knowledge of the native structure, since our clustering only relies on properties of the pseudo-Boltzmann ensemble.}
+
+As expected, compared to the average MCC over associated mono-probing analyses, a joint consideration of similar conditions only induced limited progress over single probing analyses (between -1\% and +1\%), leading to an modest average improvement (+0.2\% average MCC).
 This is consistent with the notion that supplementing a probing-based modeling with compatible conditions provides very little additional information, leading to a limited contribution of similar conditions.
@@
 Having established the redundant nature of similar conditions, and their limited contribution to downstream modeling, we turn to an analysis of conditions across clusters.
-We run \OurTool{} to assess the quality of predictions informed by condition chosen across the 8 clusters.We consider any possible combination, except for the outlier cluster~\Bull{B}, for which we only consider \NAICE as a representative.
-
-The results, summarized in Table~\ref{fig:differentconds}, demonstrate a clear advantage of including diverse conditions, leading to an average MCC value of 77\% in the absence of condition from cluster \Bull{B}.
-Those performances compare favorably against the 72\% MCC achieved on average by individual conditions, and reveal a remarkable stability, including a 5\% standard deviation, and a degradation in only two out the 24 tested cases ($\sim$1\% decay in both cases).
-A similar trend can be observed in the, arguably disruptive presence of \NAICE, inducing an 72\% average MCC, albeit with an increased standard deviation of 9\%.
-
-Finally, the consideration of four triplets of diverse conditions, reported in Table~\ref{fig:best}, allow to obtain the best prediction, achieving 85.3\% MCC.
-Meanwhile, considering the 16 conditions together yields a 80\% MCC value, increasing to 83\% in the absence of the four outlier conditions of cluster \Bull{B}.
-These results confirm that the strategy used by \OurTool{} is able to exploit the complementary information provided by multiple conditions, contributing to better predictions.
+We executed \OurTool{} to assess the quality of predictions informed by conditions chosen across the 8 clusters.
+We considered any possible combination, except for the outlier cluster~\Bull{B}, for which we only considered \NAICE as a representative.
+
+The results, summarized in Table~\ref{fig:differentconds}, demonstrate a clear advantage of including diverse conditions, with an average MCC value of 77\% in the absence of condition from cluster \Bull{B}.
+Those performances compare favorably against the 72\% MCC achieved on average by individual conditions \Rev{(P-value$=3\times 10^{-5}$)}, and reveal a remarkable stability (5\% stdev), and a degradation in only two out the 24 tested cases ($\sim$1\% \Rev{loss of MCC} in both cases).
+A similar trend can be observed in the disruptive presence of \NAICE, inducing an 72\% average MCC, albeit with an increased standard deviation of 9\%.
+
+Finally, considering the 16 conditions together induced a 80\% MCC value, increasing to 83\% in the absence of cluster \Bull{B}.
+\Rev{Overall, these results support the notion that the strategy used} by \OurTool{} is able to exploit the complementary information provided by multiple conditions, contributing to better predictions.
 
@@
 %
-	\begin{table}
-	\newcommand{\A}{45}
-	\newcommand{\B}{1.2em}
-	\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-	\centering
-	\begin{adjustbox}{max width=\linewidth}
-		\begin{tabular}{@{}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}cc@{}}
-			\toprule
-			\rotatebox{\A}{\Bull{C} \OneMSevILUMg} & \rotatebox{\A}{\Bull{D} \OneMSevILUThree} & \rotatebox{\A}{\Bull{A} \OneMSevILU} & \rotatebox{\A}{\Bull{F} \DMSMg} & \rotatebox{\A}{\Bull{E} \NMIAMgCE} & \rotatebox{\A}{\Bull{G} \CMCTMg} & \rotatebox{\A}{\Bull{H} \BzCNMg} &      MCC      & Mono-Avg \\ \midrule
-			$\bullet$             &                                  &            $\bullet$            &           $\bullet$            &                                   &                                 &                                 & \textbf{.853} &   .77    \\
-			$\bullet$             &                                  &            $\bullet$            &                                &             $\bullet$             &                                 &                                 & \textbf{.853} &   .81    \\
-			$\bullet$             &                                  &                                 &           $\bullet$            &                                   &            $\bullet$            &                                 & \textbf{.853} &   .73    \\
-			$\bullet$             &            $\bullet$             &                                 &                                &                                   &                                 &            $\bullet$            & \textbf{.853} &   .73    \\ \bottomrule
-		\end{tabular}
-	\end{adjustbox}\\
-
-	\caption{Best triplets of conditions, achieving maximal obtained MCC value.}\label{fig:best}
-\end{table}
+%	\begin{table}
+%	\newcommand{\A}{45}
+%	\newcommand{\B}{1.2em}
+%	\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+%
+%	\centering
+%	\begin{adjustbox}{max width=\linewidth}
+%		\begin{tabular}{@{}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}>{\centering}p{\B}cc@{}}
+%			\toprule
+%			\rotatebox{\A}{\Bull{C} \OneMSevILUMg} & \rotatebox{\A}{\Bull{D} \OneMSevILUThree} & \rotatebox{\A}{\Bull{A} \OneMSevILU} & \rotatebox{\A}{\Bull{F} \DMSMg} & \rotatebox{\A}{\Bull{E} \NMIAMgCE} & \rotatebox{\A}{\Bull{G} \CMCTMg} & \rotatebox{\A}{\Bull{H} \BzCNMg} &      MCC      & Mono-Avg \\ \midrule
+%			$\bullet$             &                                  &            $\bullet$            &           $\bullet$            &                                   &                                 &                                 & \textbf{.853} &   .77    \\
+%			$\bullet$             &                                  &            $\bullet$            &                                &             $\bullet$             &                                 &                                 & \textbf{.853} &   .81    \\
+%			$\bullet$             &                                  &                                 &           $\bullet$            &                                   &            $\bullet$            &                                 & \textbf{.853} &   .73    \\
+%			$\bullet$             &            $\bullet$             &                                 &                                &                                   &                                 &            $\bullet$            & \textbf{.853} &   .73    \\ \bottomrule
+%		\end{tabular}
+%	\end{adjustbox}\\
+%
+%	\caption{Best triplets of conditions, achieving maximal obtained MCC value.}\label{fig:best}
+%\end{table}
 %\begin{table}
@@
 \subsection*{Supplementing wild-type profiles with random single mutants increases prediction quality}
-In this final analysis, we decided to test the capacity of \OurTool to extract and exploit complementary information produced by the systematic probing of single-point mutants.
-This approach may appear misguided, since single point mutants are expected to adopt different structures than the wild-type sequence.
-In fact, the conformational changes induced by mutations precisely constitute the signal that is exploited by downstream modeling pipelines~\cite{Cordero2015}.
-However, those changes are typically local, and the limited influence of outliers on \OurTool{} raises a hope that the benefit of including independently-produced conditions will outweigh the impact of noise introduced by local changes.
-
-We generated 100 uniformly-distributed subsets consisting of 1, 2, 3 and 10 mutants extracted from the Cheng~\emph{et al} dataset~\cite{Cheng2017}, which we supplemented with the original wild-type sequence (WT).
+\Rev{In this final analysis, we tested the capacity of \OurTool to extract and exploit complementary information produced by the systematic probing of single-point mutants using the Mutate-and-Map (MaM) protocol~\cite{Cordero2015}.
+The joint analysis of several mutants may appear ill-advised, as single point mutants are expected to adopt different structures than the wild-type sequence (WT) , an expectation that is at the core of downstream analyses~\cite{Cordero2015}.
+However, those changes are typically local, and the limited influence of outliers on \OurTool{} substantiates the hope that the benefits of including independently-produced conditions will outweigh the cost of noise, introduced by local changes.}
+
+We generated 100 uniformly-distributed subsets consisting of 1, 2, 3 and 10 mutants extracted from the Cheng dataset~\cite{Cheng2017}, which we each supplemented with the WT sequence.
 For each set of probing profiles (\Cond{DMS}{mg}{MaP}{ilu}{}), we executed \OurTool{} with default parameters, using a sample size of 1~000 structures. 
 We also analyzed the WT sequence alone, reproducing the analysis 100 times to compare the variability across sets of mutants to the one induced by stochastic aspects of our method. 
-Finally, we performed a restricted analysis focusing on the WT, supplemented with 2 probing profiles selected from the 20 most similar conditions to the WT , as assessed by the ensemble distance metrics.
+Finally, we performed a restricted analysis focusing on the WT, supplemented with 2 probing profiles selected from the 20 most similar conditions to the WT , as assessed by the ensemble distance metric.
 By comparing our results with those obtained for unrestricted pairs of mutants, we assess the influence of structure variability on the quality of our predictions.
@@
 
-For the WT, the average MCC achieved by our predictions is of 61.83\%, comparable to the worst condition of our previous dataset.
-Our predictions are generally stable and, in 94\% out of the 100 runs, yield a structure having MCC between 60.5\% and 61\%.
-The remaining 6\% of the runs present improved MCC values, ranging from 76\% to 78\%, suggesting the existence of an alternative conformer in the pseudo-Boltzmann ensemble.
+For the WT alone, the predictions of \OurTool were associated with 61.8\% MCC, comparable to the worst condition of our previous dataset.
+Our predictions were generally stable and, in 94/100 runs, resulted in a structure having MCC between 60.5\% and 61\%.
+The remaining 6 runs showed improved MCC values, ranging from 76\% to 78\% and suggesting the existence of an alternative conformation in the pseudo-Boltzmann ensemble.
 %%\as{ This value can not be fixed since we have  0.6134: 12\%, 0.7597: 2\%, 0.7649: 1\%, 0.7783: 1\%, 0.7665: 1\%, 0.7531: 1\% , is it ok Yann not to specify the improvemnt percentage ?}
@@
 
-When the WT and a single random mutant are jointly considered, the dispersion of the results increases, and MCC values ranging from 38.9\% to 85\% can be observed reached.%min(Multiprobing_1M)
-These values, however, are not typical, and the average MCC in the presence of a single mutant increases to 64.9\%,%mean(Multiprobing_1M)
-suggesting a positive contribution from the additional mutant profile.
-
-This trend is confirmed when two random mutants are considered in addition to the WT. The average MCC across runs now increases to 66.6\%. %mean(Multiprobing_2M).
-Interestingly, the dispersion of the MCC is lower for two mutants than for a single one, with low values that are now confined to 57\%.
-This results suggests  a role of \emph{tie-breaker} played by the third condition, allowing to mitigate the deleterious effect of some erroneous outliers.
-However, such outliers still play some disruptive role, and restricting our choice of mutants within the 20 conditions most similar to the WT further improves the MCC value by 2.8\% to 67.7\%, with MCC values that now almost reach 60\%.
-
-The average MCC again increases in the presence of 3 mutants, to reach an average of 69.3\% MCC.
-For 10 mutants, the average MCC and overall distribution remains highly similar to that obtained by considering three mutants, suggesting peeking performances.
-These results support a notion of complementarity between the probing data produced across reasonably similar mutants, leading to a gradual decrease of the noise in the data.
-However, the bimodal nature of the distribution also suggests the existence of at least two conformations that are both supported by the WT and the mutant profiles.
+When the WT and a single random mutant (WT+1M) were jointly considered, the dispersion of the MCCs increased, and values ranging from 38.9\% to 85\% were observed.%min(Multiprobing_1M)
+\Rev{The average MCC in the presence of a single mutant increased to 64.9\%,%mean(Multiprobing_1M)
+suggesting a significant positive contribution from the additional mutant profile (P-value$=1\times 10^{-3}$).}
+
+This trend is confirmed when two random mutants were considered in addition to the WT (WT+2M), with an average MCC across runs increasing to 66.6\%. %mean(Multiprobing_2M).
+Interestingly, the dispersion of the MCC was lower for WT+2M than for WT+1M, with lower MCC values above 57\%.
+This results suggest  a role of \emph{tie-breaker} for the third condition, allowing to mitigate the deleterious effect of some outlier.
+However, outliers still played a disruptive role, and restricting chosen mutants to the 20 reactivity profiles most similar to the WT further improved the average MCC 2.8\% to 67.7\%,  \Rev{ a significant improvement compared to WT+1M (P-value$=2.4\times10^{-2}$).}
+
+The average MCC again increased in the presence of 3 mutants, to reach an average MCC of 69.3\%, \Rev{ a substantial improvement that nevertheless fails to reach the level of statistical significance (P-value$=11\times 10^{-2}$).} For 10 mutants, the average MCC and overall distribution remains highly similar to that obtained by considering three mutants, suggesting \Rev{a plateau in the performances}.
+Overall, these results support a notion of complementarity between the probing data produced across reasonably similar mutants, leading to a gradual decrease of the signal to noise ratio.
+The bimodal nature of the distribution also suggests the existence of two conformations, equally supported by the WT and the mutant profiles.
 
@@
 
-	\caption{Prediction accuracy (MCC) from Mutate-and-Map data.
-We consider the WT, supplemented with a set of $k=$1, 2, 3, or 10 single-point mutants, selected uniformly at random ({\sf WT+$k$M}).
-For 2 mutants, we also considered selecting mutants from the 20 most similar to the WT ({\sf WT+2M close}).}\label{fig:variantanalysis}
+	\caption{\Rev{Distributions of predictive accuracies (MCC) of structures predicted by \OurTool{}, informed by Mutate-and-Map data, for \didy{} (Cheng dataset~\cite{Cheng2017}).
+Predictions were repeated 100 times, informed both by the wild-type sequence and reactivity profile (WT), and with randomly-selected reactivity profiles associated with 1, 2, 3, or 10 single-point mutants ({\sf WT+$k$M}).
+Pairs of mutants restricted to the 20 most similar to the WT were also considered ({\sf WT+2M close}).
+} }\label{fig:variantanalysis}
 \end{figure}
@@
 We introduce \OurTool{}, an integrative method for the prediction of secondary structure models from multiple probing profiles.
-Based on the simple premise that good structural models should be stable and supported by most probing experiments, it uses a combination of stochastic backtrack in the pseudo-Boltzmann ensemble, coupled with a structural clustering to elect dominant conformations.
-This strategy is fast, stable and leads to predictions that clearly benefit from the availability of multiple probing profiles, as demonstrated by detailed analyses performed of four datasets.
+Based on the simple premise that good structural models should be \Rev{thermodynamically} stable and supported by most probing experiments, it uses a combination of stochastic backtrack in the pseudo-Boltzmann ensemble, coupled with a structural clustering to elect dominant conformations.
+This strategy is fast, \Rev{generally stable}, and leads to predictions that clearly benefit from the availability of multiple probing profiles, as demonstrated by detailed analyses performed of four datasets.
 
@@
 Surprisingly, it also appears to produce the most accurate secondary structure models when informed by a single condition, as demonstrated by our reanalysis on the Hajdin\etal\cite{Hajdin2013} benchmark.
-We attribute this unexpected result to the adoption of a sampling/clustering approach, which seems to eliminate spurious outliers in the conformational space.
-In particular, the dominance of \OurTool is surprising since, while independently developed, our method shares the sampling/clustering philosophy of the recently introduced \Software{RSample}~\cite{Spasic2017}.
-We attribute these increased performances in part to a less stringent contribution of the pseudo-potentials, and to our adaptive choice of the number of clusters.
-Indeed, even in cases where a single dominant conformation is expected, considering multiple clusters seems to create more opportunities for discarding noisy random unstable conformations.
+We attribute this unexpected result to the adoption of a sampling/clustering approach, \Rev{which appears to eliminate spurious outliers from the conformational space even when a single conformation dominates.}
+\Rev{In this context, the good performance of \OurTool in comparison with \Software{RSample}~\cite{Spasic2017} may appear surprising.
+Indeed, while independently developed and used~\cite{Deforges2017}, both methods share a common sampling/clustering scheme.
+However, \Software{RSample}~\cite{Spasic2017} reverts to a MEA~\cite{Lu2009} computation whenever a single conformation is suspected.
+The MEA structure coincides with the centroid of the whole Boltzmann ensemble, suggesting that the increased performances of \OurTool could be at least partially attributed to the use of sampling, and to our adaptive choice of the number of clusters.}
+Moreover, even in cases where a single dominant conformation is expected, considering multiple clusters seems to create more opportunities for discarding noisy random unstable conformations.
 Such conformations naturally occur in a Boltzmann-distributed subset of structures, and may end up \emph{polluting} the centroid structures by artificially inflating the cluster diversity.
 %The impact on \Software{RSample} predictions of such a \emph{noise} seems worsened by the fact that it returns the MEA, the centroid of the whole ensemble, whenever a dominant conformation is detected.
-
-Despite its stochastic foundations, \OurTool{} is typically stable in its predictions.
-We report in Figure~\ref{fig:reproducibility} the MCC values achieved over  10 repetitions of our software for all probing datasets produced in the case of didymium.
-Interestingly, the worst most notable exception to the general observed stability is observed for the probing-free prediction.
-This confirms the widespread interpretation of SHAPE-induced pseudo-potentials as focusing predictions onto a subset of the Boltzmann ensemble.
-A similar behavior is observed for the other datasets considered in our study, as seen in Supp.
-Tables S2 and S5.
-%In the interest of reproducibility, ipanemap reports, and accepts as an option, the random seed used both within the main software and its non-deterministic dependencies, leading to identical repeats of the experiments.
 
@@
 It follows that a modeler cannot objectively favor \emph{a priori} a probing condition/reagent~\cite{Yu2018}.
-In other words, multiple experimental probing profiles, even combined in a pairwise fashion, were typically observed to mitigate the empirical risk of a misprediction, and should be considered.
-
-A future work include solving the dilemma of whether or not to include Mg$^{\text{2+}}$ ions during the probing, in the perspective of a downstream computational prediction.
+In other words, multiple experimental probing profiles, even combined in a pairwise fashion, \Rev{can be used jointly to mitigate the empirical risk of a misprediction}.
+
+\Rev{Current limitations of \OurTool include the lack of an explicit support for pseudoknots, which will be addressed in a future version, using  polynomial-time dynamic programming~\cite{Janssen2015}, or a iterative heuristic alternative~\cite{Hajdin2013}.
+The inherently stochastic nature of the sampling scheme may also be challenging for certain types of analysis.
+However, despite its stochastic foundations, \OurTool{} is typically stable in its predictions, as can be seen in Figure~\ref{fig:reproducibility} for all probing datasets produced over our \didy case study.
+Interestingly, the most notable exception to the general stability is observed for the probing-free prediction, consistent with the widespread interpretation of SHAPE-induced pseudo-potentials as focusing predictions onto a subset of the Boltzmann ensemble.
+A similar behavior is observed for the other datasets considered in our study (see Supp.
+Tables S2 and S5).
+%In the interest of reproducibility, ipanemap reports, and accepts as an option, the random seed used both within the main software and its non-deterministic dependencies, leading to identical repeats of the experiments.
+Finally, the current clustering method induce a growth of time and memory requirements that scales quadratically with the number of samples.
+While this allows a joint analysis of a dozen of reactivity profiles in less than an hour on a personal computer, we expect the ability to push the number of samples will lead to better, even more reproducible, predictions.
+To that purpose, we will explore embedding techniques and linear-time community detection algorithms as alternatives to MBkM~\cite{Sculley2010}.}
+
+Future extensions also include \Rev{addressing} the dilemma of whether or not to include Mg$^{\text{2+}}$ ions during the probing, in the perspective of a downstream computational prediction.
 On the one hand, RNA reaches its native 3D structure in the presence of Mg$^{\text{2+}}$ ions~\cite{Knapp1989}.
@@
 Therefore, probing the RNA without Mg$^{\text{2+}}$ could provide a signal that is less deceptive, and thus more informative, to predicting algorithms.
-Unfortunately, destabilizing tertiary contacts increases the probability of misfolding, or the coexistence in solution of multiple conformations.
-In this work, we probed a Didymium ribozyme which contains many tertiary motifs, systematically in the presence and absence of Mg$^{\text{2+}}$ (except for CMCT and DMS probes).
-Remarkably, the most accurate prediction was obtained with 1M7 and NMIA, irrespectively of the presence/absence of Mg$^{\text{2+}}$, this despite the fact that the reactivity profiles substantially diverged between the two conditions.
+Unfortunately, destabilizing tertiary contacts increase the probability of misfolding, or the coexistence in solution of multiple conformations.
+In this work we probed \didy, which contains many tertiary motifs, in the presence and absence of Mg$^{\text{2+}}$.
+Remarkably, the most accurate predictions were obtained with 1M7 and NMIA \Rev{irrespectively of the presence/absence of Mg$^{\text{2+}}$, and despite the fact that reactivity profiles substantially diverged between the two conditions (see Supp.
+Fig.
+S4)}.
 However, for other reagents/protocols, predictions appeared positively impacted (+$10\%$ average MCC) by the presence of Mg$^{\text{2+}}$, painting a mixed picture that would deserve future investigations.
@@
 
-	\caption{MCC of predicted structures in presence/absence of Mg$^{2+}$, using both RT stops-inducing and mutation-inducing protocols for a collection of reagents, over 10 runs of \OurTool{}.
+	\caption{Stochastic variability of \OurTool{} predictions for \didy{} dataset.
+MCC of predicted structures over 10 runs of \OurTool{}, in presence/absence of Mg$^{2+}$, using both \Rev{SHAPE CE and MaP} protocols for a collection of reagents.
 }\label{fig:reproducibility}
@@
 \section*{Acknowledgements}
-The authors wish to express their gratitude to Ronny Lorenz for developing a probing-informed version of \Software{RNAsubopt}, allowing to sample the pseudo-Boltzmann ensemble, Nathalie Chamond for helpful suggestions and Benoit Masquida for the gift of DiLCRz plasmid and helpful discussions about protocols and DiLCRz structure .
+The authors wish to express their gratitude to Ronny Lorenz for developing a probing-informed version of \Software{RNAsubopt}, allowing to sample the pseudo-Boltzmann ensemble, Nathalie Chamond for helpful suggestions and Benoit Masquida for the gift of \didy{} plasmid and helpful discussions about protocols and \didy{} structure .
 
 \section*{Funding}
-This work, including AS and DA phD fellowships, has been supported by La Fondation pour la Recherche Medicale ([FRM
+This work, including AS and DA \Rev{PhD} fellowships, has been supported by La Fondation pour la Recherche Medicale ([FRM
 DBI20141423337]).

### 2019-12-22 SuppMat.tex
Updates from Overleaf

@@
 \usepackage{inconsolata}
+\usepackage{subcaption}
 
@@
 \input{macros.tex}
+
+\renewcommand{\Def}[1]{{\bf #1}}
 
@@
 
+\newpage
+
 \section{Multiprobing benchmark (Cordero~\emph{et al}~\cite{Cordero2012} dataset)}
-\subsection{MCC values of individual predictions}
+\subsection{GM values of individual predictions}
 
@@
 				\toprule
-				&        & \multicolumn{3}{l}{\Software{Rsample}}    & \multicolumn{3}{l}{\Software{IPANEMAP}}   \\
+				&        & \multicolumn{3}{l}{\Software{RSample}}    & \multicolumn{3}{l}{\Software{IPANEMAP}}   \\
 				{RNA} & Length & Sensitivity & PPV    & \multicolumn1c{GM} & Sensitivity & PPV    & \multicolumn1c{GM} \\ \midrule
@@
 		\end{adjustbox}\\}
-\caption{\OurTool~ prediction performance compared to \Software{Rsample} tool}
+\caption{\OurTool~ prediction performance compared to \Software{RSample} tool}
 \end{table}
@@
 	\begin{adjustbox}{max width=1\textwidth,max totalheight=\textheight,keepaspectratio}
-		\begin{tabular}{llHHHHHHHHHHHlHH}
+		\begin{tabular}{crHHHHHHHHHHHrHH}
 			\toprule
-			\textbf{RNA} & Length &  \multicolumn1c{Run1} &\multicolumn1c{Run2}&\multicolumn1c{Run3}&\multicolumn1c{Run4}&\multicolumn1c{Run5}&\multicolumn1c{Run6}&\multicolumn1c{Run7}&\multicolumn1c{Run8}&\multicolumn1c{Run9}&\multicolumn1c{Run10}& \multicolumn1c{Average}& \multicolumn1c{Sd}&\multicolumn1c{\textbf{MFE-RNAfold}} & \multicolumn1c{\textbf{MEA-RNAfold}}\\
+			\textbf{RNA} & Length &  \multicolumn1c{Run1} &\multicolumn1c{Run2}&\multicolumn1c{Run3}&\multicolumn1c{Run4}&\multicolumn1c{Run5}&\multicolumn1c{Run6}&\multicolumn1c{Run7}&\multicolumn1c{Run8}&\multicolumn1c{Run9}&\multicolumn1c{Run10}& \multicolumn1c{Average}& \multicolumn1c{Std Dev}&\multicolumn1c{\textbf{MFE-RNAfold}} & \multicolumn1c{\textbf{MEA-RNAfold}}\\
 			\midrule
@@
 %}
+\newpage
+\section{Case-study: Lariat capping ribozyme of {\itshape Didymium iridis}}
+\subsection{Validation of accessibility profiles: Agreement with native and predicted structures}
+We assess and validate the overall quality of probing data by defining a natural notion of {\bf agreement} between a reactivity profile and a structure.
+To that purpose, we consider a {\bf proportion $\alpha$ of the most and least reactive positions}, initially considering $\alpha=30\%$.
+Then, we investigate whether positions associated with extreme reactivity values are indeed paired or unpaired within the native structure.
 
-\section{Case-study: Lariat capping ribozyme of {\itshape Didymium iridis}}
-\subsection{Correlation analysis of reactivities}
-\subsection{Compatibility of accessibility profiles with native and predicted structures}
-We assess and validate the overall quality of probing data using a basic notion of distance between discretized profiles and structure.
-More precisely, given a secondary structure $S$ of length $n$, we define the {\bf 1D projection} of $S$ as the vector $v_S$ such that
+\paragraph{Agreement metrics.}More precisely, given a secondary structure $S$ of length $n$, we define the {\bf 1D projection} of $S$ as the vector $v_S$ such that
 $$ \forall i\in[1,n], v_S(i)=\begin{cases} 1 &\text{if $i$ unpaired in $S$}\\ 0 &\text{otherwise.} \end{cases} $$
+Given a vector $r$ of reactivities, we define a {\bf discrete reactivity vector} $d_{r}^{\alpha}$ as
+$$ \forall i\in[1,n], d_{r}^{\alpha}(i)=\begin{cases} 1 &\text{if $r_i$ within the $\alpha$ largest reactivity values},\\ 
+0 & \text{if $r_i$ within the $\alpha$ lowest reactivity values},\\
+1/2 &\text{otherwise (including if not available).} \end{cases} $$
+By focusing on the highest and lowest values, we avoid the influence of outliers, and hope to partially mitigate the natural variability of probing reactivities.
 
-Given a vector $R$ of reactivities, we define a {\bf discrete reactivity} vector $d_R$ as
-$$ \forall i\in[1,n], d_R(i)=\begin{cases} 1 &\text{if $R[i]$ within the $30\%$ largest reactivity values}\\ 
-0 & \text{if $R[i]$ within the $30\%$ lowest reactivity values}\\
-1/2 &\text{otherwise.} \end{cases} $$
-By focusing on the largest and lowest values, we avoid the influence of outliers, and hope to partially mitigate the natural variability of probing reactivities.
+We now define the {\bf normalized divergence} $||r,S||$ of a reactivity profile $r$ from a structure $S$ as
+$$ ||r,S|| = \frac{\sum_{i=1}^n |v_S(i)-d_r(i)|-(1-2\alpha)\times \frac{n}{2}}{n\times 2\alpha}.$$
+A correcting term counterbalances the $1/2$ associated with medium values, leading to an overall divergence of $(1-2\alpha)\times \frac{n}{2}$ regardless of the quality of the profile.
+The divisor term of $n\times 2\alpha$ accounts for the fact that, after compensating for the contribution of the $[\alpha,1-\alpha]$ percentile, only a proportion $2\alpha$ of the original values is effectively covered by the sum.
+Implementing those two corrections leads to a final value scaling between $0$ (all highly reactive pos.
+unpaired, all poorly reactive pos.
+paired) and $1$ (all highly reactive pos.
+paired, all poorly reactive pos.
+unpaired).
 
-We now define the {\bf normalized distance} $||R,S||$ of a reactivity profile $R$ and a structure $S$ as
-$$ ||R,S|| = \frac{\sum_{i=1}^n |v_S(i)-d_R(i)|-40\%\times \frac{n}{2}}{n\times 60\%}.$$
-Note that exactly 40\% of the positions are assigned to $1/2$, leading to $|v_S(i)-d_R(i)|=1/2$ irrespectively of the value taken by $v_S(i)$.
-The overall contribution of $40\%\times \frac{n}{2}$ needs to be corrected for, thus the subtraction.
-The divisor term of $n\times 60\%$ accounts for the fact that, after compensating for the contribution of the $[30,70]$ percentile, only $60\%$ of the original values are effectively covered by the sum.
-We thus renormalize by $n\times 60\%$ to obtain a final value scaling between $0$ (all highly reactive pos.
-unpaired, all lowly reactive pos.
-paired) and $1$ (all highly reactive pos.
-paired, all lowly reactive pos.
-unpaired).
-\begin{table}[H]
- \Heatset{min=20,  
-	max=60,   
-	max colour=BadCol, % colour at maximum
-	min colour=Aquamarine3,      % colour at minimum
-	Min colour=Aquamarine3, % colour for values below min
-	Max colour=BadCol   % colour for values above max
+Finally, we simply define the {\bf agreement of a reactivity profile $r$ with a structure $S$} as
+$$ \text{Agreement}(r,S) = 1- ||r,S||.$$
+
+\begin{figure}[t!]
+	\Heatset{min=30,  
+		max=80,
+		min colour=BadCol,      % colour at minimum
+		Min colour=BadCol, % colour for values below min
+		max colour=Aquamarine3, % colour at maximum
+		Max colour=Aquamarine3   % colour for values above max
 	}
@@
 	}
-{\centering
-\begin{tabular}{llHHS}
+		{\centering\resizebox{.5\linewidth}{!}{\begin{tabular}{@{}lllHHS}
 			\toprule
-	         &                    &      \multicolumn{2}{c}{Distance of reactivity (\%)}      &                       \\
-	Cluster  & Condition          & \multicolumn1c{to prediction} & \multicolumn1c{to native} & \multicolumn1c{MCC\%} \\ \toprule
-	\Bull{C} & \OneMSevILUMg      & 23                            & 29                        & 85                    \\
-	\Bull{A} & \OneMSevILU        & 32                            & 40                        & 84                    \\
-	\Bull{A} & \NMIAMg            & 26                            & 27                        & 81                    \\
-	\Bull{A} & \NMIA              & 30                            & 38                        & 80                    \\
-	\Bull{C} & \OneMSevILUThreeMg & 42                            & 38                        & 74                    \\
-	\Bull{E} & \NMIAMgCE          & 46                            & 39                        & 73                    \\
-	\Bull{G} & \NAIMg             & 38                            & 39                        & 73                    \\
-	\Bull{H} & \BzCNMg            & 38                            & 29                        & 71                    \\
-	\Bull{E} & \OneMSevMgCE       & 48                            & 42                        & 70                    \\
-	\Bull{G} & \CMCTMg            & 49                            & 45                        & 70                    \\
-	\Bull{D} & \OneMSevILUThree   & 33                            & 39                        & 64                    \\
-	\Bull{F} & \DMSMg             & 38                            & 38                        & 63                    \\
-	\Bull{B} & \NMIACE            & 31                            & 38                        & 61                    \\
-	\Bull{B} & \OneMSevCE         & 38                            & 41                        & 60                    \\
-	\Bull{B} & \BzCN              & 51                            & 46                        & 60                    \\
-	\Bull{B} & \NAICE             & 41                            & 44                        & 60                    \\ \bottomrule
-\end{tabular}\\}
-\caption{Distance of the predicted structure from the considered probing data.}
-\end{table}
+			    &          &                    & \multicolumn{2}{{@{}c@{}}}{Agreement$(R,S)$ (\%)}                                  &                               \\
+			Num &  Cluster        & Condition $R$      & \multicolumn1{@{}c}{$S:=\mathcal{N}$} & \multicolumn1{@{}c}{$S:=\mathcal{P}$} & \multicolumn1{c@{}}{MCC\%} \\ \toprule
+			1   & \Bull{C} & \OneMSevILUMg      & 70                                   & 77                                   & 85                            \\
+			2   & \Bull{A} & \OneMSevILU        & 60                                   & 68                                   & 84                            \\
+			3   & \Bull{A} & \NMIAMg            & 73                                   & 74                                   & 81                            \\
+			4   & \Bull{A} & \NMIA              & 62                                   & 70                                   & 80                            \\
+			5   & \Bull{C} & \OneMSevILUThreeMg & 62                                   & 58                                   & 74                            \\
+			6   & \Bull{E} & \NMIAMgCE          & 61                                   & 54                                   & 73                            \\
+			7   & \Bull{G} & \NAIMg             & 61                                   & 62                                   & 73                            \\
+			8   & \Bull{H} & \BzCNMg            & 70                                   & 62                                   & 71                            \\
+			9   & \Bull{E} & \OneMSevMgCE       & 58                                   & 52                                   & 70                            \\
+			10  & \Bull{G} & \CMCTMg            & 55                                   & 51                                   & 70                            \\
+			11  & \Bull{D} & \OneMSevILUThree   & 61                                   & 67                                   & 64                            \\
+			12  & \Bull{F} & \DMSMg             & 62                                   & 62                                   & 63                            \\
+			13  & \Bull{B} & \NMIACE            & 62                                   & 69                                   & 61                            \\
+			14  & \Bull{B} & \OneMSevCE         & 59                                   & 62                                   & 60                            \\
+			15  & \Bull{B} & \BzCN              & 54                                   & 49                                   & 60                            \\
+			16  & \Bull{B} & \NAICE             & 56                                   & 59                                   & 60                            \\ \bottomrule
+		\end{tabular}}\\}
 
-{\newcommand{\MyScale}{.26}
+%	\end{subfigure}\quad
+%	\begin{subfigure}[b]{0.52\textwidth}
+%		{\includegraphics[width=\textwidth,clip,trim=0 0 1.8cm 1.5cm]{graphs/didy/Agreement-Distribution.pdf}}
+%		
+%		\caption{Agreement distribution with random reactivity profiles.\label{fig:distribution_agreement}}
+%		\label{fig:mouse}
+%	\end{subfigure}
+	\caption{Analysis of agreements for reactivity profiles generated in various experimental conditions with both native ($\mathcal{N}$) and \OurTool{}-predicted ($\mathcal{P}$) structure.\label{tab:experiment_agreement}}
+\end{figure}
+
+\paragraph{Results.}We computed, and report in Table~\ref{tab:experiment_agreement}, the agreement of our 16 reactivity profiles with both the native and \OurTool{} prediction when considering the $\alpha$ percentile of highest/lowest reactivities ($\alpha=30\%$).
+A cursory inspection of the table reveals a {\bf good overall agreement of the measured reactivities with the native structure}, involving $61\%$ of the positions on average.
+This agreement increases to an average $70\%$ when $\alpha$ is set to $15\%$, indicating a good correlation between raw reactivities values and paired/unpaired status in the native structure.
+Also, expectedly, the agreement to the native structure appears to be positively correlated with the MCC induced by the reactivity profile.
+However, this relationship is not strict, as revealed by the $60\%$ agreement to the native observed for \OneMSevILU.
+While superficially surprising, this observation is consistent with the notion that probing data do not need to inform on many different positions to improve the quality of predictions.
+
+We tested the {\bf statistical significance} of our observation by computing  the agreement distribution in the absence of structural information.
+To that purpose, we generated random reactivity profiles, drawing a random reactivity uniformly in $[0,1]$ for each position.
+The resulting distribution can be perfectly fitted to a Normal distribution with mean agreement $\mu=49.96\%$ and standard deviation $\sigma=4.51\%$.
+{\bf All reactivity profiles showed better agreement with the native structure than a random reactivity profile}, by a margin of \Def{up to five standard deviations}, and always at least one standard deviation, better than the expected value of random profiles.
+This leads to \Def{significantly enriched agreements} (P-values(Agreement)$<$5\%) for 13 out of our 16 reactivity profiles, the remaining outliers having associated P-values of 18.5\%, 13.2\% and 9\% respectively.
+Overall, these observations support the notion of informative probing experiments.
+
+	\newcommand{\SR}[4]{{\begin{minipage}{.48\textwidth}\centering 
+			\includegraphics[angle=-90,width=\textwidth, trim=0 1.2cm 0 .5cm,clip]{graphs/didy/reactivities/#1}\\
+			#2 -- #4 (Cluster #3)\end{minipage}}}
 
 \begin{figure}	
+	{\centering \SR{1M7_Ilu_d1_Mg.pdf}{1    }{\Bull{C}  }{\OneMSevILUMg  }
+	\SR{1M7_Ilu_d1.pdf}{2    }{\Bull{A}  }{\OneMSevILU    }\\    
+	\SR{NMIA_Mg_IT.pdf}{3    }{\Bull{A}  }{\NMIAMg         }   
+	\SR{NMIA_IT.pdf}{4    }{\Bull{A}  }{\NMIA            }\\  
+	\SR{1M7_Ilu_d3_Mg.pdf}{5    }{\Bull{C}  }{\OneMSevILUThreeMg} 
+	\SR{NMIA_Mg_CE.pdf}{6    }{\Bull{E}  }{\NMIAMgCE          }\\
+	\SR{NAI_Mg_CE.pdf}{7    }{\Bull{G}  }{\NAIMg             }
+	\SR{BzCN_Mg_CE.pdf}{8    }{\Bull{H}  }{\BzCNMg            }\\}
+\caption{Projections of reactivities onto native structure (conditions 1-8 of 16).}
+\end{figure}
+\begin{figure}
+	\SR{1M7_Mg_CE.pdf}{9    }{\Bull{E}  }{\OneMSevMgCE       }
+	\SR{CMCT.pdf}{10   }{\Bull{G}  }{\CMCTMg            }
+	\SR{1M7_Ilu_d3.pdf}{11   }{\Bull{D}  }{\OneMSevILUThree   }
+	\SR{DMS.pdf}{12   }{\Bull{F}  }{\DMSMg             }
+	\SR{NMIA_CE.pdf}{13   }{\Bull{B}  }{\NMIACE            }
+	\SR{1M7_CE.pdf}{14   }{\Bull{B}  }{\OneMSevCE         }
+	\SR{BzCN_CE.pdf}{15   }{\Bull{B}  }{\BzCN              }
+	\SR{NAI_CE.pdf}{16   }{\Bull{B}  }{\NAICE   }
+\caption{Projections of reactivities onto native structure (conditions 9-16 of 16).}
+\end{figure}
+
+\newcommand{\MyScale}{.26}
+
+\subsection{Correlation analysis of raw reactivities profiles}
+\begin{figure}[H]
+	\includegraphics[width=.54\textwidth]{graphs/didy/reactivity_correlation.pdf}\hfill
+	\includegraphics[width=.45\textwidth]{graphs/didy/bi_clustering.pdf}
+	\caption{Correlation of reactivity profiles (left) and induced (pseudo-)ensemble distance (right).
+Conditions clustered in our analysis due to their impact of the pseudo-Boltzmann ensemble (black boxes, right) do not necessarily correlate very well in term of raw reactivity (white boxes, left).\label{fig:profile_correlation}}
+\end{figure}
+
+We compared raw reactivity data produced across various conditions, in order to test their overall compatibility, investigate their main determinants and compare their impacts on predictions.
+
+\paragraph{Correlation metrics and rationale.}To that purpose, we consider the Pearson product-moment \Def{correlation coefficients} $C_{r,r'}$ of two reactivity profile $r$ and $r'$, defined as
+$$ C_{r,r'} = \frac{\sum^n_{i=1}\left(r_i - \mu_r\right)\left(r'_i - \mu_{r'}\right)}{\sqrt{\sum ^n _{i=1}\left(r_i - \mu_{r}\right)^2} \sqrt{\sum ^n _{i=1}\left(r'_i - \mu_{r'}\right)^2}},$$ 
+where $\mu_{r}$ and $\mu_{r'}$  denote the mean reactivity within $r$ and $r'$ respectively.
+
+The choice of the \Def{correlation coefficient as a measurement of pairwise similarity} for probing profiles is motivated by differences in the distributions of reactivities across experimental conditions and protocols.
+Indeed, $C_{r,r'}$ remains unchanged by the application of any rescaling, through an affine transformation with positive slope, to either $r$, $r'$ (or both).
+Moreover, this metrics does not scale with the number of positions.
+Thus, we can restrict the sums and mean values to positions for which reactivities are available in both experiments, and still compare the coefficient with another pair of reactivity profiles.
+This property is important in order to tolerate \Def{missing values} introduced by experimental artifacts and reagent properties, while still being able to compare the coefficients.
+
+\paragraph{Results.} We computed the correlation coefficient for all pairs of reactivity profiles, and report the results in the left matrix of Figure~\ref{fig:profile_correlation}.
+For the sake of comparison, we also remind in this Figure the Ensemble Distance matrix, indicating the difference in impact on the computational predictions of each pair of conditions.
+
+We first observe that almost all of the pairs of reactivity profiles show a positive correlation, as expected from their observation of a common structure.
+A notable exception is \OneMSevMgCE, having average negative correlation with all its peers (except for \NMIAMgCE).
+Its outlier status is confirmed by its large ensemble distance to other conditions, yet does not prevent it to achieve a respectable MCC of 70\%.
+Finally, the perfectly negative correlation between \CMCTMg{} and \DMSMg{} can be dismissed as a computational artifact, since CMCT and DMS react with different nucleotides, and the correlation cannot be computed.
+
+Moreover, we observe that conditions clustered from the Ensemble Distance metrics (due to their impact of the pseudo-Boltzmann ensemble) do not always correlate in term of raw reactivity, with very poor correlations occurring within clustered conditions.
+This essentially reflects the fact that similar profiles may induce very similar pseudo-Boltzmann ensembles, \emph{ie} different premises sometimes imply the same conclusions.
+
+Conversely, one observes very few strong correlations outside of clusters.
+In fact, the only two exceptions are \OneMSevILU and \OneMSevILUThree which, despite belonging to different clusters and achieving different qualities of prediction, correlate almost perfectly with each other, and with other reactivity profiles produced using 1M7/SHAPEMap.
+Other natural candidates for explaining the reactivity correlations (Map vs CE, +Mg vs -Mg\ldots) do not appear to be supported by this analysis.
+
+Overall, this analysis confirms the consistency of the produced probing profiles (positive correlation), but also suggests a very delicate effect of reactivity data on the downstream predictions, with very different profiles leading to similar predictions (\OneMSevILU), and similar profiles inducing very different ones (\OneMSevILU vs \OneMSevILUThree).
+
+\subsection{Dominant conformations within main clusters}
+
+\begin{figure}[H]
 {\centering
@@
 \end{figure}
+
+\subsection{Prediction accuracy for pairs of conditions}
 
@@
 
+\subsection{Histograms of prediction accuracy informed by single-point mutants reactivities (Mutate-and-Map protocol)}
 \begin{figure}[H]

### 2019-12-22 graphs/didy/reactivities/exportPDF.py

@@ -1,0 +1,6 @@
+import os,sys
+
+for f in os.listdir("."):
+    if f.lower().endswith(".svg"):
+        outf = f.replace(".svg",".pdf").replace(".SVG",".pdf")
+        os.system("inkscape -z --export-area-drawing --export-margin=5 --export-pdf=%s %s"%(outf,f))

### 2019-12-23 Response-NAR-Round1.tex
Updates from Overleaf

@@
 
-	 Of course we have carried out control experiments without any modification reageant for each of the reported experiments, as now explicitly specified in the material and methods section.
-	 \TODOBruno{J'écris ça, mais je suis plus trop sûr, tu peux vérifier: \\ \didy{} was also probed with DMS (DiMethylSulfate) and CMCT, this time only in presence of Mg$^\text{2+}$, resulting in a total of 16 \Def{probing conditions}, a term we use in the following to denote a combination of probing technology, reagent and presence/absence of Mg$^\text{2+}$ (+ sequencing technology).
-For each probing condition, three experiments were performed in presence/absence of the reagent and in a denatured setting, following classic protocols~\cite{Wilkinson2006,Smola2015}.}}
+	 Of course we have carried out control experiments without any modification reageant for each of the reported experiments, as now explicitly specified in the material and methods section.}
 
@@
 	and commit to its continued accessibility (but will also pursue public venues for better guarantees of a continued availability). 
-	\TODO{Add minimal page and reactivities, including scripts/instructions to compute reactivities.}
-
+	\TODO{Add minimal page and reactivities, including scripts/instructions to compute reactivities.\\ Bruno: non je pense que c'est inutile.
+Les reactivité sont déja données en supplémentary matérial, pour ce qui est de donner un moyen de les calculer, tout est dans l'article de weeks sur shappemapper qu'on cite.
+Par contre il faut déposer le script qui permet de traiter les fichier ion-torrent si ce n'est déja fait}
 	}
@@
 		The performance of the algorithm using this probabilistic potential for SHAPE, DMS, and CMCT reactivities is given in Table S6 and is identical to
-		results obtained for the standard potential with slope ( m ) and intercept b ) optimi zed through grid
+		results obtained for the standard potential with slope $m$ and intercept $b$ optimized through grid
 		search .}
 		We also note the relative stagnation of performances of state-of-the-art tools integrating SHAPE data, despite an increasing sophistication of the pseudo-free-energy models.
-So, while we believe progress can be made in the integration of reactivities, we do not expect them to radically change the quality of predicted models, nor to invalidate the findings of this manuscript.
+So, while we believe progress can be made in the integration of reactivities, we do not expect them to radically change the quality of predicted models, nor to invalidate the findings described in this manuscript.
 
@@
 
-	\Answer{\TODO{Clarify in manuscript what conditions means in the context of the study}}
+	\Answer{
+	  In the revision, we clarified several times, including at the very beginning of the Methods section, that the term "condition" encompasses a wide variety of contexts (reagent, probing technology, ionic concentrations, and sometimes even structurally-homologous mutants) leading to the production of a reactivity profile.
+	}
 
@@
 		In the future, for the sake of reproducibility, we also wish to report, and letting the user specify, a \Def{random seed}, which will  allow the exact reproduction of an analysis.
-However, such a feature is currently not supported by \Software{RNAsubopt}, so we are working towards its inclusion to the \Software{ViennaRNA} package.
+However, random seeds are currently not supported by \Software{RNAsubopt}, so we are working towards their inclusion within the \Software{ViennaRNA} package.
 
@@
 	\Answer{The significance (or lack thereof) of claims in the context of RNA benchmarks is, of course, a legitimate concern, and we thank the reviewer for pointing it out.
-The way it is used in \cite{Spasic2017} should probably be taken with a grain of salt due to the dubious obedience of produced data to classic assumptions (\emph{eg} normality of distribution), but it nevertheless represents the best option available at the time.
+The way it is used in \cite{Xu2012} should probably be taken with a grain of salt due to the dubious obedience of produced data to classic assumptions (\emph{eg} normality of distribution), but it nevertheless represents the best option available at this time.
 
@@
 
-	\Answer{
-		\TODOBruno{Did we produce replicates?}
+	\Answer{We understand those reservations but, again, we need to stress that we already  {\bfseries explicitly describe this study as a case study}, and do not make any claim regarding the superiority of a given technology or probe.
+
+	We also did reproduce in triplicate all of our SHAPE CE probing experiments (but not our SHAPE-MaP ones due to the relatively high cost and length of those experiments in our wet lab), and observed minor variations at the reactivity level, leaving us to only consider a single representative (average of available reactivities) for each experiment.
+We include in this revision the reactivities measured over each of the 3 replicates for each condition.
 	}

### 2019-12-23 IPANEMAP-2019.tex
Updates from Overleaf

@@
 and
-$^{2}$\,CitCoM, Cibles th\'erapeutiques et conception de M\'edicaments, CNRS, Universit\'e de Paris, 4 avenue de l'observatoire, 75006 PARIS, France}
+$^{2}$\,CNRS UMR 8038, CitCoM, Universit\'e de Paris, 4 avenue de l'observatoire, 75006 Paris, France}
 % Affiliation must include:
@@
 
-\noindent We introduce \OurTool{}, an automated method for the computational modeling of RNA structure from multiple probing reactivity profiles.
-Input profiles can result from experiments based on diverse protocols, reagents, or even a collection of mutants, and are jointly analyzed to predict the dominant conformations of an RNA.
-
-\noindent \OurTool{} combines sampling, unsupervised learning, and multi-optimization, to produce secondary structure models that are both stable and well-supported by experimental evidences.
-The analysis of several reactivity, both publicly available and specifically produced experimentally, demonstrates the good performances of \OurTool, even in a mono probing setting.
-It confirms the potential of integrating multiple sources of probing data, leading to recommendations for the experimental design of informative probing assays.
+\noindent We introduce \OurTool{}, an automated method for the modeling of RNA structure from multiple probing reactivity profiles.
+Input profiles can result from experiments based on diverse protocols, reagents, \Rev{or collection of variants}, and are jointly analyzed to predict the dominant conformations of an RNA.
+
+\noindent \OurTool{} combines sampling, \Rev{clustering}, and multi-optimization, to produce secondary structure models that are both stable and well-supported by experimental evidences.
+The analysis of multiple reactivity profiles, both publicly available and produced in our study, demonstrates the good performances of \OurTool, even in a mono probing setting.
+It confirms the potential of integrating multiple sources of probing data, informing the design of informative probing assays.
 
@@
 
-Reagents used within \Rev{selective 2 -hydroxyl acylation analyzed by primer extension (SHAPE)~\cite{Wilkinson2006} protocols} represent a popular class of small molecules.
+Reagents used within \Rev{selective 2' -hydroxyl acylation analyzed by primer extension (SHAPE)~\cite{Wilkinson2006} protocols} represent a popular class of probes.
 They react with the 2’-hydroxyl of flexible ribose~\cite{McGinnis2012}, although the exact properties observed by SHAPE remain the object of ongoing investigations~\cite{McGinnis2012,Sexton2017,Hurst2018,Mlynsky2018,Frezza2019,Busan2019}.
-As ribose flexibility is proportional to the degree of freedom of the nucleotide, it is assumed that SHAPE reagents discriminate nucleotides in stable interactions from others.
+As ribose flexibility is proportional to the degree of freedom of the nucleotide, it is assumed that SHAPE reagents discriminate nucleotides involved in stable interactions.
 Different SHAPE reagents are endowed with different dynamics, that can help differentiate Watson-Crick base pairs from more dynamic tertiary contacts~\cite{Gherghe2008,Steen2012,Rice2014,Busan2019}.
 Other chemical probes, harbouring different chemical reactivities, have been developed.
-Amongst the most popular, DiMethyl Sulfate (DMS) is a small molecule that methylates N1 of adenines and N3 of cytosine if they are not involved in a hydrogen bond.
-Similarly, 1-cyclohexyl-3-(2-morpholinoethyl)carbodiimide metho-p-toluene sulfonate (CMCT) reacts on the Watson-Crick face of guanosine (N1) and uracyl (N3) when not engaged in a hydrogen bond~\cite{Ehresmann1987,Brunel2000}.
-These two reagents therefore mostly reveal Watson-Crick base pairing, but also any other type of contact involving the same edge.
-This diversity of probes not only increases coverage of the different positions and structural contexts in an RNA, but it also provides different qualitative information\Rev{, including in an \emph{in vivo} setting~\cite{Wang2019}}.
-It has been widely used since the very early days of RNA structure probing~\cite{Moazed1986, Ehresmann1987, Romaniuk1988, Butcher1994, Brunel2000,  Cordero2015, Somarowthu2015, Gross2017}.
+Amongst the most popular, DiMethyl Sulfate (DMS) is a small molecule that methylates Adenines and Cytosines if not involved in a hydrogen bond, while   CMCT reacts with the Watson-Crick face of unpaired Guanosines and Uracyls~\cite{Ehresmann1987,Brunel2000}.
+These two reagents not only reveal Watson-Crick base pairing, but also other types of contacts involving the same edge.
+The diversity of probes\Rev{, some of which are usable \emph{in vivo},} not only increases coverage of the different positions and structural contexts, \Rev{but also provides different qualitative information.
+\Rev{Modeling can also benefit from a joint analysis of reactivity profiles of single-point mutants, assuming structural homology~\cite{Cordero2012a}}.
+Integration of multiple sources of probing to improve structure prediction  has thus been widely used since the very early days of RNA structure probing~\cite{Moazed1986, Ehresmann1987, Romaniuk1988, Butcher1994, Brunel2000,  Cordero2015, Somarowthu2015, Gross2017} but, to the best of our knowledge, has never been fully automated.}
 
@@
 Computationally, the reactivity of a nucleotide is typically used as a proxy \Rev{to assess the unpaired nature of individual nucleotides}.
-The last couple of decades have nevertheless seen a series of paradigm shifts in the ways probing information is integrated, somehow mirroring the evolution of {\em ab initio} methods for secondary structure prediction.
+The past couple of decades have nevertheless seen a series of paradigm shifts in the ways probing information is integrated, somehow mirroring the evolution of {\em ab initio} methods for secondary structure prediction.
 The seminal work of Mathews~\cite{Mathews2004} used cutoffs to transform reactivity values into hard constraints.
-Depending on the reagent/enzyme, significantly unreactive or reactive positions were forced to either paired or remain unpaired respectively within predicted structures.
-However, such \emph{hard} constraints turned out to be very sensitive to the choice of the cutoff, easily leading to artificially unpaired predictions or unsatisfiable sets of constraints.
-With the advent of SHAPE probing, subsequent methods lifted the requirement of thresholds and integrated reactivities directly as pseudo-energies, supplementing the Turner nearest-neighbor energy model~\cite{Turner2010} within an (pseudo) energy minimization scheme~\cite{Deigan2009,Zarringhalam2012,Washietl2012}.
-Such methods can be seen as optimizing a tradeoff between the thermodynamic stability of a model, assessed by the Turner energy model, and its compatibility with the reactivity profile.
-As a consequence, popular packages for secondary structure prediction, such as \Software{RNAStructure}~\cite{Mathews2004} or the \Software{Vienna} package~\cite{Lorenz2011}, now routinely accept SHAPE-like reactivity profiles as input of their main structure prediction methods.
+Depending on the used reagent/enzyme, significantly unreactive or reactive positions were forced to remain paired or  unpaired within predicted models.
+However, such \emph{hard constraints} can be overly  sensitive to the choice of the cutoff, leading to artificially unpaired predictions or unsatisfiable sets of constraints.
+With the advent of SHAPE probing~\cite{Smola2015}, new methods~\cite{Deigan2009,Zarringhalam2012,Washietl2012} lifted the requirement of a threshold, supplementing  Turner nearest-neighbor energy model~\cite{Turner2010} with pseudo-energies derived from the reactivities.
+They  then performed a (pseudo) energy minimization, optimizing a tradeoff between the thermodynamic stability and its compatibility with the reactivity profile.
+Popular packages for secondary structure prediction, such as \Software{RNAStructure}~\cite{Mathews2004} or the \Software{Vienna} package~\cite{Lorenz2011}, now routinely accept SHAPE-like reactivity profiles as input of their main structure prediction methods.
 \Rev{Besides free-energy optimization, such methods notably} include the joint folding of aligned RNAs~\cite{Lavender2015}, partition function, statistical sampling and Maximal Expected Accuracy predictions~\cite{Spasic2017}\ldots
@@
 \subsection*{The \OurTool{} method}
-We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signals produced by various chemical probes \Rev{to predict one or several secondary structure models for a given RNA.}
-
-It performs a structural clustering across multiple sets of structures sampled in different experimental conditions, and ultimately returns a set of structures that are representative of the dominant conformations supported across conditions.
+We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signals produced by various \Rev{probing experiments to predict one or several secondary structure models for a given RNA.
+It takes as input one or several reactivity profiles produced in various \Def{conditions}, broadly defined to denote the conjunction of a reagent, a probing technology, ionic concentrations and, in extreme cases, structurally homologous mutants.}
+
+It performs a structural clustering across multiple sets of structures sampled in different experimental conditions, and ultimately returns a set of structures representing dominant conformations supported across conditions.
 Its underlying rationale is that the prominent presence of a stable secondary structure within the  (pseudo-)Boltzmann ensembles induced by multiple experimental conditions should increase its likelihood to be (one of) the native structure(s) for a given RNA.
@@
 \didy{} was probed with different SHAPE reagents: 1M7 (1-methyl-7 nitrosatoic anhydre), NMIA (N-Methyl Isatoic Anhydre), BzCN (Benzoyl cyanide) and NAI (2-methylnicotinic acid imidazolide) in presence and absence of Mg$^\text{2+}$.
-\Rev{\didy{} was also probed with DMS (DiMethylSulfate) and CMCT, this time only in presence of Mg$^\text{2+}$, resulting in a total of 16 \Def{probing conditions}, a term we use in the following to denote a combination of probing technology, reagent and presence/absence of Mg$^\text{2+}$ (+ sequencing technology).
+\Rev{\didy{} was also probed with DMS (DiMethylSulfate) and CMCT, in presence of Mg$^\text{2+}$, resulting in a total of 16 \Def{probing conditions}, a term we use in the following to denote a combination of probing technology, reagent and presence/absence of Mg$^\text{2+}$ (+ sequencing technology).
 For each probing condition, three experiments were performed in presence/absence of the reagent, and in a denatured context, following classic SHAPE protocols~\cite{Wilkinson2006,Smola2015}.}
@@
 	\caption{Dot-bracket representations of the native structure and \OurTool{} predictions for the glycine riboswitch sequence of the Cordero\etal\cite{Cordero2012} data set, supplemented with an increasing collection of probing conditions.
-Positions with green and red backgrounds indicate correctly and incorrectly predicted base pairs.}\label{fig:glycine_example}
+\Rev{Please note that, throughout the manuscript, by "conditions" we mean a combination of the nature of the probe, the probing technology, the ions present, the temperature, or any other "environmental" variation} Positions with green and red backgrounds indicate correctly and incorrectly predicted base pairs.}\label{fig:glycine_example}
 \end{table*}
@@
 Interestingly, the dispersion of the MCC was lower for WT+2M than for WT+1M, with lower MCC values above 57\%.
-This results suggest  a role of \emph{tie-breaker} for the third condition, allowing to mitigate the deleterious effect of some outlier.
+This results suggest \Rev{adding a second mutant profile has  a role of \emph{tie-breaker}}, allowing to mitigate the deleterious effect of some outlier.
 However, outliers still played a disruptive role, and restricting chosen mutants to the 20 reactivity profiles most similar to the WT further improved the average MCC 2.8\% to 67.7\%,  \Rev{ a significant improvement compared to WT+1M (P-value$=2.4\times10^{-2}$).}

### 2019-12-26 Response-NAR-Round1.tex

@@
 \tableofcontents
-\listoftodos
 
@@
 First of all, the authors would like to express their deepest gratitude for the quality and thoroughness of the reports written by the reviewers, covering a remarkably comprehensive range of aspects of the submitted manuscript.
-We have individually addressed each of the \ref{comment:last} comments to the best of our ability, but would like to start this note with higher-level answers to recurrent questions:
-
-\TODO[Yann]{Faire résumé}
-Points géneraux :
+We have individually addressed each of the \ref{comment:last} comments extracted from the first round of review, each to the best of our ability, but would like to start this response with higher-level answers to recurrent remarks:
 \begin{itemize}
-\item On a probablement pas été super clairs, mais :
-Pas de comparaison entre protocoles, plutôt un test, rendu possible par IPANEMAP de la complémentarité des protocoles
-\item Pas optimisation du traitement des donnés de sondage, mais traitement uniforme
-\item Approche multi probing nous parait susceptible d’être robuste à l’incorporation d’une réactivité moins informative.
+	\item It was not within the scope of this research to establish/compare the intrinsic predictive potentials of individual probing reagents/protocols, and we did not make any specific claim in that respect.
+We understand that some unfortunate presentation choices of  our part may have suggested stronger stances than intended (\emph{e.g.} giving the impression that the performances of SHAPE CE are representative of SHAPE technologies based on RT stops).
+This was not our intention, or message, and we have attempted in this revision  to correct any such misleading statement, and overall optics;
+	\item Nevertheless, all of our analyses, including data produced specifically for the validation of the method, support the notion than combining multiple (3+) probing experiments mitigates the risks of mispredictions due to outliers, and frequently leads to significant improvements as assessed by standard statistical tests suggested by Reviewer \#4;
+	\item To the best of our knowledge, \OurTool currently represents the only general method that addresses the automatic modeling of RNA structure from multiple sources of probing data.
+\OurTool also seems to represent a competitive option for mono-probing predictions, but this observation is merely an element of validation, and (partially) substantiates the hope of an expected good quality of predictions in a multiple setting;
+	\item We prefer to process various sources of probing data homogeneously rather than optimize our performances (potentially artificially due to overfitting over scarce data) by learning reagent/protocol specific conversions of reactivities to pseudo-energies.
+Accordingly, our method only relies on a preliminary calibration with respect to the parameters $m$ and $b$ in the \cite{Deigan2009} pseudo-energy conversion model, which proved robust enough for us not to question this initial choice.
 \end{itemize}
+
+Over the course of this extensive round of revisions, we have:
+\begin{itemize}
+  \item Rewritten extensively our manuscript, notably including the bulk of our method section;
+  \item Assessed the significance of our main quantitative claims as per \cite{Xu2012}, and included a new paragraph describing the methodology;
+  \item Performed additional tests to ascertain the quality of reactivities produced for \didy: We analyzed the correlations of reactivity profiles at the reactivity level (Supp.
+Sec.
+3.2); We also the sigifnicantly better agreement for all produced reactivity profiles than expected at random (Supp.
+Sec.
+3.1);
+  \item Improved the distribution, documentation, and streamlined the installation of \OurTool (although we could not reach a \Software{bioconda} distribution yet);
+  \item Fixed numerous typos and unfortunate turns of sentences.
+\end{itemize}
+
+We hope that, upon reviews, the editor and reviewers will share our enthusiasm for \OurTool, and consent to its dissemination to the wide readership of the journal.
 
@@
 	These histograms are very similar to those deemed as exemplary of a successful experiment by \cite{Smola2015},. 
-	In addition, those SHAPE-MaP results are consistent with the X-ray structure (see Supplementary Figure S2 and S3), and their use with our prediction software was shown to improve the accuracy of the obtained model(s).
+	In addition, those SHAPE-MaP results are consistent with the X-ray structure (see Supplementary Figure S2 and S3), and their use with our prediction software was shown to improve the accuracy of the output model(s).
 
@@
 	and commit to its continued accessibility (but will also pursue public venues for better guarantees of a continued availability). 
-	\TODO{Add minimal page and reactivities, including scripts/instructions to compute reactivities.\\ Bruno: non je pense que c'est inutile.
-Les reactivité sont déja données en supplémentary matérial, pour ce qui est de donner un moyen de les calculer, tout est dans l'article de weeks sur shappemapper qu'on cite.
-Par contre il faut déposer le script qui permet de traiter les fichier ion-torrent si ce n'est déja fait}
-	}
-
+	}
 	\end{enumerate}
@@
 	\end{itemize}
+	Additional tests can be found in the manuscripts, along with a new section that describes the precise tests carried out and their rationale, further substantiating our claims.
 	}
@@
 We wish we could do more (e.g.
-test on a wider dataset) but this is simply unrealistic, and way beyond the standards enforced for the dissemination of a new {\bf method} in structural bioinformatics.
+test on a wider dataset) but given the diversity of probes, a systematic quantitative assessment of individual probes is simply unrealistic, and way beyond the standards enforced for the dissemination of a new {\bf method} in structural bioinformatics.
 
@@
 
-	\Answer{We understand those reservations but, again, we need to stress that we already  {\bfseries explicitly describe this study as a case study}, and do not make any claim regarding the superiority of a given technology or probe.
-
-	We also did reproduce in triplicate all of our SHAPE CE probing experiments (but not our SHAPE-MaP ones due to the relatively high cost and length of those experiments in our wet lab), and observed minor variations at the reactivity level, leaving us to only consider a single representative (average of available reactivities) for each experiment.
-We include in this revision the reactivities measured over each of the 3 replicates for each condition.
+	\Answer{We understand those reservations but, again, we need to stress that we already  {\bfseries explicitly describe this study as a case study}, and do not make any claim regarding the superiority of a given technology or probe from this study.
+
+	We also did reproduce in triplicate all of our SHAPE CE probing experiments, and observed minor variations at the reactivity level, leaving us to only consider a single representative (average of available reactivities) for each experiment.
+We now include in this revision the reactivities measured over each of the 3 replicates for each condition.
+For SHAPE-MaP protocols,we could not duplicate them due to the relatively high cost and length of those experiments in our wet lab.
 	}
@@
 	\Answer{We agree, and in fact do not draw any conclusion from these results.
-We only report those values as references, to support the hypothesis that the MCCs achieved by \OurTool are probably more representative of the reactivity profiles/condition than of a specific predictive model.}
-
-	\Comment{Additionally, there is strong A/C bias in DMS data obtained by stop-based protocols. I don’t know if it was accounted for during normalization step, if normalization was indeed applied. }
-
-	\Answer{
-		We were unsure how to interpret this remark: \Ab{} and \Cb{} are indeed the only reactive nucleotides, so the normalization step is indeed restricted to positions featuring those two nucleotides.
+We only report the MEA/MFE values as references, to support the hypothesis that the MCCs achieved by \OurTool are probably more representative of the reactivity profiles/condition than of a specific predictive model.}
+
+	\Comment{Additionally, there is strong A/C bias in DMS data obtained by stop-based protocols. I don’t know if it was accounted for during normalization step, if normalization was indeed applied.}
+
+	\Answer{We were unsure how to interpret this remark: \Ab{} and \Cb{} are indeed the only reactive nucleotides, so the normalization step is indeed restricted to positions featuring those two nucleotides.
 If the reviewer is referring to the observations of \cite{Sexton2017} of opposite {\em \Ab{} vs \Cb} biases for RT-based and mutations-based probing respectively, we did not account for it.
@@
 However, despite its stochastic foundations, \OurTool{} is typically stable in its predictions, as can be seen in Figure~\ref{fig:reproducibility} for all probing datasets produced over our \didy case study.
+
 			Interestingly, the most notable exception to the general stability is observed for the probing-free prediction, consistent with the widespread interpretation of SHAPE-induced pseudo-potentials as focusing predictions onto a subset of the Boltzmann ensemble.
@@
 
-	\Answer{\TODOAfaf{Can you confirm that a similar normalization was used?}}
+	\Answer{
+	All reactivities but the \cite{Cordero2012a} were renormalized using the 2\%-8\% approach of \cite{Deigan2009}.
+	For \cite{Cordero2012a}, we used the data directly as we found the reactivities to follow the expected distribution.
+
+	However it is important to note that our analysis does not rely on a direct differential analysis at the reactivity level, except in our newly added Supplementary Section 3.2.
+The latter features a cross-analysis based on the Pearson correlation, precisely chosen as its results are impervious to affine transforms of the data.
+}
 

### 2019-12-26 IPANEMAP-2019.tex
Updates from Overleaf

@@
 	{\centering
-			\begin{adjustbox}{max width=\linewidth}
+			\begin{adjustbox}{max width=.8\linewidth}
 		\begin{tabular}{@{}lccHHH}
@@
 		\end{tabular}
-		\end{adjustbox}}
+		\end{adjustbox}\\}
 
@@
 			\multicolumn{4}{@{}l@{}}{}                                    & \multicolumn{4}{@{}c@{}}{MCC\%}\\
-			\multicolumn{4}{@{}l@{}}{Conditions in individual clusters}                                    & \MCB{-NAI} & \MCB{+NAI} \\
+			\multicolumn{4}{@{}l@{}}{Conditions in individual clusters ($+\mathcal{M}$)}                                    & \MCB{-NAI} & \MCB{+NAI} \\
 			\Bull{A} & \Bull{C} & \Bull{E} & \Bull{G}   & \MC{\Q Joint} & \MC{\Q Mono}  & \MC{\Q Joint}       & \MC{\Q Mono}      \\ \midrule
@@
 	\end{adjustbox}\\
-	\caption{Predictive performances of conditions across clusters, coupled with impact assessment of NAI.
-For each experiment, a single condition is chosen from each cluster except for cluster \Bull{B}.
-Clusters \Bull{D}, \Bull{F} and \Bull{H}, each consisting of a single condition, contribute their sole condition, leading to the list of conditions  $\mathcal{M}$:=\{\OneMSevILUThree, \BzCNMg and \DMSMg\}.
-		For each combination, the MCC of the predicted structural model is reported, as predicted from the conditions in the presence/absence of \NAICE ('Joint' columns).
-For each prediction, the average MCC value of mono conditions is reported as a baseline ('Mono' columns).
+	\caption{Predictive performances of conditions across clusters.
+In each line, a single condition is chosen from each cluster except for cluster \Bull{B}.
+\Rev{Clusters consisting of a single condition (\Bull{D}, \Bull{F} and \Bull{H}) contribute additional conditions  $ :=\{\OneMSevILUThree, \BzCNMg, \DMSMg\}$.
+		The MCC of the predicted structure is reported ('Joint' columns) and compared to the average MCC of mono conditions, in all cases in presence/absence of \NAICE.}
 	}\label{fig:differentconds}

### 2020-06-29 Response-NAR-Round2.tex
Updates from Overleaf

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[9pt,hyperref]{article} % use larger type; default would be 10pt
\setcounter{secnumdepth}{1}

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%\%\% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%\%\% PAGE DIMENSIONS
\usepackage[margin=1.3in]{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} \% for example, change the margins to 2 inches all round
% \geometry{landscape} \% set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} \% Activate to begin paragraphs with an empty line rather than an indent

%\%\% PACKAGES
\usepackage{hyperref}
\hypersetup{
	colorlinks, linkcolor={red!80!black},
	citecolor={blue}, urlcolor={red!50!black}
}

\usepackage[authoryear]{natbib}
\usepackage{amsmath} % for much better looking tables
\usepackage{xspace}
\usepackage{url}
\usepackage{soul}
\usepackage[table,x11names,dvipsnames]{xcolor}
\usepackage{amssymb} % for much better looking tables
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{ntheorem} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
\usepackage{mdframed}
\usepackage{relsize}
\usepackage{pgf}
\usepackage{cabin}
\usepackage{enumitem}
\usepackage{tgtermes}
\usepackage{collcell}
\usepackage{adjustbox}% to resize figures

%\usepackage{tgadventor}
% These packages are all incorporated in the memoir class to one degree or another...

%\%\% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%\%\% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%\%\% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

\newcommand{\FAnswer}[1]{{\color{black!95}#1}}
\newcommand{\IAnswer}[1]{\FAnswer{\hfill $\to$ #1}}
\newcommand{\PAnswer}[1]{\FAnswer{\\#1}}

\newcommand{\Answer}[1]{\noindent{{{\sffamily Response:} }{\FAnswer{#1}}}}
\newcommand{\Comment}[1]{\item{\color{red!40!black}{{\sffamily Comment:} }{#1}}}
\newcommand{\Quote}[1]{\begin{mdframed}[backgroundcolor=black!2,linecolor=gray!50, linewidth=.6mm]#1\end{mdframed}}

\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\TODO}[2][All]{{\todo[color=blue!10,inline]{\color{black}{TODO (#1)}: #2}}}
\newcommand{\TODOBruno}[2][Bruno]{{\todo[color=red!10,inline]{\color{black}{TODO (#1)}: #2}}}
\newcommand{\TODOAfaf}[2][Afaf]{{\todo[color=Aquamarine3,inline]{\color{black}{TODO (#1)}: #2}}}

%\%\% END Article customizations

%\%\% The "real" document content comes below...

%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%
%\%                                          \%\%
%\% Enter the authors here                   \%\%
%\%                                          \%\%
%\% Specify information, if available,       \%\%
%\% in the form:                             \%\%
%\%   <key>={<id1>,<id2>}                    \%\%
%\%   <key>=                                 \%\%
%\% Comment or delete the keys which are     \%\%
%\% not used. Repeat \author command as much \%\%
%\% as required.                             \%\%
%\%                                          \%\%
%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%

\input{macros.tex}

\title{Response to second round of reviews\\[.3em]\OurTool{}:  Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles}
\author{
Afaf Saaidi \and
Delphine Allouche \and
Mireille Regnier \and
Bruno Sargueil \and
Yann Ponty}

\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

\tableofcontents

\section{General comments}

\section{Specific comments}

	\subsection{Reviewer \#2}
	\begin{enumerate}[resume]
\Comment{The revised manuscript is much clearer now. However, I am still puzzled over the performance comparison between IPANEMAP and Rsample over Hajdin et al.’s data. The Hajdin dataset consists of 24 RNAs, but only 16 of them were considered in the Rsample paper (per its supplementary information). As I previously commented, IPANEMAP’s performance gains are really dramatic for the few sequences that were not considered by Spasic et al. For example, 5SrEcoli, signal recognition particle, and E. coli tRNA(phe).

Nevertheless, I also looked at a few of the sequences considered by both Spasic et al. and the authors. Below are a few examples of inconsistencies between these two benchmarks:

[..]

- Can the authors reconcile these discrepancies, or alternatively, am I missing something in the way this benchmark was set up?

Please note the following:
A. I did not check all the RNAs benchmarked by Spasic et al. so there might be additional discrepancies. 
B. Reported values were consistent between the two works for some RNAs which I did review. 
C. Given these discrepancies, I am not sure what to conclude from the performances reported for those sequences benchmarked in this manuscript but not by Spasic et al.

Maybe the predictions of these RNAs are very sensitive to the thermodynamic model parameters, I don’t know. If this is indeed the case, it is a good idea to benchmark IPANEMAP against more than a single method, similarly to what Spasic et al. did. I think this creates a more balanced picture of current structure prediction capabilities.}

    \Answer{We are greatly indebted to this reviewer for his/her}

	\Comment{Additionally, referring to the authors comment on page 24: “we object to such an adversarial view of this particular analysis (and, more generally, our work)…”  I would like to clarify I had no such intention. My critique aimed at strengthening or solidifying the reported results. I am sorry the authors interpreted my requests for better controls and comparisons as a negative opinion of their work.}

	\Answer{Actually, in addition to being wrong regarding the benchmark, we expressed ourselves poorly in this instance: Having done more than our fair share of reviewing and editorial activities...}
	\end{enumerate}

\section{Referee: 3}

	\subsection{Minor Revisions:}
		\begin{enumerate}[resume]
	\Comment{On page 4, in “Cordero dataset”, the text states that the reactivity is set to 0 for nucleotide identities that are unreactive to a specific probing agent.  But 0 is a reactivity that receives a negative pseudoenergy free change, encouraging base pairing.  It would be much better to give a 0 pseudo free energy change.  Is it possible the manuscript should really state a 0 free energy change?  If not, the authors should add some text to justify this choice.}

    \Answer{WE set the pseudo-energy to 0 (and not the reactivity, as erroneously stated in the manuscript). \TODO{correct}}

	\Comment{On page 4, the text states “a standard denaturation protocol”, but I think the manuscript should state “a standard annealing protocol”.  It looks like the intent was to anneal the structure to find its equilibrium structure at 37 degrees C, not to leave the structure denatured.}

    \Answer{Indeed, real denomination is denaturation/renaturation protocol, currently the standard nomenclature. \TODO{correct}}
	\Comment{On page 5, in statistical significance, “usual significance threshold” should really state the “usual type I error rate”.  Alpha is the probability of falsely rejecting the null hypothesis (type I error).  “Threshold” doesn’t convey this meaning.  }

    \Answer{\TODO{correct}}
	\Comment{On page 8, “barely significant” should just be changed to “but significant”. The improvements may be modest, but they are significant.  I don’t agree that something can be “barely” significant.}

    \Answer{\TODO{be polite}}
	\Comment{On page 10, “Considering three conditions yields improvement over the average performance of the triplet” confuses me.  I think what was intended is “Considering three conditions yields improvement over the average performance of the doublet [/pair]” }
    \Answer{\TODO{The average performance of prediction informed by individual conditions is lower than that of predictions simultaneously informed by the three conditions}}

	\subsection{Typos:}
	\Comment{On page 1, “different dynamics, that can help” should be “different dynamics that can help”.}

	\Answer{}
	\Comment{On page 3, “using an iterative heuristics” should be “using an iterative heuristic”.  Also, “we elect the most promising cluster” should be “we select the most promising cluster”.}

	\Answer{}
	\Comment{On page 6, “Spacic” should be “Spasic”.}

	\Answer{}

	Suggestions:
	\Comment{On page 1, there is a new phrase about probes “some of which are usable in vivo”.  My recollections is that Zaug and Cech were the first to demonstrate in vivo probing of RNA structure, and it is probably worth citing the paper [RNA. 1995 1:363.].}

	\Answer{\TODO{Ok Paillard PMID 15355993}}
	\Comment{Page 1 claims that using multiple data sources has never been fully automated.  I agree that it has never been automated when using quantitative data and soft constraints.  On the other hand, the deprecated hard constraint method (reference 2) was fully automated and used multiple probing methods.  The phrase should be made more specific.}

	\Answer{\TODO{}}

	\end{enumerate}

\bibliographystyle{abbrvnat}
\bibliography{biblio}
\end{document}

### 2020-06-29 Response-NAR-Round1.tex
Updates from Overleaf

@@
 Sec.
-3.2); We also the sigifnicantly better agreement for all produced reactivity profiles than expected at random (Supp.
+3.2); We also found significantly better agreement, with the reference structure of \cite{Meyer2014}, of all produced reactivity profiles than expected at random (Supp.
 Sec.
@@
 
-We hope that, upon reviews, the editor and reviewers will share our enthusiasm for \OurTool, and consent to its dissemination to the wide readership of the journal.
+We hope that, upon reviews, the editor and reviewers will share our enthusiasm for \OurTool, and consent to its dissemination to the wide readership of Nucleic Acids Research.
 
@@
 		\end{itemize} having respectively a similar reagent, technology, or both, as a common trait.
+However, neither of those determinants seem universally supported by the observation of the correlation matrix.
 
@@
 	\Answer{
-	All reactivities but the \cite{Cordero2012a} were renormalized using the 2\%-8\% approach of \cite{Deigan2009}.
-	For \cite{Cordero2012a}, we used the data directly as we found the reactivities to follow the expected distribution.
-
-	However it is important to note that our analysis does not rely on a direct differential analysis at the reactivity level, except in our newly added Supplementary Section 3.2.
-The latter features a cross-analysis based on the Pearson correlation, precisely chosen as its results are impervious to affine transforms of the data.
+	All reactivities were (re)normalized using the boxplot of \cite{Deigan2009}.
+
+	It is important to clarify that our analysis does not rely on a direct differential analysis at the reactivity level, except in our newly added Supplementary Section 3.2.
+The latter features a cross-analysis based on the Pearson correlation, precisely chosen as its results are impervious to preprocessing consisting of an affine transforms of the data.
 }
@@
 	\Answer{We agree that SHAPE CE should not be mischaracterized as a representative for all SHAPE protocols based on RT stops (notoriously including SHAPE-Seq).
-Our description was misleading in that respect, so we modified and removed the corresponding column in Table 2.}
+Our descriptions were sometimes misleading in that respect, so we modified and removed the corresponding column in Table 2 and tried to remove any sentence that would involuntarily suggest the superiority of a technology/reagent based on our observations.}
 
@@
 		We were happy with the overall reproducibility of our results with 1~000 structures, which has been the standard practice in the field since the original \cite{Ding2003} paper.
-Moreover, our computational requirements scale quadratically with the number of structure, so we recommend to stay with this number, which represents in our opinion a good tradeoff between reproducibility and efficiency.
-
-		To address this question more precisely, we reexecuted the \didy benchmark using a sample size of 2~000, and observed a minor non-significant variation in the MCC (ie the results did not allow to refute the hypothesis of equal expected MCCs for both experiments).}
+Moreover, our computational requirements scale quadratically with the number of structure, so we recommended a sample size of 1~000, which represents in our opinion a good tradeoff between reproducibility and efficiency.
+
+		To address this question more precisely, we reexecuted the \didy benchmark using a sample size of 2~000, and observed a minor non-significant variation in the MCC (\emph{i.e.} the results did not allow to refute the hypothesis of equal expected MCCs for both experiments).}
 

### 2020-06-29 IPANEMAP-2019.tex

@@
 A dataset was gathered by Hajdin\etal\cite{Hajdin2013} to validate the predictive capacities of probing data-driven predictions.
-It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (1M7-\SH) is available.
-This dataset includes sequences originating from a variety of organisms, and spans lengths ranging from $34$ nts to $530$ nts, with a focus on riboswitches and complex RNA architectures.
+It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (1M7-\SH) was made available.
+This dataset includes sequences originating from a variety of organisms, and spans lengths ranging from $34$ nts to $530$ nts, with a focus on riboswitches and complex RNA architectures \Rev{(full list in Supp.
+Table S3)}.
 
@@
 {\sf G} and {\sf U}), setting the reactivity to 0 for other positions.
-This assumption allowed to decrease the noise generated by reactivities associated with non-targeted nucleotides, leading to more accurate predictions (data not shown).
+This allowed to decrease the noise generated by reactivities associated with non-targeted nucleotides, leading to more accurate predictions (data not shown).
 
@@
 
-The Matthews Correlation Coefficient (MCC) is a classic metric for assessing the quality of a predicted structure $S$, identified by a set of base pairing positions, based on its agreement with an accepted reference (native) structure $\Ref$.
-It can be interpreted as a compromise \Rev{between the main metrics derived from the confusion matrix}, and is defined as
+The Matthews Correlation Coefficient (MCC) is a classic metric for assessing the quality of a predicted structure $S$, identified by a set of base pairing positions, compared to an accepted reference (native) structure $\Ref$.
+It represents a compromise \Rev{between the main metrics derived from the confusion matrix}, and is defined as
 $$\text{MCC}(S\mid \Ref)=\frac{\text{TP} \times \text{TN} -\text{FP} \times \text{FN}}{\sqrt[]{(\text{TP}+\text{FP})(\text{TP}+\text{FN})(\text{TN}+\text{FP})(\text{TN}+\text{FN})}},$$
-where TP, FP, TN, FN classically represent the correctly/erroneously predicted and correctly/erroneously omitted base pairs in $S$ with respect to $\Ref$.
+where TP, FP, TN, FN represent the correctly/erroneously predicted and correctly/erroneously omitted base pairs in $S$ with respect to $\Ref$.
+\Rev{MCC is a stringent metric, taking values between -100\% and 100\%, 0\% being the expected MCC of a "coin tossing" random predictor.}
 
@@
 \subsection*{Assessment of statistical significance}
-Following Xu\etal\cite{Xu2012}, we assessed the statistical significance of observed changes in predictive performances, using a two-tailed Student $t$-test with unequal variance, using a classic significance threshold $\alpha=5\%$.
+Following Xu\etal\cite{Xu2012}, we assessed the statistical significance of observed changes in predictive performances using a two-tailed Student $t$-test (unequal variance) with usual significance threshold $\alpha=5\%$.
 When comparing performances within a given dataset, we used the paired version of the test.
@@
 Intuitively, the $t$-test estimates the probability that an observed change in performance can be attributed to chance alone.
-It considers the hypothesis $H_0$ that the two distributions have equal mean and computes a \Def{P-value}, \emph{i.e.} the likelihood of the observed values under $H_0$.
-If the P-value is lower than $\alpha$, then $H_0$ is deemed \Def{refuted}, and the observed difference in average performance is considered as significant.
+It considers the hypothesis $H_0$ that the two distributions have equal mean and computes a \Def{P-value}, \emph{i.e.} the likelihood of the observed data under $H_0$.
+If the P-value is lower than $\alpha$, then $H_0$ is deemed \Def{refuted}, and the observed difference in performances is considered significant.
 
@@
 As a control experiment, we also report ${\tt RNAfold}$ predictions in presence/absence of probing data, both in energy minimization (MFE) and maximum expected accuracy (MEA) modes.
-Figure~\ref{cordero1} shows the averaged GM values for all combinations of tools and probing data (Detailed results can be found in Supp.
+Figure~\ref{cordero1} shows the averaged GM values for all combinations of tools and probing data (Detailed results in Supp.
 Tables S1 and S2).
@@
 Note that the Pareto-optimality always implies a single returned cluster/structure for \OurTool{} in a mono probing setting.}
-We computed and report in Figure~\ref{fig:Vsrsample} the individual Geometric Means (GM) associated with predicted structures.
-Sensitivity and PPV reported in Supp.
-Table S4, while Supp.
-Table S5 reports the stability of predictions across 10 runs of \OurTool.
+We computed and report in Figure~\ref{fig:Vsrsample} the individual Geometric Means (GM) of predicted structures.
+\Rev{Sensitivity and PPV can be found in Supp.
+Table S4, and Supp.
+Table S5 reports the stability of predictions over 10 independent runs of \OurTool.}
 
@@
 However, the two methods differ substantially whenever a single conformation is detected, in which case \Software{RSample} switches to a simple MEA prediction, whereas \OurTool{} always proceeds with an iterative clustering.
-This allows \OurTool{} to consider a larger number of clusters, which we believe helps eliminating structural outliers, leading to the election of a better final centroid structure.
+This allows \OurTool{} to consider a larger number of clusters, which we believe helps eliminating structural outliers, leading to the election of an improved final centroid structure.
 
@@
 As described in the Material and Methods section, we probed \didy{}  in 16 different conditions using different SHAPE and classical reagents, in presence/absence of Mg$^\text{2+}$ ions, using either the standard premature stop or the newly developed mutation induction technologies to map the modification sites.
-These experimental results were then used as input of \OurTool{}, either individually or combined,  for secondary structure prediction.
+\Rev{We observed an overall good agreement of reactivities with the 3D model proposed by Meyer~\etal\cite{Meyer2014} (see  Supp.
+Figures S1, S2 and S3).} Reactivities were then used as input of \OurTool{}, either individually or combined,  for secondary structure prediction.
 
@@
 Figure~\ref{fig:PCA} summarizes the pairwise distance, projected onto a 2D surface by a principal component analysis.
+\Rev{A systematic cross-analysis of raw reactivities revealed a positive, if modest, correlation between conditions, but did not support a clear partition of profiles based on probing technology or presence/absence of Mg$^{\text{2+}}$ (see Supp.
+Figure S4).}
 
@@
 \Rev{Interestingly the MCC value cannot be trivially anticipated from the compatibility of the reactivity scores with the predicted structure, or even with the native structure (see Supp.
-Table S6).}
+Table S1).}
  These predictive capacities clearly outperformed those achieved in the absence of probing data (51\% MCC), and can be interpreted as indicative of predictions of good overall quality.
-Indeed, it is worth reminding that the MCC is a stringent metric, taking values between -100\% and 100\%, 0\% being the expected MCC of a "coin tossing" random predictor.
 The lowest MCC value of 60\% is, in particular, equally consistent with Sensitivity/PPV values of 60\%/60\%, 80\%/46\% or 45\%/82\%.
@@
 We then computed the associated MCC, and compared it with the MCC of the worst performing condition (Min), best-performing condition (Max) and average MCC over the single conditions experiments.
-A summary of the results over pairs is shown in Figure~\ref{fig:pairwise} (see details in Supp.
-Table S7).
+A summary of the results over pairs is reported in Figure~\ref{fig:pairwise} (details in Supp.
+Table S6).
 
@@
 To test this hypothesis, we executed \OurTool{} on every subset of conditions within clusters, and report in Table~\ref{fig:similarconds} the quality (MCC) of predictions for any subset of similar conditions (see Supp.
-Figure~S1 for associated models).
+Figure~S5 for associated models).
 \Rev{Note that such an analysis does not use knowledge of the native structure, since our clustering only relies on properties of the pseudo-Boltzmann ensemble.}
@@
 	\caption{Predictive performances of conditions across clusters.
-In each line, a single condition is chosen from each cluster except for cluster \Bull{B}.
+A single condition is chosen from each cluster except for \Bull{B}.
 \Rev{Clusters consisting of a single condition (\Bull{D}, \Bull{F} and \Bull{H}) contribute additional conditions  $ :=\{\OneMSevILUThree, \BzCNMg, \DMSMg\}$.
-		The MCC of the predicted structure is reported ('Joint' columns) and compared to the average MCC of mono conditions, in all cases in presence/absence of \NAICE.}
+		The MCC of the predicted structure is reported ('Joint' columns) and compared to the average of mono conditions ('Mono' columns), in presence/absence of \NAICE.}
 	}\label{fig:differentconds}
@@
 \Rev{In this final analysis, we tested the capacity of \OurTool to extract and exploit complementary information produced by the systematic probing of single-point mutants using the Mutate-and-Map (MaM) protocol~\cite{Cordero2015}.
-The joint analysis of several mutants may appear ill-advised, as single point mutants are expected to adopt different structures than the wild-type sequence (WT) , an expectation that is at the core of downstream analyses~\cite{Cordero2015}.
-However, those changes are typically local, and the limited influence of outliers on \OurTool{} substantiates the hope that the benefits of including independently-produced conditions will outweigh the cost of noise, introduced by local changes.}
+The joint analysis of several mutants may appear error-prone, as single point mutants are expected to adopt different structures than the wild-type sequence (WT), an expectation that is at the core of downstream analyses~\cite{Cordero2015}.
+However, such changes are typically local, and the robustness of \OurTool{} to outlier conditions nourishes the hope that the benefits of including independently-produced conditions will outweigh the cost of noise introduced by local changes.}
 
@@
 We report in Figure~\ref{fig:variantanalysis} the distribution of MCC values associated with the first prediction returned by \OurTool (see Supp.
-Figure S2 for details).
+Figure S6 for details).
 
@@
 However, despite its stochastic foundations, \OurTool{} is typically stable in its predictions, as can be seen in Figure~\ref{fig:reproducibility} for all probing datasets produced over our \didy case study.
-Interestingly, the most notable exception to the general stability is observed for the probing-free prediction, consistent with the widespread interpretation of SHAPE-induced pseudo-potentials as focusing predictions onto a subset of the Boltzmann ensemble.
+Interestingly, the most notable exception to the general stability is observed for the probing-free prediction, consistent with the widespread interpretation of SHAPE-induced pseudo-potentials as \emph{focusing} the predictions onto a subset of the Boltzmann ensemble.
 A similar behavior is observed for the other datasets considered in our study (see Supp.
@@
 Remarkably, the most accurate predictions were obtained with 1M7 and NMIA \Rev{irrespectively of the presence/absence of Mg$^{\text{2+}}$, and despite the fact that reactivity profiles substantially diverged between the two conditions (see Supp.
-Fig.
-S4)}.
+Figure S4)}.
 However, for other reagents/protocols, predictions appeared positively impacted (+$10\%$ average MCC) by the presence of Mg$^{\text{2+}}$, painting a mixed picture that would deserve future investigations.

### 2020-06-29 IPANEMAP-2019.tex
Updates from Overleaf

@@
 \noindent We introduce \OurTool{}, an automated method for the modeling of RNA structure from multiple probing reactivity profiles.
-Input profiles can result from experiments based on diverse protocols, reagents, \Rev{or collection of variants}, and are jointly analyzed to predict the dominant conformations of an RNA.
-
-\noindent \OurTool{} combines sampling, \Rev{clustering}, and multi-optimization, to produce secondary structure models that are both stable and well-supported by experimental evidences.
+Input profiles can result from experiments based on diverse protocols, reagents, \RevA{or collection of variants}, and are jointly analyzed to predict the dominant conformations of an RNA.
+
+\noindent \OurTool{} combines sampling, \RevA{clustering}, and multi-optimization, to produce secondary structure models that are both stable and well-supported by experimental evidences.
 The analysis of multiple reactivity profiles, both publicly available and produced in our study, demonstrates the good performances of \OurTool, even in a mono probing setting.
@@
 
-Reagents used within \Rev{selective 2' -hydroxyl acylation analyzed by primer extension (SHAPE)~\cite{Wilkinson2006} protocols} represent a popular class of probes.
+Reagents used within \RevA{selective 2' -hydroxyl acylation analyzed by primer extension (SHAPE)~\cite{Wilkinson2006} protocols} represent a popular class of probes.
 They react with the 2’-hydroxyl of flexible ribose~\cite{McGinnis2012}, although the exact properties observed by SHAPE remain the object of ongoing investigations~\cite{McGinnis2012,Sexton2017,Hurst2018,Mlynsky2018,Frezza2019,Busan2019}.
@@
 Other chemical probes, harbouring different chemical reactivities, have been developed.
-Amongst the most popular, DiMethyl Sulfate (DMS) is a small molecule that methylates Adenines and Cytosines if not involved in a hydrogen bond, while   CMCT reacts with the Watson-Crick face of unpaired Guanosines and Uracyls~\cite{Ehresmann1987,Brunel2000}.
+Amongst the most popular, DiMethyl Sulfate (DMS) is a small molecule that methylates Adenines and Cytosines if not involved in a hydrogen bond, while   CMCT reacts with the Watson-Crick face of unpaired Guanosines and Uracils~\cite{Ehresmann1987,Brunel2000}.
 These two reagents not only reveal Watson-Crick base pairing, but also other types of contacts involving the same edge.
-The diversity of probes\Rev{, some of which are usable \emph{in vivo},} not only increases coverage of the different positions and structural contexts, \Rev{but also provides different qualitative information.
-\Rev{Modeling can also benefit from a joint analysis of reactivity profiles of single-point mutants, assuming structural homology~\cite{Cordero2012a}}.
+The diversity of probes\RevA{, some of which are usable \emph{in vivo},} not only increases coverage of the different positions and structural contexts, \RevA{but also provides different qualitative information.
+\RevA{Modeling can also benefit from a joint analysis of reactivity profiles of single-point mutants, assuming structural homology~\cite{Cordero2012a}}.
 Integration of multiple sources of probing to improve structure prediction  has thus been widely used since the very early days of RNA structure probing~\cite{Moazed1986, Ehresmann1987, Romaniuk1988, Butcher1994, Brunel2000,  Cordero2015, Somarowthu2015, Gross2017} but, to the best of our knowledge, has never been fully automated.}
@@
 
-Computationally, the reactivity of a nucleotide is typically used as a proxy \Rev{to assess the unpaired nature of individual nucleotides}.
+Computationally, the reactivity of a nucleotide is typically used as a proxy \RevA{to assess the unpaired nature of individual nucleotides}.
 The past couple of decades have nevertheless seen a series of paradigm shifts in the ways probing information is integrated, somehow mirroring the evolution of {\em ab initio} methods for secondary structure prediction.
@@
 Popular packages for secondary structure prediction, such as \Software{RNAStructure}~\cite{Mathews2004} or the \Software{Vienna} package~\cite{Lorenz2011}, now routinely accept SHAPE-like reactivity profiles as input of their main structure prediction methods.
-\Rev{Besides free-energy optimization, such methods notably} include the joint folding of aligned RNAs~\cite{Lavender2015}, partition function, statistical sampling and Maximal Expected Accuracy predictions~\cite{Spasic2017}\ldots
+\RevA{Besides free-energy optimization, such methods notably} include the joint folding of aligned RNAs~\cite{Lavender2015}, partition function, statistical sampling and Maximal Expected Accuracy predictions~\cite{Spasic2017}\ldots
 
@@
 \subsection*{The \OurTool{} method}
-We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signals produced by various \Rev{probing experiments to predict one or several secondary structure models for a given RNA.
+We introduce the Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles (\OurTool{}), a novel approach that integrates the signals produced by various \RevA{probing experiments to predict one or several secondary structure models for a given RNA.
 It takes as input one or several reactivity profiles produced in various \Def{conditions}, broadly defined to denote the conjunction of a reagent, a probing technology, ionic concentrations and, in extreme cases, structurally homologous mutants.}
@@
 The data-driven predicted structures are then gathered in one sample, serving as input for the clustering step.
-\OurTool{} proceeds, then, with an iterative clustering that ends once \Rev{a stopping criterion is} reached.
+\OurTool{} proceeds, then, with an iterative clustering that ends once \RevA{a stopping criterion is} reached.
 This step allows to identify the adequate number of clusters $k$ to be considered.
-The $k$ resulting clusters are then \Rev{represented} by their centroid structures.
+The $k$ resulting clusters are then \RevA{represented} by their centroid structures.
 Clusters figuring on the 2D-Pareto frontier are considered to be optimal and subsequently their corresponding centroids are reported as the predicted structure through \OurTool.}\label{fig:approach}
@@
 using $m=1.3$ and  $b=-0.4$.
-\Rev{Those values were halved in comparison to those recommended by Deigan\etal\cite{Deigan2009}, following a grid search optimization on the Cordero dataset, and based on the rationale that lower absolute values for pseudo-energy bonuses increase the expected overlap between pseudo-Boltzmann ensembles.} Those pseudo energy contributions effectively guide our predictions towards a subset of structures that are in good agreement with probing data.
+\RevA{Those values were halved in comparison to those recommended by Deigan\etal\cite{Deigan2009}, following a grid search optimization on the Cordero dataset, and based on the rationale that lower absolute values for pseudo-energy bonuses increase the expected overlap between pseudo-Boltzmann ensembles.} Those pseudo energy contributions effectively guide our predictions towards a subset of structures that are in good agreement with probing data.
 For each condition $d$ in $\mathcal{D}$, we use the soft constraints framework~\cite{Lorenz2016} of the \Software{RNAfold} software~\cite{Lorenz2011} to produce a random (multi)set $\mathcal{S}_d$ of $\NumStruct$ structures in the pseudo-Boltzmann ensemble.
 
-\Rev{\paragraph{Clustering across conditions.} In order to infer recurrent conformations across sampled sets, \OurTool{} agglomerate structure (multi)sets while keeping track of their condition of origin, and \Def{clusters} with respect to the \Def{base-pair distance}, the number of base pairs differing between two structures.
+\RevA{\paragraph{Clustering across conditions.} In order to infer recurrent conformations across sampled sets, \OurTool{} agglomerate structure (multi)sets while keeping track of their condition of origin, and \Def{clusters} with respect to the \Def{base-pair distance}, the number of base pairs differing between two structures.
 A clustering algorithm then partitions the (multi)set of sampled structures into \emph{clusters}, (multi)sets of structures such that the accumulated sum of distances over clusters is minimized.
@@
 A cluster is deemed \Def{significantly populated} if its stability exceeds a predefined threshold $\epsilon$. 
-We set  $\epsilon =\Rev{|\mathcal{D}|}/3$ by default, such that at most three clusters are deemed significantly populated, and used as our primary candidates.
+We set  $\epsilon =\RevA{|\mathcal{D}|}/3$ by default, such that at most three clusters are deemed significantly populated, and used as our primary candidates.
 Finally, we consider two clusters to be \Def{highly similar} if their centroid structures differ by at most $\delta$ base pairs ($\delta=1$ by default), allowing the identification of clusters in the presence of minor variations.
@@
 \paragraph{Filtering the promising conformer(s).} Finally, we \Def{elect the most promising cluster(s)}, and return their centroid(s).
-While the final number of clusters may potentially be large, only a handful of clusters are expected to represent structures that \Rev{are} both stable, and supported by a large number of conditions.
+While the final number of clusters may potentially be large, only a handful of clusters are expected to represent structures that \RevA{are} both stable, and supported by a large number of conditions.
 The remaining clusters are indeed probably artifacts of the clustering method, but nevertheless useful to filter out \emph{noisy} structures.
 
-\Rev{We postulate that a perfect cluster should have large stability, as defined above, and be representative of several conditions.} In the presence of a set of experimental conditions $\mathcal{D}$, we consider that a cluster $C$ supports a given condition $d$ when its probability within $d$ exceeds a given threshold $\tau$.
+\RevA{We postulate that a perfect cluster should have large stability, as defined above, and be representative of several conditions.} In the presence of a set of experimental conditions $\mathcal{D}$, we consider that a cluster $C$ supports a given condition $d$ when its probability within $d$ exceeds a given threshold $\tau$.
 The number of conditions supporting a cluster $C$ is defined as
 $$\text{Support}(C) = \left|\left\{d\in \mathcal{D} \mid \left(\textstyle\sum_{S\in C\cap \mathcal{S}_d} \mathbb{P}_d(S)\right)\ge \tau \right\}\right|.$$
-The value of $\tau$ is set to $1/(\NumClust+1)$ \Rev{with $\NumClust$ the final number of clusters},  ensuring at least one supporting cluster for each condition.
+The value of $\tau$ is set to $1/(\NumClust+1)$ \RevA{with $\NumClust$ the final number of clusters},  ensuring at least one supporting cluster for each condition.
 
@@
 It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (1M7-\SH) was made available.
-This dataset includes sequences originating from a variety of organisms, and spans lengths ranging from $34$ nts to $530$ nts, with a focus on riboswitches and complex RNA architectures \Rev{(full list in Supp.
+This dataset includes sequences originating from a variety of organisms, and spans lengths ranging from $34$ nts to $530$ nts, with a focus on riboswitches and complex RNA architectures \RevA{(full list in Supp.
 Table S3)}.
@@
 We annotated the secondary structure elements using the \Software{DSSR} software from the 3DNA suite~\cite{Lu2015}.
-\Rev{Non-canonical base pairs were removed, and a non-pseudoknotted secondary structure was extracted as the maximum subset of non-crossing base pairs~\citep{Smit2008}.}
+\RevA{Non-canonical base pairs were removed, and a non-pseudoknotted secondary structure was extracted as the maximum subset of non-crossing base pairs~\citep{Smit2008}.}
 
@@
 
-\Rev{To systematically assess the potential of multiple sources of probing, we considered a difficult example, the \textit{Didymium iridis} Lariat Capping ribozyme (\didy{}).
+\RevA{To systematically assess the potential of multiple sources of probing, we considered a difficult example, the \textit{Didymium iridis} Lariat Capping ribozyme (\didy{}).
 The native structure of \didy{}, shown in Figure~\ref{fig:didy} is highly complex, and features two pseudoknots which cannot be explicitly modeled by most computational methods, making \didy{} a challenging target for secondary structure prediction.}
@@
 \didy{} was probed with different SHAPE reagents: 1M7 (1-methyl-7 nitrosatoic anhydre), NMIA (N-Methyl Isatoic Anhydre), BzCN (Benzoyl cyanide) and NAI (2-methylnicotinic acid imidazolide) in presence and absence of Mg$^\text{2+}$.
-\Rev{\didy{} was also probed with DMS (DiMethylSulfate) and CMCT, in presence of Mg$^\text{2+}$, resulting in a total of 16 \Def{probing conditions}, a term we use in the following to denote a combination of probing technology, reagent and presence/absence of Mg$^\text{2+}$ (+ sequencing technology).
+\RevA{\didy{} was also probed with DMS (DiMethylSulfate) and CMCT, in presence of Mg$^\text{2+}$, resulting in a total of 16 \Def{probing conditions}, a term we use in the following to denote a combination of probing technology, reagent and presence/absence of Mg$^\text{2+}$ (+ sequencing technology).
 For each probing condition, three experiments were performed in presence/absence of the reagent, and in a denatured context, following classic SHAPE protocols~\cite{Wilkinson2006,Smola2015}.}
@@
 The Matthews Correlation Coefficient (MCC) is a classic metric for assessing the quality of a predicted structure $S$, identified by a set of base pairing positions, compared to an accepted reference (native) structure $\Ref$.
-It represents a compromise \Rev{between the main metrics derived from the confusion matrix}, and is defined as
+It represents a compromise \RevA{between the main metrics derived from the confusion matrix}, and is defined as
 $$\text{MCC}(S\mid \Ref)=\frac{\text{TP} \times \text{TN} -\text{FP} \times \text{FN}}{\sqrt[]{(\text{TP}+\text{FP})(\text{TP}+\text{FN})(\text{TN}+\text{FP})(\text{TN}+\text{FN})}},$$
 where TP, FP, TN, FN represent the correctly/erroneously predicted and correctly/erroneously omitted base pairs in $S$ with respect to $\Ref$.
-\Rev{MCC is a stringent metric, taking values between -100\% and 100\%, 0\% being the expected MCC of a "coin tossing" random predictor.}
+\RevA{MCC is a stringent metric, taking values between -100\% and 100\%, 0\% being the expected MCC of a "coin tossing" random predictor.}
 
@@
 
-\Rev{
+\RevA{
 \subsection*{Assessment of statistical significance}
@@
 
-\subsection*{Considering multiple probing conditions \Rev{appears to} improve the quality of predictions}\label{sec:cordero}
+\subsection*{Considering multiple probing conditions \RevA{appears to} improve the quality of predictions}\label{sec:cordero}
 
@@
 	\caption{Dot-bracket representations of the native structure and \OurTool{} predictions for the glycine riboswitch sequence of the Cordero\etal\cite{Cordero2012} data set, supplemented with an increasing collection of probing conditions.
-\Rev{Please note that, throughout the manuscript, by "conditions" we mean a combination of the nature of the probe, the probing technology, the ions present, the temperature, or any other "environmental" variation} Positions with green and red backgrounds indicate correctly and incorrectly predicted base pairs.}\label{fig:glycine_example}
+\RevA{Please note that, throughout the manuscript, by "conditions" we mean a combination of the nature of the probe, the probing technology, the ions present, the temperature, or any other "environmental" variation} Positions with green and red backgrounds indicate correctly and incorrectly predicted base pairs.}\label{fig:glycine_example}
 \end{table*}
@@
 
-	\caption{\Rev{Predictive performances (GM) of \OurTool{} on the Cordero~\cite{Cordero2012a} dataset consisting of reactivities associated with the CMCT, NMIA and DMS probing for 6 reference RNAs.
+	\caption{\RevA{Predictive performances (GM) of \OurTool{} on the Cordero~\cite{Cordero2012a} dataset consisting of reactivities associated with the CMCT, NMIA and DMS probing for 6 reference RNAs.
 (Pseudo) MFE and MEA structures, predicted using \Software{RNAfold}~\cite{Lorenz2011}, are included as references in the case of a single condition, and in the absence of probing.}}
@@
 Remarkably, the combination of the three conditions leads to the best overall predictions, averaging $78.5\%$ GM.
-This \Rev{seemingly} improves by $8.5\%$, $2.7\%$ and $1.4\%$ over the average predictions achieved using CMCT, DMS and NMIA respectively.
+This \RevA{seemingly} improves by $8.5\%$, $2.7\%$ and $1.4\%$ over the average predictions achieved using CMCT, DMS and NMIA respectively.
 %CMCT/DMS/NMIA :0.6988333333/0.7588333333/0.771
 Table~\ref{fig:glycine_example} illustrates the incremental refinement of \OurTool predictions for a glycine riboswitch upon increasing the number of considered probing conditions.
-\Rev{However, these observations merely suggest an improvement, as statistical significance cannot be established on such a small dataset.}
+\RevA{However, these observations merely suggest an improvement, as statistical significance cannot be established on such a small dataset.}
 Still, they suggests some level of complementarity between probing conditions, as already suggested by recent analyses~\citep{Brunel2000,Steen2012,Rice2014,Yu2018}  and supported by further analyses in this paper.
@@
 We assessed the predictive capacities of \OurTool in a classic setting where a single probing condition is available, and compared its performances against \Software{RSample} software recently introduced by Spacic\etal\cite{Spasic2017}, \Software{RSample} relies on a sampling/clustering method developed independently from the current work.
-It was shown to perform \Rev{comparably as} a comprehensive collection of state-of-the-art methods in probing-guided structure prediction, including \Software{RME}~\citep{Wu2015}, \Software{RNAprob}~\citep{Deng2016}, \Software{RNAprobing}~\citep{Washietl2012}, \Software{RNAsc}~\citep{Zarringhalam2012} and the \Software{fold} utility from the \Software{RNAStructure} suite~\citep{Reuter2010}.
+It was shown to perform \RevA{comparably as} a comprehensive collection of state-of-the-art methods in probing-guided structure prediction, including \Software{RME}~\citep{Wu2015}, \Software{RNAprob}~\citep{Deng2016}, \Software{RNAprobing}~\citep{Washietl2012}, \Software{RNAsc}~\citep{Zarringhalam2012} and the \Software{fold} utility from the \Software{RNAStructure} suite~\citep{Reuter2010}.
 We considered the Hajdin dataset~\cite{Hajdin2013}, which consists of 24 RNAs of lengths ranging from 47 to 500 nts, all believed to fold into a unique documented conformation.
@@
 	\includegraphics[width=\linewidth]{graphs/RsampleVsIPANEMAP/Accuracy}
-	\caption{\Rev{Comparison of predictive performances (GM) achieved by  \OurTool{} and \Software{Rsample} over the Hajdin\etal\cite{Hajdin2013} data set, consisting of 24 RNAs which associated SHAPE reactivities.
+	\caption{\RevA{Comparison of predictive performances (GM) achieved by  \OurTool{} and \Software{Rsample} over the Hajdin\etal\cite{Hajdin2013} data set, consisting of 24 RNAs which associated SHAPE reactivities.
 The length of each individual RNA is indicated within square brackets.} \label{fig:Vsrsample}}
@@
 
-\Rev{We used the default number of sampled structures for \Software{RSample}, namely 10~000 to compute correction factors and 1~000 for the clustering phase, and \OurTool (1~000 structures).
+\RevA{We used the default number of sampled structures for \Software{RSample}, namely 10~000 to compute correction factors and 1~000 for the clustering phase, and \OurTool (1~000 structures).
 For all RNAs, a single structure was returned by both software.
@@
 We computed and report in Figure~\ref{fig:Vsrsample} the individual Geometric Means (GM) of predicted structures.
-\Rev{Sensitivity and PPV can be found in Supp.
+\RevA{Sensitivity and PPV can be found in Supp.
 Table S4, and Supp.
@@
 A first observation is that \OurTool generally outperforms \Software{RSample}, averaging $80\%$ GM as opposed to $68\%$ for its competitor.
-\Rev{The superiority on \OurTool on this benchmark appears statistically significant according to a two-tailed paired $t$-test (P-value$=6\times10^{-3}$) with observed data.}
+\RevA{The superiority on \OurTool on this benchmark appears statistically significant according to a two-tailed paired $t$-test (P-value$=6\times10^{-3}$) with observed data.}
 Out of the 24 considered RNAs, \OurTool outperforms its competitor for 16 out of the 24 RNAs, achieves similar performances for 2 of them, and is dominated for the remaining 6 cases.
@@
 
-	\caption{\Rev{Model of the Lariat capping ribozyme from {\itshape Didymium iridis} (A -- PDB 4P8Z:A by Meyer\etal\cite{Meyer2014}), secondary structure, as annotated by \Software{DSSR}~\cite{Lu2015} (B) and \OurTool first-ranking prediction (C.), obtained from the integration of 8 maximally-diverging probing conditions (\NMIAMg, \OneMSevILUThreeMg, \NAIMg, \OneMSevCE, \CMCTMg,\OneMSevILUThree, and \OneMSevMgCE).}\label{fig:didy}}
+	\caption{\RevA{Model of the Lariat capping ribozyme from {\itshape Didymium iridis} (A -- PDB 4P8Z:A by Meyer\etal\cite{Meyer2014}), secondary structure, as annotated by \Software{DSSR}~\cite{Lu2015} (B) and \OurTool first-ranking prediction (C.), obtained from the integration of 8 maximally-diverging probing conditions (\NMIAMg, \OneMSevILUThreeMg, \NAIMg, \OneMSevCE, \CMCTMg,\OneMSevILUThree, and \OneMSevMgCE).}\label{fig:didy}}
 \end{figure*}
 To assess the potential of our method, we considered the Lariat capping ribozyme from \textit{Didymium Iridis} (\didy{}) whose 3D structure has been modeled at 3.85 \AA{} from X-ray diffraction data~\cite{Meyer2014}.
-\Rev{This RNA contains a large diversity of interactions, including helical domains, tertiary pairings, a kissing loop and a pseudoknot (Figure~\ref{fig:didy}).
+\RevA{This RNA contains a large diversity of interactions, including helical domains, tertiary pairings, a kissing loop and a pseudoknot (Figure~\ref{fig:didy}).
 Its automated modeling is therefore a challenging case for computational  structure prediction methods.
@@
 As described in the Material and Methods section, we probed \didy{}  in 16 different conditions using different SHAPE and classical reagents, in presence/absence of Mg$^\text{2+}$ ions, using either the standard premature stop or the newly developed mutation induction technologies to map the modification sites.
-\Rev{We observed an overall good agreement of reactivities with the 3D model proposed by Meyer~\etal\cite{Meyer2014} (see  Supp.
+\RevA{We observed an overall good agreement of reactivities with the 3D model proposed by Meyer~\etal\cite{Meyer2014} (see  Supp.
 Figures S1, S2 and S3).} Reactivities were then used as input of \OurTool{}, either individually or combined,  for secondary structure prediction.
@@
 Figure~\ref{fig:PCA} summarizes the pairwise distance, projected onto a 2D surface by a principal component analysis.
-\Rev{A systematic cross-analysis of raw reactivities revealed a positive, if modest, correlation between conditions, but did not support a clear partition of profiles based on probing technology or presence/absence of Mg$^{\text{2+}}$ (see Supp.
+\RevA{A systematic cross-analysis of raw reactivities revealed a positive, if modest, correlation between conditions, but did not support a clear partition of profiles based on probing technology or presence/absence of Mg$^{\text{2+}}$ (see Supp.
 Figure S4).}
@@
 \end{bulletlist}
-The predicted clusters are generally consistent with the ordering resulting from a spectral biclustering, implemented in \Software{scikit-learn} and executed with default parameters, as illustrated by \Rev{Figure~\ref{fig:biclustering}}.
+The predicted clusters are generally consistent with the ordering resulting from a spectral biclustering, implemented in \Software{scikit-learn} and executed with default parameters, as illustrated by \RevA{Figure~\ref{fig:biclustering}}.
 
@@
 The predictive performances of \OurTool averages a 70.5\% MCC, compared to 68.5\% and 68\% average MCC for MEA and MFE predictions respectively.
-\Rev{This suggests modest, barely significant, improvements compared to classic prediction paradigms (P-values of 2\% and 14\% respectively).}
+\RevA{This suggests modest, barely significant, improvements compared to classic prediction paradigms (P-values of 2\% and 14\% respectively).}
 The stability of predictions across conditions seem comparable for all \OurTool, MEA and MFE-driven predictions (std dev.
@@
 Across the sixteen conditions, we observed a large discrepancy in the capacity of the experimental setup to inform predictions, with MCCs ranging from above 80\% (\NMIA, \NMIAMg, \OneMSevILU and \OneMSevILUMg) to around 60\% (\NAICE, \BzCN, \OneMSevCE and \NMIACE).
-\Rev{Interestingly the MCC value cannot be trivially anticipated from the compatibility of the reactivity scores with the predicted structure, or even with the native structure (see Supp.
+\RevA{Interestingly the MCC value cannot be trivially anticipated from the compatibility of the reactivity scores with the predicted structure, or even with the native structure (see Supp.
 Table S1).}
@@
 Unsurprisingly, the presence/absence of Magnesium ions during the probing was observed to impact the predictions, with an observed drop from an average 73\% MCC in the presence of Mg$^\text{2+}$ to 67\% MCC in the absence of Mg$^\text{2+}$.
-However, while the poorly performing conditions in Cluster \Bull{B} \Rev{were} probed in the absence of Mg$^\text{2+}$, some of the most informative profiles (\OneMSevILU and \NMIA) were obtained in absence of Mg$^\text{2+}$.
-\Rev{Even more drastic changes of performances were observed when comparing the MCCs of SHAPE-MaP experiments (avg 78\%) with the historical SHAPE CE technology (avg 66\%).}
+However, while the poorly performing conditions in Cluster \Bull{B} \RevA{were} probed in the absence of Mg$^\text{2+}$, some of the most informative profiles (\OneMSevILU and \NMIA) were obtained in absence of Mg$^\text{2+}$.
+\RevA{Even more drastic changes of performances were observed when comparing the MCCs of SHAPE-MaP experiments (avg 78\%) with the historical SHAPE CE technology (avg 66\%).}
 
@@
 	%data are from multiprobing7 
-	\caption{\Rev{Predictive performances (MCC) of \OurTool{}, averaged over 10 runs, for the \didy{} dataset, consisting of 16 reactivity profiles.
+	\caption{\RevA{Predictive performances (MCC) of \OurTool{}, averaged over 10 runs, for the \didy{} dataset, consisting of 16 reactivity profiles.
 (Pseudo) MFE and MEA structures, predicted using \Software{RNAfold}~\cite{Lorenz2011}, are included for reference.
@@
 More precisely, the average MCC value of pair-informed predictions remained at 70.7\%, improving only by 0.2\% over the average MCC of mono-probing predictions.
-Pair-informed predictions were marginally more reliable, with the median MCC \Rev{increasing} to 72.5\% MCC from 70.5\% for mono-probing predictions.
+Pair-informed predictions were marginally more reliable, with the median MCC \RevA{increasing} to 72.5\% MCC from 70.5\% for mono-probing predictions.
 However, outliers strongly impacted the overall picture, and ignoring the conditions in cluster \Bull{B} increased the average MCC to 75.1\%, while unsurprisingly decreasing the standard deviation (8.6\% $\to$ 6.5\%).
@@
 Compared to bi-probing, MCC values achieved by integrating triplets of conditions indicated a clear gain over the average performances of individual conditions, supporting a notion of complementary between experiments.
-Indeed tri-probing experiments induced an average +1.7\% MCC improvement (median +2.6) over the average of mono-probing prediction in the triplet \Rev{(P-value$=1.4\times 10^{-9}$)}.
+Indeed tri-probing experiments induced an average +1.7\% MCC improvement (median +2.6) over the average of mono-probing prediction in the triplet \RevA{(P-value$=1.4\times 10^{-9}$)}.
 The gain of quality was widespread, and observed for 65\% of the triplets.
@@
 
-The enhanced performances of tri-probing over bi-probing \Rev{was deemed statistically significant (P-value$=2\times 10^{-2}$), and} consistent with the \emph{voting} principle underlying the clustering used in \OurTool.
-\Rev{Indeed, in a bi-probing setting, a single outlier (\emph{e.g.} \NAICE) may entirely determine the final structure due to its Boltzmann ensemble being very tightly concentrated around a, presumably erroneous, centroid structure.
+The enhanced performances of tri-probing over bi-probing \RevA{was deemed statistically significant (P-value$=2\times 10^{-2}$), and} consistent with the \emph{voting} principle underlying the clustering used in \OurTool.
+\RevA{Indeed, in a bi-probing setting, a single outlier (\emph{e.g.} \NAICE) may entirely determine the final structure due to its Boltzmann ensemble being very tightly concentrated around a, presumably erroneous, centroid structure.
 In the presence of three or more conditions, however, clusters resulting from an outlier condition are typically expected to be dominated in the Pareto front, by compatible structures originating from alternative conditions.
@@
 Figure~S5 for associated models).
-\Rev{Note that such an analysis does not use knowledge of the native structure, since our clustering only relies on properties of the pseudo-Boltzmann ensemble.}
+\RevA{Note that such an analysis does not use knowledge of the native structure, since our clustering only relies on properties of the pseudo-Boltzmann ensemble.}
 
@@
 The results, summarized in Table~\ref{fig:differentconds}, demonstrate a clear advantage of including diverse conditions, with an average MCC value of 77\% in the absence of condition from cluster \Bull{B}.
-Those performances compare favorably against the 72\% MCC achieved on average by individual conditions \Rev{(P-value$=3\times 10^{-5}$)}, and reveal a remarkable stability (5\% stdev), and a degradation in only two out the 24 tested cases ($\sim$1\% \Rev{loss of MCC} in both cases).
+Those performances compare favorably against the 72\% MCC achieved on average by individual conditions \RevA{(P-value$=3\times 10^{-5}$)}, and reveal a remarkable stability (5\% stdev), and a degradation in only two out the 24 tested cases ($\sim$1\% \RevA{loss of MCC} in both cases).
 A similar trend can be observed in the disruptive presence of \NAICE, inducing an 72\% average MCC, albeit with an increased standard deviation of 9\%.
@@
 Finally, considering the 16 conditions together induced a 80\% MCC value, increasing to 83\% in the absence of cluster \Bull{B}.
-\Rev{Overall, these results support the notion that the strategy used} by \OurTool{} is able to exploit the complementary information provided by multiple conditions, contributing to better predictions.
+\RevA{Overall, these results support the notion that the strategy used} by \OurTool{} is able to exploit the complementary information provided by multiple conditions, contributing to better predictions.
 
@@
 A single condition is chosen from each cluster except for \Bull{B}.
-\Rev{Clusters consisting of a single condition (\Bull{D}, \Bull{F} and \Bull{H}) contribute additional conditions  $ :=\{\OneMSevILUThree, \BzCNMg, \DMSMg\}$.
+\RevA{Clusters consisting of a single condition (\Bull{D}, \Bull{F} and \Bull{H}) contribute additional conditions  $ :=\{\OneMSevILUThree, \BzCNMg, \DMSMg\}$.
 		The MCC of the predicted structure is reported ('Joint' columns) and compared to the average of mono conditions ('Mono' columns), in presence/absence of \NAICE.}
@@
 \subsection*{Supplementing wild-type profiles with random single mutants increases prediction quality}
-\Rev{In this final analysis, we tested the capacity of \OurTool to extract and exploit complementary information produced by the systematic probing of single-point mutants using the Mutate-and-Map (MaM) protocol~\cite{Cordero2015}.
+\RevA{In this final analysis, we tested the capacity of \OurTool to extract and exploit complementary information produced by the systematic probing of single-point mutants using the Mutate-and-Map (MaM) protocol~\cite{Cordero2015}.
 The joint analysis of several mutants may appear error-prone, as single point mutants are expected to adopt different structures than the wild-type sequence (WT), an expectation that is at the core of downstream analyses~\cite{Cordero2015}.
@@
 When the WT and a single random mutant (WT+1M) were jointly considered, the dispersion of the MCCs increased, and values ranging from 38.9\% to 85\% were observed.%min(Multiprobing_1M)
-\Rev{The average MCC in the presence of a single mutant increased to 64.9\%,%mean(Multiprobing_1M)
+\RevA{The average MCC in the presence of a single mutant increased to 64.9\%,%mean(Multiprobing_1M)
 suggesting a significant positive contribution from the additional mutant profile (P-value$=1\times 10^{-3}$).}
@@
 Interestingly, the dispersion of the MCC was lower for WT+2M than for WT+1M, with lower MCC values above 57\%.
-This results suggest \Rev{adding a second mutant profile has  a role of \emph{tie-breaker}}, allowing to mitigate the deleterious effect of some outlier.
-However, outliers still played a disruptive role, and restricting chosen mutants to the 20 reactivity profiles most similar to the WT further improved the average MCC 2.8\% to 67.7\%,  \Rev{ a significant improvement compared to WT+1M (P-value$=2.4\times10^{-2}$).}
-
-The average MCC again increased in the presence of 3 mutants, to reach an average MCC of 69.3\%, \Rev{ a substantial improvement that nevertheless fails to reach the level of statistical significance (P-value$=11\times 10^{-2}$).} For 10 mutants, the average MCC and overall distribution remains highly similar to that obtained by considering three mutants, suggesting \Rev{a plateau in the performances}.
+This results suggest \RevA{adding a second mutant profile has  a role of \emph{tie-breaker}}, allowing to mitigate the deleterious effect of some outlier.
+However, outliers still played a disruptive role, and restricting chosen mutants to the 20 reactivity profiles most similar to the WT further improved the average MCC 2.8\% to 67.7\%,  \RevA{ a significant improvement compared to WT+1M (P-value$=2.4\times10^{-2}$).}
+
+The average MCC again increased in the presence of 3 mutants, to reach an average MCC of 69.3\%, \RevA{ a substantial improvement that nevertheless fails to reach the level of statistical significance (P-value$=11\times 10^{-2}$).} For 10 mutants, the average MCC and overall distribution remains highly similar to that obtained by considering three mutants, suggesting \RevA{a plateau in the performances}.
 Overall, these results support a notion of complementarity between the probing data produced across reasonably similar mutants, leading to a gradual decrease of the signal to noise ratio.
@@
 
-	\caption{\Rev{Distributions of predictive accuracies (MCC) of structures predicted by \OurTool{}, informed by Mutate-and-Map data, for \didy{} (Cheng dataset~\cite{Cheng2017}).
+	\caption{\RevA{Distributions of predictive accuracies (MCC) of structures predicted by \OurTool{}, informed by Mutate-and-Map data, for \didy{} (Cheng dataset~\cite{Cheng2017}).
 Predictions were repeated 100 times, informed both by the wild-type sequence and reactivity profile (WT), and with randomly-selected reactivity profiles associated with 1, 2, 3, or 10 single-point mutants ({\sf WT+$k$M}).
@@
 We introduce \OurTool{}, an integrative method for the prediction of secondary structure models from multiple probing profiles.
-Based on the simple premise that good structural models should be \Rev{thermodynamically} stable and supported by most probing experiments, it uses a combination of stochastic backtrack in the pseudo-Boltzmann ensemble, coupled with a structural clustering to elect dominant conformations.
-This strategy is fast, \Rev{generally stable}, and leads to predictions that clearly benefit from the availability of multiple probing profiles, as demonstrated by detailed analyses performed of four datasets.
+Based on the simple premise that good structural models should be \RevA{thermodynamically} stable and supported by most probing experiments, it uses a combination of stochastic backtrack in the pseudo-Boltzmann ensemble, coupled with a structural clustering to elect dominant conformations.
+This strategy is fast, \RevA{generally stable}, and leads to predictions that clearly benefit from the availability of multiple probing profiles, as demonstrated by detailed analyses performed of four datasets.
 
@@
 Surprisingly, it also appears to produce the most accurate secondary structure models when informed by a single condition, as demonstrated by our reanalysis on the Hajdin\etal\cite{Hajdin2013} benchmark.
-We attribute this unexpected result to the adoption of a sampling/clustering approach, \Rev{which appears to eliminate spurious outliers from the conformational space even when a single conformation dominates.}
-\Rev{In this context, the good performance of \OurTool in comparison with \Software{RSample}~\cite{Spasic2017} may appear surprising.
+We attribute this unexpected result to the adoption of a sampling/clustering approach, \RevA{which appears to eliminate spurious outliers from the conformational space even when a single conformation dominates.}
+\RevA{In this context, the good performance of \OurTool in comparison with \Software{RSample}~\cite{Spasic2017} may appear surprising.
 Indeed, while independently developed and used~\cite{Deforges2017}, both methods share a common sampling/clustering scheme.
@@
 It follows that a modeler cannot objectively favor \emph{a priori} a probing condition/reagent~\cite{Yu2018}.
-In other words, multiple experimental probing profiles, even combined in a pairwise fashion, \Rev{can be used jointly to mitigate the empirical risk of a misprediction}.
-
-\Rev{Current limitations of \OurTool include the lack of an explicit support for pseudoknots, which will be addressed in a future version, using  polynomial-time dynamic programming~\cite{Janssen2015}, or a iterative heuristic alternative~\cite{Hajdin2013}.
+In other words, multiple experimental probing profiles, even combined in a pairwise fashion, \RevA{can be used jointly to mitigate the empirical risk of a misprediction}.
+
+\RevA{Current limitations of \OurTool include the lack of an explicit support for pseudoknots, which will be addressed in a future version, using  polynomial-time dynamic programming~\cite{Janssen2015}, or a iterative heuristic alternative~\cite{Hajdin2013}.
 The inherently stochastic nature of the sampling scheme may also be challenging for certain types of analysis.
@@
 
-Future extensions also include \Rev{addressing} the dilemma of whether or not to include Mg$^{\text{2+}}$ ions during the probing, in the perspective of a downstream computational prediction.
+Future extensions also include \RevA{addressing} the dilemma of whether or not to include Mg$^{\text{2+}}$ ions during the probing, in the perspective of a downstream computational prediction.
 On the one hand, RNA reaches its native 3D structure in the presence of Mg$^{\text{2+}}$ ions~\cite{Knapp1989}.
@@
 In this work we probed \didy, which contains many tertiary motifs, in the presence and absence of Mg$^{\text{2+}}$.
-Remarkably, the most accurate predictions were obtained with 1M7 and NMIA \Rev{irrespectively of the presence/absence of Mg$^{\text{2+}}$, and despite the fact that reactivity profiles substantially diverged between the two conditions (see Supp.
+Remarkably, the most accurate predictions were obtained with 1M7 and NMIA \RevA{irrespectively of the presence/absence of Mg$^{\text{2+}}$, and despite the fact that reactivity profiles substantially diverged between the two conditions (see Supp.
 Figure S4)}.
@@
 	\caption{Stochastic variability of \OurTool{} predictions for \didy{} dataset.
-MCC of predicted structures over 10 runs of \OurTool{}, in presence/absence of Mg$^{2+}$, using both \Rev{SHAPE CE and MaP} protocols for a collection of reagents.
+MCC of predicted structures over 10 runs of \OurTool{}, in presence/absence of Mg$^{2+}$, using both \RevA{SHAPE CE and MaP} protocols for a collection of reagents.
 }\label{fig:reproducibility}
@@
 \section*{Funding}
-This work, including AS and DA \Rev{PhD} fellowships, has been supported by La Fondation pour la Recherche Medicale ([FRM
+This work, including AS and DA \RevA{PhD} fellowships, has been supported by La Fondation pour la Recherche Medicale ([FRM
 DBI20141423337]).

### 2020-07-02 IPANEMAP-2019.tex
Updates from Overleaf

@@
 
-\subsection*{\OurTool outperforms state-of-the-art predictive methods when informed by a single profile}
+\subsection*{\OurTool{} \RevB{performs comparably to} state-of-the-art predictive methods when informed by a single profile}
 %Mono-probing Mathews (Hadjim)
@@
 
-\RevA{We used the default number of sampled structures for \Software{RSample}, namely 10~000 to compute correction factors and 1~000 for the clustering phase, and \OurTool (1~000 structures).
+\RevA{We used the default numbers of sampled structures for \Software{RSample} and \OurTool{}, namely 10~000 to compute correction factors and 1~000 for the clustering phase.
 For all RNAs, a single structure was returned by both software.
@@
 \RevA{Sensitivity and PPV can be found in Supp.
-Table S4, and Supp.
+Table S4, while Supp.
 Table S5 reports the stability of predictions over 10 independent runs of \OurTool.}
+
+\RevB{%
+\OurTool{} appears to perform comparably, or even slightly favorably, to \Software{RSample}, reporting more accurate structures than its competitor for roughly two thirds of the tested RNAs, while averaging $80.1\%$ GM.
+However, the  improvement is only of around 1.5\% on average over the \citet{Hajdin2013} dataset, an improvement that is no longer significant according to a two-tailed paired T-test.
+}
 

### 2020-07-02 IPANEMAP-2019.tex
Updates from Overleaf

@@
 %\paragraph{Results description}
-A first observation is that \OurTool generally outperforms \Software{RSample}, averaging $80\%$ GM as opposed to $68\%$ for its competitor.
+%A first observation is that \OurTool generally outperforms \Software{RSample}, averaging $80\%$ GM as opposed to $68\%$ for its competitor.
 \RevA{The superiority on \OurTool on this benchmark appears statistically significant according to a two-tailed paired $t$-test (P-value$=6\times10^{-3}$) with observed data.}
-Out of the 24 considered RNAs, \OurTool outperforms its competitor for 16 out of the 24 RNAs, achieves similar performances for 2 of them, and is dominated for the remaining 6 cases.
+%Out of the 24 considered RNAs, \OurTool outperforms its competitor for 16 out of the 24 RNAs, achieves similar performances for 2 of them, and is dominated for the remaining 6 cases.
 Whenever dominant, \OurTool achieves an average GM score of $82\%$, dominating by $32.8\%$ its competitor, while being dominated by $3.5\%$ over the remaining examples (including ties).
@@
 %[1] -0.382
-Remarkably, \OurTool predicts secondary structures much more consistently than its competitor. 
-It ensures GM that are greater than 50\% in all cases, and even exceeds 60\% over all except for the yeast tRNA, encoding the Aspartic acid and HIV1 5' pseudoknot domain.
+%Remarkably, \OurTool predicts secondary structures much more consistently than its competitor. 
+%It ensures GM that are greater than 50\% in all cases, and even exceeds 60\% over all except for the yeast tRNA, encoding the Aspartic acid and HIV1 5' pseudoknot domain.
 The standard deviation of GM for \OurTool is $12.4\%$, compared to the larger value of $21\%$ for \Software{RSample} predictions.
@@
 
-This observed discrepancy is somewhat surprising, given that \OurTool and \Software{RSample} share the same sampling/clustering based modeling.
+%This observed discrepancy is somewhat surprising, given that \OurTool and \Software{RSample} share the same sampling/clustering based modeling.
 However, the two methods differ substantially whenever a single conformation is detected, in which case \Software{RSample} switches to a simple MEA prediction, whereas \OurTool{} always proceeds with an iterative clustering.

### 2020-07-02 IPANEMAP-2019.tex
Updates from Overleaf

@@
 As ribose flexibility is proportional to the degree of freedom of the nucleotide, it is assumed that SHAPE reagents discriminate nucleotides involved in stable interactions.
-Different SHAPE reagents are endowed with different dynamics, that can help differentiate Watson-Crick base pairs from more dynamic tertiary contacts~\cite{Gherghe2008,Steen2012,Rice2014,Busan2019}.
+Different SHAPE reagents are endowed with different dynamics that can help differentiate Watson-Crick base pairs from more dynamic tertiary contacts~\cite{Gherghe2008,Steen2012,Rice2014,Busan2019}.
 Other chemical probes, harbouring different chemical reactivities, have been developed.
@@
 These two reagents not only reveal Watson-Crick base pairing, but also other types of contacts involving the same edge.
-The diversity of probes\RevA{, some of which are usable \emph{in vivo},} not only increases coverage of the different positions and structural contexts, \RevA{but also provides different qualitative information.
+The diversity of probes\RevA{, some of which are usable \emph{in vivo}\RevB{~\cite{Zaug1995,Paillart2004}},} not only increases coverage of the different positions and structural contexts, \RevA{but also provides different qualitative information.
 \RevA{Modeling can also benefit from a joint analysis of reactivity profiles of single-point mutants, assuming structural homology~\cite{Cordero2012a}}.
-Integration of multiple sources of probing to improve structure prediction  has thus been widely used since the very early days of RNA structure probing~\cite{Moazed1986, Ehresmann1987, Romaniuk1988, Butcher1994, Brunel2000,  Cordero2015, Somarowthu2015, Gross2017} but, to the best of our knowledge, has never been fully automated.}
+Integration of multiple sources of probing to improve structure prediction has thus been widely used since the very early days of RNA structure probing~\cite{Moazed1986, Ehresmann1987, Romaniuk1988, Butcher1994, Brunel2000,  Cordero2015, Somarowthu2015, Gross2017} but, to the best of our knowledge, has never been fully automated \RevB{within a soft constraint framework}~\cite{Lorenz2016}.}
 
@@
 It should, at the same time, remain small enough to ensure reproducibility, while being sufficiently large to discriminate outliers and ensure consistency within each cluster.
-We determine an \Def{optimal number of clusters} $\NumClust$ using an iterative heuristics, gradually increasing the number of clusters until a significantly-populated cluster is split into two similar clusters, or a poorly populated (outlier) cluster is created.
+We determine an \Def{optimal number of clusters} $\NumClust$ using an iterative heuristic, gradually increasing the number of clusters until a significantly-populated cluster is split into two similar clusters, or a poorly populated (outlier) cluster is created.
 Namely, our iterative heuristic consists in running \CL over an increasing number $\NumClust$ of clusters, starting with $\NumClust=2$, until the following \Def{stopping criterion} is met: 1) Two significantly populated clusters have associated centroids which are highly similar; or 2) Centroid structures of significantly populated clusters from the previous iteration are highly similar to those of the current iteration.}
 
-\paragraph{Filtering the promising conformer(s).} Finally, we \Def{elect the most promising cluster(s)}, and return their centroid(s).
+\paragraph{Filtering the promising conformer(s).} Finally, we \Def{select the most promising cluster(s)}, and return their centroid(s).
 While the final number of clusters may potentially be large, only a handful of clusters are expected to represent structures that \RevA{are} both stable, and supported by a large number of conditions.
@@
 \paragraph{Cordero dataset.} 
-In order to test the integration of different probing sources, we consider a dataset introduced in Cordero\etal\cite{Cordero2012}, consisting of $6$ RNAs with known structures, for which reactivities are available for three reagents: {\tt NMIA-\SH, DMS} and {\tt CMCT}.
 Probing data were downloaded from the RMDB~\citep{Cordero2012a} on July 2017.
@@
 CMCT) probing, we restricted reactivities to positions featuring nucleotides {\sf A} and {\sf C} (resp.
-{\sf G} and {\sf U}), setting the reactivity to 0 for other positions.
+{\sf G} and {\sf U}), setting the \RevB{pseudo-energy term to 0 kcal.mol$^{\text{-1}}$} for other positions.
 This allowed to decrease the noise generated by reactivities associated with non-targeted nucleotides, leading to more accurate predictions (data not shown).
@@
 As a preliminary experiment, we verified that \didy{} generally adopts a single global architecture.
-To this end, we subjected \didy{} to a standard denaturation protocol (80\degree{}C for 2 min in H$_\text{2}$O, addition of 40 mM of HEPES at 7.5 pH, 100 mM of KCl, 5 mM of MgCl$_\text{2}$, followed by 10 min at room temperature, and 10 min at 37\degree{}C), and observed the production of a single band on a non-denaturing PAGE, strongly suggesting the adoption of a single conformation.
+To this end, we subjected \didy{} to a \RevB{standard denaturation/renaturation protocol} (80\degree{}C for 2 min in H$_\text{2}$O, addition of 40 mM of HEPES at 7.5 pH, 100 mM of KCl, 5 mM of MgCl$_\text{2}$, followed by 10 min at room temperature, and 10 min at 37\degree{}C), and observed the production of a single band on a non-denaturing PAGE, strongly suggesting the adoption of a single conformation.
 
@@
 \subsection*{Assessment of statistical significance}
-Following Xu\etal\cite{Xu2012}, we assessed the statistical significance of observed changes in predictive performances using a two-tailed Student $t$-test (unequal variance) with usual significance threshold $\alpha=5\%$.
+Following Xu\etal\cite{Xu2012}, we assessed the statistical significance of observed changes in predictive performances using a two-tailed Student $t$-test (unequal variance) with \RevB{usual type I error rate} $\alpha=5\%$.
 When comparing performances within a given dataset, we used the paired version of the test.
@@
 %* Multi conformation
-We assessed the predictive capacities of \OurTool in a classic setting where a single probing condition is available, and compared its performances against \Software{RSample} software recently introduced by Spacic\etal\cite{Spasic2017}, \Software{RSample} relies on a sampling/clustering method developed independently from the current work.
+We assessed the predictive capacities of \OurTool in a classic setting where a single probing condition is available, and compared its performances against \Software{RSample} software recently introduced by Spasic\etal\cite{Spasic2017}, \Software{RSample} relies on a sampling/clustering method developed independently from the current work.
 It was shown to perform \RevA{comparably as} a comprehensive collection of state-of-the-art methods in probing-guided structure prediction, including \Software{RME}~\citep{Wu2015}, \Software{RNAprob}~\citep{Deng2016}, \Software{RNAprobing}~\citep{Washietl2012}, \Software{RNAsc}~\citep{Zarringhalam2012} and the \Software{fold} utility from the \Software{RNAStructure} suite~\citep{Reuter2010}.
@@
 \RevB{%
-\OurTool{} appears to perform comparably, or even slightly favorably, to \Software{RSample}, reporting more accurate structures than its competitor for roughly two thirds of the tested RNAs, while averaging $80.1\%$ GM.
-However, the  improvement is only of around 1.5\% on average over the \citet{Hajdin2013} dataset, an improvement that is no longer significant according to a two-tailed paired T-test.
+\OurTool{} appears to perform similarly, or even slightly favorably in comparison with \Software{RSample}.
+It averages 80.1\% GM, and reports more accurate structural models than its competitor for 14 RNAs in the Hajdin dataset \cite{Hajdin2013}, returns structures equally as good on 2 RNAs, and is dominated by \Software{RSample} for 8 RNAs.
+However, its modest average improvement of approximately 1.5\% over the dataset is not statistically significant, according to a two-tailed paired T-test.
+Those results support the notion that \OurTool represents a competitive alternative to state-of-the-art methods for single reactivity profiles.
 }
@@
 
-\paragraph{Considering three conditions yields improvement over the average performance of the triplet.}
+\paragraph{Considering three conditions yields improvement over the average of mono conditions within the same triplet.}
 Increasing the number of conditions to three boosted the average MCC to 72.3\% and even reached 76.5\% over triplets that do not contain conditions from cluster \Bull{B}.
@@
 To the best of our knowledge, \OurTool currently represents the only available method for a joint automated consideration of multiple probing profiles. 
-Surprisingly, it also appears to produce the most accurate secondary structure models when informed by a single condition, as demonstrated by our reanalysis on the Hajdin\etal\cite{Hajdin2013} benchmark.
+%Surprisingly, it also appears to produce the most accurate secondary structure models when informed by a single condition, as demonstrated by our reanalysis on the Hajdin\etal\cite{Hajdin2013} benchmark.
 We attribute this unexpected result to the adoption of a sampling/clustering approach, \RevA{which appears to eliminate spurious outliers from the conformational space even when a single conformation dominates.}
-\RevA{In this context, the good performance of \OurTool in comparison with \Software{RSample}~\cite{Spasic2017} may appear surprising.
+%\RevA{In this context, the good performance of \OurTool in comparison with \Software{RSample}~\cite{Spasic2017} may appear surprising.
 Indeed, while independently developed and used~\cite{Deforges2017}, both methods share a common sampling/clustering scheme.
@@
 The MEA structure coincides with the centroid of the whole Boltzmann ensemble, suggesting that the increased performances of \OurTool could be at least partially attributed to the use of sampling, and to our adaptive choice of the number of clusters.}
-Moreover, even in cases where a single dominant conformation is expected, considering multiple clusters seems to create more opportunities for discarding noisy random unstable conformations.
+%Moreover, even in cases where a single dominant conformation is expected, considering multiple clusters seems to create more opportunities for discarding noisy random unstable conformations.
 Such conformations naturally occur in a Boltzmann-distributed subset of structures, and may end up \emph{polluting} the centroid structures by artificially inflating the cluster diversity.
 %The impact on \Software{RSample} predictions of such a \emph{noise} seems worsened by the fact that it returns the MEA, the centroid of the whole ensemble, whenever a dominant conformation is detected.
-
 Our analyses reveal that integrating multiple profiles of at least three conditions allows to mitigate the effect of outlier conditions and, at times, to produce better secondary structure models than the one inferred only from the most accurate profile.

### 2020-07-02 Response-NAR-Round2.tex

@@
 
-\tableofcontents
+%\tableofcontents
 
-\section{General comments}
+%\section{Specific responses}
 
-\section{Specific comments}
-
-	\subsection{Reviewer \#2}
+	\subsection{Reviewer \#4}
 	\begin{enumerate}[resume]
@@
 
-- Can the authors reconcile these discrepancies, or alternatively, am I missing something in the way this benchmark was set up?
+Can the authors reconcile these discrepancies, or alternatively, am I missing something in the way this benchmark was set up? }
 
-Please note the following:
-A. I did not check all the RNAs benchmarked by Spasic et al. so there might be additional discrepancies. 
-B. Reported values were consistent between the two works for some RNAs which I did review. 
-C. Given these discrepancies, I am not sure what to conclude from the performances reported for those sequences benchmarked in this manuscript but not by Spasic et al.
+    \Answer{We had indeed spotted these discrepancies, which we had initially attributed to code differences between the versions of \Software{RSample} used in the initial benchmark, and the one currently available online.
 
-Maybe the predictions of these RNAs are very sensitive to the thermodynamic model parameters, I don’t know.
-If this is indeed the case, it is a good idea to benchmark IPANEMAP against more than a single method, similarly to what Spasic et al.
-did.
-I think this creates a more balanced picture of current structure prediction capabilities.}
+	Following this reviewer's remark, we have corresponded with the authors of \Software{RSample}, and discovered the origin of the discrepancy.
+Namely, we had made the erroneous assumption that the format used to specify reactivities in \Software{RSample} was similar to the one used in the \Software{Vienna} package.
+Instead, the reactivity file format used in \Software{RSample} requires three columns per position, and we discovered that \Software{RSample} silently fails if a poorly-formatted file is provided.
+In this case, it \Def{reverts to a reactivity-free prediction} (essentially, an MEA prediction).
+It follows that the results reported for \Software{RSample} in the initial versions of the manuscript were those of a purely thermodynamic prediction, and did not reflect \Software{RSample}'s  performances when informed by reactivity data.
 
-    \Answer{We are greatly indebted to this reviewer for his/her}
+	When informed by suitably-formatted reactivities files, \Software{RSample} now predicts structures having accuracy that are much closer to those reported in \citet{Spasic2017}.
+\OurTool still appears to outperform \Software{RSample} slightly, reporting more accurate structures than its competitor for roughly two thirds of the tested RNAs.
+However, its average improvement in term of Geometric Mean is only of around 1.5\% over the \citet{Hajdin2013} dataset, an improvement that is no longer significant according to a two-tailed T-test with paired data.
+Upon, multiple executions of \Software{RSample}, we have witnessed minor variations, but always a slightly inferior (yet insignificant) average GM that the one of \OurTool.
+
+	We have replaced the erroneous content in our revised version, including the above Figure, and have toned down the relevant subsection to reflect our new assessment.
+Similarly, we have removed from the conclusion our discussion on the possible origins of the discrepancy in performances.
+
+	We are greatly indebted to this reviewer for following through on this point, allowing us to fix our initial (involuntary) misrepresentation of the predictive capacities of \Software{RSample}.
+
+	\begin{figure}[t]
+
+	{\centering \includegraphics[width=.8\textwidth]{graphs/RsampleVsIPANEMAP/Accuracy.pdf}\\}
+
+	\caption{Updated figure presenting the geometric means of structures predicted by \OurTool (green) and \Software{RSample} (orange -- corrected).}
+	\end{figure}}
 
@@
 
-	\Answer{Actually, in addition to being wrong regarding the benchmark, we expressed ourselves poorly in this instance: Having done more than our fair share of reviewing and editorial activities...}
+	\Answer{In addition to erring in our comparison against \Software{RSample}, we also expressed ourselves poorly in this instance.
+It is, of course, entirely legitimate to request more tests and benchmarks and, possibly (but not ideally) to express a negative opinion of our work (peer-review is a pillar of the academic world).
+We merely wished to express disagreement with what we interpreted as a request to demonstrate domination over \Software{RSample}.
+Our resistance was essentially inspired by the almost anecdotal nature of the performances of \OurTool{} in the single probing setting, here presented as a sanity check rather than a central point of our study.}
 	\end{enumerate}
 
-\section{Referee: 3}
+	\subsection{Reviewer \#3}
 
-	\subsection{Minor Revisions:}
 		\begin{enumerate}[resume]
@@
 
-    \Answer{WE set the pseudo-energy to 0 (and not the reactivity, as erroneously stated in the manuscript). \TODO{correct}}
+    \Answer{Indeed, we set the pseudo-energy to 0 (and not the reactivity, as erroneously stated in the manuscript). This is now corrected in the revised manuscript.}
 
@@
 
-    \Answer{Indeed, real denomination is denaturation/renaturation protocol, currently the standard nomenclature. \TODO{correct}}
+    \Answer{Indeed, our denomination was arguably lacking in precision.
+While the term "annealing" would be entirely acceptable, we prefer to use "denaturation/renaturation protocol", as it currently represents the standard nomenclature.}
+
 	\Comment{On page 5, in statistical significance, “usual significance threshold” should really state the “usual type I error rate”.
@@
 
-    \Answer{\TODO{correct}}
+    \Answer{We thank the reviewer for suggesting this more precise formulation, which we adopted in this revision.}
+
 	\Comment{On page 8, “barely significant” should just be changed to “but significant”.
@@
 
-    \Answer{\TODO{be polite}}
+    \Answer{We respectfully disagree with this statement: an observation that would be deemed significant for $\alpha=5\%$, but not $\alpha=1\%$ should arguably be described as more weakly supported, on a statistical level, than an observation supported for both type I error rates.
+This is reflected by the qualifier "barely" in this context.}
+
 	\Comment{On page 10, “Considering three conditions yields improvement over the average performance of the triplet” confuses me.
 I think what was intended is “Considering three conditions yields improvement over the average performance of the doublet [/pair]” }
-    \Answer{\TODO{The average performance of prediction informed by individual conditions is lower than that of predictions simultaneously informed by the three conditions}}
 
-	\subsection{Typos:}
+    \Answer{The main result of this section is that, when considering a triplet of conditions, the average performance of prediction informed by individual conditions, is lower than that of predictions simultaneously informed by the three conditions.
+We modified the title of this subsection to "Considering three conditions yields improvement over the average of single conditions in the same triplet", hoping this will be more explicit to the reader.}
+	\end{enumerate}
+
+	\subsubsection{Typos:}
+
+	\begin{enumerate}[resume]
 	\Comment{On page 1, “different dynamics, that can help” should be “different dynamics that can help”.}
 
-	\Answer{}
+	\Answer{Fixed}
 	\Comment{On page 3, “using an iterative heuristics” should be “using an iterative heuristic”.  Also, “we elect the most promising cluster” should be “we select the most promising cluster”.}
 
-	\Answer{}
+	\Answer{Fixed}
 	\Comment{On page 6, “Spacic” should be “Spasic”.}
 
-	\Answer{}
+	\Answer{Fixed}
+	\end{enumerate}
 
-	Suggestions:
+\subsubsection{Suggestions:}
+
+\begin{enumerate}[resume]
 	\Comment{On page 1, there is a new phrase about probes “some of which are usable in vivo”.
@@
 
-	\Answer{\TODO{Ok Paillard PMID 15355993}}
+	\Answer{We added a reference to this work, and to a posterior effort more similar to our context of analysis.}
 	\Comment{Page 1 claims that using multiple data sources has never been fully automated.
@@
 
-	\Answer{\TODO{}}
+	\Answer{We qualified this statement by mentioning explicitly, and adding a reference to the soft constraint framework.
+
+		However, we feel the statement holds more generally as far as the automation is concerned, since the hard constraint framework makes it virtually impossible to integrate multiple type of probing data in an automated way.
+Indeed, hard constraints are obtained from reactivities based on cutoffs, whose values are somewhat arbitrary.
+Overly liberal cutoffs will often induce contradictory constraints (i.e.
+no compatible structure) across experiments, while conservative cutoffs will lead to sparse constraints, leading to a waste of most of the probing-derived information.
+
+		In fact the tediousness, and suboptimality, of manually selecting those cutoffs while modeling HIV-1 structural elements~\citep{Deforges2017}, was one of the main motivation for this work.}
 

### 2020-07-02 SuppMat.tex
Updates from Overleaf

@@
  	\toprule
- 	RNA ID & Description                      \\ \midrule
- 	A      & Pre-Q1 riboswitch, B. subtilis      & 	M      &SARS corona virus pseudoknot       \\
+ 	RNA ID & Description                          &  \\ \midrule
+ 	A      & Pre-Q1 riboswitch, B. subtilis       & M & SARS corona virus pseudoknot          \\
  	B      & 5’ domain of 16S rRNA, E. coli       & N & 5S rRNA, E. coli                      \\
@@
  	K      & Adenine riboswitch, V. vulnificus    & W & Lysine riboswitch, T. maritima        \\
- 	L      & Fluoride riboswitch, P. syringae     & 	X      & M-Box riboswitch, B. subtilis          \\
- 	\bottomrule
+ 	L      & Fluoride riboswitch, P. syringae     & X & M-Box riboswitch, B. subtilis         \\ \bottomrule
  \end{tabular}
@@
 {RNA} & Length & Sensitivity & PPV    & \multicolumn1c{GM} & Sensitivity & PPV    & \multicolumn1c{GM} \\ \midrule
-				A     & 34     & 0.625       & 1      & 0.791              & 0.625       & 1      & 0.791              \\
-				B     & 530    & 0.6622      & 0.6203 & 0.641              & 0.8986      & 0.8062 & 0.851              \\
-				C     & 301    & 0.03        & 0.0313 & 0.031              & 0.6         & 0.6    & 0.6                \\
-				D     & 214    & 0.6349      & 0.6667 & 0.651              & 0.7302      & 0.8519 & 0.789              \\
-				E     & 500    & 0.6382      & 0.7823 & 0.707              & 0.5263      & 0.5517 & 0.539              \\
-				F     & 47     & 0.4         & 0.75   & 0.548              & 0.6         & 0.75   & 0.671              \\
-				G     & 118    & 0.7179      & 0.8    & 0.758              & 0.7692      & 0.8333 & 0.801              \\
-				H     & 76     & 0.4762      & 0.5556 & 0.514              & 0.7619      & 0.7619 & 0.762              \\
-				I     & 155    & 0.6786      & 0.8261 & 0.749              & 0.9821      & 0.9821 & 0.982              \\
-				J     & 79     & 0.7727      & 0.85   & 0.81               & 0.9545      & 0.875  & 0.914              \\
-				K     & 71     & 1           & 1      & 1                  & 1           & 0.9545 & 0.977              \\
-				L     & 66     & 0.5625      & 0.6923 & 0.624              & 0.625       & 0.7143 & 0.668              \\
-				M     & 79     & 0.6538      & 0.8095 & 0.727              & 0.6538      & 0.68   & 0.667              \\
-				N     & 120    & 0.2857      & 0.2632 & 0.274              & 0.9714      & 0.9189 & 0.945              \\
-				O     & 97     & 0.9286      & 0.9286 & 0.929              & 0.9286      & 0.8966 & 0.912              \\
-				P     & 511    & 0.9328      & 0.7762 & 0.851              & 0.8908      & 0.7681 & 0.827              \\
-				Q     & 401    & 0.5826      & 0.5583 & 0.57               & 0.7304      & 0.75   & 0.74               \\
-				R     & 425    & 0.8397      & 0.8271 & 0.833              & 0.9084      & 0.8815 & 0.895              \\
-				S     & 336    & 0.5481      & 0.6129 & 0.58               & 0.8077      & 0.866  & 0.836              \\
-				T     & 473    & 0.7986      & 0.7143 & 0.755              & 0.8819      & 0.8194 & 0.85               \\
-				U     & 412    & 0.5379      & 0.6174 & 0.576              & 0.75        & 0.8462 & 0.797              \\
-				V     & 75     & 0.619       & 0.6842 & 0.651              & 0.619       & 0.5652 & 0.591              \\
-				W     & 174    & 0.8095      & 0.8644 & 0.836              & 0.8095      & 0.9808 & 0.891              \\
-				X     & 154    & 0.875       & 0.913  & 0.894              & 0.875       & 0.913  & 0.894              \\ \bottomrule
+				A     & 34     &0.625& 1.000& 0.791& 0.625       & 1      & 0.791              \\
+				B     & 530    &0.818& 0.747& 0.781& 0.8986      & 0.8062 & 0.851              \\
+				C     & 301    &0.590& 0.590& 0.590& 0.6         & 0.6    & 0.6                \\
+				D     & 214    &0.746& 0.870& 0.806& 0.7302      & 0.8519 & 0.789              \\
+				E     & 500    &0.487& 0.503& 0.495& 0.5263      & 0.5517 & 0.539              \\
+				F     & 47     &0.600& 0.750& 0.671& 0.6         & 0.75   & 0.671              \\
+				G     & 118    &0.795& 0.912& 0.851& 0.7692      & 0.8333 & 0.801              \\
+				H     & 76     &0.952& 0.952& 0.952& 0.7619      & 0.7619 & 0.762              \\
+				I     & 155    &0.946& 0.964& 0.955& 0.9821      & 0.9821 & 0.982              \\
+				J     & 79     &0.636& 0.583& 0.609& 0.9545      & 0.875  & 0.914              \\
+				K     & 71     &1.000& 1.000& 1.000& 1           & 0.9545 & 0.977              \\
+				L     & 66     &0.625& 0.667& 0.646& 0.625       & 0.7143 & 0.668              \\
+				M     & 79     &0.654& 0.654& 0.654& 0.6538      & 0.68   & 0.667              \\
+				N     & 120    &0.857& 0.833& 0.845& 0.9714      & 0.9189 & 0.945              \\
+				O     & 97     &0.929& 0.929& 0.929& 0.9286      & 0.8966 & 0.912              \\
+				P     & 511    &0.908& 0.783& 0.843& 0.8908      & 0.7681 & 0.827              \\
+				Q     & 401    &0.670& 0.726& 0.697& 0.7304      & 0.75   & 0.74               \\
+				R     & 425    &0.817& 0.811& 0.814& 0.9084      & 0.8815 & 0.895              \\
+				S     & 336    &0.798& 0.865& 0.831& 0.8077      & 0.866  & 0.836              \\
+				T     & 473    &0.889& 0.815& 0.851& 0.8819      & 0.8194 & 0.85               \\
+				U     & 412    &0.485& 0.533& 0.508& 0.75        & 0.8462 & 0.797              \\
+				V     & 75     &1.000& 1.000& 1.000& 0.619       & 0.5652 & 0.591              \\
+				W     & 174    &0.810& 0.895& 0.851& 0.8095      & 0.9808 & 0.891              \\
+				X     & 154    &0.875& 0.913& 0.894& 0.875       & 0.913  & 0.894              \\ \bottomrule& & & 
 			\end{tabular}		
@@
 
-\begin{figure}	
+\begin{figure}[p]
 	{\centering \SR{1M7_Ilu_d1_Mg.pdf}{1    }{\Bull{C}  }{\OneMSevILUMg  }
@@
 \end{figure}
-\begin{figure}
+\begin{figure}[p]
 	\SR{1M7_Mg_CE.pdf}{9    }{\Bull{E}  }{\OneMSevMgCE       }
@@
 \newcommand{\MyScale}{.26}
-
+\newpage
 \subsection{Correlation analysis of raw reactivities profiles}

### 2020-07-03 IPANEMAP-2019.tex
Updates from Overleaf

@@
 
-\noindent We introduce \OurTool{}, an automated method for the modeling of RNA structure from multiple probing reactivity profiles.
+\noindent We introduce \OurTool{}, \RevB{the first} automated method for the modeling of RNA structure from multiple probing reactivity profiles.
 Input profiles can result from experiments based on diverse protocols, reagents, \RevA{or collection of variants}, and are jointly analyzed to predict the dominant conformations of an RNA.
@@
 
-To the best of our knowledge, \OurTool currently represents the only available method for a joint automated consideration of multiple probing profiles. 
-%Surprisingly, it also appears to produce the most accurate secondary structure models when informed by a single condition, as demonstrated by our reanalysis on the Hajdin\etal\cite{Hajdin2013} benchmark.
-We attribute this unexpected result to the adoption of a sampling/clustering approach, \RevA{which appears to eliminate spurious outliers from the conformational space even when a single conformation dominates.}
-%\RevA{In this context, the good performance of \OurTool in comparison with \Software{RSample}~\cite{Spasic2017} may appear surprising.
-Indeed, while independently developed and used~\cite{Deforges2017}, both methods share a common sampling/clustering scheme.
-However, \Software{RSample}~\cite{Spasic2017} reverts to a MEA~\cite{Lu2009} computation whenever a single conformation is suspected.
-The MEA structure coincides with the centroid of the whole Boltzmann ensemble, suggesting that the increased performances of \OurTool could be at least partially attributed to the use of sampling, and to our adaptive choice of the number of clusters.}
-%Moreover, even in cases where a single dominant conformation is expected, considering multiple clusters seems to create more opportunities for discarding noisy random unstable conformations.
-Such conformations naturally occur in a Boltzmann-distributed subset of structures, and may end up \emph{polluting} the centroid structures by artificially inflating the cluster diversity.
-%The impact on \Software{RSample} predictions of such a \emph{noise} seems worsened by the fact that it returns the MEA, the centroid of the whole ensemble, whenever a dominant conformation is detected.
 Our analyses reveal that integrating multiple profiles of at least three conditions allows to mitigate the effect of outlier conditions and, at times, to produce better secondary structure models than the one inferred only from the most accurate profile.
@@
 In other words, multiple experimental probing profiles, even combined in a pairwise fashion, \RevA{can be used jointly to mitigate the empirical risk of a misprediction}.
+
+To the best of our knowledge, \OurTool currently represents the only available method for a joint automated consideration of multiple probing profiles.
+\RevB{While hard constraints~\cite{Mathews2004} could in principle support multiple sources of probing data, their derivation from reactivities usually require a careful choice of cutoffs, making their automation virtually impossible.
+Overly liberal cutoffs will often induce contradictory constraints (i.e.
+no compatible structure) across conditions, while conservative cutoffs will lead to sparse constraints, practically wasting most of the probing-derived information.
+In fact the tediousness, and suboptimality, of the manual determination of those cutoffs while modeling HIV-1 structural elements~\citep{Deforges2017}, was one of the main motivation behind the present work.}
 
@@
 \section*{Acknowledgements}
-The authors wish to express their gratitude to Ronny Lorenz for developing a probing-informed version of \Software{RNAsubopt}, allowing to sample the pseudo-Boltzmann ensemble, Nathalie Chamond for helpful suggestions and Benoit Masquida for the gift of \didy{} plasmid and helpful discussions about protocols and \didy{} structure .
+The authors wish to express their gratitude to: Ronny Lorenz for developing a probing-informed version of \Software{RNAsubopt}, allowing to sample the pseudo-Boltzmann ensemble; Nathalie Chamond for helpful suggestions; Benoit Masquida for the gift of \didy{} plasmid and helpful discussions about protocols and \didy{} structure; \RevB{and an anonymous reviewer for helping us detect an error in our execution of \Software{RSample}}.
 

### 2020-07-03 Response-NAR-Round2.tex

@@
 %\section{Specific responses}
+The authors apologize for the delays in this revision, induced by the ongoing COVID pandemic, which has significantly disrupted the daily operation of our respective academic institutions.
+They also wish to express their gratitude for the reviewers, whose remaining remarks have allowed us to greatly improve the quality of the manuscript.
 
@@
 
-	We are greatly indebted to this reviewer for following through on this point, allowing us to fix our initial (involuntary) misrepresentation of the predictive capacities of \Software{RSample}.
+	We are {\bfseries greatly indebted} to this reviewer for following through on this point, allowing us to fix our initial (obviously involuntary) misrepresentation of the predictive capacities of \Software{RSample}, .
 
@@
 	\Answer{In addition to erring in our comparison against \Software{RSample}, we also expressed ourselves poorly in this instance.
-It is, of course, entirely legitimate to request more tests and benchmarks and, possibly (but not ideally) to express a negative opinion of our work (peer-review is a pillar of the academic world).
-We merely wished to express disagreement with what we interpreted as a request to demonstrate domination over \Software{RSample}.
-Our resistance was essentially inspired by the almost anecdotal nature of the performances of \OurTool{} in the single probing setting, here presented as a sanity check rather than a central point of our study.}
+It is, of course, entirely legitimate to request more tests and benchmarks and, possibly (albeit not ideally ;) ) to express a negative opinion of our work, peer-review being on of the pillars of the academic world.
+We merely wished to express disagreement with what we interpreted as a request to stress a dominating behavior in comparison with \Software{RSample}.
+Our resistance was mainly inspired by the almost anecdotal nature of the good performances of \OurTool{} in the single probing setting.
+This benchmark was presented as a sanity check rather than a central point of our study, as \OurTool{} has been developed to handle multiple probing.}
 	\end{enumerate}
 
+\newpage 
 	\subsection{Reviewer \#3}
@@
     \Answer{We respectfully disagree with this statement: an observation that would be deemed significant for $\alpha=5\%$, but not $\alpha=1\%$ should arguably be described as more weakly supported, on a statistical level, than an observation supported for both type I error rates.
-This is reflected by the qualifier "barely" in this context.}
+This, in our opinion, is accurately reflected by the qualifier "barely" in this context.}
 
@@
 
-    \Answer{The main result of this section is that, when considering a triplet of conditions, the average performance of prediction informed by individual conditions, is lower than that of predictions simultaneously informed by the three conditions.
+    \Answer{The main result of this subsection is that, when considering a triplet of conditions, the average performance of prediction informed by individual conditions, is lower than that of predictions simultaneously informed by the three conditions.
 We modified the title of this subsection to "Considering three conditions yields improvement over the average of single conditions in the same triplet", hoping this will be more explicit to the reader.}
@@
 
-	\Answer{We qualified this statement by mentioning explicitly, and adding a reference to the soft constraint framework.
+	\Answer{We qualified this statement by mentioning explicitly the soft constraint framework, and adding a reference to one of its first generic implementation.
 
 		However, we feel the statement holds more generally as far as the automation is concerned, since the hard constraint framework makes it virtually impossible to integrate multiple type of probing data in an automated way.
-Indeed, hard constraints are obtained from reactivities based on cutoffs, whose values are somewhat arbitrary.
+Indeed, \emph{hard constraints} are obtained from reactivities based on cutoffs, whose values are somewhat arbitrary and typically need to be calibrated for each experiment.
 Overly liberal cutoffs will often induce contradictory constraints (i.e.
